\fancyhead[RO,LE]{\textit{None}}
\fancyhead[RE,LO]{}
\section{Hour of None}
\elcol{℣. O God, {\ding{64}} make speed to save us.}{℣. Deus, {\ding{64}} in adjutórium nostrum inténde.}
\elcol{℟. O Lord, make haste to help us.}{℟. Dómine, ad adjuvándum nos festína.}
\elcol{℣. Glory be to the Father, and to the Son, and to the Holy Ghost.}{℣. Glória Patri, et Fílio, * et Spirítui Sancto:}
\elcol{℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.}{℟. Sicut erat in princípio, et nunc, et semper, * et in sǽcula s{\ae}culórum. Amen.}
\elcol{℣. Praise ye the Lord.}{℣. Laudáte Dominum.}
\elcol{℟. The Lord's Name be praised.}{℟. Sit Nomen Dómini Benedíctum.}
\subby{Rerum Deus}
\elcol{\begin{hangparas}{1.25em}{1}
\lett{O}{God,} creation’s secret force,\par
\secondline{Thyself unmoved, all motion's source,}
%\thirdline{Who from the morn till evening ray,}
\par\noindent
Who from the morn till evening ray,\par\noindent
Through all its changes guid'st the day:\\
\par\noindent
Grant us, when this short life is past,\par\noindent
The glorious evening that shall last;\par\noindent
That, by a holy death attained,\par\noindent
Eternal glory may be gained.\\
\par\noindent
O Father, that we ask be done,\par\noindent
Through Jesus Christ, thine only Son;\par\noindent
Who, with the Holy Ghost and thee,\par\noindent
Doth live and reign eternally. Amen.\\
\end{hangparas}}
{\lett{R}{erum,} Deus, tenax vigor,\par
\secondline{Immótus in te pérmanens,}
\thirdline{Lucis diúrnæ témpora}
Succéssibus detérminans:\\
\par
Largíre lumen véspere,\par
Quo vita nusquam décidat,\par
Sed prǽmium mortis sacr{\ae}\par
Perénnis instet glória.\\
\par
Præsta, Pater piíssime,\par
Patríque compar Únice,\par
Cum Spíritu Paráclito\par
Regnans per omne. Amen.\\}

\subbypsalm{Psalm 126. \textit{In convertendo}}

\lett{W}{hen} the Lord turned again the captivity of Sion : then were we like unto them that dream.\par
\secondline{\psanum{2}Then was our mouth filled with laughter : and our tongue with joy.}
\thirdline{\psanum{3}Then said they among the heathen : The Lord hath done great things for them.}
\psanum{4}Yea, the Lord hath done great things for us already : whereof we rejoice.\par
\psanum{5}Turn our captivity, O Lord : as the rivers in the south.\par
\psanum{6}They that sow in tears : shall reap in joy.\par
\psanum{7}He that now goeth on his way weeping, and beareth forth good seed : shall doubtless come again with joy, and bring his sheaves with him.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.

\subbypsalm{Psalm 127. \textit{Nisi Dominus}}

\lett{E}{xcept} the Lord build the house : their labour is but lost that build it.\par
\secondline{\psanum{2}Except the Lord keep the city : the watchman waketh but in vain.}
\thirdline{\psanum{3}It is but lost labour that ye haste to rise up early, and so late take rest, and eat the bread of carefulness : for so he giveth his beloved sleep.}
\psanum{4}Lo, children and the fruit of the womb : are an heritage and gift that cometh of the Lord.\par
\psanum{5}Like as the arrows in the hand of the giant : even so are the young children.\par
\psanum{6}Happy is the man that hath his quiver full of them : they shall not be ashamed when they speak with their enemies in the gate.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.

%MANUAL ADJUSTMENT:
\vspace{-1ex}

\subbypsalm{Psalm 128. \textit{Beati omnes}}

\lett{B}{lessed} are all they that fear the Lord : and walk in his ways.\par
\secondline{\psanum{2}For thou shalt eat the labours of thine hands : O well is thee, and happy shalt thou be.}
\thirdline{\psanum{3}Thy wife shall be as the fruitful vine : upon the walls of thine house.}
\psanum{4}Thy children like the olive-branches : round about thy table.\par
\psanum{5}Lo, thus shall the man be blessed : that feareth the Lord.\par
\psanum{6}The Lord from out of Sion shall so bless thee : that thou shalt see Jerusalem in prosperity all thy life long.\par
\psanum{7}Yea, that thou shalt see thy children's children : and peace upon Israel.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.

%MANUAL ADJUSTMENT:
\vspace{-1ex}

\subby{Chapter}
\begin{inhead}
	Ferias Throughout the Year
\end{inhead}
\elcoldent{Bear ye one another's burdens, and so fulfil the law of Christ.}{Alter alterius ónera portáte, et sic adimplébitis legem Christi.}
\begin{inhead}
	Sundays \& Feast Days
\end{inhead}
\elcoldent{One Lord, one Faith, one Baptism, one God and Father of all, who is above all, and through all, and in you all.}{Unus Dóminus, una fides, unum baptisma. Unus Deus et Pater ómnium, qui est super omnes, et per ómnia, et in ómnibus nobis.}
\begin{inhead}
	Eastertide
\end{inhead}
\elcoldent{There are three that bear witness in earth: the Spirit, and the Water, and the Blood: and these three agree in one.}{Et tres sunt, qui testimónium dant in terra: spíritus, et aqua, et sanguis: et hi tres unum sunt.}
\elcol{℟. Thanks be to God.}{℟. Deo grátias.}
\subby{Short Respond}
\begin{inhead}
	Ferias Throughout the Year
\end{inhead}
\elcol{℣. Deliver me, O Lord, and be merciful unto me.}{℣. Rédime me, et miserére mei.}

\elcol{℟. Deliver me, O Lord, and be merciful unto me.}{℟. Rédime me, et miserére mei.}

\elcol{℣. My foot standeth right; I will praise the Lord in the congregations.}{℣. Pes meus stetit in dirécto: in ecclésiis benedícam te, Dómine.}

\elcol{℟. And be merciful unto me.}{℟. Et miserére mei.}

\elcol{℣. Glory be to the Father, and to the Son, and to the Holy Ghost.}{℣. Glória Patri, et Fílio, et Spirítui Sancto.}
    
\elcol{℟. Deliver me, O Lord, and be merciful unto me.}{℟. Rédime me, et miserére mei.}

\elcol{℣. I will praise thy Name, O Lord.}{℣. Et confitébor nómini tuo, Dómine.}

\elcol{℟. For thou hast delivered me out of all my trouble.}{℟. Quóniam ex omni tribulatióne eripuísti me.}

\begin{inhead}
	Sundays, Feasts, \& in Eastertide
\end{inhead}

\elcol{℣. I call with my whole heart; hear me, O Lord (alleluia, alleluia).}{℣. Clamávi in toto corde meo, exáudi me, Dómine (allelúja, allelúja).}

\elcol{℟. I call with my whole heart; hear me, O Lord (alleluia, alleluia).}{℟. Clamávi in toto corde meo, exáudi me, Dómine (allelúja, allelúja).}

\elcol{℣. I will keep thy statutes.}{℣. Justificatiónes tuas requíram.}

\elcol{℟. Hear me, O Lord (alleluia, alleluia).}{℟. Exáudi me, Dómine (allelúja, allelúja).}

\elcol{℣. Glory be to the Father, and to the Son, and to the Holy Ghost.}{℣. Glória Patri, et Fílio, et Spirítui Sancto.}

\elcol{℟. I will call with my whole heart; hear me, O Lord (alleluia, alleluia).}{℟. Clamávi in toto corde meo, exáudi me, Dómine (allelúja, allelúja).}

\elcol{℣. Cleanse me, O Lord, from my secret faults (alleluia).}{℣. Ab occúltis meis munda me (allelúja).}

\elcol{℟. Keep thy servant also from presumptuous sins (alleluia).}{℟. Et ab aliénis parce servo tuo (allelúja).}

\subby{Suffrages}
\begin{rubric}
	The Suffrages, said kneeling, are said on all Ferias outside of Eastertide. They are omitted on Sundays, Feasts, and within Principle, Primary, and Secondary Octaves.\par
	\textsc{Note,} If omitted, the Office continues with the Collect.
\end{rubric}

\elcol{℣. Lord, have mercy upon us.}{℣. Kýrie, eléison.}
\elcol{℟. Christ, have mercy upon us.}{℟. Christe, eléison.}
\elcol{℣. Lord, have mercy upon us.}{℣. Kýrie, eléison.}
%\subsec{Lord's Prayer}
\needspace{4\baselineskip}
\elcol{\lett{O}{ur} Father, who art in heaven, Hallowed be thy Name. Thy kingdom come. Thy will be done on earth, As it is in heaven. Give us this day our daily bread. And forgive us our trespasses, As we forgive those who trespass against us. And lead us not into temptation; But deliver us from evil. Amen.}
{\lett{P}{ater} noster, qui es in c{\ae}lis, sanctificétur nomen tuum: advéniat regnum tuum: fiat volúntas tua, sicut in c{\ae}lo et in terra. Panem nostrum quotidiánum da nobis hódie: et dimítte nobis débita nostra, sicut et nos dimíttimus debitóribus nostris: et ne nos indúcas in tentatiónem: sed líbera nos a malo. Amen.}

\elcol{℣. Cast me not away in the time of age.}{℣. Ne proícias me in témpore senectútis.}

\elcol{℟. Forsake me not when my strength faileth me.}{℟. Cum defécerit virtus mea, ne derelínquas me.}

\elcol{℣. Hide not thy face from me.}{℣. Non avértas fáciem tuam a me.}

\elcol{℟. Lest I be like unto them that go down into the pit.}{℟. Et símilis ero descendéntibus in lacum.}

\elcol{℣. Quicken me, O Lord, for thy Name's sake.}{℣. Propter nomen tuum, Dómine, vivificábis me.}

\elcol{℟. And for thy righteousness' sake bring my soul out of trouble.}{℟. In {\ae}quitáte tua edúces de tribulatióne ánimam meam}

\subby{Collects}
\elcol{℣. The Lord be with you.\par
\textit{or,} O Lord, hear our prayer.}{℣. Dóminus vobíscum.\par
\textit{vel,} Dómine, exáudi oratiónem nostram.}
\elcol{℟. And with thy spirit.\par
\textit{or,} And let our cry come unto thee.}{℟. Et cum spíritu tuo.\par
\textit{vel,} Et clamor noster ad te véniat.}
\elcol{℣. Let us pray.}{℣. Orémus.}
\begin{rubric}
	The Collect(s) of the Day is here said, followed by this memorial Collect.
\end{rubric}
\needspace{4\baselineskip}
\lett{L}{ord} Jesus Christ, who for our sakes didst tread the paths of death; make known to us the way of life; that as thou wast reckoned with the transgressors in thy death, and with the rich in thy burial, so we, who are dead in trespasses and sins, may be raised up by thee to the land of true riches. Who livest and reignest with the Father and the Holy Ghost, ever one God, world without end. \textit{Amen.}\\

\elcol{℣. The Lord be with you.\par
\textit{or,} O Lord, hear our prayer.}{℣. Dóminus vobíscum.\par
\textit{vel,} Dómine, exáudi oratiónem nostram.}
\elcol{℟. And with thy spirit.\par
\textit{or,} And let our cry come unto thee.}{℟. Et cum spíritu tuo.\par
\textit{vel,} Et clamor noster ad te véniat.}
\elcol{℣. Let us bless the Lord.}{℣. Benedicámus Dómino.}
\elcol{℟. Thanks be to God.}{℟. Deo grátias}
\elcol{℣. May the souls {\ding{64}} of the faithful departed, through the mercy of God, rest in peace.}{℣. Fidélium ánim{\ae} {\ding{64}} per misericórdiam Dei requiéscant in pace.}
\needspace{4\baselineskip}
\lett{N}{ow} unto him that is able to keep us from falling and to present us faultless before the presence of his glory with exceeding joy; to the only wise God our Saviour be glory and majesty, dominion, and power, both now and ever. \textit{Amen.}


\begin{center}
	{\textsc{Here endeth the Order of None.}}
\end{center}