\fancyhead[RO,LE]{\textit{Sext}}
\fancyhead[RE,LO]{}
\section{Hour of Sext}
\elcol{℣. O God, {\ding{64}} make speed to save us.}{℣. Deus, {\ding{64}} in adjutórium nostrum inténde.}
\elcol{℟. O Lord, make haste to help us.}{℟. Dómine, ad adjuvándum nos festína.}
\elcol{℣. Glory be to the Father, and to the Son, and to the Holy Ghost.}{℣. Glória Patri, et Fílio, * et Spirítui Sancto:}
\elcol{℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.}{℟. Sicut erat in princípio, et nunc, et semper, * et in sǽcula s{\ae}culórum. Amen.}
\elcol{℣. Praise ye the Lord.}{℣. Laudáte Dominum.}
\elcol{℟. The Lord's Name be praised.}{℟. Sit Nomen Dómini Benedíctum.}
\subby{Rector Potens}
\elcol{\begin{hangparas}{1.25em}{1}
\lett{O}{God} of truth, O Lord of might,\par
\secondline{Who ord'rest time and change aright,}
\par\noindent
And send'st the early morning ray,\par\noindent
And light'st the glow of perfect day.\\
\par\noindent
Extinguish thou each sinful fire,\par\noindent
And banish every ill desire:\par\noindent
And while thou keep'st the body whole,\par\noindent
Shed forth thy peace upon the soul.\\
\par\noindent
O Father, that we ask be done,\par\noindent
Through Jesus Christ, thine only Son;\par\noindent
Who, with the Holy Ghost and thee,\par\noindent
Doth live and reign eternally. Amen.\\
\end{hangparas}}
{\lett{R}{ector} potens, verax Deus,\par
\secondline{Qui témperas rerum vices,}
\thirdline{Splendóre mane illúminas,}
Et ígnibus merídiem:\\
\par
Exstíngue flammas lítium,\par
Aufer calórem nóxium,\par
Confer salútem córporum,\par
Verámque pacem córdium.\\
\par
Pr{\ae}sta, Pater piíssime,\par
Patríque compar Únice,\par
Cum Spíritu Paráclito\par
Regnans per omne sǽculum. Amen.\\}


\subbypsalm{Psalm 123. \textit{Ad te levavi oculos meos}}
\lett{U}{nto} thee lift I up mine eyes : O thou that dwellest in the heavens.\par
\secondline{\psanum{2}Behold, even as the eyes of servants look unto the hand of their masters, and as the eyes of a maiden unto the hand of her mistress : even}
\noindent
so our eyes wait upon the Lord our God, until he have mercy upon us.\par
\psanum{3}Have mercy upon us, O Lord, have mercy upon us : for we are utterly despised.\par
\psanum{4}Our soul is filled with the scornful reproof of the wealthy : and with the despitefulness of the proud.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.

\subbypsalm{Psalm 124. \textit{Nisi quia Dominus}}
\lett{I}{f} the Lord himself had not been on our side, now may Israel say : if the Lord himself had not been on our side, when men rose up against us;\par
\secondline{\psanum{2}They had swallowed us up quick : when they were so wrathfully displeased at us.}
\thirdline{\psanum{3}Yea, the waters had drowned us : and the stream had gone over our soul.}
\psanum{4}The deep waters of the proud : had gone even over our soul.\par
\psanum{5}But praised be the Lord : who hath not given us over for a prey unto their teeth.\par
\psanum{6}Our soul is escaped even as a bird out of the snare of the fowler : the snare is broken, and we are delivered.\par
\psanum{7}Our help standeth in the Name of the Lord : who hath made heaven and earth.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.

\subbypsalm{Psalm 125. \textit{Qui confidunt}}
\lett{T}{hey} that put their trust in the Lord shall be even as the mount Sion : which may not be removed, but standeth fast for ever.\par
\secondline{\psanum{2}The hills stand about Jerusalem : even so standeth the Lord round about his people, from this time forth for evermore.}
\thirdline{\psanum{3}For the rod of the ungodly cometh not into the lot of the righteous : lest the righteous put their hand unto wickedness.}
\psanum{4}Do well, O Lord : unto those that are good and true of heart.\par
\psanum{5}As for such as turn back unto their own wickedness : the Lord shall lead them forth with the evil-doers; but peace shall be upon Israel.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.

\subby{Chapter}
\begin{inhead}
	Ferias Throughout the Year
\end{inhead}
\elcoldent{Prove all things, hold fast to that which is good: abstain from all appearance of evil.}{Omnia autem probáte: quod bonum est tenéte. Ab omni spécie mala abstinéte vos.}
\begin{inhead}
	Sundays \& Feast Days
\end{inhead}
\elcoldent{There are three that bear record in heaven: The Father, the Word, and the Holy Ghost: and these three are one.}{Quóniam tres sunt, qui testimónium dant in c{\ae}lo: Pater, Verbum, et Spíritus Sanctus: et hi tres unum sunt.}
\begin{inhead}
	Eastertide
\end{inhead}
\elcoldent{Who is he that overcometh the world, but he that believeth that Jesus is the Son of God? This is he that came by water and blood, even Jesus Christ.}{Quis est, qui vincit mundum, nisi qui credit quóniam Jesus est Fílius Dei? Hic est, qui venit per aquam et sánguinem, Jesus Christus.}
\elcol{℟. Thanks be to God.}{℟. Deo grátias.}
\subby{Short Respond}
\begin{inhead}
	Ferias Throughout the Year
\end{inhead}
\elcol{℣. I will bless the Lord at all times.}{℣. Benedícam Dóminum in omni témpore.}

\elcol{℟. I will bless the Lord at all times.}{℟. Benedícam Dóminum in omni témpore.}

\elcol{℣. His praise shall ever be in my mouth.}{℣. Semper laus ejus in ore meo.}
    
\elcol{℟. At all times.}{℟. In omni témpore.}

\elcol{℣. Glory be to the Father, and to the Son, and to the Holy Ghost.}{℣. Glória Patri, et Fílio, et Spirítui Sancto.}
    
\elcol{℟. I will bless the Lord at all times.\\}{℟. Benedícam Dóminum in omni témpore.\\}

\elcol{℣. Behold, God is my helper.}{℣. Ecce enim, Deus ádjuvat me.}

\elcol{℟. The Lord is with them that uphold my soul.}{℟. Et Dóminus suscéptor est ánim{\ae} me{\ae}.}

\begin{inhead}
	Sundays, Feasts, \& in Eastertide
\end{inhead}

\elcol{℣. O Lord, thy Word endureth for ever in heaven (alleluia, alleluia).}{℣. In {\ae}térnum, Dómine, * verbum tuum pérmanet in c{\ae}lo (allelúja, allelúja).}

\elcol{℟. O Lord, thy Word endureth for ever in heaven (alleluia, alleluia).}{℟. In {\ae}térnum, Dómine, * verbum tuum pérmanet in c{\ae}lo (allelúja, allelúja).}

\elcol{℣. Thy truth also remaineth from one generation to another.}{℣. In generatiónem et generatiónem véritas tua.}

\elcol{℟. For ever in heaven (alleluia, alleluia).}{℟. Pérmanet in c{\ae}lo (allelúja, allelúja).}

\elcol{℣. Glory be to the Father, and to the Son, and to the Holy Ghost.}{℣. Glória Patri, et Fílio, et Spirítui Sancto.}
    
\elcol{℟. O Lord, thy Word endureth for ever in heaven (alleluia, alleluia).}{℟. In {\ae}térnum, Dómine, * verbum tuum pérmanet in c{\ae}lo (allelúja, allelúja).}

\elcol{℣. The Lord is my shepherd, therefore can I lack nothing (alleluia).}{℣. Dóminus regit me, et nihil mihi déerit.}

\elcol{℟. He shall feed me in a green pasture (alleluia).}{℟. In loco páscu{\ae} ibi me collocávit (allelúja).}

\subby{Suffrages}
\begin{rubric}
	The Suffrages, said kneeling, are said on all Ferias outside of Eastertide. They are omitted on Sundays, Feasts, and within Principle, Primary, and Secondary Octaves.\par
	\textsc{Note,} If omitted, the Office continues with the Collect.
\end{rubric}

\elcol{℣. Lord, have mercy upon us.}{℣. Kýrie, eléison.}
\elcol{℟. Christ, have mercy upon us.}{℟. Christe, eléison.}
\elcol{℣. Lord, have mercy upon us.}{℣. Kýrie, eléison.}
%\subsec{Lord's Prayer}
\needspace{4\baselineskip}
\elcol{\lett{O}{ur} Father, who art in heaven, Hallowed be thy Name. Thy kingdom come. Thy will be done on earth, As it is in heaven. Give us this day our daily bread. And forgive us our trespasses, As we forgive those who trespass against us. And lead us not into temptation; But deliver us from evil. Amen.}
{\lett{P}{ater} noster, qui es in c{\ae}lis, sanctificétur nomen tuum: advéniat regnum tuum: fiat volúntas tua, sicut in c{\ae}lo et in terra. Panem nostrum quotidiánum da nobis hódie: et dimítte nobis débita nostra, sicut et nos dimíttimus debitóribus nostris: et ne nos indúcas in tentatiónem: sed líbera nos a malo. Amen.}

\elcol{℣. Hide not thy face from thy servant, for I am in trouble.}{℣. Et ne avértas fáciem tuam a púero tuo: quóniam tríbulor.}

\elcol{℟. O haste thee and hear me.}{℟. Velóciter exáudi me.}

\elcol{℣. Draw nigh unto my soul and save it.}{℣. Inténde ánim{\ae} me{\ae}, et líbera eam.}

\elcol{℟. O deliver me, because of mine enemies.}{℟. Propter inimícos meos éripe me.}

\elcol{℣. Arise, O Christ, and help us.}{℣. Exsúrge, Christe, ádjuva nos.}

\elcol{℟. And deliver us for thy Name's sake.}{℟. Et líbera nos propter nomen tuum.}

\subby{Collects}
\elcol{℣. The Lord be with you.\par
\textit{or,} O Lord, hear our prayer.}{℣. Dóminus vobíscum.\par
\textit{vel,} Dómine, exáudi oratiónem nostram.}
\elcol{℟. And with thy spirit.\par
\textit{or,} And let our cry come unto thee.}{℟. Et cum spíritu tuo.\par
\textit{vel,} Et clamor noster ad te véniat.}
\elcol{℣. Let us pray.}{℣. Orémus.}
\begin{rubric}
	The Collect(s) of the Day is here said, followed by this memorial Collect.
\end{rubric}
\needspace{4\baselineskip}
\lett{L}{ord} Jesus, who didst stretch out thine arms of love on the hard wood of the Cross, that all men might come within the reach of thy saving embrace; clothe us in thy Spirit, that we, stretching forth our hands in loving labour for others, may bring those who know thee not to the knowledge and love of thee, who with the Father and the Holy Ghost livest and reignest ever one God, world without end. \textit{Amen.}\\

\elcol{℣. The Lord be with you.\par
\textit{or,} O Lord, hear our prayer.}{℣. Dóminus vobíscum.\par
\textit{vel,} Dómine, exáudi oratiónem nostram.}
\elcol{℟. And with thy spirit.\par
\textit{or,} And let our cry come unto thee.}{℟. Et cum spíritu tuo.\par
\textit{vel,} Et clamor noster ad te véniat.}
\elcol{℣. Let us bless the Lord.}{℣. Benedicámus Dómino.}
\elcol{℟. Thanks be to God.}{℟. Deo grátias}
\elcol{℣. May the souls {\ding{64}} of the faithful departed, through the mercy of God, rest in peace.}{℣. Fidélium ánim{\ae} {\ding{64}} per misericórdiam Dei requiéscant in pace.}
\needspace{4\baselineskip}
\lett{N}{ow} unto the King eternal, immortal, invisible, the only wise God, be honour and glory, for ever and ever. \textit{Amen.}


\begin{center}
	{\textsc{Here endeth the Order of Sext.}}
\end{center}