\fancyhead[RO,LE]{\textit{Terce}}
\fancyhead[RE,LO]{}
\section{Hour of Terce}
\elcol{℣. O God, {\ding{64}} make speed to save us.}{℣. Deus, {\ding{64}} in adjutórium nostrum inténde.}
\elcol{℟. O Lord, make haste to help us.}{℟. Dómine, ad adjuvándum nos festína.}
\elcol{℣. Glory be to the Father, and to the Son, and to the Holy Ghost.}{℣. Glória Patri, et Fílio, * et Spirítui Sancto:}
\elcol{℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.}{℟. Sicut erat in princípio, et nunc, et semper, * et in sǽcula s{\ae}culórum. Amen.}
\elcol{℣. Praise ye the Lord.}{℣. Laudáte Dominum.}
\elcol{℟. The Lord's Name be praised.}{℟. Sit Nomen Dómini Benedíctum.}
\subby{Nunc Sancte}
\elcol{\begin{hangparas}{1.25em}{1}
\lett{C}{ome} Holy Ghost, with God the Son,\par
\secondline{And God the Father, ever one:}\par\noindent
%\thirdline{Shed forth thy grace within our breast,}
Shed forth thy grace within our breast,\par\noindent
And dwell with us, a ready Guest.\\
\par\noindent
By every power, by heart and tongue,\par\noindent
By act and deed, thy praise be sung;\par\noindent
Inflame with perfect love each sense,\par\noindent
That others' souls may kindle thence.\\
\par\noindent
O Father, that we ask be done,\par\noindent
Through Jesus Christ, thine only Son;\par\noindent
Who, with the Holy Ghost and thee,\par\noindent
Shall live and reign eternally. Amen.\\
\end{hangparas}}
{\lett{N}{unc,} Sancte, nobis, Spíritus,\par
\secondline{Unum Patri cum Fílio,}
\thirdline{Dignáre promptus íngeri}
Nostro refúsus péctori.\\

Os, lingua, mens, sensus, vigor\par
Confessiónem pérsonent.\par
Flamméscat igne cáritas,\par
Accéndat ardor próximos.\\

Præsta, Pater piíssime,\par
Patríque compar Únice,\par
Cum Spíritu Paráclito\par
Regnans per omne sǽculum. Amen.\\}


\subbypsalm{Psalm 120. \textit{Ad Dominum}}

\lett{W}{hen} I was in trouble I called upon the Lord : and he heard me.\par
\secondline{\psanum{2}Deliver my soul, O Lord, from lying lips : and from a deceitful tongue.}
\thirdline{\psanum{3}What reward shall be given or done unto thee, thou false tongue : even mighty and sharp arrows, with hot burning coals.}
\psanum{4}Woe is me, that I am constrained to dwell with Mesech : and to have my habitation among the tents of Kedar.\par
\psanum{5}My soul hath long dwelt among them : that are enemies unto peace.\par
\psanum{6}I labour for peace, but when I speak unto them thereof : they make them ready to battle.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.

\subbypsalm{Psalm 121. \textit{Levavi oculus}}

\lett{I}{will} lift up mine eyes unto the hills : from whence cometh my help.\par
\secondline{\psanum{2}My help cometh even from the Lord : who hath made heaven and earth.}
\thirdline{\psanum{3}He will not suffer thy foot to be moved : and he that keepeth thee will not sleep.}
\psanum{4}Behold, he that keepeth Israel : shall neither slumber nor sleep.\par
\psanum{5}The Lord himself is thy keeper : the Lord is thy defence upon thy right hand;\par
\psanum{6}So that the sun shall not burn thee by day : neither the moon by night.\par
\psanum{7}The Lord shall preserve thee from all evil : yea, it is even he that shall keep thy soul.\par
\psanum{8}The Lord shall preserve thy going out, and thy coming in : from this time forth for evermore.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.

\subbypsalm{Psalm 122. \textit{Laetatus sum}}

\lett{I}{was} glad when they said unto me : We will go into the house of the Lord.\par
\secondline{\psanum{2}Our feet shall stand in thy gates : O Jerusalem.}
\thirdline{\psanum{3}Jerusalem is built as a city : that is at unity in itself.}
\psanum{4}For thither the tribes go up, even the tribes of the Lord : to testify unto Israel, to give thanks unto the Name of the Lord.\par
\psanum{5}For there is the seat of judgement : even the seat of the house of David.\par
\psanum{6}O pray for the peace of Jerusalem : they shall prosper that love thee.\par
\psanum{7}Peace be within thy walls : and plenteousness within thy palaces.\par
\psanum{8}For my brethren and companions' sakes : I will wish thee prosperity.\par
\psanum{9}Yea, because of the house of the Lord our God : I will seek to do thee good.\par
℣. Glory be to the Father, and to the Son, and to the Holy Ghost.\par
℟. As it was in the beginning, is now, and ever shall be, world without end. Amen.

\subby{Chapter}
\begin{inhead}
	Feria Days Throughout the Year
\end{inhead}
\elcoldent{Heal me, O Lord, and I shall be healed; save me, and I shall be saved: for thou art my praise.}{Sana me, Dómine, et sanábor: salvum me fac, et salvus ero: quóniam laus mea tu es.}
\begin{inhead}
	Sundays \& Feast Days
\end{inhead}
\elcoldent{O the depth of the riches both of the wisdom and knowledge of God! How unsearchable are his judgements, and his ways past finding out!}{O altitúdo divitiárum sapiénti{\ae}, et sciénti{\ae} Dei: quam incomprehensibília sunt judícia ejus, et investigábiles vi{\ae} ejus!}
%MANUAL ADJUSTMENT:
\clearpage
\begin{inhead}
	Eastertide
\end{inhead}
\elcoldent{For whatsoever is born of God overcometh the world: and this is the victory that overcometh the world, even our faith.}{Omne, quod natum est ex Deo, vincit mundum: et h{\ae}c est victória, qu{\ae} vincit mundum, fides nostra.}
\elcol{℟. Thanks be to God.}{℟. Deo grátias.}
\subby{Short Respond}
\begin{inhead}
	Ferias Throughout the Year
\end{inhead}
\elcol{℣. Heal my soul, for I have sinned against thee.}{℣. Sana ánimam meam, Quia peccávi tibi.}

\elcol{℟. Heal my soul, for I have sinned against thee.}{℟. Sana ánimam meam, Quia peccávi tibi.}

\elcol{℣. I said, Lord, be merciful unto me.}{℣. Ego dixi: Dómine, miserére mei.}
    
\elcol{℟. For I have sinned against thee.}{℟. Quia peccávi tibi.}

\elcol{℣. Glory be to the Father, and to the Son, and to the Holy Ghost.}{℣. Glória Patri, et Fílio, et Spirítui Sancto.}
    
\elcol{℟. Heal my soul, for I have sinned against thee.\\}{℟. Sana ánimam meam, Quia peccávi tibi.\\}

\elcol{℣. Thou hast been my succour.}{℣. Adjútor meus, esto.}

\elcol{℟. Leave me not, neither forsake me, O God of my salvation.}{℟. Ne derelínquas me, Neque despícias me, Deus, salutáris meus.}

\begin{inhead}
	Sundays, Feasts, \& in Eastertide
\end{inhead}

\elcol{℣. Incline my heart, O God, unto thy testimonies (alleluia, alleluia).}{℣.  Inclína cor meum in testimónia tua (allelúja, allelúja).}

\elcol{℟. Incline my heart, O God, unto thy testimonies (alleluia, alleluia).}{℟. Inclína cor meum in testimónia tua (allelúja, allelúja).}

\elcol{℣. Turn away mine eyes, lest they behold vanity.}{℣. Avérte óculos meos ne vídeant vanitátem.}
    
\elcol{℟. Unto thy testimonies (alleluia, alleluia).}{℟. In testimónia tua (allelúja, allelúja).}

\elcol{℣. Glory be to the Father, and to the Son, and to the Holy Ghost.}{℣. Glória Patri, et Fílio, et Spirítui Sancto.}
    
\elcol{℟. Incline my heart, O God, unto thy testimonies (alleluia, alleluia).\\}{℟. Inclína cor meum in testimónia tua (allelúja, allelúja).\\}

\elcol{℣. I said, Lord, be merciful unto me (alleluia).}{℣. Ego dixi: Dómine, miserére mei (allelúja).}

\elcol{℟. For I have sinned against thee (alleluia).}{℟. Quia peccávi tibi (allelúja).}

\subby{Suffrages}
\begin{rubric}
	The Suffrages, said kneeling, are said on all Ferias outside of Eastertide. They are omitted on Sundays, Feasts, and within Principle, Primary, and Secondary Octaves.\par
	\textsc{Note,} If omitted, the Office continues with the Collect.
\end{rubric}

\elcol{℣. Lord, have mercy upon us.}{℣. Kýrie, eléison.}
\elcol{℟. Christ, have mercy upon us.}{℟. Christe, eléison.}
\elcol{℣. Lord, have mercy upon us.}{℣. Kýrie, eléison.}
%\subsec{Lord's Prayer}
\needspace{4\baselineskip}
\elcol{\lett{O}{ur} Father, who art in heaven, Hallowed be thy Name. Thy kingdom come. Thy will be done on earth, As it is in heaven. Give us this day our daily bread. And forgive us our trespasses, As we forgive those who trespass against us. And lead us not into temptation; But deliver us from evil. Amen.}
{\lett{P}{ater} noster, qui es in c{\ae}lis, sanctificétur nomen tuum: advéniat regnum tuum: fiat volúntas tua, sicut in c{\ae}lo et in terra. Panem nostrum quotidiánum da nobis hódie: et dimítte nobis débita nostra, sicut et nos dimíttimus debitóribus nostris: et ne nos indúcas in tentatiónem: sed líbera nos a malo. Amen.}

\elcol{℣. Send out thy light and thy truth, that they may lead me.}{℣. Emítte lucem tuam et veritátem tuam: ipsa me deduxérunt.}

\elcol{℟. And bring me unto thy holy hill, and to thy dwelling.}{℟. Et adduxérunt in montem sanctum tuum et in tabernácula tua.}

\elcol{℣. O let us live.}{℣. Vivificábis nos.}

\elcol{℟. And we shall call upon thy Name.}{℟. Et nomen tuum invocábimus.}

\elcol{℣. Turn us again, O Lord God of hosts.}{℣. Dómine, Deus virtútum, convérte nos.}

\elcol{℟. Show the light of thy countenance, and we shall be whole.}{℟. Et osténde fáciem tuam, et salvi érimus.}

\subby{Collects}
\elcol{℣. The Lord be with you.\par
\textit{or,} O Lord, hear our prayer.}{℣. Dóminus vobíscum.\par
\textit{vel,} Dómine, exáudi oratiónem nostram.}
\elcol{℟. And with thy spirit.\par
\textit{or,} And let our cry come unto thee.}{℟. Et cum spíritu tuo.\par
\textit{vel,} Et clamor noster ad te véniat.}
\elcol{℣. Let us pray.}{℣. Orémus.}
\begin{rubric}
	The Collect(s) of the Day is here said, followed by this memorial Collect.
\end{rubric}
\needspace{4\baselineskip}
\lett{H}{elp} us this day, O God, to serve thee devoutly and the world busily. May we do our work wisely, give succour secretly, go to our meat appetitely, sit thereat discreetly, arise temperately, please our friends duly, go to our bed merrily, and sleep surely; for the joy of our Lord Jesus Christ. \textit{Amen.}\\
%MANUAL ADJUSTMENT:
\clearpage
\elcol{℣. The Lord be with you.\par
\textit{or,} O Lord, hear our prayer.}{℣. Dóminus vobíscum.\par
\textit{vel,} Dómine, exáudi oratiónem nostram.}
\elcol{℟. And with thy spirit.\par
\textit{or,} And let our cry come unto thee.}{℟. Et cum spíritu tuo.\par
\textit{vel,} Et clamor noster ad te véniat.}
\elcol{℣. Let us bless the Lord.}{℣. Benedicámus Dómino.}
\elcol{℟. Thanks be to God.}{℟. Deo grátias}
\elcol{℣. May the souls {\ding{64}} of the faithful departed, through the mercy of God, rest in peace.}{℣. Fidélium ánim{\ae} {\ding{64}} per misericórdiam Dei requiéscant in pace.}
\needspace{4\baselineskip}
\lett{U}{nto} him that loved us, and washed us from our sins in his own blood, and hath made us kings and priests unto God and his Father, to him be glory and dominion for ever and ever. \textit{Amen.}


%\elcol{\lett{U}{nto} him that loved us, and washed us from our sins in his own blood, and hath made us kings and priests unto God and his Father, to him be glory and dominion for ever and ever. Amen.}{\lett{Q}{ui} dilexit nos, et lavit nos a peccatis nostris in sanguine suo, 6 et fecit nos regnum, et sacerdotes Deo et Patri suo: ipsi gloria et imperium in s{\ae}cula s{\ae}culorum. Amen.}
\begin{center}
	{\textsc{Here endeth the Order of Terce.}}
\end{center}