\phantomsection
\addcontentsline{toc}{chapter}{Dedication}
\fancyhead[C]{{\LARGE Dedication}}
\noindent
This \textit{Prayer Book Hymnal} is dedicated to Patriarch \textsc{John X}, \nth{171} Successor of St. Peter, Head of the Church, Patriarch of Antioch, and Defender of the Faith.\\
\par\noindent
And we are beyond thankful for our Christ-loving, God-fearing, and faithful-caring chief Bishop and Metropolitan \textsc{Saba}. For his support of the Western Rite continues our endeavours, lightens our hearts, and shows his divine charity for his flock.\\
\par\noindent
And for our faithful, compassionate, and righteous Bishop \textsc{John}, we must send up everlasting praise and thanksgiving unto the Most Holy Trinity. For through him has the daily care, discipline, and concern for the Vicariate been carried. Through him have so many of us flocked for refuge in the Holy Orthodox Catholic Church. And through him have we been intimately cared and nourished.\\
\par\noindent
Finally, we must thank our Lord and God and Saviour Jesus Christ for our beloved Priests, Deacons, and all Clergy in the Vicariate for their tireless care, concern, and work for our salvation. Chiefly among them must surely be our blessed Assistant to the Vicar General, Fr. John Fenton. He who has taken great and serious care and rigour for staying the course of the Vicariate when it is battered and afflicted from within and without.\\
\par\noindent
Let us for ever bless the Father and the Son and the Holy Ghost, together with all the Saints, chiefly of blessed Mary ever-Virgin, blessed John Baptist, and the holy Apostles Peter and Paul, for our fathers in the faith. May God, through the merits and prayers of the Saints, bless and preserve them now and for ever. Amen.
\vfill
\begin{center}
\textsc{And I say also unto thee, That thou art Peter, and upon this rock I will build my church; and the gates of hell shall not prevail against it. And I will give unto thee the keys of the kingdom of heaven: and whatsoever thou shalt bind on earth shall be bound in heaven: and whatsoever thou shalt loose on earth shall be loosed in heaven.}
\end{center}