\fancyhead[RE,LO]{}\fancyhead[RO,LE]{}
\fancyhead[C]{\LARGE Frequent Prayers}
\phantomsection
\addcontentsline{toc}{chapter}{Frequently Said Prayers during Mass}

\section*{Prayers at the Foot of the Altar}

\elcol{℣. In the Name of the Father, {\ding{64}} and of the Son, and of the Holy Ghost.}{℣. In nómine Patris, {\ding{64}} et Fílii, et Spíritus Sancti.}
\elcol{℟. Amen.}{℟. Amen.}

\elcol{℣. I will go unto the altar of God.}{℣. Introíbo ad altáre Dei.}
\elcol{℟. Even unto the God of my joy and gladness.}{℟. Ad Deum, qui l{\ae}tíficat juventútem meam.}
\par\noindent
	\centerline{\rule{0.5\textwidth}{0.4pt}}\\
\par\noindent
\elcol{℣. Give sentence with me, O God, and defend my cause against the ungodly people: O deliver me from the deceitful and wicked man.}{℣. Júdica me, Deus, et discérne causam meam de gente non sancta: ab hómine iníquo et dolóso érue me.}
\elcol{℟. For thou art the God of my strength; why hast thou put me from thee: and why go I so heavily, while the enemy oppresseth me?}{℟. Quia tu es, Deus, fortitúdo mea: quare me repulísti, et quare tristis incédo, dum afflígit me inimícus?}
\elcol{℣. O send out thy light and thy truth, that they may lead me: and bring me unto thy holy hill, and to thy dwelling.}{℣. Emítte lucem tuam et veritátem tuam: ipsa me deduxérunt, et adduxérunt in montem sanctum tuum et in tabernácula tua.}
\elcol{℟. And that I may go unto the altar of God, even the God of my joy and gladness: and upon the harp will I give thanks unto thee, O God, my God.}{℟. Et introíbo ad altáre Dei: ad Deum, qui l{\ae}tíficat juventútem meam.}
\elcol{℣. Why art thou so heavy, O my soul: and why art thou so disquieted within me?}{℣. Confitébor tibi in cíthara, Deus, Deus meus: quare tristis es, ánima mea, et quare contúrbas me?}
\elcol{℟. O put thy trust in God: for I will yet give him thanks, which is the help of my countenance, and my God.}{℟. Spera in Deo, quóniam adhuc confitébor illi: salutáre vultus mei, et Deus meus.}
\elcol{℣. Glory be to the Father, and to the Son: and to the Holy Ghost.}{℣. Glória Patri, et Fílio, et Spirítui Sancto.}
\elcol{℟. As it was in the beginning, is now, and ever shall be: world without end. Amen.}{℟. Sicut erat in princípio, et nunc, et semper: et in sǽcula s{\ae}culórum. Amen.}
\elcol{℣. I will go unto the altar of God:}{℣. Introíbo ad altáre Dei.}
\elcol{℟. Even unto the God of my joy and gladness.}{℟. Ad Deum, qui l{\ae}tíficat juventútem meam.}
\par\noindent
	\centerline{\rule{0.5\textwidth}{0.4pt}}\\
\par\noindent
\clearpage
\elcol{℣. Our help is in the Name {\ding{64}} of the Lord:}{℣. Adjutórium nostrum {\ding{64}} in nómine Dómini.}
\elcol{℟. Who hath made heaven and earth.}{℟. Qui fecit c{\ae}lum et terram.}

\elcol{\lett{I}{confess} to almighty God, to blessed Mary ever Virgin, to blessed Michael the Archangel, to blessed John Baptist, to the holy Apostles Peter and Paul, to all the Saints, and to you, brethren: that I have sinned exceedingly in thought, word, and deed: \inrub{Strike breast thrice.} by my fault, my own fault, my own most grievous fault. Therefore I beg blessed Mary ever Virgin, blessed Michael the Archangel, blessed John Baptist, the holy Apostles Peter and Paul, all the Saints, and you, brethren, to pray for me to the Lord our God.}{\lett{C}{onf\smash{í}teor} Deo omnipoténti, beát{\ae} Marí{\ae} semper Vírgini, beáto Michaéli Archángelo, beáto Joánni Baptíst{\ae}, sanctis Apóstolis Petro et Paulo, ómnibus Sanctis, et vobis, fratres: quia peccávi nimis cogitatióne, verbo et ópere: \inrub{Strike breast thrice.} mea culpa, mea culpa, mea máxima culpa. Ideo precor beátam Maríam semper Vírginem, beátum Michaélem Archángelum, beátum Joánnem Baptístam, sanctos Apóstolos Petrum et Paulum, omnes Sanctos, et vos, fratres, oráre pro me ad Dóminum, Deum nostrum.}
\elcol{℟. Almighty God have mercy upon thee, forgive thee thy sins, and bring thee to everlasting life.}{℟. Misereátur tui omnípotens Deus, et, dimíssis peccátis tuis, perdúcat te ad vitam {\ae}térnam.}
\elcol{℟. Amen.}{℟. Amen.}
\elcol{℟. I confess to almighty God, to blessed Mary ever Virgin, to blessed Michael the Archangel, to blessed John Baptist, to the holy Apostles Peter and Paul, to all the Saints, and to thee, father: that I have sinned exceedingly in thought, word, and deed: \inrub{Strike breast thrice.} by my fault, my own fault, my own most grievous fault. Therefore I beg blessed Mary ever Virgin, blessed Michael the Archangel, blessed John Baptist, the holy Apostles Peter and Paul, all the Saints, and thee, father, to pray for me to the Lord our God.}{℟. Confíteor Deo omnipoténti, beát{\ae} Marí{\ae} semper Vírgini, beáto Michaéli Archángelo, beáto Joánni Baptíst{\ae}, sanctis Apóstolis Petro et Paulo, ómnibus Sanctis, et tibi, pater: quia peccávi nimis cogitatióne, verbo et ópere: \inrub{Strike breast thrice.} mea culpa, mea culpa, mea máxima culpa. Ideo precor beátam Maríam semper Vírginem, beátum Michaélem Archángelum, beátum Joánnem Baptístam, sanctos Apóstolos Petrum et Paulum, omnes Sanctos, et te, pater, oráre pro me ad Dóminum, Deum nostrum.}
\elcol{℣. Almighty God have mercy upon you, forgive you your sins, and bring you to everlasting life.}{℣. Misereátur vestri omnípotens Deus, et, dimíssis peccátis vestris, perdúcat vos ad vitam {\ae}térnam.}
\elcol{℟. Amen.}{℟. Amen.}
\elcol{℣. The almighty and merciful Lord grant unto us pardon, {\ding{64}} absolution, and remission of our sins.\par
℟. Amen.}{℣. Indulgéntiam, {\ding{64}} absolutiónem et remissiónem peccatórum nostrórum tríbuat nobis omnípotens et miséricors Dóminus.\par
℟. Amen.}
\clearpage
\elcol{℣. Wilt thou not turn again and quicken us, O God?}{℣. Deus, tu convérsus vivificábis nos?}
\elcol{℟. That thy people may rejoice in thee.}{℟. Et plebs tua l{\ae}tábitur in te.}
\elcol{℣. O Lord, show thy mercy upon us.}{℣. Osténde nobis, Dómine, misericórdiam tuam.}
\elcol{℟. And grant us thy salvation.}{℟. Et salutáre tuum da nobis.}
\elcol{℣. O Lord, hear my prayer.}{℣. Dómine, exáudi oratiónem meam.}
\elcol{℟. And let my cry come unto thee.}{℟. Et clamor meus ad te véniat.}
\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}
\elcol{℣. Let us pray.}{℣. Orémus.}

\section*{Collect for Purity}
\lett{A}{lmighty} God, unto whom all hearts are open, all desires known, and from whom no secrets are hid; Cleanse the thoughts of our hearts by the inspiration of thy Holy Spirit, that we may perfectly love thee, and worthily magnify thy holy Name; through Christ our Lord. \textit{Amen.}
\clearpage
\section*{Ten Commandments}
\begin{center}
	\textsc{God spake these words, and said:}
\end{center}
%\begin{multicols}{2}
\par\noindent
    I am the \textsc{Lord} thy God; Thou shalt have none other gods but me.\par
    \textit{Lord, have mercy upon us, and incline our hearts to keep this law.}
    \par\noindent
    Thou shalt not make to thyself any graven image, nor the likeness of any thing that is in heaven above, or in the earth beneath, or in the water under the earth; thou shalt not bow down to them, nor worship them:
    \par\noindent
    \leftskip=2em
	{\small{for I the Lord thy God am a jealous God, and visit the sins of the fathers upon the children, unto the third and fourth generation of them that hate me; and show mercy unto thousands in them that love me and keep my commandments.}}
	\par
	\leftskip=0cm
    \textit{Lord, have mercy upon us, and incline our hearts to keep this law.}
    \par\noindent
    Thou shalt not take the Name of the Lord thy God in vain;
    \par\noindent
    \leftskip=2em
	{\small{for the Lord will not hold him guiltless, that taketh his Name in vain.}}
	\par
	\leftskip=0em
    \textit{Lord, have mercy upon us, and incline our hearts to keep this law.}
    \par\noindent
    Remember that thou keep holy the Sabbath-day.
    \par\noindent
    \leftskip=2em
	{\small{Six days shalt thou labour, and do all that thou hast to do; but the seventh day is the Sabbath of the Lord thy God. In it thou shalt do no manner of work; thou, and thy son, and thy daughter, thy man-servant, and thy maid-servant, thy cattle, and the stranger that is within thy gates. For in six days the Lord made heaven and earth, the sea, and all that in them is, and rested the seventh day: wherefore the Lord blessed the seventh day, and hallowed it.}}
	\par
	\leftskip=0em
    \textit{Lord, have mercy upon us, and incline our hearts to keep this law.}
    \par\noindent
    Honour thy father and thy mother;
    \par\noindent
    \leftskip=2em
	{\small{that thy days may be long in the land which the Lord thy God giveth thee.}}
	\par
	\leftskip=0em
    \textit{Lord, have mercy upon us, and incline our hearts to keep this law.}
    \par\noindent
    Thou shalt do no murder.\par
    \textit{Lord, have mercy upon us, and incline our hearts to keep this law.}
    \par\noindent
    Thou shalt not commit adultery.\par
    \textit{Lord, have mercy upon us, and incline our hearts to keep this law.}
\par\noindent
    Thou shalt not steal.\par
    \textit{Lord, have mercy upon us, and incline our hearts to keep this law.}
\par\noindent
    Thou shalt not bear false witness against thy neighbour.\par
    \textit{Lord, have mercy upon us, and incline our hearts to keep this law.}
    \par\noindent
    Thou shalt not covet.
    \par\noindent
    \leftskip=2em
	{\small{thy neighbour's house, thou shalt not covet thy neighbour's wife, nor his servant, nor his maid, nor his ox, nor his ass, nor any thing that is his.}}
	\par
	\leftskip=0em
	\textit{Lord, have mercy upon us, and write all these thy laws in our hearts, we beseech thee.}
\clearpage
\section*{Summary of the Law}
\begin{center}
	{\textsc{Hear what our Lord Jesus Christ saith.}}
\end{center}

\lett{T}{hou} shalt love the Lord thy God with all thy heart, and with all thy soul, and with all thy mind. This is the first and great commandment. And the second is like unto it; Thou shalt love thy neighbour as thyself. On these two commandments hang all the Law and the Prophets.

\section*{Invitation to Confession}
\lett{Y}{e} who do truly and earnestly repent you of your sins, and are in love and charity with your neighbours, and intend to lead a new life, following the commandments of God, and walking from henceforth in his holy ways; Draw near with faith, and take this holy Sacrament to your comfort; and make your humble confession to Almighty God, devoutly kneeling.

\section*{Confession}
\lett{A}{lmighty} God, Father of our Lord Jesus Christ, Maker of all things, Judge of all men; We acknowledge and bewail our manifold sins and wickedness, Which we, from time to time, most grievously have committed, By thought, word, and deed, Against thy Divine Majesty, Provoking most justly thy wrath and indignation against us. We do earnestly repent, And are heartily sorry for these our misdoings; The remembrance of them is grievous unto us; The burden of them is intolerable. Have mercy upon us, Have mercy upon us, most merciful Father; For thy Son our Lord Jesus Christ's sake, Forgive us all that is past; And grant that we may ever hereafter Serve and please thee in newness of life, To the honour and glory of thy Name; Through Jesus Christ our Lord. Amen.

\section*{Absolution}
\lett{A}{lmighty} God, our heavenly Father, who of his great mercy hath promised forgiveness of sins to all those who with hearty repentance and true faith turn unto him; Have mercy upon you; pardon {\ding{64}} and deliver you from all your sins; confirm and strengthen you in all goodness; and bring you to everlasting life; through Jesus Christ our Lord.

℟. Amen.
\clearpage
\section*{Comfortable Words}
\begin{center}
	\textsc{Hear what comfortable words our Saviour Christ saith unto all who truly turn to him.}
\end{center}
\par\noindent
Come unto me, all ye that travail and are heavy laden, and I will refresh you. \vr{Matt. 11:28}
\par\noindent
    So God loved the world, that he gave his only-begotten Son, to the end that all that believe in him should not perish, but have everlasting life. \vr{John 3:16}
    \par\noindent
    \begin{center}
		\textsc{Hear also what Saint Paul saith.}
	\end{center}
\par\noindent
    This is a true saying, and worthy of all men to be received, That Christ Jesus came into the world to save sinners. \vr{1 Tim. 1:15}
\par\noindent
    \begin{center}
		\textsc{Hear also what Saint John saith.}
	\end{center}
    \par\noindent
    If any man sin, we have an Advocate with the Father, Jesus Christ the righteous; and he is the Propitiation for our sins. \vr{1 John 2:1-2}

\section*{Distribution of the Body}
\lett{T}{he} Body of our Lord Jesus Christ, which was given for thee, preserve thy body and soul unto everlasting life. Take and eat this in remembrance that Christ died for thee, and feed on him in thy heart by faith, with thanksgiving.

\section*{Distribution of the Blood}
\lett{T}{he} Blood of our Lord Jesus Christ, which was shed for thee, preserve thy body and soul unto everlasting life. Drink this in remembrance that Christ's Blood was shed for thee, and be thankful.

\section*{Distribution of the Body \& Blood}
\lett{T}{he} Body of our Lord Jesus Christ which was given for thee, and his Blood which was shed for thee. Take them in remembrance that Christ died for thee, and feed on him in thy heart by faith with thanksgiving.