\phantomsection
\addcontentsline{toc}{chapter}{Preface}
\fancyhead[C]{{\LARGE Preface}}
\noindent
We are blessed to publish this necessary hymnal and companion to the 2024 Proposed Book of Common Prayer. The Prayer Book Hymnal supplements the Book of Common Prayer by providing liturgical elements, which may prove expedient to divine worship, along with notation for the hymns. For the convenience of the priest, the frequently said prayers of the Mass (when his BCP may not be easily accessed) are provided. For both the Office and the Mass, the most common settings and common chants are set as Office and Communion Services. Regarding the propers, the Office and Mass propers for the minor Feast Days (not found in the BCP), as well as the regular ferial hymns, are provided. Additionally, the Office propers which are omitted in the Red Letter Days are also provided. These have all been typeset to Gregorian notation, for the ease of use by cantors and the Congregation.

For all these precious liturgical elements, this book may rightly be called a hymnal. However, the chief treasure is the set of `Devotional Hymns'. For these, the best of various hymnary traditions have been chosen. Uniquely beneficial as opening and closing hymns, we hope that these provide spiritual benefits for those who use them, especially since this is the first Orthodox hymnal to include many of these musical pieces.

We hope this Prayer Book Hymnal provides benefit and utility to those living within, or interested in, the English Orthodox tradition.