%\phantomsection
%\addcontentsline{toc}{section}{Common of Saints}
\fancyhead[RO,LE]{}
\fancyhead[RE,LO]{}
\fancyhead[C]{\LARGE Common of Saints}

\subby{Common of Vigils of the Apostles}\label{CommonVigilApostles}
\fancyhead[RO,LE]{\textit{Apostolic Vigils}}

\introit
\lett{A}{s} for me, I am like a green olive-tree in the house of the Lord, my trust is in the tender mercy of my God: and I will hope in thy name, for thy saints like it well. \textit{Ps.} Why boastest thou thyself, thou tyrant: that thou canst do mischief?

\collect
\lett{G}{rant,} we beseech thee, almighty God: that the venerable solemnity of blessed \textit{N.} thine Apostle, which we here prevent, may increase our devotion and set forward our salvation. Through.
\begin{rubric}
	If, however, the preceding Collect has already been said in the Mass, or for the commemoration of a Confessor Bishop, then the following is to be said:
\end{rubric}
\lett{W}{e} beseech thee, almighty God: that as we do prevent the festival of blessed \textit{N.} thine Apostle, so he may implore thy mercy for us; that we being delivered from our iniquities, may likewise be defended against all adversities. Through.
\begin{rubric}
	\nth{2} and \nth{3} Collects as assigned for Vigils, according to the Rubrics.
\end{rubric}

\readingcitation{Epistle}{Ecclesiasticus 44:22}
%RV:
\lett{I}{n} Isaac did the Lord establish, for Abraham his father's sake, The blessing of all men, and the covenant: And he made it rest upon the head of Jacob; He acknowledged him in his blessings, And gave to him by inheritance, And divided his portions; Among twelve tribes did he part them. And he brought out of him a man of mercy, Which found favour in the sight of all flesh; A man beloved of God and men, even Moses, Whose memorial is blessed. He made him like to the glory of the saints, And magnified him in the fears of his enemies. By his words he caused the wonders to cease; He glorified him in the sight of kings; He gave him commandment for his people, And shewed him part of his glory. He sanctified him in his faithfulness and meekness; He chose him out of all flesh. He made him to hear his voice, And led him into the thick darkness, And gave him commandments face to face, Even the law of life and knowledge, That he might teach Jacob the covenant, And Israel his judgments. He exalted Aaron, a holy man like unto him, Even his brother, of the tribe of Levi. He established for him an everlasting covenant, And gave him the priesthood of the people; He beautified him with comely ornaments, And girded him about with a robe of glory.

%Traditional Text:
%\lett{T}{he} blessing of the Lord was upon the head of the righteous. Therefore did the Lord give him an heritage, and divided his portions; among the twelve tribes did he part them: and he found favour in the sight of all flesh. And he magnified him, so that his enemies stood in fear of him. By his words he caused the wonders to cease, and he made him glorious in the sight of kings, and gave him a commandment for his people, and shewed him his glory. He sanctified him in his faithfulness and meekness, and chose him out of all men. And he gave him commandments before his face, even the law of life and knowledge, and made him to be exalted. An everlasting covenant he made with him, and girded him about with the girdle of righteousness: and the Lord crowned him with the crown of glory.

\gradual{The righteous shall flourish like a palm-tree: and shall spread abroad like a cedar in Libanus in the house of the Lord. ℣. To tell of thy loving-kindness early in the morning, and of thy truth in the night-season.}

\readingcitation{Gospel}{John 15:12}
\lett{A}{t that time:} Jesus said unto his disciples: This is my commandment, That ye love one another, as I have loved you. Greater love hath no man than this, that a man lay down his life for his friends. Ye are my friends, if ye do whatsoever I command you. Henceforth I call you not servants; for the servant knoweth not what his lord doeth: but I have called you friends; for all things that I have heard of my Father I have made known unto you. Ye have not chosen me, but I have chosen you, and ordained you, that ye should go and bring forth fruit, and that your fruit should remain: that whatsoever ye shall ask of the Father in my name, he may give it you.

\offertory{Thou hast crowned him with glory and worship: thou hast made him to have dominion of the works of thy hands, O Lord.}

\secret
\lett{O}{Lord,} who didst exalt blessed \textit{N.} to be numbered among thine Apostles: grant that we thy people, who on this day, preventing his heavenly birth, do offer unto thee these holy mysteries, may be assisted by his intercession both in the making of our supplications before thee, and in the obtaining of all that we desire. Through.
\begin{rubric}
	Other Secrets as above.
\end{rubric}
\communion{His honour is great in thy salvation: glory and great worship shalt thou lay upon him, O Lord.}
\postcommunion
\lett{W}{e} beseech thee, O Lord, mercifully to hear the supplication of thy holy Apostle \textit{N.}: that thou wouldest vouchsafe unto us thy pardon, and grant us thy everlasting healing. Through.
\begin{rubric}
	Other Postcommunions as above.
\end{rubric}


\subby{Common of the Apostles}\label{CommonApostles}
\fancyhead[RO,LE]{\textit{Apostles}}

\subbysub{I Evensong}

%\begin{rubric}
%	In I Evensong, the Office Hymn is of Mattins with the following Versicle \& Antiphon.
%\end{rubric}

\gregorioscore{resources/gabc/Common/ApostlesEvensong.gabc}

℣. Their sound is gone out into all lands.

℟. And their words into the ends of the world.

\antiphon{Mag.}{They will deliver you up {\dag} to the councils, and they will scourge you in their synagogues; and ye shall be brought before governors and kings for my sake, for a testimony against them and the Gentiles.}

%\begin{multicols}{2}
%\lett{L}{et} heav'n's exultant praises ring,
%
%\secondline{And earth with joy responsive sing:}
%
%\thirdline{Th' Apostles' deeds and high estate}
%
%This festal-tide we celebrate.\\
%
%O ye who, thron'd in glory dread,
%
%Shall judge the living and the dead,
%
%True lights, the world illumining,
%
%Regard the suppliant prayer we bring.\\
%
%The gates of heav'n, at your command,
%
%To all or clos'd or open stand:
%
%May we, at your august decree,
%
%Be loos'd from our iniquity.\\
%
%The power, of old to you convey'd,
%
%Sickness and health alike obey'd:
%
%May ye our ailing souls once more
%
%To strength and holiness restore;\\
%
%That Christ, th' unerring Judge of doom,
%
%When he at time's last end shall come,
%
%May grant us, for his mercy's sake,
%
%Of joys eternal to partake.\\
%
%All laud to God the Father be;
%
%All praise, eternal Son, to thee;
%
%All glory, as is ever meet,
%
%To God the Holy Paraclete. Amen.\\

\subbysub{Mattins}

\invitatoryhymn

\gregorioscore{resources/gabc/Common/ApostlesInvitatory.gabc}

\officehymn

\gregorioscore{resources/gabc/Common/ApostlesMattins.gabc}

℣. They declared the work of God.

℟. And wisely considered of his doing.
%\end{multicols}

\antiphon{Ben.}{Ye which have forsaken all, {\dag} and followed me, shall receive an hundredfold, and shall inherit everlasting life.}

\subbysub{II Evensong}

\begin{rubric}
	The Office Hymn \& Versicle are of I Evensong with the following Antiphon.
\end{rubric}

\antiphon{Mag.}{Be ye valiant in warfare, {\dag} and contend ye with the old serpent: and ye shall receive an eternal kingdom, alleluia.}


\subby{Common of the Apostles in Eastertide}\label{CommonApostlesEaster}
\fancyhead[RO,LE]{\textit{Apostles Easter}}

\subbysub{I Evensong}
%\begin{multicols}{2}
%\lett{T}{he} Apostles' hearts were full of pain
%
%\secondline{For their dear Lord so lately slain:}
%
%\thirdline{That Lord his servants' wicked train}
%
%With bitter scorn had dar'd arraign.\\
%
%With gentle voice the Angel gave
%
%The women tidings at the grave;
%
%`Forthwith your Master shall ye see:
%
%He goes before to Galilee.'\\
%
%And while with fear and joy they press'd
%
%To tell these tidings to the rest,
%
%Their Lord, their living Lord, they meet,
%
%And see his form, and kiss his feet.\\
%
%Th' Eleven, when they hear, with speed
%
%To Galilee forthwith proceed:
%
%That there they may behold once more
%
%The Lord's dear face, as oft before.\\
%
%We pray thee, King with glory deck'd,
%
%In this our Paschal joy, protect
%
%From all that death would fain effect
%
%Thy ransom'd flock, thine own elect.

\gregorioscore{resources/gabc/Common/ApostlesEvensongEaster.gabc}

    \par\noindent
	\centerline{\rule{0.25\textwidth}{0.4pt}}
	\par\noindent
\begin{rubric}
	From Low Sunday, exclusive, until Ascension Thursday, exclusive:
\end{rubric}

\gregorioscore{resources/gabc/Common/ApostlesEvensongEasterEnding1.gabc}

%6. To thee who, dead, again dost live,
%
%All glory, Lord, thy people give;
%
%All glory, as is ever meet,
%
%To Father and to Paraclete. Amen.

\begin{rubric}
	From Ascension Thursday, inclusive, until Whitsunday, inclusive:
\end{rubric}

\gregorioscore{resources/gabc/Common/ApostlesEvensongEasterEnding2.gabc}

%6. All glory, Lord, to thee we pay,
%
%Ascending o'er the stars to-day;
%
%All glory, as is ever meet,
%
%To Father and to Paraclete. Amen.
    \par\noindent
	\centerline{\rule{0.25\textwidth}{0.4pt}}
	\par
℣. O ye holy and righteous, rejoice in the Lord, alleluia.

℟. God hath chosen you to him to be his inheritance, alleluia.
%\end{multicols}

\antiphon{Mag.}{Light perpetual {\dag} shall shine upon thy Saints, O Lord; and an ageless eternity, alleluia.}

\subbysub{Mattins}

\begin{rubric}
	The Invitatory Hymn is of I Evensong.
\end{rubric}

\officehymn

\gregorioscore{resources/gabc/Common/ApostlesMattinsEaster.gabc}

    \par\noindent
	\centerline{\rule{0.25\textwidth}{0.4pt}}
	\par\noindent
\begin{rubric}
	From Low Sunday, exclusive, until Ascension Thursday, exclusive:
\end{rubric}

\gregorioscore{resources/gabc/Common/ApostlesMattinsEasterEnding1.gabc}

\begin{rubric}
	From Ascension Thursday, inclusive, until Whitsunday, inclusive:
\end{rubric}

\gregorioscore{resources/gabc/Common/ApostlesMattinsEasterEnding2.gabc}
    \par\noindent
	\centerline{\rule{0.25\textwidth}{0.4pt}}
	\par
℣. Right dear in the sight of the Lord, alleluia.

℟. Is the death of his Saints, alleluia.

\antiphon{Ben.}{Daughters of Jerusalem, {\dag} come forth and behold the Martyrs with the diadems wherewith the Lord hath crowned them in the day of solemnity and rejoicing, alleluia, alleluia.}

\subbysub{II Evensong}

\begin{rubric}
	The Office Hymn is of I Evensong with the following Versicle \& Antiphon.
\end{rubric}

℣. Right dear in the sight of the Lord, alleluia.

℟. Is the death of his Saints, alleluia.

\antiphon{Mag.}{O ye holy and righteous, {\dag} rejoice in the Lord, alleluia: God hath chosen you to him to be his inheritance, alleluia.}


\subby{I: Common of One Martyr Bishop out of Eastertide}\label{CommonMartyrBishopI}
\fancyhead[RO,LE]{\textit{Martyr Bishop I}}

\subbysub{I Evensong}

\gregorioscore{resources/gabc/Common/MartyrEvensong.gabc}

℣. Thou hast crowned him with glory and honour, O Lord. 

℟. And madest him to have dominion of the works of thy hands.

\antiphon{Mag.}{This is a Martyr {\dag} who strove for his Master's precepts, even unto death: and feared not the words of evil men, forasmuch as he was stablished on a sure foundation.}

\subbysub{Mattins}

\begin{rubric}
	The Invitatory Hymn is of I Evensong.
\end{rubric}

\officehymn

\gregorioscore{resources/gabc/Common/MartyrMattins.gabc}

℣. The righteous shall flourish like a palm-tree.

℟. And shall spread abroad like a cedar in Libanus.

\antiphon{Ben.}{He that hateth his life {\dag} in this world shall keep it unto life eternal.}\par\noindent

\subbysub{II Evensong}

\begin{rubric}
	The Office Hymn is of I Evensong with the following Versicle \& Antiphon.
\end{rubric}

℣. The righteous shall flourish like a palm-tree.

℟. And shall spread abroad like a cedar in Libanus.

\antiphon{Mag.}{Whosoever will come after me, {\dag} let him deny himself, and take up his cross, and follow me.}

\introit
\lett{T}{he} Lord hath established a covenant of peace with him, and made him a prince: that he should have the dignity of the priesthood for ever. \textit{Ps.} Lord, remember David: and all his trouble.

\collect
\lett{A}{lmighty} God, mercifully look upon our infirmities: that whereas we are oppressed by the burden of our sins, the glorious intercession of blessed \textit{N.} thy Martyr and Bishop may be our succour and defence. Through.

\readingcitation{Epistle}{James 1:12}
\lett{D}{early beloved:} Blessed is the man that endureth temptation: for when he is tried, he shall receive the crown of life, which the Lord hath promised to them that love him. Let no man say when he is tempted, I am tempted of God: for God cannot be tempted with evil, neither tempteth he any man: but every man is tempted, when he is drawn away of his own lust, and enticed. Then when lust hath conceived, it bringeth forth sin: and sin, when it is finished, bringeth forth death. Do not err, my beloved brethren. Every good gift and every perfect gift is from above, and cometh down from the Father of lights, with whom is no variableness, neither shadow of turning. Of his own will begat he us with the word of truth, that we should be a kind of firstfruits of his creatures.

\gradall{I have found David my servant, with my holy oil have I anointed him: my hand shall hold him fast, and my arm shall strengthen him. ℣. The enemy shall not be able to do him violence, the son of wickedness shall not hurt him.}{Alleluia, alleluia. ℣. Thou art a priest for ever, after the order of Melchisedech. Alleluia.}

\begin{rubric}
{In Septuagesimatide or Lent, replacing the Alleluia:}
\end{rubric}\par\noindent
\tract{Thou hast given him his heart's desire: and hast not denied him the request of his lips. ℣. For thou hast prevented him with the blessings of goodness. ℣. And hast set a crown of pure gold upon his head.}

\readingcitation{Gospel}{Luke 14:26}
\lett{A}{t that time:} Jesus said unto the multitudes:  If any man come to me, and hate not his father, and mother, and wife, and children, and brethren, and sisters, yea, and his own life also, he cannot be my disciple. And whosoever doth not bear his cross, and come after me, cannot be my disciple. For which of you, intending to build a tower, sitteth not down first, and counteth the cost, whether he have sufficient to finish it? lest haply, after he hath laid the foundation, and is not able to finish it, all that behold it begin to mock him, saying, This man began to build, and was not able to finish. Or what king, going to make war against another king, sitteth not down first, and consulteth whether he be able with ten thousand to meet him that cometh against him with twenty thousand? Or else, while the other is yet a great way off, he sendeth an ambassage, and desireth conditions of peace. So likewise, whosoever he be of you that forsaketh not all that he hath, he cannot be my disciple.

\offertory{My truth and my mercy shall be with him: and in my name shall his horn be exalted.}

\secret
\lett{W}{e} beseech thee, O Lord, mercifully to accept this our sacrifice which we offer unto thee pleading the merits of blessed \textit{N.} thy Martyr and Bishop: that the same may avail for our perpetual succour. Through.

\communion{I have sworn once by my holiness: His seed shall endure for ever: and his seat is like as the sun before me; he shall stand fast for evermore as the moon, and as the faithful witness in heaven.}

\postcommunion
\lett{W}{e} beseech thee, O Lord our God, that like as we whom thou hast refreshed by the partaking of thy sacred gift do offer unto thee our worship; so, by the intercession of blessed \textit{N.} thy Martyr and Bishop, we may perceive the benefit of the same. Through.


\subby{II: Common of One Martyr Bishop out of Eastertide}\label{CommonMartyrBishopII}
\fancyhead[RO,LE]{\textit{Martyr Bishop II}}
\begin{rubric}
	Daily Office propers as in the First Common of a Martyr Bishop (p. \pageref{CommonMartyrBishopI}).
\end{rubric}
\introit
\lett{O}{ye} priests of the Lord, bless ye the Lord: O ye holy and humble men of heart, bless ye the Lord. \textit{Ps.} O all ye works of the Lord, bless ye the Lord: praise him, and magnify him for ever.

\collect
\lett{O}{God,} who makest us glad with the yearly solemnity of blessed \textit{N.} thy Martyr and Bishop: mercifully grant; that, as we now celebrate his birthday, so we may likewise rejoice in his protection. Through.

\readingcitation{Epistle}{2 Corinthians 1:3}
\lett{B}{rethren:} Blessed be God, even the Father of our Lord Jesus Christ, the Father of mercies, and the God of all comfort; who comforteth us in all our tribulation, that we may be able to comfort them which are in any trouble, by the comfort wherewith we ourselves are comforted of God. For as the sufferings of Christ abound in us, so our consolation also aboundeth by Christ. And whether we be afflicted, it is for your consolation and salvation, which is effectual in the enduring of the same sufferings which we also suffer: or whether we be comforted, it is for your consolation and salvation. And our hope of you is stedfast, knowing, that as ye are partakers of the sufferings, so shall ye be also of the consolation: in Christ Jesus our Lord.

\gradall{Thou hast crowned him with glory and worship. ℣. Thou hast made him to have dominion of the works of thy hands, O Lord.}{Alleluia, alleluia. ℣. This is a priest whom the Lord hath crowned. Alleluia.}

\begin{rubric}
{In Septuagesimatide or Lent, replacing the Alleluia:}
\end{rubric}\par\noindent
\tract{Blessed is the man that feareth the Lord: he hath great delight in his commandments. ℣. His seed shall be mighty upon earth: the generation of the faithful shall be blessed. ℣. Riches and plenteousness shall be in his house: and his righteousness endureth for ever.}

\readingcitation{Gospel}{Matthew 16:24}
\lett{A}{t that time:} Jesus said unto his disciples: If any man will come after me, let him deny himself, and take up his cross, and follow me. For whosoever will save his life shall lose it: and whosoever will lose his life for my sake shall find it. For what is a man profited, if he shall gain the whole world, and lose his own soul? or what shall a man give in exchange for his soul? For the Son of man shall come in the glory of his Father with his angels; and then he shall reward every man according to his works.

\offertory{I have found David my servant, with my holy oil have I anointed him: my hand shall hold him fast, and my arm shall strengthen him.}

\secret
\lett{S}{anctify,} O Lord, the gifts which we dedicate to thee: that at the intercession of blessed \textit{N.}, thy Martyr and Bishop, they may obtain for us thy gracious favour. Through.

\communion{Thou hast set, Lord, a crown of pure gold upon his head.}

\postcommunion
\lett{M}{ay} this communion, O Lord, cleanse us from guilt: and, at the intercession of blessed \textit{N.}, thy Martyr and Bishop, make us partakers of thy heavenly healing. Through.


\subby{I: Common of One Martyr not a Bishop out of Eastertide}\label{CommonMartyrNotBishopI}
\fancyhead[RO,LE]{\textit{Martyr not Bishop I}}
\begin{rubric}
	Daily Office propers as in the First Common of a Martyr Bishop (p. \pageref{CommonMartyrBishopI}).
\end{rubric}

\introit
\lett{T}{he} righteous shall rejoice: in thy strength, O Lord: exceeding glad shall he be of thy salvation: thou hast given him his heart's desire. \textit{Ps.} For thou hast prevented him with the blessings of goodness: and hast set a crown of pure gold upon his head.

\collect
\lett{G}{rant,} we beseech thee, almighty God: that we, who devoutly celebrate the birthday of blessed \textit{N.}, thy Martyr, may by his intercession be stablished in the love of thy name. Through.

\readingcitation{Epistle}{Wisdom 10:10}
%RV:
\lett{W}{isdom} guided the righteous man in straight paths; She shewed him God's kingdom, and gave him knowledge of holy things; She prospered him in his toils, and multiplied the fruits of his labour; When in their covetousness men dealt hardly with him, She stood by him and made him rich; She guarded him from enemies, And from those that lay in wait she kept him safe, And over his sore conflict she watched as judge, That he might know that godliness is more powerful than all. When a righteous man was sold, wisdom forsook him not, But from sin she delivered him; She went down with him into a dungeon, And in bonds she left him not, Till she brought him the sceptre of a kingdom, And authority over those that dealt tyrannously with him; She shewed them also to be false that had mockingly accused him, And gave him eternal glory. 

%\lett{T}{he} Lord guided the just ìn right paths, shewed him the kingdom of God, and gave him knowledge of holy things: made him rich in his travails, and multiplied the fruit of his labours. In the covetousness of such as oppressed him he stood by him, and made him rich. He defended him from his enemies, and kept him safe from those that lay in wait, and in a sore conflict he gave him the victory, that he might know that godliness is stronger than all. When the just was sold, she forsook him not, but delivered him from sin: she went down with him into the pit, and left him not in bonds, till she brought him the sceptre of the kingdom, and power against those that oppressed him: as for them that had accused him, she shewed them to be liars, and the Lord God gave him perpetual glory.

\gradall{Blessed is the man that feareth the Lord: he hath great delight in his commandments. ℣. His seed shall be mighty upon earth: the generation of the faithful shall be blessed.}{Alleluia, alleluia. ℣. Thou hast set, O Lord, a crown of pure gold upon his head. Alleluia.}

\begin{rubric}
{In Septuagesimatide or Lent, replacing the Alleluia:}
\end{rubric}\par\noindent
\tract{Thou hast given him his heart's desire: and hast not denied him the request of his lips. ℣. For thou hast prevented him with the blessings of goodness. ℣. And hast set a crown of pure gold upon his head.}

\readingcitation{Gospel}{Matthew 10:34}
\lett{A}{t that time:} Jesus said unto his disciples: 34 Think not that I am come to send peace on earth: I came not to send peace, but a sword. For I am come to set a man at variance against his father, and the daughter against her mother, and the daughter in law against her mother in law. And a man's foes shall be they of his own household. He that loveth father or mother more than me is not worthy of me: and he that loveth son or daughter more than me is not worthy of me. And he that taketh not his cross, and followeth after me, is not worthy of me. He that findeth his life shall lose it: and he that loseth his life for my sake shall find it. He that receiveth you receiveth me, and he that receiveth me receiveth him that sent me. He that receiveth a prophet in the name of a prophet shall receive a prophet's reward; and he that receiveth a righteous man in the name of a righteous man shall receive a righteous man's reward. And whosoever shall give to drink unto one of these little ones a cup of cold water only in the name of a disciple, verily I say unto you, he shall in no wise lose his reward.

\offertory{Thou hast crowned him with glory and worship: thou hast made him to have dominion of the works of thy hands, O Lord.}

\secret
\lett{W}{e} beseech thee, O Lord, to accept our prayers and oblations: and graciously hearken unto us, whom thou dost cleanse by thy heavenly mysteries. Through.

\communion{If any man will come after me, let him deny himself, and take up his cross, and follow me.}

\postcommunion
\lett{G}{rant,} we beseech thee, O Lord our God: that like as we in this life do gladly honour the memory of thy saints; so we may rejoice to behold them for ever. Through.


\subby{II: Common of One Martyr not a Bishop out of Eastertide}\label{CommonMartyrNotBishopII}
\fancyhead[RO,LE]{\textit{Martyr not Bishop II}}

\begin{rubric}
	Daily Office propers as in the First Common of a Martyr Bishop (p. \pageref{CommonMartyrBishopI}).
\end{rubric}

\introit
\lett{T}{he} righteous shall rejoice in the Lord, and put his trust in him: and all they that are true of heart shall be glad. \textit{Ps.} Hear my voice, O God, in my prayer: preserve my life from fear of the enemy.

\collect
\lett{G}{rant,} we beseech thee, almighty God: that, at the intercession of blessed \textit{N.}, thy Martyr, we may both be delivered from all adversities which may happen to the body, and from all evil thoughts, which may assault and hurt the soul. Through.

\readingcitation{Epistle}{2 Timothy 2:8}
\lett{D}{early beloved:} Remember that Jesus Christ of the seed of David was raised from the dead according to my gospel: wherein I suffer trouble, as an evil doer, even unto bonds; but the word of God is not bound. Therefore I endure all things for the elect's sakes, that they may also obtain the salvation which is in Christ Jesus with eternal glory. But thou hast fully known my doctrine, manner of life, purpose, faith, longsuffering, charity, patience, persecutions, afflictions, which came unto me at Antioch, at Iconium, at Lystra; what persecutions I endured: but out of them all the Lord delivered me. Yea, and all that will live godly in Christ Jesus shall suffer persecution.

\gradall{Though the righteous fall, he shall not be cast away: for the Lord upholdeth him with his hand. ℣. He is ever merciful, and lendeth: and his seed is blessed.}{Alleluia, alleluia. ℣. He that followeth me shall not walk in darkness: but shall have the light of eternal life. Alleluia.}

\begin{rubric}
{In Septuagesimatide or Lent, replacing the Alleluia:}
\end{rubric}\par\noindent
\tract{Blessed is the man that feareth the Lord: he hath great delight in his commandments. ℣. His seed shall be mighty upon earth: the generation of the faithful shall be blessed. ℣. Riches and plenteousness shall be in his house: and his righteousness endureth for ever.}

\readingcitation{Gospel}{Matthew 10:26}
\lett{A}{t that time:} Jesus said unto his disciples: There is nothing covered, that shall not be revealed; and hid, that shall not be known. What I tell you in darkness, that speak ye in light: and what ye hear in the ear, that preach ye upon the housetops. And fear not them which kill the body, but are not able to kill the soul: but rather fear him which is able to destroy both soul and body in hell.

Are not two sparrows sold for a farthing? and one of them shall not fall on the ground without your Father. But the very hairs of your head are all numbered. Fear ye not therefore, ye are of more value than many sparrows.

Whosoever therefore shall confess me before men, him will I confess also before my Father which is in heaven.

\offertory{Thou hast set, O Lord, a crown of pure gold upon his head: he asked life of thee, and thou gavest it him, alleluia.}

\secret
\lett{G}{rant,} O Lord, that this our bounden service may be acceptable in thy sight: that these our oblations may, by the prayers of him on whose solemnity they are offered, be made profitable unto our salvation. Through.

\communion{If any man serve me, let him follow me: and where I am, there shall also my servant be.}

\postcommunion
\lett{W}{e} beseech thee, O Lord our God, that like as we, whom thou hast refreshed by the partaking of thy sacred gift, do ofter unto thee our worship: so by the intercession of blessed \textit{N.}, thy Martyr, we may perceive the benefit of the same. Through.

\begin{rubric}
	The following Epistles and Gospel may be used instead of the ones provided above.
\end{rubric}

\readingcitation{1. Epistle}{James 1:2}\label{James12}
\lett{D}{early beloved:} Count it all joy when ye fall into divers temptations; knowing this, that the trying of your faith worketh patience. But let patience have her perfect work, that ye may be perfect and entire, wanting nothing. If any of you lack wisdom, let him ask of God, that giveth to all men liberally, and upbraideth not; and it shall be given him. But let him ask in faith, nothing wavering. For he that wavereth is like a wave of the sea driven with the wind and tossed. For let not that man think that he shall receive any thing of the Lord. A double minded man is unstable in all his ways. Let the brother of low degree rejoice in that he is exalted: but the rich, in that he is made low: because as the flower of the grass he shall pass away. For the sun is no sooner risen with a burning heat, but it withereth the grass, and the flower thereof falleth, and the grace of the fashion of it perisheth: so also shall the rich man fade away in his ways.

Blessed is the man that endureth temptation: for when he is tried, he shall receive the crown of life, which the Lord hath promised to them that love him.

\readingcitation{2. Epistle}{1 Peter 4:13}\label{1Peter413}
\lett{D}{early beloved:} Rejoice, inasmuch as ye are partakers of Christ's sufferings; that, when his glory shall be revealed, ye may be glad also with exceeding joy. If ye be reproached for the name of Christ, happy are ye; for the spirit of glory and of God resteth upon you: on their part he is evil spoken of, but on your part he is glorified. But let none of you suffer as a murderer, or as a thief, or as an evildoer, or as a busybody in other men's matters. Yet if any man suffer as a Christian, let him not be ashamed; but let him glorify God on this behalf. For the time is come that judgment must begin at the house of God: and if it first begin at us, what shall the end be of them that obey not the gospel of God? And if the righteous scarcely be saved, where shall the ungodly and the sinner appear? Wherefore let them that suffer according to the will of God commit the keeping of their souls to him in well doing, as unto a faithful Creator.

\readingcitation{Another Gospel}{John 12:24}
\lett{A}{t that time:} Jesus said unto his disciples: Verily, verily, I say unto you, Except a corn of wheat fall into the ground and die, it abideth alone: but if it die, it bringeth forth much fruit. He that loveth his life shall lose it; and he that hateth his life in this world shall keep it unto life eternal. If any man serve me, let him follow me; and where I am, there shall also my servant be: if any man serve me, him will my Father honour.

\subby{I: Common of Many Martyrs out of Eastertide}\label{CommonMartyrsI}
\fancyhead[RO,LE]{\textit{Martyrs I}}

\subbysub{I Evensong}

\gregorioscore{resources/gabc/Common/MartyrsEvensong.gabc}

℣. Be glad, O ye righteous, and rejoice in the Lord.

℟. And be joyful, all ye that are true of heart.

\antiphon{Mag.}{For theirs is the kingdom of heaven, {\dag} who despised worldly living: who have won the rewards of the kingdom, and have washed their robes in the blood of the Lamb.}

\subbysub{Mattins}

\invitatoryhymn

\gregorioscore{resources/gabc/Common/MartyrsInvitatory.gabc}

\officehymn

\gregorioscore{resources/gabc/Common/MartyrsMattins.gabc}

℣. Let the Saints be joyful with glory.

℟. Let them rejoice in their beds.

\antiphon{Ben.}{The very hairs of your head {\dag} are all numbered: fear ye not therefore, ye are of more value than many sparrows.}

\subbysub{II Evensong}

\begin{rubric}
	The Office Hymn is of I Evensong, and the Versicle is of Mattins, with the following Antiphon.
\end{rubric}

\antiphon{Mag.}{In the heavenly kingdom {\dag} rejoice the souls of the Blessed, who followed the footsteps of Christ their Master: and since for love of him they poured forth their life-blood, therefore with Christ do they exult for ever.}

\introit
\lett{L}{et} the sorrowful sighing of the prisoners, O Lord, come before thee: reward thou our neighbours seven-fold into their bosom: avenge thou the blood of thy Saints that is shed. \textit{Ps.} O God, the heathen are come into thine inheritance: thy holy temple have they defiled: and made Jerusalem an heap of stones.

\collect
\begin{inhead}
	Many Martyrs not Bishops
\end{inhead}
\lett{O}{God,} who vouchsafest unto us to celebrate the birthday of thy holy Martyrs
\textit{N. and N.}: grant that we may rejoice in the everlasting felicity of their fellowship. Through.

\begin{inhead}
	Many Martyrs Bishops
\end{inhead}
\lett{P}{rotect} us, O Lord, we beseech thee, who observe the feast of thy blessed Martyrs and Bishops \textit{N. and N.}: and grant that by their meritorious supplication we may find favour in thy sight. Through.

\readingcitation{Epistle}{Wisdom 3:1}\label{Wisdom31}
%RV:
\lett{T}{he} souls of the righteous are in the hand of God, And no torment shall touch them. In the eyes of the foolish they seemed to have died; And their departure was accounted to be their hurt, And their journeying away from us to be their ruin: But they are in peace. For even if in the sight of men they be punished, Their hope is full of immortality; And having borne a little chastening, they shall receive great good; Because God made trial of them, and found them worthy of himself. As gold in the furnace he proved them, And as a whole burnt offering he accepted them. And in the time of their visitation they shall shine forth, And as sparks among stubble they shall run to and fro. They shall judge nations, and have dominion over peoples; And the Lord shall reign over them for evermore. 

%\lett{T}{he} souls of the righteous are in the hand of God, and there shall no torment touch them. In the sight of the unwise they seemed to die: and their departure is taken for misery, And their going from us to be utter destruction: but they are in peace. For though they be punished in the sight of men, yet is their hope full of immortality. And having been a little chastised, they shall be greatly rewarded: for God proved them, and found them worthy for himself. As gold in the furnace hath he tried them, and received them as a burnt offering. And in the time of their visitation they shall shine, and run to and fro like sparks among the stubble. They shall judge the nations, and have dominion over the people, and their Lord shall reign for ever.

\gradall{God is glorious in his holy ones: fearful in praises, doing wonders. ℣. Thy right hand, O Lord, is become glorious in power: thy right hand hath dashed in pieces the enemy.}{Alleluia, alleluia. ℣. The bodies of the Saints are buried in peace, but their name liveth for evermore. Alleluia.}

\begin{rubric}
{In Septuagesimatide or Lent, replacing the Alleluia:}
\end{rubric}\par\noindent
\tract{They that sow in tears, shall reap in joy. ℣. He that now goeth on his way weeping, and beareth forth good seed. ℣. Shall doubtless come again with joy, and bring his sheaves with him.}

\readingcitation{Gospel}{Luke 21:9}
\lett{A}{t that time:} Jesus said unto his disciples: When ye shall hear of wars and commotions, be not terrified: for these things must first come to pass; but the end is not by and by. Then said he unto them, Nation shall rise against nation, and kingdom against kingdom: and great earthquakes shall be in divers places, and famines, and pestilences; and fearful sights and great signs shall there be from heaven. But before all these, they shall lay their hands on you, and persecute you, delivering you up to the synagogues, and into prisons, being brought before kings and rulers for my name's sake. And it shall turn to you for a testimony. Settle it therefore in your hearts, not to meditate before what ye shall answer: for I will give you a mouth and wisdom, which all your adversaries shall not be able to gainsay nor resist. And ye shall be betrayed both by parents, and brethren, and kinsfolks, and friends; and some of you shall they cause to be put to death. And ye shall be hated of all men for my name's sake. But there shall not an hair of your head perish. In your patience possess ye your souls.

\begin{rubric}
	After Septuagesima, \emph{Alleluia} at the end of the following Offertory is omitted.
\end{rubric}

\offertory{God is wonderful in his holy ones: even the God of Israel, he will give strength and power unto his people: blessed be God, alleluia.}

\secret
\begin{inhead}
	Many Martyrs not Bishops
\end{inhead}
\lett{W}{e} beseech thee, O Lord, that the gifts which we offer unto thee of our bounden duty and service may be acceptable unto thee for the honour of thy Just ones: and by thy mercy profitable unto us for our salvation. Through.

\begin{inhead}
	Many Martyrs Bishops
\end{inhead}
\lett{A}{ssist} us mercifully, O Lord, in these our supplications which we make before thee in remembrance of thy Saints: that we who trust not in our own righteousness may be succoured by the merits of them that have found favour in thy sight. Through.

\communion{Though they be punished in the sight of men, yet hath God proved them: as gold in the furnace hath he tried them, and received them as a burnt offering.}

\postcommunion
\begin{inhead}
	Many Martyrs not Bishops
\end{inhead}
\lett{G}{rant} to us, we beseech thee, O Lord: at the intercession of thy holy Martyrs \textit{N. and N.}; that those things which we touch with our mouths we may receive in purity of heart. Through.

\begin{inhead}
	Many Martyrs Bishops
\end{inhead}
\lett{O}{Lord,} who hast fulfilled us with saving mysteries: we beseech thee that we may be aided by the prayers of those whose festival we celebrate. Through.


\subby{II: Common of Many Martyrs out of Eastertide}\label{CommonMartyrsII}
\fancyhead[RO,LE]{\textit{Martyrs II}}
\begin{rubric}
	Daily Office propers as in the First Common of Many Martyrs (p. \pageref{CommonMartyrsI}).
\end{rubric}

\begin{rubric}
	The Collect, Secret, \& Postcommunion are of the preceding Mass.
\end{rubric}

\introit
\lett{L}{et} the people tell of the wisdom of the Saints, and let the church shew forth their praise: their names shall live for evermore. \textit{Ps.} Rejoice in the Lord, O ye righteous: for it becometh well the just to be thankful.

\readingcitation{Epistle}{Wisdom 5:15}
%RV:
\lett{B}{ut} the righteous live for ever, And in the Lord is their reward, And the care for them with the Most High. Therefore shall they receive the crown of royal dignity And the diadem of beauty from the Lord's hand; Because with his right hand shall he cover them, And with his arm shall he shield them. He shall take his jealousy as complete armour, And shall make the whole creation his weapons for vengeance on his enemies: He shall put on righteousness as a breastplate, And shall array himself with judgement unfeigned as with a helmet; He shall take holiness as an invincible shield.

%\lett{B}{ut} the righteous live for evermore; their reward also is with the Lord, and the care of them is with the most High. Therefore shall they receive a glorious kingdom, and a beautiful crown from the Lord's hand: for with his right hand shall he cover them, and with his arm shall he protect them. He shall take to him his jealousy for complete armour, and make the creature his weapon for the revenge of his enemies. He shall put on righteousness as a breastplate, and true judgment instead of an helmet. He shall take holiness for an invincible shield.

\gradall{Our soul is escaped even as a bird out of the snare of the fowler. ℣. The snare is broken, and we are delivered: our help standeth in the name of the Lord, who hath made heaven and earth.}{Alleluia, alleluia. ℣. Let the righteous be glad and rejoice before God: let them also be merry and joyful. Alleluia.}

\begin{rubric}
{In Septuagesimatide or Lent, replacing the Alleluia:}
\end{rubric}\par\noindent
\tract{They that sow in tears, shall reap in joy. ℣. He that now goeth on his way weeping, and beareth forth good seed. ℣. Shall doubtless come again with joy, and bring his sheaves with him.}

\readingcitation{Gospel}{Luke 6:17}
\lett{A}{t that time:} Jesus came down from the mountain, and stood in the plain, and the company of his disciples, and a great multitude of people out of all Jud{\ae}a and Jerusalem, and from the sea coast of Tyre and Sidon, which came to hear him, and to be healed of their diseases; and they that were vexed with unclean spirits: and they were healed. And the whole multitude sought to touch him: for there went virtue out of him, and healed them all.

And he lifted up his eyes on his disciples, and said, Blessed be ye poor: for yours is the kingdom of God. Blessed are ye that hunger now: for ye shall be filled. Blessed are ye that weep now: for ye shall laugh. Blessed are ye, when men shall hate you, and when they shall separate you from their company, and shall reproach you, and cast out your name as evil, for the Son of man’s sake. Rejoice ye in that day, and leap for joy: for, behold, your reward is great in heaven.

\begin{rubric}
	After Septuagesima, \emph{Alleluia} at the end of the following Offertory is omitted.
\end{rubric}

\offertory{Let the saints be joyful with glory, let them rejoice in their beds: let the praises of God be in their mouth, alleluia.}

\communion{I say unto you, my friends: Be not afraid of them that persecute you.}


\subby{III: Common of Many Martyrs out of Eastertide}\label{CommonMartyrsIII}
\fancyhead[RO,LE]{\textit{Martyrs III}}
\begin{rubric}
	Daily Office propers as in the First Common of Many Martyrs (p. \pageref{CommonMartyrsI}).
\end{rubric}

\introit
\lett{B}{ut} the salvation of the righteous cometh of the Lord: who is also their strength in the time of trouble. \textit{Ps.} Fret not thyself because of the ungodly: neither be thou envious against the evil doers.

\collect
\lett{O}{God,} who makest us glad with the yearly solemnity of thy holy Martyrs \textit{N. and N.}: mercifully grant; that as we do rejoice in their merits, so we may be enkindled by their example. Through.

\readingcitation{Epistle}{Hebrews 10:32}\label{Hebrews1032}
\lett{B}{rethren:} Call to remembrance the former days, in which, after ye were illuminated, ye endured a great fight of afflictions; partly, whilst ye were made a gazingstock both by reproaches and afflictions; and partly, whilst ye became companions of them that were so used. For ye had compassion of me in my bonds, and took joyfully the spoiling of your goods, knowing in yourselves that ye have in heaven a better and an enduring substance. Cast not away therefore your confidence, which hath great recompence of reward. For ye have need of patience, that, after ye have done the will of God, ye might receive the promise. For yet a little while, and he that shall come will come, and will not tarry. Now the just shall live by faith.

\gradall{The righteous cry, and the Lord heareth them: and delivereth them out of all their troubles. ℣. The Lord is nigh unto them that are of a contrite heart: and will save such as be of an humble spirit.}{Alleluia, alleluia. ℣. The noble army of Martyrs praise thee, O Lord. Alleluia.}

\begin{rubric}
{In Septuagesimatide or Lent, replacing the Alleluia:}
\end{rubric}\par\noindent
\tract{They that sow in tears, shall reap in joy. ℣. He that now goeth on his way weeping, and beareth forth good seed. ℣. Shall doubtless come again with joy, and bring his sheaves with him.}

\readingcitation{Gospel}{Luke 12:1}
\lett{A}{t that time:} Jesus said unto his disciples: Beware ye of the leaven of the Pharisees, which is hypocrisy. For there is nothing covered, that shall not be revealed; neither hid, that shall not be known. Therefore whatsoever ye have spoken in darkness shall be heard in the light; and that which ye have spoken in the ear in closets shall be proclaimed upon the housetops. And I say unto you my friends, Be not afraid of them that kill the body, and after that have no more that they can do. But I will forewarn you whom ye shall fear: Fear him, which after he hath killed hath power to cast into hell; yea, I say unto you, Fear him. Are not five sparrows sold for two farthings, and not one of them is forgotten before God? But even the very hairs of your head are all numbered. Fear not therefore: ye are of more value than many sparrows. Also I say unto you, Whosoever shall confess me before men, him shall the Son of man also confess before the angels of God.

\begin{rubric}
	After Spetuagesima, \emph{Alleluia} at the end of the following Offertory is omitted.
\end{rubric}

\offertory{The souls of the just are in the hand of God, and there shall no torment of malice touch them: in the sight of the unwise they seemed to die: but they are in peace, alleluia.}

\secret
\lett{W}{e} beseech thee mercifully to accept these our oblations: that at the intercession of thy holy Martyrs \textit{N. and N.} we may be defended against all adversities. Through.

\communion{What I tell you in darkness, that speak ye in light, saith the Lord: and what ye hear in the ear, that preach ye upon the housetops.}

\postcommunion
\lett{M}{ay} this communion, O Lord, cleanse us from guilt: and at the intercession of thy holy Martyrs \textit{N. and N.} make us partakers of thy heavenly healing. Through.

\begin{rubric}
	The following Epistles and Gospel may be used instead of the ones provided above.
\end{rubric}

\readingcitation{1. Epistle}{Wisdom 10:17}\label{Wisdom1017}
%RV:
\lett{W}{isdom} rendered unto holy men a reward of their toils; She guided them along a marvellous way, And became unto them a covering in the day-time, And a flame of stars through the night. She brought them over the Red sea, And led them through much water; But their enemies she drowned, And out of the bottom of the deep she cast them up. Therefore the righteous spoiled the ungodly; And they sang praise to thy holy name, O Lord, And extolled with one accord thy hand that fought for them, O Lord our God.

%\lett{G}{od} rendered to the righteous a reward of their labours, guided them in a marvellous way, and was unto them for a cover by day, and a light of stars in the night season; Brought them through the Red sea, and led them through much water: But he drowned their enemies, and cast them up out of the bottom of the deep. Therefore the righteous spoiled the ungodly, and praised thy holy name, O Lord, and magnified with one accord thine hand, that fought for them, O Lord our God.

\readingcitation{2. Epistle}{Romans 5:1}\label{Romans51}
\lett{B}{rethren:} Being justified by faith, we have peace with God through our Lord Jesus Christ: by whom also we have access by faith into this grace wherein we stand, and rejoice in hope of the glory of God. And not only so, but we glory in tribulations also: knowing that tribulation worketh patience; and patience, experience; and experience, hope: and hope maketh not ashamed; because the love of God is shed abroad in our hearts by the Holy Ghost which is given unto us.

\readingcitation{3. Epistle}{Romans 8:18}\label{Romans818}
\lett{B}{rethren:} I reckon that the sufferings of this present time are not worthy to be compared with the glory which shall be revealed in us. For the earnest expectation of the creature waiteth for the manifestation of the sons of God. For the creature was made subject to vanity, not willingly, but by reason of him who hath subjected the same in hope, because the creature itself also shall be delivered from the bondage of corruption into the glorious liberty of the children of God. For we know that the whole creation groaneth and travaileth in pain together until now. And not only they, but ourselves also, which have the firstfruits of the Spirit, even we ourselves groan within ourselves, waiting for the adoption, to wit, the redemption of our body: in Christ Jesus our Lord.

\readingcitation{4. Epistle}{2 Corinthians 6:4}\label{2Corinthians64}
\lett{B}{rethren:} Let us approve ourselves as the ministers of God, in much patience, in afflictions, in necessities, in distresses, in stripes, in imprisonments, in tumults, in labours, in watchings, in fastings; by pureness, by knowledge, by longsuffering, by kindness, by the Holy Ghost, by love unfeigned, by the word of truth, by the power of God, by the armour of righteousness on the right hand and on the left, by honour and dishonour, by evil report and good report: as deceivers, and yet true; as unknown, and yet well known; as dying, and, behold, we live; as chastened, and not killed; as sorrowful, yet alway rejoicing; as poor, yet making many rich; as having nothing, and yet possessing all things.

\readingcitation{5. Epistle}{Hebrews 11:33}\label{Hebrews1133}
\lett{B}{rethren:} The saints through faith subdued kingdoms, wrought righteousness, obtained promises, stopped the mouths of lions, quenched the violence of fire, escaped the edge of the sword, out of weakness were made strong, waxed valiant in fight, turned to flight the armies of the aliens: women received their dead raised to life again: and others were tortured, not accepting deliverance, that they might obtain a better resurrection: and others had trial of cruel mockings and scourgings, yea, moreover of bonds and imprisonment: they were stoned, they were sawn asunder, were tempted, were slain with the sword: they wandered about in sheepskins and goat-skins, being destitute, afflicted, tormented: (of whom the world was not worthy;) they wandered in deserts, and 1a mountains, and in dens and caves of the earth. And these all obtained a good report through faith, and were found in Christ Jesus our Lord.

\readingcitation{6. Epistle}{Revelation 7:13}\label{Revelation713}
\lett{I}{n those days:} One of the elders answered, saying unto me, What are these which are arrayed in white robes? and whence came they? And I said unto him, Sir, thou knowest. And he said to me, These are they which came out of great tribulation, and have washed their robes, and made them white in the blood of the Lamb. Therefore are they before the throne of God, and serve him day and night in his temple: and he that sitteth on the throne shall dwell among them. They shall hunger no more, neither thirst any more; neither shall the sun light on them, nor any heat. For the Lamb which is in the midst of the throne shall feed them, and shall lead them unto living fountains of waters: and God shall wipe away all tears from their eyes.

\readingcitation{1. Gospel}{Matthew 24:3}\label{Matthew243}
\lett{A}{t that time:} As Jesus sat upon the mount of Olives, the disciples came unto him privately, saying: Tell us, when shall these things be? and what shall be the sign of thy coming, and of the end of the world? And Jesus answered and said unto them, Take heed that no man deceive you. For many shall come in my name, saying, I am Christ; and shall deceive many. And ye shall hear of wars and rumours of wars: see that ye be not troubled: for all these things must come to pass, but the end is not yet. For nation shall rise against nation, and kingdom against kingdom: and there shall be famines, and pestilences, and earthquakes, in divers places. All these are the beginning of sorrows.

Then shall they deliver you up to be afflicted, and shall kill you: and ye shall be hated of all nations for my name’s sake. And then shall many be offended, and shall betray one another, and shall hate one another. And many false prophets shall rise, and shall deceive many. And because iniquity shall abound, the love of many shall wax cold. But he that shall endure unto the end, the same shall be saved.

\readingcitation{2. Gospel}{Matthew 5:1}
\lett{A}{t that time:} Jesus, seeing the multitudes, went up into a mountain, and when he was set, his disciples came unto him, and he opened mouth, and taught them, saying: Blessed are the poor in spirit: for theirs is the kingdom of heaven. Blessed are they that mourn: for they shall be comforted. Blessed are the meek: for they shall inherit the earth. Blessed are they which do hunger and thirst after righteousness: for they shall be filled. Blessed are the merciful: for they shall obtain mercy. Blessed are the pure in heart: for they shall see God. Blessed are the peacemakers: for they shall be called the children of God. Blessed are they which are persecuted for righteousness’ sake: for theirs is the kingdom of heaven. Blessed are ye, when men shall revile you, and persecute you, and shall say all manner of evil against you falsely, for my sake. Rejoice, and be exceeding glad: for great is your reward in heaven.

\readingcitation{3. Gospel}{Matthew 11:25}
\lett{A}{t that time:} Jesus answered and said: I thank thee, O Father, Lord of heaven and earth, because thou hast hid these things from the wise and prudent, and hast revealed them unto babes. Even so, Father: for so it seemed good in thy sight. All things are delivered unto me of my Father: and no man knoweth the Son, but the Father; neither knoweth any man the Father, save the Son, and he to whomsoever the Son will reveal him.

Come unto me, all ye that labour and are heavy laden, and I will give you rest. Take my yoke upon you, and learn of me; for I am meek and lowly in he art: and ye shall find rest unto your souls. For my yoke is easy, and my burden is light.

\readingcitation{4. Gospel}{Luke 11:47}\label{Luke1147}
\lett{A}{t that time:} Jesus said unto the scribes and Pharisees: Woe unto you! for ye build the sepulchres of the prophets, and your fathers killed them. Truly ye bear witness that ye allow the deeds of your fathers: for they indeed killed them, and ye build their sepulchres. Therefore also said the wisdom of God, I will send them prophets and apostles, and some of them they shall slay and persecute: that the blood of all the prophets, which was shed from the foundation of the world, may be required of this generation; from the blood of Abel unto the blood of Zacharias, which perished between the altar and the temple: verily I say unto you, It shall be required of this generation.

\readingcitation{5. Gospel}{Luke 10:16}\label{Luke1016}
\lett{A}{t that time:} Jesus said unto his disciples: He that heareth you heareth me; and he that despiseth you despiseth me; and he that despiseth me despiseth him that sent me. And the seventy returned again with joy, saying, Lord, even the devils are subject unto us through thy name. And he said unto them, I beheld Satan as lightning fall from heaven. Behold, I give unto you power to tread on serpents and scorpions, and over all the power of the enemy: and nothing shall by any means hurt you. Notwithstanding in this rejoice not, that the spirits are subject unto you; but rather rejoice, because your names are written in heaven.


\subby{Common of One Martyr in Eastertide}\label{CommonMartyrEaster}
\fancyhead[RO,LE]{\textit{Martyr in Easter}}
\begin{rubric}
	Daily Office Antiphons \& Versicles as in the Common of Apostles in Eastertide (p. \pageref{CommonApostles}), with the following Hymns.
\end{rubric}

\subbysub{I Evensong}

\gregorioscore{resources/gabc/Common/MartyrEvensongEaster.gabc}

    \par\noindent
	\centerline{\rule{0.25\textwidth}{0.4pt}}
	\par\noindent
\begin{rubric}
	From Low Sunday, exclusive, until Ascension Thursday, exclusive:
\end{rubric}

\gregorioscore{resources/gabc/Common/MartyrEvensongEasterEnding1.gabc}

\begin{rubric}
	From Ascension Thursday, inclusive, until Whitsunday, inclusive:
\end{rubric}

\gregorioscore{resources/gabc/Common/MartyrEvensongEasterEnding2.gabc}

    \par\noindent
	\centerline{\rule{0.25\textwidth}{0.4pt}}
	\par


\subbysub{Mattins}

\begin{rubric}
	The Invitatory Hymn is of I Evensong.
\end{rubric}

\officehymn

\gregorioscore{resources/gabc/Common/MartyrMattinsEaster.gabc}

    \par\noindent
	\centerline{\rule{0.25\textwidth}{0.4pt}}
	\par\noindent
\begin{rubric}
	From Low Sunday, exclusive, until Ascension Thursday, exclusive:
\end{rubric}

\gregorioscore{resources/gabc/Common/MartyrMattinsEasterEnding1.gabc}

\begin{rubric}
	From Ascension Thursday, inclusive, until Whitsunday, inclusive:
\end{rubric}

\gregorioscore{resources/gabc/Common/MartyrMattinsEasterEnding2.gabc}

    \par\noindent
	\centerline{\rule{0.25\textwidth}{0.4pt}}
	\par

\subbysub{II Evensong}

\begin{rubric}
	The Office Hymn is of I Evensong.
\end{rubric}

\introit
\lett{T}{hou} hast hidden me, O God, from the gathering together of the froward, alleluia: and from the insurrection of the wicked doers, alleluia, alleluia. \textit{Ps.} Hear my voice, O God, in my prayer: preserve my lite from fear of the enemy.

\collect
\begin{inhead}
	Martyr not Bishop
\end{inhead}
\lett{G}{rant,} we beseech thee, almighty God: that we, who devoutly celebrate the birthday of blessed \textit{N.}, thy Martyr; may by his intercession be stablished in the love of thy name. Through.
\begin{inhead}
	or
\end{inhead}
\lett{G}{rant,} we beseech thee, almighty God: that at the intercession of blessed \textit{N.} thy Martyr, we may both be delivered from all adversities which may happen to the body, and from all evil thoughts which may assault and hurt the soul. Through.

%MANUAL ADJUSTMENT:
\clearpage
\begin{inhead}
	Martyr Bishop
\end{inhead}
\lett{A}{lmighty} God, mercifully look upon our infirmities: that whereas we are oppressed by the burden of our sins, the glorious intercession of blessed \textit{N.} thy Martyr and Bishop may be our succour and defence. Through.
\begin{inhead}
	or
\end{inhead}
\lett{O}{God,} who makest us glad with the yearly solemnity of blessed \textit{N.}, thy Martyr and Bishop: mercifully grant; that as we now devoutly celebrate his birthday, so we may likewise rejoice in his protection. Through.

\readingcitation{Epistle}{Wisdom 5:1}
%RV:
\lett{T}{he} righteous man shall stand in great boldness Before the face of them that afflicted him, And them that make his labours of no account. When they see it, they shall be troubled with terrible fear, And shall be amazed at the marvel of God's salvation. They shall say within themselves repenting, And for distress of spirit shall they groan, This was he whom aforetime we had in derision, And made a parable of reproach: We fools accounted his life madness, And his end without honour: How was he numbered among sons of God? And how is his lot among saints?

%\lett{T}{he} righteous man shall stand in great boldness before the face of such as have afflicted him, and made no account of his labours. When they see it, they shall be troubled with terrible fear, and shall be amazed at the strangeness of his salvation, so far beyond all that they looked for. And they repenting and groaning for anguish of spirit shall say within themselves, This was he, whom we had sometimes in derision, and a proverb of reproach: We fools accounted his life madness, and his end to be without honour: How is he numbered among the children of God, and his lot is among the saints!

\begin{rubric}
	The Epistle from the Third Common of One Martyr not a Bishop out of Eastertide (2 Timothy 2:8) may be said instead.
\end{rubric}

\alleluia{Alleluia, alleluia. ℣. O Lord, the very heavens shall praise thy wondrous works: and thy truth in the congregation of the saints. Alleluia. ℣. Thou hast set, O Lord, a crown of pure gold upon his head. Alleluia.}

\readingcitation{Gospel}{John 15:1}
\lett{A}{t that time:} Jesus said unto his disciples: I am the true vine, and my Father is the husbandman. Every branch in me that beareth not fruit he taketh away: and every branch that beareth fruit, he purgeth it, that it may bring forth more fruit. Now ye are clean through the word which I have spoken unto you. Abide in me, and I in you. As the branch cannot bear fruit of itself, except it abide in the vine; no more can ye, except ye abide in me. I am the vine, ye are the branches: He that abideth in me, and I in him, the same bringeth forth much fruit: for without me ye can do nothing. If a man abide not in me, he is cast forth as a branch, and is withered; and men gather them, and cast them into the fire, and they are burned. If ye abide in me, and my words abide in you, ye shall ask what ye will, and it shall be done unto you.

\offertory{O Lord, the very heavens shall praise thy wondrous works: and thy truth in the congregation of the saints, alleluia, alleluia.}

\secret
\begin{inhead}
	Martyr not Bishop
\end{inhead}
\lett{W}{e} beseech thee, O Lord, accept our prayers and oblations: and graciously hearken unto us whom thou dost cleanse by thy heavenly mysteries. Through.
\begin{inhead}
	or
\end{inhead}
\lett{G}{rant,} O Lord, that our bounden service may be acceptable in thy sight: that these our oblations may, by the prayers of him on whose solemnity they are offered, be made profitable unto our salvation. Through.

\begin{inhead}
	Martyr Bishop
\end{inhead}
\lett{W}{e} beseech thee, O Lord, mercifully to accept this our sacrifice, which we offer unto thee pleading the merits of blessed \textit{N.}, thy Martyr and Bishop: that the same may effectually avail for our perpetual succour. Through.

\begin{inhead}
	or
\end{inhead}
\lett{S}{anctify,} O Lord, the gifts which we dedicate to thee: that at the intercession of blessed \textit{N.}, thy Martyr and Bishop, they may obtain for us thy gracious favour. Through,

\communion{The righteous shall rejoice in the Lord, and put his trust in him: and all they that are true of heart shall be glad, alleluia, alleluia.}

\postcommunion
\begin{inhead}
	Martyr not Bishop
\end{inhead}
\lett{G}{rant,} we beseech thee, O Lord our God: that like as we in this life do gladly honour the memory of thy Saints; so we may rejoice to behold them for ever. Through.
\begin{inhead}
	or
\end{inhead}
\lett{W}{e} beseech thee, O Lord our God: that like as we, whom thou hast refreshed by the partaking of thy sacred gift, do offer unto thee our worship; so by the intercession of blessed \textit{N.}, thy Martyr, we may perceive the benefit of the same. Through.

\begin{inhead}
	Martyr Bishop
\end{inhead}
\lett{W}{e} beseech thee, O Lord our God, that like as we, whom thou hast refreshed by the partaking of thy sacred gift, do offer unto thee our worship: so by the intercession of blessed \textit{N.}, thy Martyr and Bishop, we may perceive the benefit of the same. Through.
\begin{inhead}
	or
\end{inhead}
\lett{M}{ay} this communion, O Lord, cleanse us from guilt: and, at the intercession of blessed \textit{N.}, thy Martyr and Bishop, make us partakers of thy heavenly healing. Through,

\begin{rubric}
	The first Epistle and the Gospel of the above Mass may be said also in the following Mass, and those in the following may be said in the above, provided a special Epistle or Gospel is not appointed.
\end{rubric}

%MANUAL ADJUSTMENT:
\clearpage
\subby{Common of Many Martyrs in Eastertide}\label{CommonMartyrsEaster}
\fancyhead[RO,LE]{\textit{Martyrs in Easter}}

\begin{rubric}
	Daily Office Antiphons \& Versicles as in the Common of Apostles in Eastertide (p. \pageref{CommonApostles}), with the following Hymns.
\end{rubric}

\subbysub{I Evensong}

\gregorioscore{resources/gabc/Common/MartyrsEvensongEaster.gabc}

    \par\noindent
	\centerline{\rule{0.25\textwidth}{0.4pt}}
	\par\noindent
\begin{rubric}
	From Low Sunday, exclusive, until Ascension Thursday, exclusive:
\end{rubric}

\gregorioscore{resources/gabc/Common/MartyrsEvensongEasterEnding1.gabc}

\begin{rubric}
	From Ascension Thursday, inclusive, until Whitsunday, inclusive:
\end{rubric}

\gregorioscore{resources/gabc/Common/MartyrsEvensongEasterEnding2.gabc}

    \par\noindent
	\centerline{\rule{0.25\textwidth}{0.4pt}}
	\par

\subbysub{Mattins}

\begin{rubric}
	The Invitatory Hymn is from the First Common of Many Martyrs (p. \pageref{CommonMartyrsI}).
\end{rubric}

\begin{rubric}
	The Office Hymn is as in I Evensong.
\end{rubric}

\subbysub{II Evensong}

\begin{rubric}
	The Hymn is as in I Evensong.
\end{rubric}

\introit
\lett{T}{hy} saints give thanks unto thee, O Lord: they shew the glory of thy kingdom, alleluia, alleluia. \textit{Ps.} I will magnify thee, O God, my King: and I will praise thy name for ever and ever.

\collect
\begin{inhead}
	Many Martyrs not Bishops
\end{inhead}
\lett{O}{God,} who vouchsafest unto us to celebrate the birthday of thy holy Martyrs \textit{N. and N.}: grant that we may rejoice in the everlasting felicity of their fellowship. Through.
\begin{inhead}
	or
\end{inhead}
\lett{O}{God,} who makest us glad with the yearly solemnity of thy holy Martyrs \textit{N. and N.}: mercifully grant that as we rejoice in their merits, so we may be enkindled by their example. Through.

\begin{inhead}
	Martyrs Bishops
\end{inhead}
\lett{P}{rotect} us, O Lord, we beseech thee, who observe the feast of thy blessed Martyrs and Bishops \textit{N. and N.}: and grant that by their meritorious supplication we may ever find favour in thy sight. Through.

\readingcitation{Epistle}{1 Peter 1:3}
\lett{B}{lessed} be the God and Father of our Lord Jesus Christ, which according to his abundant mercy hath begotten us again unto a lively hope by the resurrection of Jesus Christ from the dead, to an inheritance incorruptible, and undefiled, and that fadeth not away, reserved in heaven for you, who are kept by the power of God through faith unto salvation ready to be revealed in the last time. Wherein ye greatly rejoice, though now for a season, if need be, ye are in heaviness through manifold temptations: that the trial of your faith, being much more precious than of gold that perisheth, though it be tried with fire, might be found unto praise and honour and glory at the appearing of Jesus Christ our Lord.

\alleluia{Alleluia, alleluia. ℣. Thy saints, O Lord, shall grow as the lily, and as the odour of balsam shall they be before thee. Alleluia. ℣. Right dear in the sight of the Lord is the death of his saints. Alleluia.}

\readingcitation{Gospel}{John 15:5}
\lett{A}{t that time:} Jesus said unto his disciples: I am the vine, ye are the branches: He that abideth in me, and I in him, the same bringeth forth much fruit: for without me ye can do nothing. If a man abide not in me, he is cast forth as a branch, and is withered; and men gather them, and cast them into the fire, and they are burned. If ye abide in me, and my words abide in you, ye shall ask what ye will, and it shall be done unto you. Herein is my Father glorified, that ye bear much fruit; so shall ye be my disciples. As the Father hath loved me, so have I loved you: continue ye in my love. If ye keep my commandments, ye shall abide in my love; even as I have kept my Father’s commandments, and abide in his love. These things have I spoken unto you, that my joy might remain in you, and that your joy might be full.

\offertory{Be glad, O ye righteous, and rejoice in the Lord: and be joyful, all ye that are true of heart, alleluia, alleluia.}

\secret
\begin{inhead}
	Martyrs not Bishops
\end{inhead}
\lett{W}{e} beseech thee, O Lord, that the gifts which we offer unto thee of our bounden duty and service may be acceptable unto thee for the honour of thy Just ones: and by thy mercy profitable unto us for our salvation. Through.

\begin{inhead}
	or
\end{inhead}

\lett{W}{e} beseech thee, O Lord, mercifully to accept these our oblations: that at the intercession of thy holy Martyrs \textit{N. and N.} we may be defended against all adversities. Through.

\begin{inhead}
	Martyrs Bishops
\end{inhead}
\lett{A}{ssist} us mercifully, O Lord, in these our supplications, which we make before thee in remembrance of thy Saints: that we who trust not in our own righteousness may be succoured by the merits of them that have found favour in thy sight. Through.

\communion{Rejoice in the Lord, O ye righteous, alleluia: for it becometh well the just to be thankful, alleluia.}

\postcommunion
\begin{inhead}
	Martyrs not Bishops
\end{inhead}
\lett{G}{rant} to us, we beseech thee, O Lord: at the intercession of thy holy Martyrs \textit{N. and N.}; that those things which we touch with the mouth, we may receive in purity of heart. Through.

\begin{inhead}
	or
\end{inhead}

\lett{M}{ay} this communion, O Lord, cleanse us from guilt: and, at the intercession of thy holy Martyrs \textit{N. and N.}, make us partakers of thy heavenly healing. Through.

\begin{inhead}
	Martyrs Bishops
\end{inhead}
\lett{O}{Lord,} who hast fulfilled us with saving mysteries: we beseech thee that we may be aided by the prayers of those whose festival we celebrate. Through.

\begin{rubric}
	The following Epistle and Gospel may be used instead of the ones provided above.
\end{rubric}

\readingcitation{Another Epistle}{Revelation 19:1}\label{Revelation191}
\lett{I}{n those days:} After these things I, John, heard a great voice of much people in heaven, saying, Alleluia; Salvation, and glory, and honour, and power, unto the Lord our God: for true and righteous are his judgments: for he hath judged the great whore, which did corrupt the earth with her fornication, and hath avenged the blood of his servants at her hand. And again they said, Alleluia. And her smoke rose up for ever and ever. And the four and twenty elders and the four beasts fell down and worshipped God that sat on the throne, saying, Amen; Alleluia.

And a voice came out of the throne, saying, Praise our God, all ye his servants, and ye that fear him, both small and great. And I heard as it were the voice of a great multitude, and as the voice of many waters, and as the voice of mighty thunderings, saying, Alleluia: for the Lord God omnipotent reigneth. Let us be glad and rejoice, and give honour to him: for the marriage of the Lamb is come, and his wife hath made herself ready. And to her was granted that she should be arrayed in fine linen, clean and white: for the fine linen is the righteousness of saints.

And he saith unto me, Write, Blessed are they which are called unto the marriage supper of the Lamb. 

\readingcitation{Another Gospel}{John 16:20}\label{John1620}
\lett{A}{t that time:} Jesus said unto his disciples: Verily, verily, I say unto you, That ye shall weep and lament, but the world shall rejoice: and ye shall be sorrowful, but your sorrow shall be turned into joy. A woman when she is in travail hath sorrow, because her hour is come: but as soon as she is delivered of the child, she remembereth no more the anguish, for joy that a man is born into the world. And ye now therefore have sorrow: but I will see you again, and your heart shall rejoice, and your joy no man taketh from you.

\begin{rubric}
	In the Common of Confessors, and Virgins, and in other Masses of Eastertide, all is said as in the rest of the year, a double Alleluia being added in the Introit before the Psalm, and at the end of the Offertory and Communion one Alleluia, where it is not given; and the Gradual being omitted, two ℣℣. are said with four Alleluias, as is set down in their places.
\end{rubric}


\subby{I: Common of a Confessor Bishop}\label{CommonConfessorBishopI}
\fancyhead[RO,LE]{\textit{Confessor Bishop I}}

\subbysub{I Evensong}

\gregorioscore{resources/gabc/Common/ConfessorBishopEvensong.gabc}

℣. The Lord loved him and adorned him.

℟. He clothed him with a robe of glory.

\antiphon{Mag.}{O thou Priest and Bishop, {\dag} thou workers of mighty works, thou good shepherd of the people, pray unto the Lord for us.}

\subbysub{Mattins}

\begin{rubric}
	The Invitatory Hymn is from I Evensong.
\end{rubric}

\officehymn

\gregorioscore{resources/gabc/Common/ConfessorBishopMattins.gabc}

℣. The Lord guided the righteous in right paths.

℟. And shewed him the kingdom of God.

\antiphon{Ben.}{Well done, {\dag} thou good and faithful servant: thou hast been faithful over a few things, I will make thee ruler over many things, saith the Lord.}

\subbysub{II Evensong}

\begin{rubric}
	The Office Hymn is of I Evensong, and the Versicle is of Mattins, with the following Antiphon.
\end{rubric}

\antiphon{Mag.}{I will liken him {\dag} unto a wise man, which built his house upon a rock.}

\introit
\lett{T}{he} Lord hath established a covenant of peace with him, and made him a prince: that he should have the dignity of the priesthood for ever (Alleluia, alleluia). \textit{Ps.} Lord, remember David: and all his trouble.

\collect
\lett{G}{rant,} we beseech thee, almighty God: that the venerable solemnity of blessed \textit{N.}, thy Confessor and Bishop may increase our devotion and set forward our salvation. Through.

\readingcitation{Epistle}{Ecclesiasticus 44:16}
\lett{E}{noch} pleased the Lord, and was translated, Being an example of repentance to all generations. Noah was found perfect and righteous; In the season of wrath he was taken in exchange for the world; Therefore was there left a remnant unto the earth, When the flood came. Everlasting covenants were made with him, That all flesh should no more be blotted out by a flood. Abraham was a great father of a multitude of nations; And there was none found like him in glory; Who kept the law of the Most High, And was taken into covenant with him: In his flesh he established the covenant; And when he was proved, he was found faithful. Therefore he assured him by an oath, That the nations should be blessed in his seed; That he would multiply him as the dust of the earth, And exalt his seed as the stars, And cause them to inherit from sea to sea, And from the River unto the utmost part of the earth. In Isaac also did he establish likewise, for Abraham his father’s sake, The blessing of all men, and the covenant: And he made it rest upon the head of Jacob; He acknowledged him in his blessings, And gave to him by inheritance, And divided his portions; Among twelve tribes did he part them. And he brought out of him a man of mercy, Which found favour in the sight of all flesh.

%Traditional Text:
%\lett{B}{ehold,} a great priest, who in his days pleased God, and was found just; in the time of wrath he was made a means of reconciliation. There was none like unto him who kept the law of the Most High. Therefore he assured him by an oath that he would increase him among his people. He established him with the blessing of all men and the covenant and made it rest upon his head. He acknowledged him in his blessing: he preserved for him his mercy: so that he found favour in the sight of the Lord. He made him glorious in the sight of kings, and gave him a crown of glory. An everlasting covenant he made with him, and gave him a great priesthood, and clothed him with a robe of glory. To execute the office of the priesthood, and bless the people in his name, and to offer him a worthy incense and a sweet savour.

\gradall{Behold, a great priest, who in his days pleased God. ℣. There was none found like unto him who kept the law of the Most High.}{Alleluia, alleluia. ℣. Thou art a Priest for ever after the order of Melchisedech. Alleluia.}

\begin{rubric}
{In Septuagesimatide or Lent, replacing the Alleluia:}
\end{rubric}\par\noindent
\tract{Blessed is the man that feareth the Lord: he hath great delight in his commandments. ℣. His seed shall be mighty upon earth: the generation of the faithful shall be blessed. ℣. Riches and plenteousness shall be in his house: and his righteousness endureth for ever.}

\begin{rubric}
{In Eastertide, replacing the Lesser Alleluia:}
\end{rubric}\par\noindent
\alleluia{Alleluia, alleluia. ℣. Thou art a priest for ever after the order of Melchisedech, alleluia. ℣. This is a priest whom the Lord hath crowned. Alleluia.}

\readingcitation{Gospel}{Matthew 25:14}
\lett{A}{t that time:} Jesus spake this parable unto his disciples: A man travelling into a far country, called his own servants, and delivered unto them his goods. And unto one he gave five talents, to another two, and to another one; to every man according to his several ability; and straightway took his journey. Then he that had received the five talents went and traded with the same, and made them other five talents. And likewise he that had received two, he also gained other two. But he that had received one went and digged in the earth, and hid his lord’s money. After a long time the lord of those servants cometh, and reckoneth with them. And so he that had received five talents came and brought other five talents, saying, Lord, thou deliveredst unto me five talents: behold, I have gained beside them five talents more. His lord said unto him, Well done, thou good and faithful servant: thou hast been faithful over a few things, I will make thee ruler over many things: enter thou into the joy of thy lord. He also that had received two talents came and said, Lord, thou deliveredst unto me two talents: behold, I have gained two other talents beside them. His lord said unto him, Well done, good and faithful servant; thou hast been faithful over a few things, I will make thee ruler over many things: enter thou into the joy of thy lord.

\offertory{I have found David my servant, with my holy oil have I anointed him; my hand shall hold him fast, and my arm shall strengthen him (Alleluia).}

\secret
\lett{W}{e} beseech thee, O Lord, that we, remembering with gladness the merits of thy Saints, may in all places feel the succour of their intercession. Through.

\communion{A faithful and wise steward, whom the lord hath made ruler over his household: to give them their portion of meat in due season (Alleluia).}

\postcommunion
\lett{G}{rant,} we beseech thee, almighty God: that we, shewing forth our thankfulness for the gifts which we have received, may, at the intercession of blessed \textit{N.}, thy Confessor and Bishop, obtain yet more abundant mercies. Through.


\subby{II: Common of a Confessor Bishop}\label{CommonConfessorBishopII}
\fancyhead[RO,LE]{\textit{Confessor Bishop II}}
\begin{rubric}
	Daily Office propers are as in the First Common of a Confessor Bishop (p. \pageref{CommonConfessorBishopI}).
\end{rubric}

\introit
\lett{L}{et} thy priests, O Lord, be clothed with righteousness, and let thy saints sing with joyfulness: for thy servant David's sake, turn not away the presence of thine Anointed (Alleluia, alleluia). \textit{Ps.} Lord, remember David: and all his trouble.

\collect
\lett{W}{e} beseech thee, O Lord, graciously to hear the prayers which we offer unto thee on the solemnity of blessed \textit{N.}, thy Confessor and Bishop: that, like as he was found worthy to do thee faithful service, so by his merits and intercession we may be absolved from all our sins. Through.

\readingcitation{Epistle}{Hebrews 7:23}
\lett{B}{rethren:} They were many priests, because they were not suffered to continue by reason of death: but this man, because he continueth ever, hath an unchangeable priesthood. Wherefore he is able also to save them to the uttermost that come unto God by him, seeing he ever liveth to make intercession for them.

For such an high priest became us, who is holy, harmless, undefiled, separate from sinners, and made higher than the heavens; who needeth not daily, as those high priests, to offer up sacrifice, first for his own sins, and then for the people’s: for this he did once, when he offered up himself, Jesus Christ, our Lord.

\gradall{I will deck her priests with health: and her saints shall rejoice and sing. ℣. There shall I make the horn of David to flourish: I have ordained a lantern for mine Anointed.}{Alleluia, alleluia. ℣. The Lord sware, and will not repent: Thou art a priest for ever after the order of Melchisedech. Alleluia.}

\begin{rubric}
{In Septuagesimatide or Lent, replacing the Alleluia:}
\end{rubric}\par\noindent
\tract{Blessed is the man that feareth the Lord: he hath great delight in his commandments. ℣. His seed shall be mighty upon earth: the generation of the faithful shall be blessed. ℣. Riches and plenteousness shall be in his house: and his righteousness endureth for ever.}

\begin{rubric}
{In Eastertide, replacing the Lesser Alleluia:}
\end{rubric}\par\noindent
\alleluia{Alleluia, alleluia. ℣. The Lord sware, and will not repent: Thou art a priest for ever after the order of Melchisedech. Alleluia. ℣. The Lord loved him, and adorned him: and clothed him with a robe of glory. Alleluia.}

\readingcitation{Gospel}{Matthew 24:42}
\lett{A}{t that time:} Jesus said unto his disciples: Watch, for ye know not what hour your Lord doth come. But know this, that if the goodman of the house had known in what watch the thief would come, he would have watched, and would not have suffered his house to be broken up. Therefore be ye also ready: for in such an hour as ye think not the Son of man cometh. Who then is a faithful and wise servant, whom his lord hath made ruler over his household, to give them meat in due season? Blessed is that servant, whom his lord when he cometh shall find so doing. Verily I say unto you, That he shall make him ruler over all his goods.

\offertory{My truth and my mercy shall be with him: and in my name shall his horn be exalted (Alleluia).}

\secret
\lett{W}{e} beseech thee, O Lord, that our devout observance of the yearly solemnity of blessed \textit{N.}, thy Confessor and Bishop, may render us acceptable unto thy loving-kindness: that this service of propitiation, which we duly offer, may be profitable unto him for the reward of blessedness, and obtain for us the gifts of thy grace. Through.

\communion{Blessed is the servant, whom the lord when he cometh shall find watching: verily I say unto you, He shall make him ruler over all his goods (Alleluia).}

\postcommunion
\lett{O}{God,} who rewardest the souls of them that put their trust in thee: vouchsafe; that we who keep the solemn festival of blessed \textit{N.}, thy Confessor and Bishop, may by his prayers obtain thy merciful pardon. Through.

\begin{rubric}
	The following Epistles and Gospels may be used instead of the ones provided above.
\end{rubric}

\readingcitation{1. Epistle}{Hebrews 5:1}
\lett{B}{rethren:} Every high priest taken from among men is ordained for men in things pertaining to God, that he may offer both gifts and sacrifices for sins: who can have compassion on the ignorant, and on them that are out of the way; for that he himself also is compassed with infirmity. And by reason hereof he ought, as for the people, so also for himself, to offer for sins. And no man taketh this honour unto himself, but he that is called of God, as was Aaron.

\readingcitation{2. Epistle}{Hebrews 13:7}
\lett{B}{rethren:} Remember them which have the rule over you, who have spoken unto you the word of God: whose faith follow, considering the end of their conversation. Jesus Christ the same yesterday, and to day, and for ever. Be not carried about with divers and strange doctrines. For it is a good thing that the heart be established with grace; not with meats, which have not profited them that have been occupied therein.

We have an altar, whereof they have no right to eat which serve the tabernacle. For the bodies of those beasts, whose blood is brought into the sanctuary by the high priest for sin, are burned without the camp. Wherefore Jesus also, that he might sanctify the people with his own blood, suffered without the gate. Let us go forth therefore unto him without the camp, bearing his reproach. For here have we no continuing city, but we seek one to come. By him therefore let us offer the sacrifice of praise to God continually, that is, the fruit of our lips giving thanks to his name. But to do good and to communicate forget not: for with such sacrifices God is well pleased. Obey them that have the rule over you, and submit yourselves: for they watch for your souls, as they that must give account.

\readingcitation{1. Gospel}{Luke 11:33}
\lett{A}{t that time:} Jesus said unto his disciples: No man, when he hath lighted a candle, putteth it in a secret place, neither under a bushel, but on a candlestick, that they which come in may see the light. The light of the body is the eye: therefore when thine eye is single, thy whole body also is full of light; but when thine eye is evil, thy body also is full of darkness. Take heed therefore that the light which is in thee be not darkness. If thy whole body therefore be full of light, having no part dark, the whole shall be full of light, as when the bright shining of a candle doth give thee light.

\readingcitation{2. Gospel}{Mark 13:33}\label{Mark1333}
\lett{A}{t that time:} Jesus said unto his disciples: Take ye heed, watch and pray: for ye know not when the time is. For the Son of man is as a man taking a far journey, who left his house, and gave authority to his servants, and to every man his work, and commanded the porter to watch. Watch ye therefore: for ye know not when the master of the house cometh, at even, or at midnight, or at the cockcrowing, or in the morning: lest coming suddenly he find you sleeping. And what I say unto you I say unto all, Watch.

\subby{Common of Doctors}\label{CommonDoctors}
\fancyhead[RO,LE]{\textit{Doctors}}

\begin{rubric}
	The Daily Office propers are from the closest relevant Common, except for the \emph{Magnificat} Antiphon, as followeth.
\end{rubric}

\antiphon{Mag.}{O Teacher right excellent, {\dag} O light of Holy Church, O blessed \textit{N.}, lover of the divine law: intercede for us unto the Son of God.}

\introit
\lett{I}{n} the midst of the Church he opened his mouth: and the Lord filled him with the spirit of wisdom and of understanding: and clothed him with a robe of glory. (Alleluia, alleluia.) \textit{Ps.} It is a good thing to give thanks unto the Lord: and to sing praises unto thy name, O most Highest.

\collect
\lett{O}{God,} who didst give blessed \textit{N.} unto thy people to be a minister of everlasting salvation: grant, we beseech thee; that as we have learned of him the doctrine of life on earth, so we may be found worthy to have him for our advocate in heaven. Through.

\readingcitation{Epistle}{2 Timothy 4:1}\label{2Timothy41}
\lett{D}{early beloved:} I charge thee therefore before God, and the Lord Jesus Christ, who shall judge the quick and the dead at his appearing and his kingdom; preach the word; be instant in season, out of season; reprove, rebuke, exhort with all longsuffering and doctrine. For the time will come when they will not endure sound doctrine; but after their own lusts shall they heap to themselves teachers, having itching ears; and they shall turn away their ears from the truth, and shall be turned unto fables. But watch thou in all things, endure afflictions, do the work of an evangelist, make full proof of thy ministry.\par
For I am now ready to be offered, and the time of my departure is at hand. I have fought a good fight, I have finished my course, I have kept the faith: henceforth there is laid up for me a crown of righteousness, which the Lord, the righteous judge, shall give me at that day: and not to me only, but unto all them also that love his appearing.

\gradall{The mouth of the righteous is exercised in wisdom, and his tongue will be talking of judgment. ℣. The law of his God is in his heart: and his goings shall not slide.}{Alleluia, alleluia. ℣. The Lord loved him, and adorned him: and clothed him with a robe of glory. Alleluia.}

\begin{rubric}
{In Septuagesimatide or Lent, replacing the Alleluia:}
\end{rubric}\par\noindent
\tract{Blessed is the man that feareth the Lord: he hath great delight in his commandments. ℣. His seed shall be mighty upon earth: the generation of the faithful shall be blessed. ℣. Riches and plenteousness shall be in his house: and his righteousness endureth for ever.}

\begin{rubric}
{In Eastertide, replacing the Lesser Alleluia:}
\end{rubric}\par\noindent
\alleluia{Alleluia, alleluia. ℣. The Lord loved him, and adorned him: and clothed him with a robe of glory. Alleluia. ℣. The righteous shall grow as the lily; and flourish for ever before the Lord. Alleluia.}

\readingcitation{Gospel}{Matthew 5:13}
\lett{A}{t that time:} Jesus said unto his disciples: Ye are the salt of the earth: but if the salt have lost his savour, wherewith shall it be salted? it is thenceforth good for nothing, but to be cast out, and to be trodden under foot of men. Ye are the light of the world. A city that is set on an hill cannot be hid. Neither do men light a candle, and put it under a bushel, but on a candlestick; and it giveth light unto all that are in the house. Let your light so shine before men, that they may see your good works, and glorify your Father which is in heaven.\par
Think not that I am come to destroy the law, or the prophets: I am not come to destroy, but to fulfil. For verily I say unto you, Till heaven and earth pass, one jot or one tittle shall in no wise pass from the law, till all be fulfilled. Whosoever therefore shall break one of these least commandments, and shall teach men so, he shall be called the least in the kingdom of heaven: but whosoever shall do and teach them, the same shall be called great in the kingdom of heaven.

\offertory{The righteous shall flourish like a palm-tree: and shall spread abroad like a cedar in Libanus (Alleluia).}

\secret
\begin{inhead}
	Doctor \& Bishop
\end{inhead}
\lett{M}{ay} the devout prayers of thy Bishop and Doctor, Saint \textit{N.}, never fail to succour us, O Lord: that they may render our oblations acceptable in thy sight; and may ever obtain for us thy merciful pardon. Through.

\begin{inhead}
	Doctor not a Bishop
\end{inhead}
\lett{M}{ay} the devout prayers of thy Confessor and Doctor, Saint \textit{N.}, never fail to succour us, O Lord: that they may render our oblations acceptable in thy sight; and may ever obtain for us thy merciful pardon. Through.

\communion{A faithful and wise steward, whom his lord hath made ruler over his household: to give them their portion of meat in due season (Alleluia).}

\postcommunion
\begin{inhead}
	Doctor \& Bishop
\end{inhead}
\lett{W}{e} beseech thee, O Lord, that blessed \textit{N.}, thy Bishop and illustrious Doctor, may stand before thee as our advocate: that these thy sacrifices may avail for our salvation. Through.

\begin{inhead}
	Doctor not a Bishop
\end{inhead}
\lett{W}{e} beseech thee, O Lord, that blessed \textit{N.}, thy Confessor and illustrious Doctor, may stand before thee as our advocate: that these thy sacrifices may avail for our salvation. Through.

\begin{rubric}
	The following Epistle may be used instead of the one provided above.
\end{rubric}
\readingcitation{Another Epistle}{Ecclesiasticus 39:5}\label{Ecclesiasticus395}
%RV:
\lett{T}{he} just will apply his heart to resort early to the Lord that made him, And will make supplication before the Most High, And will open his mouth in prayer, And will make supplication for his sins. If the great Lord will, He shall be filled with the spirit of understanding: He shall pour forth the words of his wisdom, And in prayer give thanks unto the Lord. He shall direct his counsel and knowledge, And in his secrets shall he meditate. He shall shew forth the instruction which he hath been taught, And shall glory in the law of the covenant of the Lord. Many shall commend his understanding; And so long as the world endureth, it shall not be blotted out: His memorial shall not depart, And his name shall live from generation to generation. Nations shall declare his wisdom, And the congregation shall tell out his praise. 
%\lett{T}{he} just will give his heart to resort early to the Lord that made him, and will pray before the most High, and will open his mouth in prayer, and make supplication for his sins. When the great Lord will, he shall be filled with the spirit of understanding: he shall pour out wise sentences, and give thanks unto the Lord in his prayer. He shall direct his counsel and knowledge, and in his secrets shall he meditate. He shall shew forth that which he hath learned, and shall glory in the law of the covenant of the Lord. Many shall commend his understanding; and so long as the world endureth, it shall not be blotted out; his memorial shall not depart away, and his name shall live from generation to generation. Nations shall shew forth his wisdom, and the congregation shall declare his praise.

\subby{I: Common of a Confessor not a Bishop}\label{CommonConfessorNotBishopI}
\fancyhead[RO,LE]{\textit{Confessor not Bishop I}}

\subbysub{I Evensong}

\begin{rubric}
	The Hymn is from the First Common of a Confessor Bishop (p. \pageref{CommonConfessorBishopI}), with the following.
\end{rubric}
	
	℣. The Lord loved him and adorned him.

	℟. He clothed him with a robe of glory.
	
\antiphon{Mag.}{I will liken him {\dag} unto a wise man, which built his house upon a rock.}

\subbysub{Mattins}

\begin{rubric}
	The Invitatory Hymn is from I Evensong of the First Common of a Confessor Bishop (p. \pageref{CommonConfessorBishopI})
\end{rubric}

\officehymn

\gregorioscore{resources/gabc/Common/ConfessorNotBishopMattins.gabc}

℣. The Lord guided the righteous in right paths.

℟. And shewed him the kingdom of God.

\antiphon{Ben.}{Well done, good and faithful servant, {\dag} thou hast been faithful over a few things, I will make thee ruler over many things: enter thou into the joy of thy Lord.}

\subbysub{II Evensong}

\begin{rubric}
	The Hymn is from I Evensong of the First Common of a Confessor Bishop (p. \pageref{CommonConfessorBishopI}), with the following.
\end{rubric}

℣. The Lord guided the righteous in right paths.

℟. And shewed him the kingdom of God.

\antiphon{Mag.}{The world, and all things earthly, {\dag} this man despised, triumphant: he laid up treasures in heaven by word and deed.}

\introit
\lett{T}{he} mouth of the righteous is exercised in wisdom, and his tongue will be talking of judgment: the law of his God is in his heart. (Alleluia, alleluia.) \textit{Ps.} Fret not thyself because of the ungodly: neither be thou envious against the evildoers.

\collect
\lett{O}{God,} who makest us glad with the yearly solemnity of blessed \textit{N.}, thy Confessor: mercifully grant; that, as we now celebrate his birthday, so we may follow the example of his life. Through.

\readingcitation{Epistle}{Ecclesiasticus 31:8}
%RV:
\lett{B}{lessed} is the man that is found without blemish, And that goeth not after gold. Who is he? and we will call him blessed: For wonderful things hath he done among his people. Who hath been tried thereby, and found perfect? Then let him glory. Who hath had the power to transgress, and hath not transgressed? And to do evil, and hath not done it? His goods shall be made sure, And the congregation shall declare his alms. 

%\lett{B}{lessed} is the man that is found without blemish, and hath not gone after gold. Who is he? and we will call him blessed: for wonderful things hath he done among his people. Who hath been tried thereby, and found perfect? then let him glory. Who might offend, and hath not offended? or done evil, and hath not done it? His goods shall be established, and the congregation shall declare his alms.

\gradall{The righteous shall flourish like a palm-tree: and shall spread abroad like a cedar in Libanus in the house of the Lord. ℣. To tell of thy loving-kindness early in the morning, and of thy truth in the night-season.}{Alleluia, alleluia. ℣. Blessed is the man that endureth temptation: for when he is tried, he shall receive the crown of life. Alleluia.}

\begin{rubric}
{In Septuagesimatide or Lent, replacing the Alleluia:}
\end{rubric}\par\noindent
\tract{Blessed is the man that feareth the Lord: he hath great delight in his commandments. ℣. His seed shall be mighty upon earth: the generation of the faithful shall be blessed. ℣. Riches and plenteousness shall be in his house: and his righteousness endureth for ever.}

\begin{rubric}
{In Eastertide, replacing the Lesser Alleluia:}
\end{rubric}\par\noindent
\alleluia{Alleluia, alleluia. ℣. Blessed is the man that endureth temptation: for when he is tried, he shall receive the crown of life. Alleluia. ℣. The Lord loved him, and adorned him: and clothed him with a robe of glory. Alleluia.}

\readingcitation{Gospel}{Luke 12:35}
\lett{A}{t that time:} Jesus said unto his disciples: Let your loins be girded about, and your lights burning; and ye yourselves like unto men that wait for their lord, when he will return from the wedding; that when he cometh and knocketh, they may open unto him immediately. Blessed are those servants, whom the lord when he cometh shall find watching: verily I say unto you, that he shall gird himself, and make them to sit down to meat, and will come forth and serve them. And if he shall come in the second watch, or come in the third watch, and find them so, blessed are those servants. And this know, that if the goodman of the house had known what hour the thief would come, he would have watched, and not have suffered his house to be broken through. Be ye therefore ready also: for the Son of man cometh at an hour when ye think not.

\offertory{My truth and my mercy shall be with him: and in my name shall his horn be exalted (Alleluia).}

\secret
\lett{G}{rant,} we beseech thee, O Lord, that we who, trusting in this our sacrifice of praise, do offer it before thee to the honour of thy Saints: may by the same be delivered from all evils both in this life and in that which is to come. Through.

\communion{Blessed is the servant whom the lord when he cometh shall find watching: verily I say unto you, he shall make him ruler over all his goods (Alleluia).}

\postcommunion
\lett{O}{Lord,} our God, who hast refreshed us with heavenly meat and drink, we humbly beseech thee: that we may be defended by the prayers of him in whose memory we have received the same. Through.


\subby{II: Common of a Confessor not a Bishop}\label{CommonConfessorNotBishopII}
\fancyhead[RO,LE]{\textit{Confessor not Bishop II}}

\begin{rubric}
	Daily Office propers are as in the First Common of a Confessor not a Bishop (p. \pageref{CommonConfessorNotBishopI}).
\end{rubric}

\introit
\lett{T}{he} just shall flourish like a palm-tree: and shall spread abroad like a cedar in Libanus: planted in the house of the Lord: in the courts of the house of our God. (Alleluia, alleluia.) \textit{Ps.} It is a good thing to give thanks unto the Lord: and to sing praises unto thy name, O most Highest.

\collect
\lett{A}{ssist} us mercifully, O Lord, in these our supplications which we make before thee on the solemnity of blessed \textit{N.}, thy Confessor: that we, who put not our trust in our own righteousness, may be succoured by the prayers of him who found favour in thy sight. Through.

\readingcitation{Epistle}{1 Corinthians 4:9}
\lett{B}{rethren:} We are made a spectacle unto the world, and to angels, and to men. We are fools for Christ's sake, but ye are wise in Christ; we are weak, but ye are strong; ye are honourable, but we are despised. Even unto this present hour we both hunger, and thirst, and are naked, and are buffeted, and have no certain dwellingplace; and labour, working with our own hands: being reviled, we bless; being persecuted, we suffer it: being defamed, we intreat: we are made as the filth of the world, and are the offscouring of all things unto this day. I write not these things to shame you, but as my beloved sons I warn you: in Christ Jesus our Lord.

\gradall{The mouth of the righteous is exercised in wisdom, and his tongue will be talking of judgment. ℣. The law of his God is in his heart: and his goings shall not slide.}{Alleluia, alleluia. ℣. Blessed is the man that feareth the Lord: he hath great delight in his commandments. Alleluia.}

\begin{rubric}
{In Septuagesimatide or Lent, replacing the Alleluia:}
\end{rubric}\par\noindent
\tract{Blessed is the man that feareth the Lord: he hath great delight in his commandments. ℣. His seed shall be mighty upon earth: the generation of the faithful shall be blessed. ℣. Riches and plenteousness shall be in his house: and his righteousness endureth for ever.}

\begin{rubric}
{In Eastertide, replacing the Lesser Alleluia:}
\end{rubric}\par\noindent
\alleluia{Alleluia, alleluia. ℣. Blessed is the man that feareth the Lord: he hath great delight in his commandments. Alleluia. ℣. The righteous shall grow as the lily: and flourish for ever before the Lord. Alleluia.}

\readingcitation{Gospel}{Luke 12:32}
\lett{A}{t that time:} Jesus said unto his disciples: Fear not, little flock; for it is your Father's good pleasure to give you the kingdom. Sell that ye have, and give alms; provide yourselves bags which wax not old, a treasure in the heavens that faileth not, where no thief approacheth, neither moth corrupteth. For where your treasure is, there will your heart be also.

\offertory{The righteous shall rejoice in thy strength, O Lord, exceeding glad shall he be of thy salvation: thou hast given him his heart's desire (Alleluia).}

\secret
\lett{G}{rant} to us, we beseech thee, almighty God: that this oblation of our humble service may be acceptable in thy sight to the honour of thy Saints, and may cleanse us both in body and soul. Through.

\communion{Verily I say unto you: that ye which have forsaken all, and followed me, shall receive an hundredfold, and shall inherit everlasting life (Alleluia).}

\postcommunion
\lett{W}{e} beseech thee, almighty God: that we, who have received this heavenly food, may at the intercession of blessed \textit{N.}, thy Confessor, be thereby defended against all adversities. Through.

\begin{rubric}
	The following Epistle and Gospel may be used instead of the ones provided above.
\end{rubric}
\readingcitation{Another Epistle}{Philippians 3:7}\label{Philippians37}
\lett{B}{rethren:} What things were gain to me, those I counted loss for Christ. Yea doubtless, and I count all things but loss for the excellency of the knowledge of Christ Jesus my Lord: for whom I have suffered the loss of all things, and do count them but dung, that I may win Christ, and be found in him, not having mine own righteousness, which is of the law, but that which is through the faith of Christ, the righteousness which is of God by faith: that I may know him, and the power of his resurrection, and the fellowship of his sufferings, being made conformable unto his death; if by any means I might attain unto the resurrection of the dead. Not as though I had already attained, either were already perfect: but I follow after, if that I may apprehend that for which also I am apprehended of Christ Jesus.

\readingcitation{Another Gospel}{Luke 19:12}\label{Luke1912}
\lett{A}{t that time:} Jesus spake this parable unto his disciples: A certain nobleman went into a far country to receive for himself a kingdom, and to return. And he called his ten servants, and delivered them ten pounds, and said unto them, Occupy till I come. But his citizens hated him, and sent a message after him, saying, We will not have this man to reign over us. And it came to pass, that when he was returned, having received the kingdom, then he commanded these servants to be called unto him, to whom he had given the money, that he might know how much every man had gained by trading. Then came the first, saying, Lord, thy pound hath gained ten pounds. And he said unto him, Well, thou good servant: because thou hast been faithful in a very little, have thou authority over ten cities. And the second came, saying, Lord, thy pound hath gained five pounds. And he said likewise to him, Be thou also over five cities. And another came, saying, Lord, behold, here is thy pound, which I have kept laid up in a napkin: for I feared thee, because thou art an austere man: thou takest up that thou layedst not down, and reapest that thou didst not sow. And he saith unto him, Out of thine own mouth will I judge thee, thou wicked servant. Thou knewest that I was an austere man, taking up that I laid not down, and reaping that I did not sow: wherefore then gavest not thou my money into the bank, that at my coming I might have required mine own with usury? And he said unto them that stood by, Take from him the pound, and give it to him that hath ten pounds. (And they said unto him, Lord, he hath ten pounds.) For I say unto you, That unto every one which hath shall be given; and from him that hath not, even that he hath shall be taken away from him.

\subby{Common of Abbots}\label{CommonAbbots}
\fancyhead[RO,LE]{\textit{Abbots}}

\begin{rubric}
	Daily Office propers are as in the First Common of a Confessor not a Bishop (p. \pageref{CommonConfessorNotBishopI}).
\end{rubric}

%\subsubsec{Daily Office Propers}\par\noindent
%\antiphon{Mag.}{I will liken him {\dag} unto a wise man, which built his house upon a rock.}

%\antiphon{Ben.}{Well done, good and faithful servant, {\dag} thou hast been faithful over a few things, I will make thee ruler over many things: enter thou into the joy of the Lord.}

%\antiphon{Mag.}{The world, and all things earthly, {\dag} this man despised, triumphant: he laid up treasures in heaven by word and deed.}
\introit
\lett{T}{he} mouth of the righteous is exercised in wisdom, and his tongue will be talking of judgment: the law of his God is in his heart. (Alleluia, alleluia) \textit{Ps.} Fret not thyself because of the ungodly: neither be thou envious against the evil doers.

\collect
\lett{W}{e} beseech thee, O Lord, that the prayers of the blessed Abbot, \textit{N.} may commend us unto thee: that by his advocacy thou wouldest vouchsafe unto us those things to which by our own merits we cannot obtain. Through.

\readingcitation{Epistle}{Ecclesiasticus 45:1}
%RV:
\lett{B}{eloved} of God and men, even Moses, Whose memorial is blessed. He made him like to the glory of the saints, And magnified him in the fears of his enemies. By his words he caused the wonders to cease; He glorified him in the sight of kings; He gave him commandment for his people, And shewed him part of his glory. He sanctified him in his faithfulness and meekness; He chose him out of all flesh. He made him to hear his voice, And led him into the thick darkness, And gave him commandments face to face, Even the law of life and knowledge, 

%\lett{B}{eloved} of God and men, whose memorial is blessed. He made him like to the glorious saints, and magnified him, so that his enemies stood in fear of him. By his words he caused the wonders to cease, and he made him glorious in the sight of kings, and gave him a commandment for his people, and shewed him part of his glory. He sanctified him in his faithfuless and meekness, and chose him out of all men. He made him to hear his voice, and brought him into the dark cloud, and gave him commandments before his face, even the law of life and knowledge.

\gradall{Thou hast prevented him, O Lord, with the blessings of goodness: and hast set a crown of pure gold upon his head. ℣. He asked life of thee, and thou gavest him a long life, even for ever and ever.}{Alleluia, alleluia. ℣. The righteous shall flourish like a palm-tree: and shall spread abroad like a cedar in Libanus. Alleluia.}

\begin{rubric}
{In Septuagesimatide or Lent, replacing the Alleluia:}
\end{rubric}\par\noindent
\tract{Blessed is the man that feareth the Lord: he hath great delight in his commandments. ℣. His seed shall be mighty upon earth: the generation of the faithful shall be blessed. ℣. Riches and plenteousness shall be in his house: and his righteousness endureth for ever.}

\begin{rubric}
{In Eastertide, replacing the Lesser Alleluia:}
\end{rubric}\par\noindent
\alleluia{Alleluia, alleluia. ℣. The righteous shall flourish like a palm-tree: and shall spread abroad like a cedar in Libanus. Alleluia. ℣. The righteous shall grow as the lily: and flourish for ever before the Lord. Alleluia.}

\readingcitation{Gospel}{Matthew 19:27}
\lett{A}{t that time:} Peter said unto Jesus, Behold, we have forsaken all, and followed thee; what shall we have therefore? And Jesus said unto them, Verily I say unto you, That ye which have followed me, in the regeneration when the Son of man shall sit in the throne of his glory, ye also shall sit upon twelve thrones, judging the twelve tribes of Israel. And every one that hath forsaken houses, or brethren, or sisters, or father, or mother, or wife, or children, or lands, for my name's sake, shall receive an hundredfold, and shall inherit everlasting life.

\offertory{Thou hast given him his heart's desire, O Lord, and hast not denied him the request of his lips: thou hast set a crown of pure gold upon his head. (Alleluia.)}

\secret
\lett{W}{e} beseech thee, O Lord, that thy holy Abbot \textit{N.} may intercede for us: that this sacrifice which we offer and present upon thy holy altar may be profitable unto us for our salvation. Through.

\communion{A faithful and wise steward, whom his lord hath made ruler over his household, to give them their portion of meat in due season (Alleluia.)}

\postcommunion
\lett{L}{et} thy sacrament, O Lord, which we have now received and the prayers of the blessed Abbot, \textit{N.}, effectually defend us: that we may both imitate the example of his conversation, and receive the succour of his intercession. Through.


\subby{I: Common of a Virgin Martyr}\label{CommonVirginMartyrI}
\fancyhead[RO,LE]{\textit{Virgin Martyr I}}

\subbysub{I Evensong}

\begin{rubric}
	The Hymn is of Mattins with the following Versicle \& Antiphon.
\end{rubric}

℣. In thy grace and in thy beauty.

℟. Go forth, ride prosperously, and reign.

\antiphon{Mag.}{Come thou Bride of Christ, {\dag} receive the crown which the Lord hath prepared for thee for ever.}\par

\subbysub{Mattins}

\invitatoryhymn

\gregorioscore{resources/gabc/Common/VirginMartyrInvitatory.gabc}

\officehymn

\gregorioscore{resources/gabc/Common/VirginMartyrMattins.gabc}

℣. Full of grace are thy lips.

℟. Because God hath blessed thee for ever.

\antiphon{Ben.}{The kingdom of heaven {\dag} is likened unto a merchantman seeking goodly pearls; who, when he had found one pearl of great price, went and sold all that he had, and bought it.}\par

\subbysub{II Evensong}

\begin{rubric}
	The Hymn \& Versicle are of Mattins with the following Antiphon.
\end{rubric}

\antiphon{Mag.}{Come thou Bride of Christ, {\dag} receive the crown which the Lord hath prepared for thee for ever.}

\introit
\lett{I}{will} speak of thy testimonies, even before kings, and will not be ashamed: and my delight shall be in thy commandments, which I have loved exceedingly (Alleluia, alleluia). \textit{Ps.} Blessed are those that are undefiled in the way: and walk in the law of the Lord.

\collect
\lett{O}{God,} who among the manifold works of thy power hast bestowed even upon the weakness of women the victory of martyrdom: mercifully grant; that we, who celebrate the birthday of blessed \textit{N.} thy Virgin and Martyr, may by her example be drawn nearer unto thee. Through.

\readingcitation{Epistle}{Ecclesiasticus 51:1}
%RV:
\lett{I}{will} give thanks unto thee, O Lord, O King, And will praise thee, O God my Saviour: I do give thanks unto thy name: For thou wast my protector and helper, And didst deliver my body out of destruction, And out of the snare of a slanderous tongue, From lips that forge lies, And wast my helper before them that stood by; And didst deliver me, according to the abundance of thy mercy, and greatness of thy name, From the gnashings of teeth ready to devour, Out of the hand of such as sought my life, Out of the manifold afflictions which I had; From the choking of a fire on every side, And out of the midst of fire which I kindled not; Out of the depth of the belly of the grave, And from an unclean tongue, And from lying words, The slander of an unrighteous tongue unto the king. My soul drew near even unto death, And my life was near to the grave beneath. They compassed me on every side, And there was none to help me. I was looking for the succour of men, And it was not. And I remembered thy mercy, O Lord, And thy working which hath been from everlasting, How thou deliverest them that wait for thee, And savest them out of the hand of the enemies, O Lord our God.

%\lett{I}{will} thank thee, O Lord and King, and praise thee, O God my Saviour. I do give praise unto thy name: for thou art my defender and helper, and hast preserved my body from destruction, and from the snare of the slanderous tongue, and from the lips that forge lies, and hast been mine helper against mine adversaries. And hast delivered me, according to the multitude of thy mercies and greatness of thy name, from the teeth of them that were ready to devour me, and out of the hands of such as sought after my life, and from the manifold afflictions which I had: from the choking of fire on every side, and from the midst of the fire which I kindled not: from the depth of the belly of hell, from an unclean tongue and from lying words. By an accusation to the king from an unrighteous tongue my soul drew near even unto death: but thou deliverest such as wait for thee, and savest them out of the hands of their enemies, O Lord our God.

\gradall{Thou hast loved righteousness, and hated iniquity. ℣. Wherefore God, even thy God, hath anointed thee with the oil of gladness.}{Alleluia, alleluia. ℣. The Virgins that be her fellows shall be brought unto the King: they that bear her company shall be brought unto thee with joy. Alleluia.}

\begin{rubric}
{In Septuagesimatide or Lent, replacing the Alleluia:}
\end{rubric}\par\noindent
\tract{Come, Spouse of Christ, receive the crown, which the Lord hath prepared for thee for ever: for love of whom thou didst shed thy blood. ℣. Thou hast loved righteousness, and hated iniquity: wherefore God, even thy God, hath anointed thee with the oil of gladness above thy fellows. ℣. In thy comeliness and in thy beauty, go forth, proceed prosperously and reign.}

\begin{rubric}
{In Eastertide, replacing the Lesser Alleluia:}
\end{rubric}\par\noindent
\alleluia{Alleluia, alleluia. ℣. The Virgins that be her fellows shall be brought unto the King: they that bear her company shall be brought unto thee with joy. Alleluia. ℣. In thy comeliness and in thy beauty go forth, proceed prosperously and reign. Alleluia.}

\readingcitation{Gospel}{Matthew 25:1}
\lett{A}{t that time:} Jesus spake this parable unto his disciples: The kingdom of heaven be likened unto ten virgins, which took their lamps, and went forth to meet the bridegroom. And five of them were wise, and five were foolish. They that were foolish took their lamps, and took no oil with them: but the wise took oil in their vessels with their lamps. While the bridegroom tarried, they all slumbered and slept. And at midnight there was a cry made, Behold, the bridegroom cometh; go ye out to meet him. Then all those virgins arose, and trimmed their lamps. And the foolish said unto the wise, Give us of your oil; for our lamps are gone out. But the wise answered, saying, Not so; lest there be not enough for us and you: but go ye rather to them that sell, and buy for yourselves. And while they went to buy, the bridegroom came; and they that were ready went in with him to the marriage: and the door was shut. Afterward came also the other virgins, saying, Lord, Lord, open to us. But he answered and said, Verily I say unto you, I know you not. Watch therefore, for ye know neither the day nor the hour.

\offertory{The Virgins that be her fellows shall be brought unto the King: they that bear her company shall be brought unto thee, with joy and gladness: and shall enter into the palace of the Lord the King (Alleluia).}

\secret
\lett{R}{eceive,} O Lord, the gifts which we offer on the solemnity of blessed \textit{N.}, thy Virgin and Martyr: through whose advocacy we trust to be delivered. Through.

\communion{Let the proud be confounded, for they go wickedly about to destroy me: but I will be occupied in thy commandments and in thy statutes, that I be not ashamed (Alleluia).}

\postcommunion
\lett{M}{ay} the mysteries which we have received be for our succour, O Lord: and at the intercession of blessed \textit{N.}, thy Virgin and Martyr, cause us to rejoice in thy continual protection. Through.


\subby{II: Common of a Virgin Martyr}\label{CommonVirginMartyrII}
\fancyhead[RO,LE]{\textit{Virgin Martyr II}}

\begin{rubric}
	Daily Office propers as in the First Common of a Virgin Martyr (p. \pageref{CommonVirginMartyrI}).
\end{rubric}

\introit
\lett{T}{he} ungodly laid wait for me to destroy me: O Lord, I will consider thy testimonies: I see that all things come to an end: but thy commandment is exceeding broad (Alleluia, alleluia). \textit{Ps.} Blessed are those that are undefiled in the way: and walk in the law of the Lord.

\collect
\lett{W}{e} beseech thee, O Lord, that, like as blessed \textit{N.}, thy Virgin and Martyr, was ever found pleasing unto thee, both by the merit of her chastity, and by her confession of thy power: so she may implore for us thy pardon. Through.

\readingcitation{Epistle}{Ecclesiasticus 51:9}
%RV:
\lett{O}{Lord my God,} I lifted up my supplication from the earth, And prayed for deliverance from death. I called upon the Lord, the Father of my Lord, That he would not forsake me in the days of affliction, In the time when there was no help against the proud. I will praise thy name continually, And will sing praise with thanksgiving; And my supplication was heard: For thou savedst me from destruction, And deliveredst me from the evil time: Therefore will I give thanks and praise unto thee, And bless thy name, O Lord our God.

%\lett{O}{Lord} my God, thou has lifted up my habitation upon the earth, and I prayed for deliverance from death. I called upon the Lord, the Father of my Lord, that he would not leave me in the days of my trouble, and in the time of the proud, when there was no help. I will praise thy name continually, and will sing praise with thanksgiving, and so my prayer was heard. For thou savedst me from destruction, and deliveredst me from the evil time. Therefore will I give thanks, and praise thee, O Lord our God.

\gradall{God shall help her with his countenance: God is in the midst of her, therefore shall she not be removed. ℣. The rivers of the flood thereof shall make glad the city of God: the holy place of the tabernacle of the Most Highest.}{Alleluia, alleluia. ℣. This is a wise Virgin, and one of the number of the prudent. Alleluia.}

\begin{rubric}
{In Septuagesimatide or Lent, replacing the Alleluia:}
\end{rubric}\par\noindent
\tract{Come, Spouse of Christ, receive the crown which the Lord hath prepared for thee for ever: for love of whom thou didst shed thy blood. ℣. Thou hast loved righteousness, and hated iniquity: wherefore God, even thy God, hath anointed thee with the oil of gladness above thy fellows. ℣. In thy comeliness and in thy beauty go forth, proceed prosperously and reign.}

\begin{rubric}
{In Eastertide, replacing the Lesser Alleluia:}
\end{rubric}\par\noindent
\alleluia{Alleluia, alleluia. ℣. This is a wise Virgin, and one of the number of the prudent. Alleluia. ℣. O how beautiful is the chaste generation with its glory. Alleluia.}

\readingcitation{Gospel}{Matthew 13:44}
\lett{A}{t that time:} Jesus spake this parable unto his disciples: The kingdom of heaven is like unto treasure hid in a field; the which when a man hath found, he hideth, and for joy thereof goeth and selleth all that he hath, and buyeth that field.

Again, the kingdom of heaven is like unto a merchant man, seeking goodly pearls: who, when he had found one pearl of great price, went and sold all that he had, and bought it.

Again, the kingdom of heaven is like unto a net, that was cast into the sea, and gathered of every kind: which, when it was full, they drew to shore, and sat down, and gathered the good into vessels, but cast the bad away. So shall it be at the end of the world: the angels shall come forth, and sever the wicked from among the just, and shall cast them into the furnace of fire: there shall be wailing and gnashing of teeth.

Jesus saith unto them, Have ye understood all these things? They say unto him, Yea, Lord. Then said he unto them, Therefore every scribe which is instructed unto the kingdom of heaven is like unto a man that is an householder, which bringeth forth out of his treasure things new and old.

\offertory{Full of grace are thy lips, because God hath blessed thee for ever and ever (Alleluia).}

\secret
\lett{G}{raciously} receive, O Lord, through the merits of blessed \textit{N.}, thy Virgin and Martyr, the sacrifices which we offer unto thee: and grant that they may avail for our continual help. Through.

\communion{I deal with the thing that is lawful and right, O Lord, let the proud do me no wrong: I hold straight all thy commandments, and all false ways I utterly abhor (Alleluia).}

\postcommunion
\lett{O}{Lord} our God, who hast fulfilled us with the bounty of thy heavenly gift: we beseech thee, that, at the intercession of blessed \textit{N.}, thy Virgin and Martyr, we may ever live by the partaking of the same. Through.

\begin{rubric}
	The Gospel from the Common of Many Virgin Martyrs (p. \pageref{CommonVirginMartyrs}) may be used instead of the one provided above.
\end{rubric}


\subby{Common of Many Virgin Martyrs}\label{CommonVirginMartyrs}
\fancyhead[RO,LE]{\textit{Virgin Martyrs}}

\begin{rubric}
The propers as in the Second Common of a Virgin Martyr (p. \pageref{CommonVirginMartyrII}), except for the following.
\end{rubric}

\antiphon{Mag.}{O ye wise Virgins, {\dag} arise and trim your lamps: for behold the Bridegroom cometh; go ye out to meet him.}

\collect
\lett{G}{rant,} we beseech thee, O Lord our God, that we may at all times so devoutly honour the triumphs of thy holy Virgins and Martyrs \textit{N. and N.}: that, although we cannot worthily shew forth their praises, yet we may continually honour them with lowly service. Through.

\readingcitation{Epistle}{1 Corinthians 7:25}
\lett{B}{rethren:} Concerning virgins I have no commandment of the Lord: yet I give my judgment, as one that hath obtained mercy of the Lord to be faithful. I suppose therefore that this is good for the present distress, I say, that it is good for a man so to be. Art thou bound unto a wife? seek not to be loosed. Art thou loosed from a wife? seek not a wife. But and if thou marry, thou hast not sinned; and if a virgin marry, she hath not sinned. Nevertheless such shall have trouble in the flesh: but I spare you.

But this I say, brethren, the time is short: it remaineth, that both they that have wives be as though they had none; and they that weep, as though they wept not; and they that rejoice, as though they rejoiced not; and they that buy, as though they possessed not; and they that use this world, as not abusing it: for the fashion of this world passeth away.

But I would have you without carefulness. He that is unmarried careth for the things that belong to the Lord, how he may please the Lord: but he that is married careth for the things that are of the world, how he may please his wife. There is difference also between a wife and a virgin. The unmarried woman careth for the things of the Lord, that she may be holy both in body and in spirit: in Christ Jesus our Lord.

\readingcitation{Gospel}{Matthew 19:3}
\lett{A}{t that time:} The Pharisees came unto Jesus, tempting him, and saying unto him, Is it lawful for a man to put away his wife for every cause? And he answered and said unto them, Have ye not read, that he which made them at the beginning made them male and female, and said, For this cause shall a man leave father and mother, and shall cleave to his wife: and they twain shall be one flesh? Wherefore they are no more twain, but one flesh. What therefore God hath joined together, let not man put asunder. They say unto him, Why did Moses then command to give a writing of divorcement, and to put her away? He saith unto them, Moses because of the hardness of your hearts suffered you to put away your wives: but from the beginning it was not so. And I say unto you, Whosoever shall put away his wife, except it be for fornication, and shall marry another, committeth adultery: and whoso marrieth her which is put away doth commit adultery.

His disciples say unto him, If the case of the man be so with his wife, it is not good to marry. But he said unto them, All men cannot receive this saying, save they to whom it is given. For there are some eunuchs, which were so born from their mother's womb: and there are some eunuchs, which were made eunuchs of men: and there be eunuchs, which have made themselves eunuchs for the kingdom of heaven's sake. He that is able to receive it, let him receive it.

\secret
\lett{O}{Lord,} we beseech thee, look down upon these gifts, which we offer on thine altars on this festival of thy holy Virgins and Martyrs \textit{N. and N.}: that as by these blessed mysteries thou hast bestowed glory upon them; so likewise of thy bounty thou wouldest vouchsafe to us thy pardon. Through.

\postcommunion
\lett{G}{rant} to us, we beseech thee, O Lord, at the intercession of thy holy Virgins and Martyrs, \textit{N. and N.}: that those things which we touch with our lips we may receive in purity of heart. Through.

\subby{I: Common of a Virgin Only}\label{CommonVirginOnlyI}
\fancyhead[RO,LE]{\textit{Virgin I}}

\begin{rubric}
	Daily Office propers as in the First Common of a Virgin Martyr (p. \pageref{CommonVirginMartyrI}).
\end{rubric}

\introit
\lett{T}{hou} hast loved righteousness, and hated iniquity: wherefore God, even thy God, hath anointed thee with the oil of gladness above thy fellows (Alleluia, alleluia). \textit{Ps.} My heart is inditing of a good matter: I speak of the things which I have made unto the King.

\collect
\lett{G}{raciously} hear us, O God of our salvation: that, like as we do rejoice in the festival of blessed \textit{N.}, thy Virgin; so we may be instructed in all godly and devout affection. Through.

\readingcitation{Epistle}{2 Corinthians 10:17}
\lett{B}{rethren:} He that glorieth, let him glory in the Lord. For not he that commendeth himself is approved, but whom the Lord commendeth. Would to God ye could bear with me a little in my folly: and indeed bear with me. For I am jealous over you with godly jealousy: for I have espoused you to one husband, that I may present you as a chaste virgin to Christ.

\gradall{In thy comeliness and in thy beauty go forth, proceed prosperously and reign. ℣. Because of the word of truth, of meekness, and righteousness: and thy right hand shall teach thee terrible things.}{Alleluia, alleluia. ℣. The Virgins that be her fellows shall be brought unto the King: they that bear her company shall be brought unto thee with joy. Alleluia.}

\begin{rubric}
{In Septuagesimatide or Lent, replacing the Alleluia:}
\end{rubric}\par\noindent
\tract{Hearken, O daughter, and consider, incline thine ear: so shall the King have pleasure in thy beauty. ℣. The rich among the people shall make their supplication before thee: kings' daughters were among thy honourable women. ℣. The Virgins that be her fellows shall be brought unto the King: they that bear her company shall be brought unto thee. ℣. With joy and gladness shall they be brought: and shall enter into the King's palace.}

\begin{rubric}
{In Eastertide, replacing the Lesser Alleluia:}
\end{rubric}\par\noindent
\alleluia{Alleluia, alleluia. ℣. The Virgins that be her fellows shall be brought unto the King: they that bear her company shall be brought unto them with joy. Alleluia. ℣. In thy comeliness and in thy beauty go forth, proceed prosperously and reign. Alleluia.}

\readingcitation{Gospel}{Matthew 25:1}
\lett{A}{t that time:} Jesus spake this parable unto his disciples: the kingdom of heaven shall be likened unto ten virgins, which took their lamps, and went forth to meet the bridegroom. And five of them were wise, and five were foolish. They that were foolish took their lamps, and took no oil with them: but the wise took oil in their vessels with their lamps. While the bridegroom tarried, they all slumbered and slept. And at midnight there was a cry made, Behold, the bridegroom cometh; go ye out to meet him. Then all those virgins arose, and trimmed their lamps. And the foolish said unto the wise, Give us of your oil; for our lamps are gone out. But the wise answered, saying, Not so; lest there be not enough for us and you: but go ye rather to them that sell, and buy for yourselves. And while they went to buy, the bridegroom came; and they that were ready went in with him to the marriage: and the door was shut. Afterward came also the other virgins, saying, Lord, Lord, open to us. But he answered and said, Verily I say unto you, I know you not. Watch therefore, for ye know neither the day nor the hour.

\offertory{Kings' daughters were among thy honourable women, upon thy right hand did stand the queen in a vesture of gold, wrought about with divers colours (Alleluia).}

\secret
\lett{G}{rant,} O Lord, that like as thy dedicated people do acknowledge that in tribulation they have been succoured by the merits of thy Saints: so this oblation, which they offer unto thee in honour of the same, may be acceptable in thy sight. Through.

\communion{The five wise virgins took oil in their vessels with their lamps: and at midnight there was a cry made: Behold the bridegroom cometh: go ye out to meet Christ the Lord.}

\postcommunion
\lett{O}{Lord,} who hast satisfied thy family with sacred gifts: we beseech thee; that we may at all times be comforted by the intercession of her whose festival we celebrate. Through.

%MANUAL ADJUSTMENT:
\clearpage
\subby{II: Common of a Virgin Only}\label{CommonVirginOnlyII}
\fancyhead[RO,LE]{\textit{Virgin II}}

\begin{rubric}
	Daily Office propers as in the First Common of a Virgin Martyr (p. \pageref{CommonVirginMartyrI}), except for the following Invitatory Hymn.
\end{rubric}

\gregorioscore{resources/gabc/Common/VirginMartyrInvitatory.gabc}

\introit
\lett{A}{ll} the rich among the people shall make their supplication before thee: the Virgins that be her fellows shall be brought unto the King: they that bear her company shall be brought unto thee with joy and gladness (Alleluia, alleluia). \textit{Ps.} My heart is inditing of a good matter: I speak of the things which I have made unto the King.

\collect
\lett{G}{raciously} hear us, O God of our salvation: that, like as we do rejoice in the festival of blessed \textit{N.}, thy Virgin, so we may be instructed in all godly and devout affection. Through.

\readingcitation{Epistle}{1 Corinthians 7:25}
\lett{B}{rethren:} Concerning virgins I have no commandment of the Lord: yet I give my judgment, as one that hath obtained mercy of the Lord to be faithful. I suppose therefore that this is good for the present distress, I say, that it is good for a man so to be. Art thou bound unto a wife? seek not to be loosed. Art thou loosed from a wife? seek not a wife. But and if thou marry, thou hast not sinned; and if a virgin marry, she hath not sinned. Nevertheless such shall have trouble in the flesh: but I spare you.

But this I say, brethren, the time is short: it remaineth, that both they that have wives be as though they had none; and they that weep, as though they wept not; and they that rejoice, as though they rejoiced not; and they that buy, as though they possessed not; and they that use this world, as not abusing it: for the fashion of this world passeth away.

But I would have you without carefulness. He that is unmarried careth for the things that belong to the Lord, how he may please the Lord: but he that is married careth for the things that are of the world, how he may please his wife. There is difference also between a wife and a virgin. The unmarried woman careth for the things of the Lord, that she may be holy both in body and in spirit: in Christ Jesus our Lord.

\gradall{The King shall have pleasure in thy beauty, for he is thy Lord God. ℣. Hearken, O daughter, and consider, incline thine ear.}{Alleluia, alleluia. ℣. This is a wise Virgin, and one of the number of the prudent. Alleluia.}

\begin{rubric}
{In Septuagesimatide or Lent, replacing the Alleluia:}
\end{rubric}\par\noindent
\tract{The King shall have pleasure in thy beauty. ℣. The rich among the people shall make their supplication before thee: kings' daughters were among thy honourable women. ℣. The Virgins that be her fellows shall be brought unto the King: they that bear her company shall be brought unto thee. ℣. With joy and gladness shall they be brought: and shall enter into the King's palace.}

\begin{rubric}
{In Eastertide, replacing the Lesser Alleluia:}
\end{rubric}\par\noindent
\alleluia{Alleluia, alleluia. ℣. This is a wise Virgin, and one of the number of the prudent. Alleluia. ℣. O how beautiful is the chaste generation with its glory. Alleluia.}

\readingcitation{Gospel}{Matthew 13:44}
\lett{A}{t that time:} Jesus spake this parable unto his disciples: The kingdom of heaven is like unto treasure hid in a field; the which when a man hath found, he hideth, and for joy thereof goeth and selleth all that he hath, and buyeth that field.

Again, the kingdom of heaven is like unto a merchant man, seeking goodly pearls: who, when he had found one pearl of great price, went and sold all that he had, and bought it.

Again, the kingdom of heaven is like unto a net, that was cast into the sea, and gathered of every kind: which, when it was full, they drew to shore, and sat down, and gathered the good into vessels, but cast the bad away. So shall it be at the end of the world: the angels shall come forth, and sever the wicked from among the just, and shall cast them into the furnace of fire: there shall be wailing and gnashing of teeth.

Jesus saith unto them, Have ye understood all these things? They say unto him, Yea, Lord. Then said he unto them, Therefore every scribe which is instructed unto the kingdom of heaven is like unto a man that is an householder, which bringeth forth out of his treasure things new and old.

\offertory{The Virgins that be her fellows shall be brought unto the King: they that bear her company shall be brought unto thee, with joy and gladness: and shall enter into the palace of the Lord the King (Alleluia).}

\secret
\lett{G}{rant,} O Lord, that like as thy dedicated people do acknowledge that in tribulation they have been succoured by the merits of thy Saints: so this oblation, which they do offer unto thee in honour of the same, may be acceptable in thy sight. Through.

\communion{The kingdom of heaven is like unto a merchant man, seeking goodly pearls: who when he had found one pearl of great price, gave all that he had, and bought it (Alleluia).}

\postcommunion
\lett{O}{Lord,} who hast satisfied thy family with sacred gifts: we beseech thee, that we may at all times be comforted by the intercession of her whose festival we celebrate. Through.

\begin{rubric}
	The following Gospel may be used instead of the one provided above.
\end{rubric}
\readingcitation{Another Gospel}{Matthew 25:1}
\lett{A}{t that time:} Jesus spake this parable unto his disciples: The kingdom of heaven be likened unto ten virgins, which took their lamps, and went forth to meet the bridegroom. And five of them were wise, and five were foolish. They that were foolish took their lamps, and took no oil with them: but the wise took oil in their vessels with their lamps. While the bridegroom tarried, they all slumbered and slept. And at midnight there was a cry made, Behold, the bridegroom cometh; go ye out to meet him. Then all those virgins arose, and trimmed their lamps. And the foolish said unto the wise, Give us of your oil; for our lamps are gone out. But the wise answered, saying, Not so; lest there be not enough for us and you: but go ye rather to them that sell, and buy for yourselves. And while they went to buy, the bridegroom came; and they that were ready went in with him to the marriage: and the door was shut. Afterward came also the other virgins, saying, Lord, Lord, open to us. But he answered and said, Verily I say unto you, I know you not. Watch therefore, for ye know neither the day nor the hour.


\subby{Common of a Martyr not a Virgin}\label{CommonMartyrNotVirgin}
\fancyhead[RO,LE]{\textit{Martyr not Virgin}}

\subbysub{I Evensong}

\begin{rubric}
	The Hymn is of Mattins with the following Versicle \& Antiphon.
\end{rubric}

℣. In thy grace and in thy beauty.

℟. Go forth, ride prosperously, and reign.

\antiphon{Mag.}{The kingdom of heaven {\dag} is likened unto a merchantman seeking goodly pearls; who, when he had found one pearl of great price, went and sold all that he had, and bought it.}\par\noindent

\subbysub{Mattins}

\invitatoryhymn

\gregorioscore{resources/gabc/Common/MartyrNotVirginInvitatory.gabc}

\officehymn

\gregorioscore{resources/gabc/Common/MartyrNotVirginMattins.gabc}

℣. Full of grace are thy lips.

℟. Because God hath blessed thee for ever.

\antiphon{Ben.}{Give her {\dag} of the fruit of her hands; and let her own works praise her in the gates.}\par\noindent

\subbysub{II Evensong}

\begin{rubric}
	The Hymn \& Versicle are of Mattins with the following Antiphon.
\end{rubric}

\antiphon{Mag.}{She stretched out {\dag} her hand to the poor; yea, she reacheth forth her hands to the needy; she eateth not the bread of idleness.}

\introit
\lett{T}{he} ungodly laid wait for me to destroy me: O Lord, I will consider thy testimonies: I see that all things come to an end: but thy commandment is exceeding broad (Alleluia). \textit{Ps.} Blessed are those that are undefiled in the way: and walk in the law of the Lord.

%Roman Collect with no intercession.
\collect
\lett{O}{God,} who among the manifold works of thy power hast bestowed even upon the weakness of women the victory of martyrdom: mercifully grant; that we, who celebrate the birthday of blessed \textit{N.}, thy Martyr, may by her example draw nearer unto thee. Through.

\readingcitation{Epistle}{Ecclesiasticus 51:1}
%RV:
\lett{I}{will} give thanks unto thee, O Lord, O King, And will praise thee, O God my Saviour: I do give thanks unto thy name: For thou wast my protector and helper, And didst deliver my body out of destruction, And out of the snare of a slanderous tongue, From lips that forge lies, And wast my helper before them that stood by; And didst deliver me, according to the abundance of thy mercy, and greatness of thy name, From the gnashings of teeth ready to devour, Out of the hand of such as sought my life, Out of the manifold afflictions which I had; From the choking of a fire on every side, And out of the midst of fire which I kindled not; Out of the depth of the belly of the grave, And from an unclean tongue, And from lying words, The slander of an unrighteous tongue unto the king. My soul drew near even unto death, And my life was near to the grave beneath. They compassed me on every side, And there was none to help me. I was looking for the succour of men, And it was not. And I remembered thy mercy, O Lord, And thy working which hath been from everlasting, How thou deliverest them that wait for thee, And savest them out of the hand of the enemies, O Lord our God.

%\lett{I}{will} thank thee, O Lord and King, and praise thee, O God my Saviour. I do give praise unto thy name: for thou art my defender and helper, and hast preserved my body from destruction, and from the snare of the slanderous tongue, and from the lips that forge lies, and hast been mine helper against mine adversaries. And hast delivered me, according to the multitude of thy mercies and greatness of thy name, from the teeth of them that were ready to devour me, and out of the hands of such as sought after my life, and from the manifold afflictions which I had: from the choking of fire on every side, and from the midst of the fire which I kindled not: from the depth of the belly of hell, from an unclean tongue and from lying words. By an accusation to the king from an unrighteous tongue my soul drew near even unto death: but thou deliverest such as wait for thee, and savest them out of the hands of their enemies, O Lord our God.

\gradall{Thou hast loved righteousness and hated iniquity. ℣. Wherefore God, even thy God, hath anointed thee with the oil of gladness.}{Alleluia, alleluia. ℣. In thy comeliness and in thy beauty go forth, proceed prosperously and reign. Alleluia.}

\begin{rubric}
{In Septuagesimatide or Lent, replacing the Alleluia:}
\end{rubric}\par\noindent
\tract{Come, Spouse of Christ, receive the crown which the Lord hath prepared for thee for ever: for love of whom thou didst shed thy blood. ℣. Thou hast loved righteousness and hated iniquity: wherefore God, even thy God, hath anointed thee with the oil of gladness, above thy fellows. ℣. In thy comeliness and in thy beauty go forth, proceed prosperously and reign.}

\begin{rubric}
{In Eastertide, replacing the Lesser Alleluia:}
\end{rubric}\par\noindent
\alleluia{Alleluia, alleluia. ℣. In thy comeliness and in thy beauty, go forth, proceed prosperously and reign. Alleluia. ℣. Because of the word of truth, of meekness, and righteousness: and thy right hand shall teach thee terrible things. Alleluia.}

\readingcitation{Gospel}{Matthew 13:44}
\lett{A}{t that time:} Jesus spake this parable unto his disciples: The kingdom of heaven is like unto treasure hid in a field; the which when a man hath found, he hideth, and for joy thereof goeth and selleth all that he hath, and buyeth that field.

Again, the kingdom of heaven is like unto a merchant man, seeking goodly pearls: who, when he had found one pearl of great price, went and sold all that he had, and bought it.

Again, the kingdom of heaven is like unto a net, that was cast into the sea, and gathered of every kind: which, when it was full, they drew to shore, and sat down, and gathered the good into vessels, but cast the bad away. So shall it be at the end of the world: the angels shall come forth, and sever the wicked from among the just, and shall cast them into the furnace of fire: there shall be wailing and gnashing of teeth.

Jesus saith unto them, Have ye understood all these things? They say unto him, Yea, Lord. Then said he unto them, Therefore every scribe which is instructed unto the kingdom of heaven is like unto a man that is an householder, which bringeth forth out of his treasure things new and old.

\begin{rubric}
	After Septuagesima the \emph{Alleluia} at the end of the following Offertory is omitted.
\end{rubric}
\offertory{Full of grace are thy lips: because God hath blessed thee for ever and ever, alleluia.}

\secret
\lett{R}{eceive,} O Lord, the gifts which we offer on the solemnity of blessed \textit{N.}, thy Martyr: through whose advocacy we trust to be delivered. Through.

\communion{Princes have persecuted me without a cause: but my heart standeth in awe of thy word: I am as glad of thy word, as one that findeth great spoils (Alleluia).}

\postcommunion
\lett{M}{ay} the mysteries which we have received, O Lord, be for our succour: and, at the intercession of blessed \textit{N.} thy Martyr, cause us to rejoice in thy continual protection. Through.


\subby{Common of Many Martyrs not Virgins}\label{CommonMartyrsNotVirgins}
\fancyhead[RO,LE]{\textit{Martyrs not Virgins}}

\begin{rubric}
	The propers as in the Common of a Martyr not a Virgin (p. \pageref{CommonMartyrNotVirgin}), except for the following.
\end{rubric}

℣. Thou hast crowned them with glory and honour, O Lord.

℟. And madest them to have dominion of the works of thy hands.

\antiphon{Mag. \& Ben.}{For theirs is the kingdom of heaven, {\dag} who despised worldly living: who have won the rewards of the kingdom, and have washed their robes in the blood of the Lamb.}\par\noindent

\collect
\lett{G}{rant,} we beseech thee, O Lord our God, that we may at all times so devoutly honour the triumphs of thy holy Martyrs \textit{N. and N.}: that although we cannot worthily shew forth their praises, yet we may continually honour them with lowly service. Through.

\secret
\lett{O}{Lord,} we beseech thee, look down upon these gifts which we offer on thine altars on this festival of thy holy Martyrs \textit{N. and N.}: that as by these blessed mysteries thou hast bestowed glory upon them; so likewise of thy bounty thou wouldest vouchsafe us thy pardon. Through.

\postcommunion
\lett{G}{rant} to us, we beseech thee, O Lord, at the intercession of thy holy Martyrs \textit{N. and N.}: that those things which we touch with our lips, we may receive in purity of heart. Through.


\subby{Common of Neither Virgin nor Martyr}\label{CommonNeitherVirginMartyr}
\fancyhead[RO,LE]{\textit{Not Virgin Nor Martyr}}

\begin{rubric}
	Daily Office propers as in the Common of a Martyr not a Virgin (p. \pageref{CommonMartyrNotVirgin}).
\end{rubric}

%\antiphon{Mag.}{The kingdom of heaven {\dag} is likened unto a merchantman seeking goodly pearls; who, when he had found one pearl of great price, went and sold all that he had, and bought it.}\par\noindent
%\antiphon{Ben.}{Give her {\dag} of the fruit of her hands; and let her own works praise her in the gates.}\par\noindent
%\antiphon{Mag.}{She stretched out {\dag} her hand to the poor; yea, she reacheth forth her hands to the needy; she eateth not the bread of idleness.}

\introit
\lett{I}{know,} O Lord, that thy judgments are right, and that thou of very faithfulness hast caused me to be troubled: my flesh trembleth for fear of thee, and I am afraid of thy judgments (Alleluia, alleluia). \textit{Ps.} Blessed are those that are undefiled in the way: and walk in the law of the Lord.

\collect
\lett{G}{raciously} hear us, O God of our salvation: and grant that, like as we do rejoice in the festival of blessed \textit{N.}; so we may be instructed in all godly and devout affection. Through.

\readingcitation{Epistle}{Proverbs 31:10}
\lett{W}{ho} can find a virtuous woman? for her price is far above rubies. The heart of her husband doth safely trust in her, so that he shall have no need of spoil. She will do him good and not evil all the days of her life. She seeketh wool, and flax, and worketh willingly with her hands. She is like the merchants' ships; she bringeth her food from afar. She riseth also while it is yet night, and giveth meat to her household, and a portion to her maidens. She considereth a field, and buyeth it: with the fruit of her hands she planteth a vineyard. She girdeth her loins with strength, and strengtheneth her arms. She perceiveth that her merchandise is good: her candle goeth not out by night. She layeth her hands to the spindle, and her hands hold the distaff. She stretcheth out her hand to the poor; yea, she reacheth forth her hands to the needy. She is not afraid of the snow for her household: for all her household are clothed with scarlet. She maketh herself coverings of tapestry; her clothing is silk and purple. Her husband is known in the gates, when he sitteth among the elders of the land. She maketh fine linen, and selleth it; and delivereth girdles unto the merchant. Strength and honour are her clothing; and she shall rejoice in time to come. She openeth her mouth with wisdom; and in her tongue is the law of kindness. She looketh well to the ways of her household, and eateth not the bread of idleness. Her children arise up, and call her blessed; her husband also, and he praiseth her. Many daughters have done virtuously, but thou excellest them all. Favour is deceitful, and beauty is vain: but a woman that feareth the Lord, she shall be praised. Give her of the fruit of her hands; and let her own works praise her in the gates.

\gradall{Full of grace are thy lips: because God hath blessed thee for ever. ℣. Because of the word of truth, of meekness, and righteousness: and thy right hand shall teach thee terrible things.}{Alleluia, alleluia. ℣. In thy comeliness and in thy beauty go forth, ride prosperously, and reign. Alleluia.}

\begin{rubric}
{In Septuagesimatide or Lent, replacing the Alleluia:}
\end{rubric}\par\noindent
\tract{Come, Spouse of Christ, receive the crown, which the Lord hath prepared for thee for ever. ℣. Thou hast loved righteousness, and hated iniquity: wherefore God, even thy God, hath anointed thee with the oil of gladness above thy fellows. ℣. In thy comeliness and in thy beauty go forth, ride prosperously, and reign.}

\begin{rubric}
{In Eastertide, replacing the Lesser Alleluia:}
\end{rubric}\par\noindent
\alleluia{Alleluia, alleluia. ℣. In thy comeliness and in thy beauty, go forth, ride prosperously, and reign. Alleluia. ℣. Because of the word of truth, of meekness, and righteousness: and thy right hand shall teach thee terrible things. Alleluia.}

\readingcitation{Gospel}{Matthew 13:44}
\lett{A}{t that time:} Jesus spake this parable unto his disciples: The kingdom of heaven is like unto treasure hid in a field; the which when a man hath found, he hideth, and for joy thereof goeth and selleth all that he hath, and buyeth that field.

Again, the kingdom of heaven is like unto a merchant man, seeking goodly pearls: who, when he had found one pearl of great price, went and sold all that he had, and bought it.

Again, the kingdom of heaven is like unto a net, that was cast into the sea, and gathered of every kind: which, when it was full, they drew to shore, and sat down, and gathered the good into vessels, but cast the bad away. So shall it be at the end of the world: the angels shall come forth, and sever the wicked from among the just, and shall cast them into the furnace of fire: there shall be wailing and gnashing of teeth.

Jesus saith unto them, Have ye understood all these things? They say unto him, Yea, Lord. Then said he unto them, Therefore every scribe which is instructed unto the kingdom of heaven is like unto a man that is an householder, which bringeth forth out of his treasure things new and old.

\offertory{Full of grace are thy lips: because God hath blessed thee for ever and ever (Alleluia).}

\secret
\lett{G}{rant,} O Lord, that like as thy faithful people do acknowledge that in tribulation they have been succoured by the merits of thy Saints: so this oblation, which they here do offer unto thee in honour of the same, may be acceptable in thy sight. Through.

\communion{Thou hast loved righteousness, and hated iniquity: wherefore God, even thy God, hath anointed thee with the oil of gladness above thy fellows (Alleluia).}

\postcommunion
\lett{O}{Lord,} who hast satisfied thy family with sacred gifts: we beseech thee that we may at all times be comforted by the intercession of her whose festival we celebrate. Through.

\begin{rubric}
	The following Epistle may be used for a Widow instead of the Epistle provided above.
\end{rubric}
\readingcitation{Another Epistle}{1 Timothy 5:3}\label{1Timothy53}
\lett{D}{early beloved:} Honour widows that are widows indeed. But if any widow have children or nephews, let them learn first to shew piety at home, and to requite their parents: for that is good and acceptable before God. Now she that is a widow indeed, and desolate, trusteth in God, and continueth in supplications and prayers night and day. But she that liveth in pleasure is dead while she liveth. And these things give in charge, that they may be blameless. But if any provide not for his own, and specially for those of his own house, he hath denied the faith, and is worse than an infidel. Let not a widow be taken into the number under threescore years old, having been the wife of one man, well reported of for good works; if she have brought up children, if she have lodged strangers, if she have washed the saints' feet, if she have relieved the afflicted, if she have diligently followed every good work.

\subby{Common of the Dedication of a Church}\label{CommonDedication}
\fancyhead[RO,LE]{\textit{Church Dedication}}

\subbysub{I Evensong}

\gregorioscore{resources/gabc/Common/DedicationEvensong.gabc}

℣. This is the Lord's house, firmly builded.

℟. It is well founded on the solid rock.

\antiphon{Mag.}{The Lord hath hallowed {\dag} his tabernacle: for this is the house of the Lord, wherein men may call upon his Name; whereof it is written, My Name shall be therein, saith the Lord.}

\subbysub{Mattins}

\begin{rubric}
	The Invitatory Hymn is from I Evensong.
\end{rubric}

\officehymn

\gregorioscore{resources/gabc/Common/DedicationMattins.gabc}

℣. This is the Lord's house, firmly builded.

℟. It is well founded on the solid rock.

\antiphon{Ben.}{Zacchaeus, {\dag} make haste, and come down; for today I must abide at thy house: and he made haste, and came down, and received him joyfully into his house. This day is salvation come to this house form the Lord, alleluia.}

\subbysub{II Evensong}

\begin{rubric}
	The Hymn is of I Evensong with the following Versicle \& Antiphon.
\end{rubric}

℣. Holiness becometh thine house, O Lord.

℟. For ever.

\antiphon{Mag.}{O how dreadful {\dag} is this place! truly this is none other but the house of God, and the gate of heaven.}

\introit
\lett{H}{ow} dreadful is this place: this is the house of God, and the gate of heaven: and men shall call it the palace of God (Alleluia, alleluia). \textit{Ps.} O how amiable are thy dwellings, thou Lord of hosts! My soul hath a desire and longing to enter into the courts of the Lord.

\collect
\lett{O}{God,} who year by year renewest unto us the day of consecration of this thy holy temple, and dost ever bring us again in safety to thy sacred mysteries: graciously hear the prayers of thy people, and grant; that whosoever entereth this temple to ask for blessings may rejoice in the obtaining of all his petitions. Through.

\begin{rubric}
	On the day itself of the Dedication of a Church and through the Octave, and when the Collect is to be varied, is said the following.
\end{rubric}
\lett{O}{God,} who thyself invisible, containest all things, but dost for the salvation of mankind shew forth visibly the signs of thy power: enlighten this temple with the power of thine in-dwelling, and grant; that all they, who come together here to ask thy mercy, may, in whatsoever tribulation they call upon thee, obtain the blessings of thy consolation. Through.

\readingcitation{Epistle}{Revelation 21:2}
\lett{I}{n those days:} I saw the holy city, new Jerusalem, coming down from God out of heaven, prepared as a bride adorned for her husband. And I heard a great voice out of heaven saying, Behold, the tabernacle of God is with men, and he will dwell with them, and they shall be his people, and God himself shall be with them, and be their God. And God shall wipe away all tears from their eyes; and there shall be no more death, neither sorrow, nor crying, neither shall there be any more pain: for the former things are passed away. And he that sat upon the throne said, Behold, I make all things new. 

\gradall{This dwelling is God's handywork, it is a mystery beyond all price, that cannot be spoken against. ℣. O God, before whom standeth the choir of Angels, graciously hear the prayers of thy servants.}{Alleluia, alleluia. ℣. I will worship toward thy holy temple: and praise thy name. Alleluia.}

\begin{rubric}
{In Septuagesimatide or Lent, replacing the Alleluia:}
\end{rubric}\par\noindent
\tract{They that put their trust in the Lord shall be even as the mount Sion: which may not be removed, but standeth fast for ever. ℣. The hills stand about Jerusalem, even so standeth the Lord round about his people, from this time forth for evermore.}

\begin{rubric}
{In Eastertide, replacing the Lesser Alleluia:}
\end{rubric}\par\noindent
\alleluia{Alleluia, alleluia. ℣. I will worship toward thy holy temple: and praise thy name. Alleluia. ℣. The house of the Lord is well founded upon a sure rock. Alleluia.}

\readingcitation{Gospel}{Luke 19:1}
\lett{A}{t that time:} Jesus entered and passed through Jericho. And, behold, there was a man named Zacch{\ae}us, which was the chief among the publicans, and he was rich. And he sought to see Jesus who he was; and could not for the press, because he was little of stature. And he ran before, and climbed up into a sycomore tree to see him: for he was to pass that way. And when Jesus came to the place, he looked up, and saw him, and said unto him, Zacch{\ae}us, make haste, and come down; for to day I must abide at thy house. And he made haste, and came down, and received him joyfully. And when they saw it, they all murmured, saying, That he was gone to be guest with a man that is a sinner. And Zacch{\ae}us stood, and said unto the Lord; Behold, Lord, the half of my goods I give to the poor; and if I have taken any thing from any man by false accusation, I restore him fourfold. And Jesus said unto him, This day is salvation come to this house, forsomuch as he also is a son of Abraham. For the Son of man is come to seek and to save that which was lost.

\begin{rubric}
	After Septuagesima the \emph{Alleluia} at the end of the following Offertory is omitted.
\end{rubric}
\offertory{O Lord God, in the uprightness of mine heart I have willingly offered all these things; and now I have seen with joy thy people, which are present here: O God of Israel, keep this imagination of their heart, alleluia.}

\secret

\begin{inhead}
	In the Dedication itself
\end{inhead}
\lett{O}{Lord,} we beseech thee, mercifully hear our prayers: that all we, who are gathered within the precincts of this temple, whose dedication year by year we celebrate, may perfectly serve thee, both in body and soul; that, like as we do here present our offerings unto thee, so we may by thee be enabled to attain unto the rewards of everlasting felicity. Through.

\begin{inhead}
	Outside the dedicated Church itself
\end{inhead}
\lett{O}{Lord,} we beseech thee, mercifully hear our prayers: that, like as we do here present our offerings unto thee, so we may by thee be enable to attain unto the rewards of everlasting felicity. Through.

\begin{inhead}
	On the day itself of the Dedication of a Church, and through the Octave, and when the secret is to be varied, is said the following.
\end{inhead}
\lett{O}{God,} who art the author of the gifts to be consecrated unto thee, pour forth upon this house of prayer thy blessing: that all who shall call upon thy name therein may perceive the succour of thy defence. Through.

\begin{rubric}
	The common Preface, even in Lent, although there shall occur some Commemoration or common Octave with a Preface of their own, which are not of some Mystery of our Lord.
\end{rubric}

\communion{My house shall be called the house of prayer, saith the Lord: in it every one that asketh receiveth; and he that seeketh findeth: and to him that knocketh it shall be opened (Alleluia).}

\postcommunion
\lett{O}{God,} who of elect and living stones dost fashion for thy majesty an everlasting habitation: assist the supplications of us thy people; that, like as thy Church increaseth in visible habitations, so it may grow and prosper in spiritual advancement. Through.

\begin{rubric}
	On the day itself of the Dedication of a Church, and through the Octave, or where the Postcommunion is to be varied, is said the following.
\end{rubric}
\lett{W}{e} beseech thee, almighty God: that in this place, which we, though unworthy, have dedicated to thy name, thou wouldest incline thy merciful ears to all that call upon thee. Through.

\begin{rubric}
	On the day of the Dedication itself, Mass is said as above with the second Prayers; and the Collect of the Mystery or Saint in whose honour the church is dedicated is added to the first under one conclusion.
\end{rubric}

\begin{rubric}
	During the Octave, Mass is said as above; but within the Octave the \nth{2} Collect is of St. Mary, from the Votive Mass of the Season: \nth{3} against the persecutors of the Church or for the Chief Bishop.
\end{rubric}

\begin{rubric}
	On the day of the Dedication of an Altar, Mass is said as on the Dedication of a Church, except the Prayers following: to which likewise in the case of a fixed Altar the Prayer of the Mystery or Saint in whose honour the Altar is dedicated is added under one conclusion.
\end{rubric}

\collect
\lett{O}{God,} who from the whole assembly of thy Saints dost prepare for thyself an everlasting habitation: grant to this building the increase of heavenly grace; that like as we do here venerate with devout affection the relics of thy Saints, so we may ever be aided by their merits. Through.

\secret
\lett{W}{e} beseech thee, thy Lord our God, let thy Holy Spirit come down upon this altar: that he may sanctify the gifts of thy people, and graciously cleanse the hearts of them that receive the same. Through . . . in the unity of the same Holy Spirit.

\postcommunion
\lett{A}{lmighty} and everlasting God, sanctify with the power of thy heavenly benediction this altar, which we have dedicated to the honour of thy name: and shew forth upon all that put their trust in thee the bounty of thy succour; that they may here obtain the grace of thy sacraments and the fulfilment of their desires. Through.

\subby{Common of Feasts of the Blessed Virgin Mary}\label{CommonBVM}
\fancyhead[RO,LE]{\textit{Blessed Virgin Mary}}

\subbysub{I Evensong}

\gregorioscore{resources/gabc/Common/BVMEvensong.gabc}

℣. Vouchsafe that I may praise thee, O holy Virgin.

℟. Give me strength against thine enemies.

\antiphon{Mag.}{O holy Mary, {\dag} help thou the suffering, strengthen the faint-hearted, comfort the sorrowful; pray for the people, entreat for the clergy, intercede for all womankind vowed unto God: may all acknowledge the help of thy prayer, who celebrate thy holy festival.}

\subbysub{Mattins}

\begin{inhead}
	Invitatory Hymn
\end{inhead}

\gregorioscore{resources/gabc/Common/BVMInvitatory.gabc}

\begin{inhead}
	Office Hymn
\end{inhead}

\gregorioscore{resources/gabc/Common/BVMMattins.gabc}

℣. Full of grace are thy lips.

℟. Because God hath blessed thee for ever.

\antiphon{Ben.}{Blessed art thou, {\dag} O Mary, for thou hast believed: and there shall be a performance in thee of those things which were told thee from the Lord, alleluia.}

\subbysub{II Evensong}

\begin{rubric}
	The Hymn \& Versicle are of I Evensong with the following Antiphon.
\end{rubric}

\antiphon{Mag.}{All generations {\dag} shall call me blessed: for God hath regarded the lowliness of his handmaiden.}

\introit
\lett{H}{ail,} O Mother most holy, who in childbirth didst bring forth the Monarch: him who o'er heaven and earth reigneth for ever and ever (Alleluia, alleluia). \textit{Ps.} My heart is inditing of a good matter: I speak of the things which I have made unto the King.

\collect
\lett{G}{rant,} we beseech thee, O Lord God, that we thy servants may enjoy perpetual health of mind and of body: and, at the glorious intercession of blessed Mary ever Virgin, be delivered from present sadness, and rejoice in everlasting gladness. Through.

\readingcitation{Epistle}{Ecclesiasticus 24:9}
%RV:
\lett{H}{e} created me from the beginning before the world; And to the end I shall not fail. In the holy tabernacle I ministered before him; And so was I established in Sion. In the beloved city likewise he gave me rest; And in Jerusalem was my authority. And I took root in a people that was glorified, Even in the portion of the Lord’s own inheritance. 

%\lett{H}{e} created me from the beginning before the world, and I shall never fail, in the holy tabernacle I served before him. And so was I established in Sion, likewise in the beloved city he gave me rest, and in Jerusalem was my power. And I took root in an honourable people, even in the portion of the Lord's inheritance, and my abode is in the full assembly of the saints.

\gradall{Blessed and venerable art thou, O Virgin Mary: who without spot wast found Mother of the Saviour. ℣. Virgin Mother of God, he whom the whole world containeth not, being made man lay hid in thy womb.}{Alleluia, alleluia. ℣. After childbirth, O Virgin, thou didst remain inviolate: Mother of God, intercede for us. Alleluia.}
\begin{rubric}
	In Advent, the following Alleluia Verse replaces the one above.
\end{rubric}
\alleluia{Alleluia, alleluia. ℣. Hail, Mary, full of grace; the Lord is with thee: blessed art thou among women. Alleluia.}

\begin{rubric}
{In Septuagesimatide or Lent, replacing the Alleluia:}
\end{rubric}\par\noindent
\tract{Rejoice, O Virgin Mary, alone thou hast destroyed all heresies. ℣. Who didst believe the words of the Archangel Gabriel. ℣. Whilst a Virgin thou didst bring forth him who is God and man: and after childbirth, O Virgin, inviolate didst remain. ℣. Mother of God intercede for us.}

\begin{rubric}
{In Eastertide, replacing the Lesser Alleluia:}
\end{rubric}\par\noindent
\alleluia{Alleluia, alleluia. ℣. Now hath blossomed Jesse's rod: a Virgin bears both man and God: God restoreth peace to men, high and low are one again. Alleluia. ℣. Hail, Mary, full of grace, the Lord is with thee: blessed art thou among women. Alleluia.}

\readingcitation{Gospel}{Luke 11:27}
\lett{A}{t that time:} As Jesus spake to the multitudes, a certain woman of the company lifted up her voice, and said unto him: Blessed is the womb that bare thee, and the paps which thou hast sucked. But he said: Yea rather, blessed are they that hear the word of God, and keep it.

\offertory{Hail, Mary, full of grace; the Lord is with thee: blessed art thou among women, and blessed is the fruit of thy womb (Alleluia).}

\secret
\lett{T}{hrough} thy mercy, O Lord, and the intercession of blessed Mary ever Virgin, may this oblation avail for our prosperity and peace, both now and for ever. Through.

\communion{Blessed is the womb of the Virgin Mary, that bore the Son of the everlasting Father (Alleluia).}

\postcommunion
\lett{G}{rant,} we beseech thee, O Lord: that we, who have received these aids to our salvation, may at all times and in all places be protected through the advocacy of blessed Mary ever Virgin; in whose honour we have made these offerings to thy majesty. Through.

%MANUAL ADJUSTMENT:
\clearpage
\subby{Office of Saint Mary on Saturday}\label{BVMSaturday}
\fancyhead[RO,LE]{\textit{Saturday BVM}}

\begin{rubric}
	The Hymns are as in the Common of the Blessed Virgin Mary (p. \pageref{CommonBVM}).
\end{rubric}

\subbysub{Through the Year}

\begin{inhead}
	Evensong
\end{inhead}

℣. Full of grace are thy lips.

℟. Because God hath blessed thee for ever.

\antiphon{Mag.}{O blessed Mother {\dag} and spotless Virgin, thou glorious Queen of the world, intercede for us to the Lord.}

\lett{G}{rant,} we beseech thee, O Lord God, that we thy servants may enjoy perpetual health of mind and of body: and, at the glorious intercession of blessed Mary ever Virgin, be delivered from present sadness, and rejoice in everlasting gladness. Through.
\needspace{4\baselineskip}
\lett{D}{efend} us, we beseech thee, O Lord, from all perils of mind and body: and at the intercession of blessed Joseph, of thy blessed Apostles Peter and Paul, of blessed \textit{N.} and of all Saints, graciously bestow upon us both peace and safety: that all adversity and error being done away, thy Church may serve thee in untroubled freedom. Through.

\begin{inhead}
	Mattins
\end{inhead}

℣. Blessed art thou amongst women.

℟. And blessed is the fruit of thy womb.

\antiphon{Ben.}{O ever blessed Mother of God, {\dag} Mary ever Virgin, temple of the Godhead, hallowed shrine of the Holy Spirit, thou only, above all others, wast acceptable to our Lord Jesus Christ: pray for the people, entreat for the clergy, intercede for all womankind vowed unto God.}

\lett{G}{rant,} we beseech thee, O Lord God, that we thy servants may enjoy perpetual health of mind and of body: and, at the glorious intercession of blessed Mary ever Virgin, be delivered from present sadness, and rejoice in everlasting gladness. Through.
\needspace{4\baselineskip}
\lett{D}{efend} us, we beseech thee, O Lord, from all perils of mind and body: and at the intercession of blessed Joseph, of thy blessed Apostles Peter and Paul, of blessed \textit{N.} and of all Saints, graciously bestow upon us both peace and safety: that all adversity and error being done away, thy Church may serve thee in untroubled freedom. Through.

\subbysub{After Christmas}

\begin{rubric}
	From Saturday after the Epiphany Octave through Saturday before Purification.
\end{rubric}

\begin{inhead}
    Evensong
\end{inhead}

℣. Full of grace are thy lips.

℟. Because God hath blessed thee for ever.

\antiphon{Mag.}{Great {\dag} is the mystery of the inheritance: the womb of her that knew not man is become the temple of the Godhead: by taking flesh of her, he was no way defiled: all the nations shall gather, saying: Glory be to thee, O Lord.}

\needspace{4\baselineskip}
\lett{O}{God,} who by the virgin child-bearing of blessed Mary hast bestowed upon mankind the rewards of eternal salvation: grant, we beseech thee; that we may perceive her intercession for us, through whom we have been counted worthy to receive the author of life, Jesus Christ thy Son, our Lord. Who liveth.
\needspace{4\baselineskip}
\lett{D}{efend} us, we beseech thee, O Lord, from all perils of mind and body: and at the intercession of blessed Joseph, of thy blessed Apostles Peter and Paul, of blessed \textit{N.} and of all Saints, graciously bestow upon us both peace and safety: that all adversity and error being done away, thy Church may serve thee in untroubled freedom. Through.

\begin{inhead}
	Mattins
\end{inhead}

℣. Blessed art thou amongst women.

℟. And blessed is the fruit of thy womb.

\antiphon{Ben.}{A great and wondrous mystery {\dag} is made known to us this day; a new thing is wrought in both natures: God is made man; that which was, remained, and that which was not, he assumed; suffering no confusion, nor yet division.}

\needspace{4\baselineskip}
\lett{O}{God,} who by the virgin child-bearing of blessed Mary hast bestowed upon mankind the rewards of eternal salvation: grant, we beseech thee; that we may perceive her intercession for us, through whom we have been counted worthy to receive the author of life, Jesus Christ thy Son, our Lord. Who liveth.
\needspace{4\baselineskip}
\lett{D}{efend} us, we beseech thee, O Lord, from all perils of mind and body: and at the intercession of blessed Joseph, of thy blessed Apostles Peter and Paul, of blessed \textit{N.} and of all Saints, graciously bestow upon us both peace and safety: that all adversity and error being done away, thy Church may serve thee in untroubled freedom. Through.

\subbysub{Eastertide}

\begin{inhead}
	Evensong
\end{inhead}

℣. Full of grace art thy lips, alleluia.

℟. Because God hath blessed thee for ever, alleluia.

\antiphon{Mag.}{O Queen of heaven, {\dag} be joyful, alleluia; Because he whom so meetly thou barest, alleluia, Hath arisen, as he promised, alleluia: Pray for us to the Father, alleluia.}

\needspace{4\baselineskip}
\lett{G}{rant,} we beseech thee, O Lord God, that we thy servants may enjoy perpetual health of mind and of body: and, at the glorious intercession of blessed Mary ever Virgin, be delivered from present sadness, and rejoice in everlasting gladness. Through.
\needspace{4\baselineskip}
\lett{O}{God,} who didst will that thy Son should for us undergo the burden of the Cross, that he might deliver us from the power of the enemy: grant unto us thy servants that we may attain unto the joy of his resurrection. Through the same.

\begin{inhead}
	Mattins
\end{inhead}

\begin{rubric}
	Mattins is as in Evensong.
\end{rubric}

\subby{Seasonal Prayers}
\fancyhead[RO,LE]{\textit{Seasonal Prayers}}

   \subbysub{Of St. Mary in Advent}\label{SPMaryInAdvent}
   \begin{inhead}
   	Collect
   \end{inhead}
   \lett{O}{God,} who wast pleased that thy Word should take flesh of the womb of the blessed Virgin Mary at the message of an Angel: grant to thy humble servants; that we, who believe her to be truly the Mother of God, may be aided by her intercession in thy sight. Through the same.
  \begin{inhead}
  	Secret
  \end{inhead}
   \lett{W}{e} beseech thee, O Lord, stablish in our hearts the mysteries of thy true religion: that, as we believe thy Son, conceived of a Virgin, to be very God and very man; so by the power of his life-giving resurrection we may be found worthy to attain unto everlasting felicity. Through the same.
   \begin{inhead}
   	Postcommunion
   \end{inhead}
   \lett{W}{e} beseech thee, O Lord, pour thy grace into our hearts: that, as we have known the incarnation of thy Son Christ by the message of an Angel; so by his cross and passion we may be brought unto the glory of his resurrection. (Through the same.)


    \subbysub{Of St. Mary after Christmas}\label{SPMaryPostChristmas}
   \begin{inhead}
   	Collect
   \end{inhead}
    \lett{O}{God,} who by the virgin child-bearing of blessed Mary hast bestowed upon mankind the rewards of eternal salvation: grant, we beseech thee; that we may perceive her intercession for us, through whom we have been counted worthy to receive the author of life, Jesus Christ thy Son, our Lord. (Who liveth.)
  \begin{inhead}
  	Secret
  \end{inhead}
    \lett{T}{hrough} thy mercy, O Lord, and the intercession of blessed Mary ever Virgin, may this oblation avail for our prosperity and peace both now and for ever. (Through.)
   \begin{inhead}
   	Postcommunion
   \end{inhead}
    \lett{M}{ay} this communion, O Lord, cleanse us from guilt: and at the intercession of blessed Mary the Virgin Mother of God, make us partakers of thy heavenly healing. (Through the same.)
    
    
    \subbysub{Of St. Mary in Eastertide}\label{SPMaryInEaster}
   \begin{inhead}
   	Collect
   \end{inhead}
    \lett{G}{rant,} we beseech thee, O Lord God, that we thy servants may enjoy perpetual health of mind and of body: and, at the glorious intercession of blessed Mary ever Virgin, be delivered from present sadness, and rejoice in everlasting gladness. (Through.)
  \begin{inhead}
  	Secret
  \end{inhead}
    \lett{T}{hrough} thy mercy, O Lord, and the intercession of blessed Mary ever Virgin, may this oblation avail for our prosperity and peace, both now and ever. (Through.)
   \begin{inhead}
   	Postcommunion
   \end{inhead}
    \lett{G}{rant,} we beseech thee, O Lord: that we who have received these aids to our salvation may at all times and in all places be protected by the advocacy of blessed Mary ever Virgin; in whose honour we have made these offerings to thy majesty. (Through.)\\

    \subbysub{Of Saints}\label{SPSaints}
   \begin{inhead}
   	Collect
   \end{inhead}
    \lett{D}{efend} us, O Lord, we beseech thee, from all dangers of mind and of body: and at the intercession of the blessed and glorious Mother of God, Mary ever Virgin, with blessed Joseph, thy blessed Apostles Peter and Paul, blessed \emph{N.}, and all the Saints, graciously grant us health and peace; that all adversities and errors being done away, thy Church may serve thee in freedom and quietness. (Through the same.)
  \begin{inhead}
  	Secret
  \end{inhead}
    \lett{G}{raciously} hear us, O God our Saviour: and by the power of this sacrament defend us from all enemies of mind and of body; granting us grace in this world, and glory in that which is to come. (Through.)
   \begin{inhead}
   	Postcommunion
   \end{inhead}
    \lett{W}{e} beseech thee, O Lord, that the gift now offered in this divine sacrament may cleanse and defend us: that at the intercession of blessed Mary the Virgin Mother of God, of blessed Joseph, of thy blessed Apostles Peter and Paul, of blessed \emph{N.} and of all the Saints; it may set us free from every perverse way, and deliver us from all adversities. (Through the same.)
    
   \subbysub{Against the Persecutors of the Church}\label{SPAgainst}
   \begin{inhead}
   	Collect
   \end{inhead}
   \lett{W}{e} beseech thee, O Lord, mercifully to hear the prayers of thy Church: that all adversities and errors being done away, she may serve thee in freedom and quietness. Through.
  \begin{inhead}
  	Secret
  \end{inhead}
   \lett{D}{efend} us, O Lord, who wait upon thy mysteries: that we, cleaving fast to things heavenly, may serve thee both in body and soul. Through.
   \begin{inhead}
   	Postcommunion
   \end{inhead}
   \lett{W}{e} beseech thee, O Lord our God: that we, whom thou makest to rejoice in the partaking of heavenly things, may by thee be defended against all earthly perils. Through.
   
   \subbysub{For the Chief Bishop}\label{SPChiefBishop}
   \begin{inhead}
   	Collect
   \end{inhead}
   \lett{O}{God,} the pastor and ruler of all the faithful, mercifully look upon thy servant \emph{N.}, whom thou hast been pleased to set as pastor over thy Church; grant him, we beseech thee, to be in word and conversation a wholesome example to the people committed to his charge; that he with them may attain unto everlasting life. Through.
  \begin{inhead}
  	Secret
  \end{inhead}
   \lett{L}{ook} favourably, we beseech thee, O Lord, upon the gifts which we offer: and guide with thy continual protection thy servant \textit{N.}, whom thou hast been pleased to set as pastor over thy Church. Through.
   \begin{inhead}
   	Postcommunion
   \end{inhead}
   \lett{W}{e} beseech thee, O Lord, that the divine sacrament which we have here received may be our protection: and together with the flock entrusted to his charge ever save and defend thy servant \textit{N.}; whom thou hast been pleased to set as pastor over thy Church. Through.
   
   \subbysub{Of the Holy Ghost}\label{SPHolyGhost}
   \begin{inhead}
   	Collect
   \end{inhead}
   \lett{G}{od,} who didst teach the hearts of thy faithful people, by the sending to them the light of thy Holy Spirit: grant us by the same Spirit to have a right judgment in all things; and evermore to rejoice in his holy comfort. (Through . . . in the unity of the same Holy Spirit.)
  \begin{inhead}
  	Secret
  \end{inhead}
   \lett{S}{anctify,} we beseech thee, O Lord, the gifts which we offer: and cleanse our hearts by the enlightening of the Holy Spirit. (Through . . . in the unity of the same Holy Spirit.)
   \begin{inhead}
   	Postcommunion
   \end{inhead}
   \lett{P}{our} thy Holy Spirit upon us, O Lord, and cleanse our hearts: that they may be made fruitful by the inward sprinkling of his dew. (Through . . . in the unity of the same Holy Spirit.)

    \subbysub{For the Living and Departed}\label{SPLivingDeparted}
   \begin{inhead}
   	Collect
   \end{inhead}
   \lett{A}{lmighty} and everlasting God, who hast dominion both of the living and of the dead, and hast mercy upon all who thou foreknowest will be thine in faith and works: we humbly beseech thee; that all those for whom we are minded to pour forth our prayers, whether in this present world they still be held in the flesh, or being delivered from the body have passed into that which is to come, may at the intercession of all thy Saints obtain of thy bountiful goodness the remission of all their sins. Through.
  \begin{inhead}
  	Secret
  \end{inhead}
   \lett{O}{God,} to whom alone is known the number of the elect, whom thou hast appointed unto heavenly felicity: grant, we beseech thee: that at the intercession of all thy Saints, the names of all those who have been commended to our prayers, and of all thy faithful people, may ever remain written in the book of those that are predestined to everlasting blessedness. Through.
   \begin{inhead}
   	Postcommunion
   \end{inhead}
   \lett{A}{lmighty} and merciful God, cleanse us, we pray thee, by the sacraments which we have received: and, at the intercession of all thy Saints, grant; that this thy sacrament may not bring upon us guilt to our punishment but pardon to our salvation: that it may cleanse away sins, that it may confirm the feeble, that it may defend us against every danger in this world: that it may obtain for thy faithful people both living and dead the remission of all their sins. Through.