\phantomsection
\addcontentsline{toc}{section}{Gregorian Propers}
\fancyhead[C]{\LARGE Gregorian Propers}
\fancyhead[RO,LE]{\textit{Of Saints}}
\fancyhead[RE,LO]{}

\subby{St. Andrew}
\fancyhead[RE,LO]{St. Andrew}
\begin{inhead}
    {Second Class Double\\
30 November}
\end{inhead}

%Considered for BCP:
%\collect
%\lett{W}{e} humbly entreat thy majesty, O Lord: that as blessed Andrew the Apostle was to thy Church a preacher and governor; so he may be a perpetual intercessor for us in thy sight. Through.


\readingcitation{Epistle}{Romans 10:9}
\lett{B}{rethren:} With the heart man believeth unto righteousness; and with the mouth confession is made unto salvation. For the scripture saith, Whosoever believeth on him shall not be ashamed. For there is no difference between the Jew and the Greek: for the same Lord over all is rich unto all that call upon him. For whosoever shall call upon the name of the Lord shall be saved. How then shall they call on him in whom they have not believed? and how shall they believe in him of whom they have not heard? and how shall they hear without a preacher? And how shall they preach, except they be sent? as it is written, How beautiful are the feet of them that preach the gospel of peace, and bring glad tidings of good things! But they have not all obeyed the gospel. For Esaias saith, Lord, who hath believed our report? So then faith cometh by hearing, and hearing by the word of God. But I say, Have they not heard? Yes verily, their sound went into all the earth, and their words unto the ends of the world.

\subby{St. Thomas}
\fancyhead[RE,LO]{St. Thomas}
\begin{inhead}
    {Second Class Double\\
21 December}
\end{inhead}
%Considered for BCP:
%\collect
%\lett{G}{rant} us, we beseech thee, O Lord, to glory in the solemnity of thy blessed Apostle Thomas: that we may ever be succoured by his protection; and follow his faith with worthy devotion. Through.
%\begin{rubric}
%    Commemoration of the Feria, except in Conventual Masses if an Ember Day occur.
%\end{rubric}

%\readingcitation{Epistle}{Ephesians 2:19}
%\lett{B}{rethren:} Now ye are no more strangers and foreigners, but fellowcitizens with the saints, and of the household of God; And are built upon the foundation of the apostles and prophets, Jesus Christ himself being the chief corner stone; In whom all the building fitly framed together groweth unto an holy temple in the Lord: In whom ye also are builded together for an habitation of God through the Spirit.

\readingcitation{Gospel}{John 20:24}
\lett{A}{t that time:} Thomas, one of the twelve, called Didymus, was not with them when Jesus came. The other disciples therefore said unto him, We have seen the Lord. But he said unto them, Except I shall see in his hands the print of the nails, and put my finger into the print of the nails, and thrust my hand into his side, I will not believe. And after eight days again his disciples were within, and Thomas with them: then came Jesus, the doors being shut, and stood in the midst, and said, Peace be unto you. Then saith he to Thomas, Reach hither thy finger, and behold my hands; and reach hither thy hand, and thrust it into my side: and be not faithless, but believing. And Thomas answered and said unto him, My Lord and my God. Jesus saith unto him, Thomas, because thou hast seen me, thou hast believed: blessed are they that have not seen, and yet have believed.

\subby{Conversion of St. Paul}
\fancyhead[RE,LO]{Conversion of St. Paul}
\begin{inhead}
    {Greater Double\\
25 January}
\end{inhead}
\collect
\lett{O}{God,} who by the preaching of blessed Paul the Apostle hast taught the whole world: grant us, we beseech thee; that we, who this day celebrate his Conversion, may through his example draw nearer unto thee. Through.
\needspace{4\baselineskip}
\lett{O}{God,} who didst bestow upon thy blessed Apostle Peter the keys of the kingdom of heaven, and didst appoint unto him the high priesthood of binding and loosing: vouchsafe; that by the help of his intercession we may be delivered from the bonds of our iniquities. Who livest.

\readingcitation{Gospel}{Matthew 19:27}
\lett{A}{t that time:} Peter said unto Jesus: Behold, we have forsaken all, and followed thee; what shall we have therefore? And Jesus said unto them, Verily I say unto you, That ye which have followed me, in the regeneration when the Son of man shall sit in the throne of his glory, ye also shall sit upon twelve thrones, judging the twelve tribes of Israel. And every one that hath forsaken houses, or brethren, or sisters, or father, or mother, or wife, or children, or lands, for my name's sake, shall receive an hundredfold, and shall inherit everlasting life.

\subby{Purification of the Blessed Virgin Mary}
\fancyhead[RE,LO]{Purification of the B.V.M.}
\begin{inhead}
    {Second Class Double\\
2 February}
\end{inhead}

\readingcitation{Epistle}{Malachi 3:1}
\lett{B}{ehold,} I will send my messenger, and he shall prepare the way before me: and the Lord, whom ye seek, shall suddenly come to his temple, even the messenger of the covenant, whom ye delight in: behold, he shall come, saith the \textsc{Lord} of hosts. But who may abide the day of his coming? and who shall stand when he appeareth? for he is like a refiner's fire, and like fullers' soap: And he shall sit as a refiner and purifier of silver: and he shall purify the sons of Levi, and purge them as gold and silver, that they may offer unto the \textsc{Lord} an offering in righteousness. Then shall the offering of Judah and Jerusalem be pleasant unto the \textsc{Lord}, as in the days of old, and as in former years.

\readingcitation{Gospel}{Luke 2:22}
\lett{A}{t that time:} When the days of her purification according to the law of Moses were accomplished, they brought him to Jerusalem, to present him to the Lord; (As it is written in the law of the Lord, Every male that openeth the womb shall be called holy to the Lord;) And to offer a sacrifice according to that which is said in the law of the Lord, A pair of turtledoves, or two young pigeons. And, behold, there was a man in Jerusalem, whose name was Simeon; and the same man was just and devout, waiting for the consolation of Israel: and the Holy Ghost was upon him. And it was revealed unto him by the Holy Ghost, that he should not see death, before he had seen the Lord's Christ. And he came by the Spirit into the temple: and when the parents brought in the child Jesus, to do for him after the custom of the law, Then took he him up in his arms, and blessed God, and said, Lord, now lettest thou thy servant depart in peace, according to thy word: For mine eyes have seen thy salvation, Which thou hast prepared before the face of all people; A light to lighten the Gentiles, and the glory of thy people Israel.

%\subby{St. Matthias}
%\fancyhead[RE,LO]{St. Matthias}
%\begin{inhead}
%    {Second Class Double\\
%24 February (25 February in a Leap Year)}
%\end{inhead}
%Considered for BCP:
%\collect
%\lett{O}{God,} who didst join blessed Matthias to the college of thine Apostles: grant, we beseech thee; that through his mediation we may ever perceive thy tender mercy towards us. Through.
%\begin{rubric}
%    In Lent, Commemoration is made of the Feria in non-conventual Masses.
%\end{rubric}


\subby{St. Benedict}
\fancyhead[RE,LO]{St. Benedict}
\begin{inhead}
    {Second Class Double\\
21 March}
\end{inhead}

\collect
\lett{W}{e} beseech thee, O Lord, that the prayers of the blessed Abbot, Benedict may commend us unto thee: that by his advocacy thou wouldest vouchsafe unto us those things to which by our own merits we cannot attain. Through.
\begin{rubric}
    In Lent, Commemoration is made of the Feria.
\end{rubric}
\readingcitation{Epistle}{Ecclesiasticus 45:1}
\lett{B}{eloved} of God and men, whose memorial is blessed. He made him like to the glorious saints, and magnified him, so that his enemies stood in fear of him. By his words he caused the wonders to cease, and he made him glorious in the sight of kings, and gave him a commandment for his people, and shewed him part of his glory. He sanctified him in his faithfuless and meekness, and chose him out of all men. He made him to hear his voice, and brought him into the dark cloud, and gave him commandments before his face, even the law of life and knowledge.

\gradtr{Thou hast prevented him, O Lord, with the blessings of goodness: thou hast set a crown of pure gold upon his head. ℣. He asked life of thee, and thou gavest him a long life, even for ever and ever.}{Blessed is the man that feareth the Lord: he hath great delight in his commandments. ℣. His seed shall be mighty upon earth: the generation of the faithful shall be blessed. ℣. Riches and plenteousness shall be in his house: and his righteousness endureth for ever.}

\readingcitation{Gospel}{Matthew 19:27}
\lett{A}{t that time:} Peter said unto Jesus: Behold, we have forsaken all, and followed thee; what shall we have therefore? And Jesus said unto them, Verily I say unto you, That ye which have followed me, in the regeneration when the Son of man shall sit in the throne of his glory, ye also shall sit upon twelve thrones, judging the twelve tribes of Israel. And every one that hath forsaken houses, or brethren, or sisters, or father, or mother, or wife, or children, or lands, for my name's sake, shall receive an hundredfold, and shall inherit everlasting life.

\secret
\lett{W}{e} beseech thee, O Lord, that thy holy Abbot Benedict may intercede for us: that this sacrifice which we offer and present upon thy holy altar may be profitable unto us for our salvation. Through.
\begin{rubric}
    In Lent, Commemoration is made of the Feria.
\end{rubric}
\communion{A faithful and wise steward, whom his lord hath made ruler over his household, to give them their portion of meat in due season.}
\postcommunion
\lett{L}{et} thy sacrament, O Lord, which we have now received and the prayers of the blessed Abbot, Benedict, effectually defend us: that we may both imitate the example of his conversation, and receive the succour of his intercession. Through.
\begin{rubric}
    In Lent, Commemoration is made of the Feria.
\end{rubric}
\begin{rubric}
    In Lent, the Last Gospel is of the Feria.
\end{rubric}

\subby{Annunciation of the Blessed Virgin Mary}
\fancyhead[RE,LO]{Annunciation of the B.V.M.}
\begin{inhead}
    {First Class Double\\
25 March}
\end{inhead}

%Considered for BCP:
%\collect
%\lett{O}{God,} who wast pleased that thy Word should take flesh of the womb of the Blessed Virgin Mary at the message of an Angel: grant to thy humble servants; that we who believe her to be truly the Mother of God may be aided by her intercession in thy sight. Through the same.
%\begin{rubric}
%    In Lent, Commemoration is made of the Feria in non-conventual Masses.
%\end{rubric}

\readingcitation{Gospel}{Luke 1:26}
\lett{A}{t that time:} The angel Gabriel was sent from God unto a city of Galilee, named Nazareth, To a virgin espoused to a man whose name was Joseph, of the house of David; and the virgin's name was Mary. And the angel came in unto her, and said, Hail, thou that art highly favoured, the Lord is with thee: blessed art thou among women. And when she saw him, she was troubled at his saying, and cast in her mind what manner of salutation this should be. And the angel said unto her, Fear not, Mary: for thou hast found favour with God. And, behold, thou shalt conceive in thy womb, and bring forth a son, and shalt call his name \divineName{Jesus}. He shall be great, and shall be called the Son of the Highest: and the Lord God shall give unto him the throne of his father David: And he shall reign over the house of Jacob for ever; and of his kingdom there shall be no end. Then said Mary unto the angel, How shall this be, seeing I know not a man? And the angel answered and said unto her, The Holy Ghost shall come upon thee, and the power of the Highest shall overshadow thee: therefore also that holy thing which shall be born of thee shall be called the Son of God. And, behold, thy cousin Elisabeth, she hath also conceived a son in her old age: and this is the sixth month with her, who was called barren. For with God nothing shall be impossible. And Mary said, Behold the handmaid of the Lord; be it unto me according to thy word. 

\postcommunion
\lett{W}{e} beseech thee, O Lord, pour thy grace into our hearts; that, as we have known the incarnation of thy Son Jesus Christ by the message of an Angel, so by his cross and passion we may be brought unto the glory of his resurrection. Through the same.

\begin{rubric}
%CHECK:
    If any Feast fall during Holy Week or the Easter Octave, it is transferred to the next available day.
\end{rubric}

\subby{St. George}
\fancyhead[RE,LO]{St. George}
\begin{inhead}
    {First Class Double with a Common Octave\\
23 April}
\end{inhead}
%Sarum-Roman: 0.2

\collect
%Roman collect with indirect invocation of a saint.
\lett{O}{God,} who makest us glad by the merits and intercession of blessed George, thy Martyr: mercifully grant; that we, who through him ask for thy benefits may by the gift of thy grace obtain the same. Through.

\readingcitation{Epistle}{2 Timothy 2:8}
\lett{D}{early beloved:} Remember that Jesus Christ of the seed of David was raised from the dead according to my gospel: Wherein I suffer trouble, as an evil doer, even unto bonds; but the word of God is not bound. Therefore I endure all things for the elect's sakes, that they may also obtain the salvation which is in Christ Jesus with eternal glory. But thou hast fully known my doctrine, manner of life, purpose, faith, longsuffering, charity, patience, Persecutions, afflictions, which came unto me at Antioch, at Iconium, at Lystra; what persecutions I endured: but out of them all the Lord delivered me. Yea, and all that will live godly in Christ Jesus shall suffer persecution.

\gradual{Blessed is the man that feareth the Lord: he hath great delight in his commandments. ℣. His seed shall be mighty upon earth: the generation of the faithful shall be blessed.}

\tract{Thou hast given him his heart's desire, and hast not denied him the request of his lips. ℣. For thou shalt prevent him with the blessing of goodness. ℣. Thou shalt set a crown of pure gold upon his head.}

\begin{rubric}
    If it be Easter, the Alleluia Verse is said instead of the Gradual \& Tract.
\end{rubric}

\alleluia{Alleluia, alleluia. ℣. O Lord, the very heavens shall praise thy wondrous works: and thy truth in the congregation of the Saints. Alleluia. ℣. Thou hast set, O Lord, a crown of pure gold upon his head. Alleluia.}

\readingcitation{Gospel}{John 15:1}
\lett{A}{t that time:} Jesus said to his disciples: I am the true vine, and my Father is the husbandman. Every branch in me that beareth not fruit he taketh away: and every branch that beareth fruit, he purgeth it, that it may bring forth more fruit. Now ye are clean through the word which I have spoken unto you. Abide in me, and I in you. As the branch cannot bear fruit of itself, except it abide in the vine; no more can ye, except ye abide in me. I am the vine, ye are the branches: He that abideth in me, and I in him, the same bringeth forth much fruit: for without me ye can do nothing. If a man abide not in me, he is cast forth as a branch, and is withered; and men gather them, and cast them into the fire, and they are burned. If ye abide in me, and my words abide in you, ye shall ask what ye will, and it shall be done unto you.
\offertory{O Lord, the very heavens shall praise thy wondrous works: and thy truth in the congregation of the saints, (alleluia, alleluia.)}
\secret
\lett{S}{anctify,} O Lord, the gifts which we offer: and at the intercession of blessed George, thy Martyr, cleanse us thereby from the defilements of our iniquities. Through.
\begin{rubric}
    In Lent, Commemoration is made of the Feria in non-conventual Masses.
\end{rubric}
\communion{The righteous shall rejoice in the Lord, and put his trust in him : and all they that are true of heart shall be glad, (alleluia, alleluia.)}
\postcommunion
\lett{W}{e} humbly beseech thee, almighty God: that as thou dost refresh us with thy sacraments, so at the intercession of blessed George, thy Martyr, thou wouldest grant unto us to do thee worthy and acceptable service. Through.
\begin{rubric}
    In Lent, Commemoration is made of the Feria in non-conventual Masses.
\end{rubric}
\begin{rubric}
    In Lent, the Last Gospel is of the Feria in non-conventual Masses.
\end{rubric}

\subby{St. Mark}
\fancyhead[RE,LO]{St. Mark}
\begin{inhead}
    {Second Class Double\\
25 April}
\end{inhead}
%S-R: N/A
%Latest Date of Gregorian Easter
\collect
%Considered for BCP:
%\lett{O}{God,} who hast exalted blessed Mark thine Evangelist by the grace of the preaching of the Gospel: grant, we beseech thee; that we may ever both profit by his learning, and be defended by his prayer. Through.
%\begin{rubric}
%    In Lent, Commemoration is made of the Feria in non-conventual Masses.
%\end{rubric}

\readingcitation{Epistle}{Ezekiel 1:10}
\lett{A}{s} for the likeness of their faces, they four had the face of a man, and the face of a lion, on the right side: and they four had the face of an ox on the left side; they four also had the face of an eagle. Thus were their faces: and their wings were stretched upward; two wings of every one were joined one to another, and two covered their bodies. And they went every one straight forward: whither the spirit was to go, they went; and they turned not when they went. As for the likeness of the living creatures, their appearance was like burning coals of fire, and like the appearance of lamps: it went up and down among the living creatures; and the fire was bright, and out of the fire went forth lightning. And the living creatures ran and returned as the appearance of a flash of lightning.

\readingcitation{Gospel}{Luke 10:1}
\lett{A}{t that time:} The Lord appointed other seventy also, and sent them two and two before his face into every city and place, whither he himself would come. Therefore said he unto them, The harvest truly is great, but the labourers are few: pray ye therefore the Lord of the harvest, that he would send forth labourers into his harvest. Go your ways: behold, I send you forth as lambs among wolves. Carry neither purse, nor scrip, nor shoes: and salute no man by the way. And into whatsoever house ye enter, first say, Peace be to this house. And if the son of peace be there, your peace shall rest upon it: if not, it shall turn to you again. And in the same house remain, eating and drinking such things as they give: for the labourer is worthy of his hire. Go not from house to house. And into whatsoever city ye enter, and they receive you, eat such things as are set before you: And heal the sick that are therein, and say unto them, The kingdom of God is come nigh unto you.

\subby{Sts. Philip \& James}
\fancyhead[RE,LO]{Sts. Philip \& James}
\begin{inhead}
    {Second Class Double\\
1 May}
\end{inhead}

\collect
\lett{O}{God,} who makest us to rejoice in the yearly solemnity of thine Apostles Philip and James: grant, we beseech thee; that, as we rejoice in their merits, so we may be taught by their example. Through.

\readingcitation{Epistle}{Wisdom 5:1}
\lett{T}{he} righteous man shall stand in great boldness before the face of such as have afflicted him, and made no account of his labours. When they see it, they shall be troubled with terrible fear, and shall be amazed at the strangeness of his salvation, so far beyond all that they looked for. And they repenting and groaning for anguish of spirit shall say within themselves, This was he, whom we had sometimes in derision, and a proverb of reproach: We fools accounted his life madness, and his end to be without honour: How is he numbered among the children of God, and his lot is among the saints!

%DOUBLE CHECK GOSPEL


\subby{St. Barnabas}
\fancyhead[RE,LO]{St. Barnabas}
\begin{inhead}
    {Greater Double\\
11 June}
\end{inhead}
%S-R: N/A
%Sarum considered for BCP.
\collect
\lett{O}{God,} who dost gladden us by the merits and intercession of blessed Barnabas thine Apostle: mercifully grant; that we, who through him ask thy benefits, may obtain them by the gift of thy grace. Through.
\readingcitation{Epistle}{Acts 11:21}%DOUBLE CHECK:
\lett{I}{n those days:} A great number believed, and turned unto the Lord. Tidings of these things came unto the ears of the church which was in Jerusalem: and they sent forth Barnabas, that he should go as far as Antioch. Who, when he came, and had seen the grace of God, was glad, and exhorted them all, that with purpose of heart they would cleave unto the Lord. For he was a good man, and full of the Holy Ghost and of faith: and much people was added unto the Lord. Then departed Barnabas to Tarsus, for to seek Saul. And when he had found him, he brought him unto Antioch. And it came to pass that a whole year they assembled themselves with the church, and taught much people. And the disciples were called Christians first in Antioch. %
Now there were in the church that was at Antioch certain prophets and teachers; as Barnabas, and Simeon that was called Niger, and Lucius of Cyrene, and Manaen, which had been brought up with Herod the tetrarch, and Saul. As they ministered to the Lord, and fasted, the Holy Ghost said, Separate me Barnabas and Saul for the work whereunto I have called them. And when they had fasted and prayed, and laid their hands on them, they sent them away.

\readingcitation{Gospel}{Matthew 10:16}
\lett{A}{t that time:} Jesus said unto his disciples: Behold, I send you forth as sheep in the midst of wolves: be ye therefore wise as serpents, and harmless as doves. But beware of men: for they will deliver you up to the councils, and they will scourge you in their synagogues; And ye shall be brought before governors and kings for my sake, for a testimony against them and the Gentiles. But when they deliver you up, take no thought how or what ye shall speak: for it shall be given you in that same hour what ye shall speak. For it is not ye that speak, but the Spirit of your Father which speaketh in you. And the brother shall deliver up the brother to death, and the father the child: and the children shall rise up against their parents, and cause them to be put to death. And ye shall be hated of all men for my name's sake: but he that endureth to the end shall be saved.

\subby{St. John Baptist}\label{StJohnBaptist}
\fancyhead[RE,LO]{St. John Baptist}
\begin{inhead}
    {First Class Double with a Common Octave\\
24 June}
%(25 June is the last possible day of Eastertide)
\end{inhead}

\collect
\lett{O}{God,} who hast made this day honourable unto us by the nativity of blessed John: grant unto thy people the grace of spiritual joys: and direct the minds of all the faithful into the way of everlasting salvation. Through.

\readingcitation{Epistle}{Isaiah 49:1}
\lett{L}{isten,} O isles, unto me; and hearken, ye people, from far; The \divineName{Lord} hath called me from the womb; from the bowels of my mother hath he made mention of my name. And he hath made my mouth like a sharp sword; in the shadow of his hand hath he hid me, and made me a polished shaft; in his quiver hath he hid me; And said unto me, Thou art my servant, O Israel, in whom I will be glorified. Then I said, I have laboured in vain, I have spent my strength for nought, and in vain: yet surely my judgment is with the \divineName{Lord}, and my work with my God. And now, saith the \divineName{Lord} that formed me from the womb to be his servant, to bring Jacob again to him, Though Israel be not gathered, yet shall I be glorious in the eyes of the \divineName{Lord}, and my God shall be my strength. And he said, It is a light thing that thou shouldest be my servant to raise up the tribes of Jacob, and to restore the preserved of Israel: I will also give thee for a light to the Gentiles, that thou mayest be my salvation unto the end of the earth. Thus saith the \divineName{Lord}, the Redeemer of Israel, and his Holy One, to him whom man despiseth, to him whom the nation abhorreth, to a servant of rulers, Kings shall see and arise, princes also shall worship, because of the \divineName{Lord} that is faithful, and the Holy One of Israel, and he shall choose thee.

\readingcitation{Gospel}{Luke 1:57}
\lett{E}{lisabeth's} full time came that she should be delivered; and she brought forth a son. And her neighbours and her cousins heard how the Lord had shewed great mercy upon her; and they rejoiced with her. And it came to pass, that on the eighth day they came to circumcise the child; and they called him Zacharias, after the name of his father. And his mother answered and said, Not so; but he shall be called John. And they said unto her, There is none of thy kindred that is called by this name. And they made signs to his father, how he would have him called. And he asked for a writing table, and wrote, saying, His name is John. And they marvelled all. And his mouth was opened immediately, and his tongue loosed, and he spake, and praised God. And fear came on all that dwelt round about them: and all these sayings were noised abroad throughout all the hill country of Judaea. And all they that heard them laid them up in their hearts, saying, What manner of child shall this be! And the hand of the Lord was with him. And his father Zacharias was filled with the Holy Ghost, and prophesied, saying, Blessed be the Lord God of Israel; for he hath visited and redeemed his people.

\subby{Sts. Peter \& Paul}
\fancyhead[RE,LO]{Sts. Peter \& Paul}
\begin{inhead}
    {First Class Double with a Common Octave\\
29 June}
\end{inhead}

\collect
\lett{O}{God,} who hast hallowed this day by the martyrdom of thine Apostles Peter and Paul: grant unto thy Church in all things to follow the commandment of those; through whom she received the beginning of religion. Through.

\subby{St. Mary Magdalene}
\fancyhead[RE,LO]{St. Mary Magdalene}
\begin{inhead}
    {Greater Double\\
22 July}
\end{inhead}

\collect
\lett{W}{e} beseech thee, O Lord, that as thou didst hearken to the prayers of blessed Mary Magdalene, and on the fourth day didst raise her brother Lazarus from the dead to life: so we may be succoured by her intercession. Who livest.

\readingcitation{Epistle}{Song of Songs 3:2}
\lett{I}{will} rise, and go about the city in the streets, and in the broad ways I will seek him whom my soul loveth: I sought him, but I found him not. The watchmen that go about the city found me: to whom I said, Saw ye him whom my soul loveth? It was but a little that I passed from them, but I found him whom my soul loveth: I held him, and would not let him go, until I had brought him into my mother's house, and into the chamber of her that conceived me. I charge you, O ye daughters of Jerusalem, by the roes, and by the hinds of the field, that ye stir not up, nor awake my love, till he please. Set me as a seal upon thine heart, as a seal upon thine arm: for love is strong as death; jealousy is cruel as the grave: the coals thereof are coals of fire, which hath a most vehement flame. Many waters cannot quench love, neither can the floods drown it: if a man would give all the substance of his house for love, it would utterly be contemned.

\subby{St. James}
\fancyhead[RE,LO]{St. James}
\begin{inhead}
    {Second Class Double\\
25 July}
\end{inhead}

\collect
\lett{B}{e} thou, O Lord, the sanctifier and defender of thy people: that being aided by the protection of thine Apostle James, they may both please thee by their conversation, and serve thee with a quiet mind. Through.

\readingcitation{Epistle}{1 Corinthians 4:9}
\lett{B}{rethren:} I think that God hath set forth us the apostles last, as it were appointed to death: for we are made a spectacle unto the world, and to angels, and to men. We are fools for Christ's sake, but ye are wise in Christ; we are weak, but ye are strong; ye are honourable, but we are despised. Even unto this present hour we both hunger, and thirst, and are naked, and are buffeted, and have no certain dwellingplace; And labour, working with our own hands: being reviled, we bless; being persecuted, we suffer it: Being defamed, we intreat: we are made as the filth of the world, and are the offscouring of all things unto this day. I write not these things to shame you, but as my beloved sons I warn you. For though ye have ten thousand instructors in Christ, yet have ye not many fathers: for in Christ Jesus I have begotten you through the gospel.

\readingcitation{Gospel}{Matthew 20:20}
\lett{A}{t that time:} There came unto Jesus the mother of Zebedee's children with her sons, worshipping him, and desiring a certain thing of him. And he said unto her, What wilt thou? She saith unto him, Grant that these my two sons may sit, the one on thy right hand, and the other on the left, in thy kingdom. But Jesus answered and said, Ye know not what ye ask. Are ye able to drink of the cup that I shall drink of, and to be baptized with the baptism that I am baptized with? They say unto him, We are able. And he saith unto them, Ye shall drink indeed of my cup, and be baptized with the baptism that I am baptized with: but to sit on my right hand, and on my left, is not mine to give, but it shall be given to them for whom it is prepared of my Father.

\subby{Transfiguration of Our Lord Jesus Christ}
\fancyhead[RE,LO]{Transfiguration}
\begin{inhead}
    {Second Class Double\\
6 August}
\end{inhead}

\introit
\lett{T}{he} lightnings shone upon the ground: the earth was moved, and shook withal. \textit{Ps.} O how amiable are thy dwellings, thou Lord of hosts! My soul hath a desire and longing to enter into the courts of the Lord.
\collect
\lett{O}{God,} who in the glorious Transfiguration of thine only-begotten Son didst confirm the mysteries of the faith by the testimony of the fathers, and in the voice proceeding from the bright cloud didst wondrously foreshow the perfect adoption of thy sons; mercifully vouchsafe; that we, being made fellow-heirs of the same the King of glory, may likewise attain to be partakers of that glory. Through the same.
\readingcitation{Epistle}{2 Peter 1:16}
%The 1928 BCP reading begins three verses earlier than the Roman.
\lett{D}{early beloved:} We have not followed cunningly devised fables, when we made known unto you the power and coming of our Lord Jesus Christ, but were eyewitnesses of his majesty. For he received from God the Father honour and glory, when there came such a voice to him from the excellent glory, This is my beloved Son, in whom I am well pleased. And this voice which came from heaven we heard, when we were with him in the holy mount. We have also a more sure word of prophecy; whereunto ye do well that ye take heed, as unto a light that shineth in a dark place, until the day dawn, and the day star arise in your hearts.

\gradall{Thou art fairer than the children of men: full of grace are thy lips. ℣. My heart is inditing of a good matter: I speak of the things which I have made unto the King.}{Alleluia, alleluia. ℣. He is the brightness of the everlasting light, the unspotted mirror, and the image of his goodness. Alleluia.}

\readingcitation{Gospel}{Matthew 17:1}
\lett{A}{t that time:} Jesus taketh Peter, James, and John his brother, and bringeth them up into an high mountain apart, And was transfigured before them: and his face did shine as the sun, and his raiment was white as the light. And, behold, there appeared unto them Moses and Elias talking with him. Then answered Peter, and said unto Jesus, Lord, it is good for us to be here: if thou wilt, let us make here three tabernacles; one for thee, and one for Moses, and one for Elias. While he yet spake, behold, a bright cloud overshadowed them: and behold a voice out of the cloud, which said, This is my beloved Son, in whom I am well pleased; hear ye him. And when the disciples heard it, they fell on their face, and were sore afraid. And Jesus came and touched them, and said, Arise, and be not afraid. And when they had lifted up their eyes, they saw no man, save Jesus only. And as they came down from the mountain, Jesus charged them, saying, Tell the vision to no man, until the Son of man be risen again from the dead.

\offertory{Riches and plenteousness shall be in his house: and his righteousness endureth for ever, alleluia.}

\secret
\lett{S}{anctify,} we beseech thee, O Lord, by the glorious Transfiguration of thine only-begotten Son, the gifts which we offer unto thee: and by the splendour of his brightness, cleanse us from the defilements of our iniquities. Through the same.

\communion{Tell the vision which ye have seen to no man, until the Son of man be risen again from the dead.}

\postcommunion
\lett{G}{rant,} we beseech thee, almighty God: that we, who celebrate in solemn worship the most holy mysteries of the Transfiguration of thy Son, may attain thereto in pureness of mind and understanding. Through the same.

\subby{Assumption of the Blessed Virgin Mary}
\fancyhead[RE,LO]{Assumption of the B.V.M.}
\begin{inhead}
    {First Class Double with Common Octave\\
15 August}
\end{inhead}

\introit
\lett{A}{great} wonder appeared in heaven: a woman clothed with the sun, and the moon under her feet, and upon her head a crown of twelve stars. \textit{Ps.} O sing unto the Lord a new song: for he hath done marvellous things.
\collect
\lett{A}{lmighty} and everlasting God, who didst assume the immaculate Virgin Mary, mother of thy Son, in body and soul to the glory of heaven: grant, we beseech thee; that, being ever intent on things above, we may be worthy to be partakers of her glory. Through the same.
%Old Roman: {\lett{F}{orgive,} O Lord, we beseech thee, the offences of thy servants: that we, who of ourselves can do nothing which is acceptable unto thee; may be saved by the intercession of the Mother of thy Son our Lord. Who liveth.}
\readingcitation{Epistle}{Judith 13:17}
\lett{B}{lessed} be thou, O our God, which hast this day brought to nought the enemies of thy people. O daughter, blessed art thou of the most high God above all the women upon the earth; and blessed be the Lord God, which hath created the heavens and the earth, which hath directed thee to the cutting off of the head of the chief of our enemies. For this thy confidence shall not depart from the heart of men, which remember the power of God for ever. And God turn these things to thee for a perpetual praise, to visit thee in good things because thou hast not spared thy life for the affliction of our nation, but hast revenged our ruin, walking a straight way before our God.

\gradall{Hearken, O daughter, and consider, and incline thine ear, so shall the king have pleasure in thy beauty. ℣. All glorious the king's daughter entereth in, her clothing is of wrought gold.}{Alleluia, alleluia. ℣. Mary is taken up into heaven: the host of Angels rejoice. Alleluia.}

\readingcitation{Gospel}{Luke 1:41}

\lett{A}{t that time:} And it came to pass, that, when Elisabeth heard the salutation of Mary, the babe leaped in her womb; and Elisabeth was filled with the Holy Ghost: and she spake out with a loud voice, and said, Blessed art thou among women, and blessed is the fruit of thy womb. And whence is this to me, that the mother of my Lord should come to me? For, lo, as soon as the voice of thy salutation sounded in mine ears, the babe leaped in my womb for joy. And blessed is she that believed: for there shall be a performance of those things which were told her from the Lord. And Mary said, My soul doth magnify the Lord, and my spirit hath rejoiced in God my Saviour. For he hath regarded the low estate of his handmaiden: for, behold, from henceforth all generations shall call me blessed. For he that is mighty hath done to me great things; and holy is his name. And his mercy is on them that fear him from generation to generation.

\offertory{I will put enmity between thee and the Woman, and between thy seed and her Seed.}

\secret
\lett{L}{et} this oblation of our bounden duty ascend unto thee, O Lord, and, at the intercession of the most blessed Virgin Mary, assumed into heaven, may our hearts, enkindled with the fire of charity, continually yearn after thee. Through.

\communion{All generations shall call me blessed, for he that is mighty hath done to me great things.}
\postcommunion
\lett{G}{rant,} we beseech thee, O Lord: that we, who have received the sacraments of salvation; may, through the merits and intercession of the blessed Virgin Mary, assumed into heaven, be brought unto the glory of the resurrection. Through.

\begin{rubric}
    Within the Octave, Mass is said as on the Feast. And the \nth{2} Collect is of the Holy Ghost (p. \pageref{SPHolyGhost}) \& \nth{3} Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}

\subby{St. Bartholomew}
\fancyhead[RE,LO]{St. Bartholomew}
\begin{inhead}
    {Second Class Double\\
24 August}
\end{inhead}

\collect
\lett{A}{lmighty} and everlasting God, who hast given us on this day in the festival of blessed Bartholomew thine Apostle venerable and holy joy: grant unto thy Church, we beseech thee; both to love that which he believed, and to preach that which he taught. Through.

\readingcitation{Epistle}{1 Corinthians 12:27}
\lett{B}{rethren:} Ye are the body of Christ, and members in particular. And God hath set some in the church, first apostles, secondarily prophets, thirdly teachers, after that miracles, then gifts of healings, helps, governments, diversities of tongues. Are all apostles? are all prophets? are all teachers? are all workers of miracles? Have all the gifts of healing? do all speak with tongues? do all interpret? But covet earnestly the best gifts.


\readingcitation{Gospel}{Luke 6:12}
\lett{A}{t that time:} Jesus went out into a mountain to pray, and continued all night in prayer to God. And when it was day, he called unto him his disciples: and of them he chose twelve, whom also he named apostles; Simon, (whom he also named Peter,) and Andrew his brother, James and John, Philip and Bartholomew, Matthew and Thomas, James the son of Alphaeus, and Simon called Zelotes, And Judas the brother of James, and Judas Iscariot, which also was the traitor. And he came down with them, and stood in the plain, and the company of his disciples, and a great multitude of people out of all Judaea and Jerusalem, and from the sea coast of Tyre and Sidon, which came to hear him, and to be healed of their diseases; And they that were vexed with unclean spirits: and they were healed. And the whole multitude sought to touch him: for there went virtue out of him, and healed them all.

\subby{St. Matthew}
\fancyhead[RE,LO]{St. Matthew}
\begin{inhead}
    {Second Class Double\\
21 September}
\end{inhead}

%Considered for BCP:
%\collect
%\lett{M}{ay} we be assisted, O Lord, by the prayers of blessed Matthew, the Apostle and Evangelist: that those things, which of ourselves we cannot obtain, may be vouchsafed unto us by his intercession. Through.

\readingcitation{Epistle}{Ezekiel 1:10}
\lett{T}{he} likeness of the faces of the four animals: they four had the face of a man, and the face of a lion, on the right side: and they four had the face of an ox on the left side; they four also had the face of an eagle. Thus were their faces: and their wings were stretched upward; two wings of every one were joined one to another, and two covered their bodies. And they went every one straight forward: whither the spirit was to go, they went; and they turned not when they went. As for the likeness of the living creatures, their appearance was like burning coals of fire, and like the appearance of lamps: it went up and down among the living creatures; and the fire was bright, and out of the fire went forth lightning. And the living creatures ran and returned as the appearance of a flash of lightning.

\readingcitation{Gospel}{Matthew 9:9}
\lett{A}{t that time:} As Jesus passed forth from thence, he saw a man, named Matthew, sitting at the receipt of custom: and he saith unto him, Follow me. And he arose, and followed him. And it came to pass, as Jesus sat at meat in the house, behold, many publicans and sinners came and sat down with him and his disciples. And when the Pharisees saw it, they said unto his disciples, Why eateth your Master with publicans and sinners? But when Jesus heard that, he said unto them, They that be whole need not a physician, but they that are sick. But go ye and learn what that meaneth, I will have mercy, and not sacrifice: for I am not come to call the righteous, but sinners.

\subby{St. Michael}
\fancyhead[RE,LO]{St. Michael}
\begin{inhead}
    {First Class Double\\
29 September}
\end{inhead}

\readingcitation{Epistle}{Revelation 1:1}

\lett{I}{n those days:} God shewed the things which must shortly come to pass; and he sent and signified it by his angel unto his servant John: Who bare record of the word of God, and of the testimony of Jesus Christ, and of all things that he saw. Blessed is he that readeth, and they that hear the words of this prophecy, and keep those things which are written therein: for the time is at hand. John to the seven churches which are in Asia: Grace be unto you, and peace, from him which is, and which was, and which is to come; and from the seven Spirits which are before his throne; And from Jesus Christ, who is the faithful witness, and the first begotten of the dead, and the prince of the kings of the earth. Unto him that loved us, and washed us from our sins in his own blood.

\subby{St. Luke}
\fancyhead[RE,LO]{St. Luke}
\begin{inhead}
    {Second Class Double\\
18 October}
\end{inhead}

%Considered for BCP:
%\collect
%\lett{W}{e} beseech thee, O Lord: that thy holy Evangelist Luke may intercede for us: who for the honour of thy name continually bare in his own body the mortification of the Cross. Through.

\readingcitation{Epistle}{2 Corinthians 8:16}
\lett{B}{rethren:} I give thanks to God, which put the same earnest care into the heart of Titus for you. For indeed he accepted the exhortation; but being more forward, of his own accord he went unto you. And we have sent with him the brother, whose praise is in the gospel throughout all the churches; And not that only, but who was also chosen of the churches to travel with us with this grace, which is administered by us to the glory of the same Lord, and declaration of your ready mind: Avoiding this, that no man should blame us in this abundance which is administered by us: Providing for honest things, not only in the sight of the Lord, but also in the sight of men. And we have sent with them our brother, whom we have oftentimes proved diligent in many things, but now much more diligent, upon the great confidence which I have in you. Whether any do enquire of Titus, he is my partner and fellowhelper concerning you: or our brethren be enquired of, they are the messengers of the churches, and the glory of Christ. Wherefore shew ye to them, and before the churches, the proof of your love, and of our boasting on your behalf.

\readingcitation{Gospel}{Luke 10:1}
\lett{A}{t that time:} The Lord appointed other seventy also, and sent them two and two before his face into every city and place, whither he himself would come. Therefore said he unto them, The harvest truly is great, but the labourers are few: pray ye therefore the Lord of the harvest, that he would send forth labourers into his harvest. Go your ways: behold, I send you forth as lambs among wolves. Carry neither purse, nor scrip, nor shoes: and salute no man by the way. And into whatsoever house ye enter, first say, Peace be to this house. And if the son of peace be there, your peace shall rest upon it: if not, it shall turn to you again. And in the same house remain, eating and drinking such things as they give: for the labourer is worthy of his hire. Go not from house to house. And into whatsoever city ye enter, and they receive you, eat such things as are set before you: And heal the sick that are therein, and say unto them, The kingdom of God is come nigh unto you.

\subby{Sts. Simon and Jude}
\fancyhead[RE,LO]{Sts. Simon \& Jude}
\begin{inhead}
    {Second Class Double\\
28 October}
\end{inhead}

\collect
\lett{O}{God,} who through thy blessed Apostles Simon and Jude hast suffered us to come unto the knowledge of thy name: grant us by advancing in virtue to celebrate their everlasting glory, and by the celebration thereof to advance in virtue. Through.

\readingcitation{Epistle}{Ephesians 4:7}
\lett{B}{rethren:} Unto every one of us is given grace according to the measure of the gift of Christ. Wherefore he saith, When he ascended up on high, he led captivity captive, and gave gifts unto men. (Now that he ascended, what is it but that he also descended first into the lower parts of the earth? He that descended is the same also that ascended up far above all heavens, that he might fill all things.) And he gave some, apostles; and some, prophets; and some, evangelists; and some, pastors and teachers; For the perfecting of the saints, for the work of the ministry, for the edifying of the body of Christ: Till we all come in the unity of the faith, and of the knowledge of the Son of God, unto a perfect man, unto the measure of the stature of the fulness of Christ:

\readingcitation{Gospel}{John 15:17}
\lett{A}{t that time:} Jesus said unto his disciples: 17 These things I command you, that ye love one another. If the world hate you, ye know that it hated me before it hated you. If ye were of the world, the world would love his own: but because ye are not of the world, but I have chosen you out of the world, therefore the world hateth you. Remember the word that I said unto you, The servant is not greater than his lord. If they have persecuted me, they will also persecute you; if they have kept my saying, they will keep yours also. But all these things will they do unto you for my name's sake, because they know not him that sent me. If I had not come and spoken unto them, they had not had sin: but now they have no cloke for their sin. He that hateth me hateth my Father also. If I had not done among them the works which none other man did, they had not had sin: but now have they both seen and hated both me and my Father. But this cometh to pass, that the word might be fulfilled that is written in their law, They hated me without a cause.

\subby{All Hallows}
\fancyhead[RE,LO]{All Hallows}
\begin{inhead}
    {Second Class Double\\
1 November}
\end{inhead}

%Considered for BCP:
%\collect
%\lett{A}{lmighty} and everlasting God, who in one solemnity hast vouchsafed unto us to venerate the merits of all thy Saints: we beseech thee; that, at the intercession of so great a multitude, thou wouldest bestow upon us, who entreat thee, the abundance of thy mercy. Through.

%RE-CHECK:
\readingcitation{Epistle}{Revelation 7:2}
\lett{I}{n those days:} Behold, I, John, saw another angel ascending from the east, having the seal of the living God: and he cried with a loud voice to the four angels, to whom it was given to hurt the earth and the sea, Saying, Hurt not the earth, neither the sea, nor the trees, till we have sealed the servants of our God in their foreheads. And I heard the number of them which were sealed: and there were sealed an hundred and forty and four thousand of all the tribes of the children of Israel. Of the tribe of Juda were sealed twelve thousand. Of the tribe of Reuben were sealed twelve thousand. Of the tribe of Gad were sealed twelve thousand. Of the tribe of Aser were sealed twelve thousand. Of the tribe of Nepthalim were sealed twelve thousand. Of the tribe of Manasses were sealed twelve thousand. Of the tribe of Simeon were sealed twelve thousand. Of the tribe of Levi were sealed twelve thousand. Of the tribe of Issachar were sealed twelve thousand. Of the tribe of Zabulon were sealed twelve thousand. Of the tribe of Joseph were sealed twelve thousand. Of the tribe of Benjamin were sealed twelve thousand. After this I beheld, and, lo, a great multitude, which no man could number, of all nations, and kindreds, and people, and tongues, stood before the throne, and before the Lamb, clothed with white robes, and palms in their hands; And cried with a loud voice, saying, Salvation to our God which sitteth upon the throne, and unto the Lamb. And all the angels stood round about the throne, and about the elders and the four beasts, and fell before the throne on their faces, and worshipped God, Saying, Amen: Blessing, and glory, and wisdom, and thanksgiving, and honour, and power, and might, be unto our God for ever and ever. Amen.

\readingcitation{Gospel}{Matthew 5:1}
\lett{A}{t that time:} Jesus, seeing the multitudes, went up into a mountain: and when he was set, his disciples came unto him: And he opened his mouth, and taught them, saying, Blessed are the poor in spirit: for theirs is the kingdom of heaven. Blessed are they that mourn: for they shall be comforted. Blessed are the meek: for they shall inherit the earth. Blessed are they which do hunger and thirst after righteousness: for they shall be filled. Blessed are the merciful: for they shall obtain mercy. Blessed are the pure in heart: for they shall see God. Blessed are the peacemakers: for they shall be called the children of God. Blessed are they which are persecuted for righteousness' sake: for theirs is the kingdom of heaven. Blessed are ye, when men shall revile you, and persecute you, and shall say all manner of evil against you falsely, for my sake. Rejoice, and be exceeding glad: for great is your reward in heaven.


