\fancyhead[RO,LE]{\textit{Of Season}}
\fancyhead[RE,LO]{}

   \fancyhead[RE,LO]{Advent I}
   \subby{First Sunday of Advent}\label{AdventI}
\begin{inhead}
    {First Class Semidouble}
\end{inhead}

\collect
\lett{S}{tir} up, we beseech thee, O Lord, thy power and come: that by thy protection we may be found worthy to be set free from the dangers of our sins which beset us, and to be saved by thy deliverance. Who livest.
\begin{rubric}
    This Collect is to be repeated every day, with the other Collects in Advent, until Christmas Day.
\end{rubric}
\begin{rubric}
    \nth{2} Collect is \emph{Of St. Mary in Advent} (p. \pageref{SPMaryInAdvent}) \& \nth{3} \emph{Against the Persecutors of the Church} (p. \pageref{SPAgainst}) or \emph{For the Chief Bishop} (p. \pageref{SPChiefBishop}).
\end{rubric}
\readingcitation{Epistle}{Romans 13:8}
\lett{B}{rethren:} Know the time, that now it is high time to awake out of sleep: for now is our salvation nearer than when we believed. The night is far spent, the day is at hand: let us therefore cast off the works of darkness, and let us put on the armour of light. Let us walk honestly, as in the day; not in rioting and drunkenness, not in chambering and wantonness, not in strife and envying. But put ye on the Lord Jesus Christ.

\readingcitation{Gospel}{Luke 21:25}
\lett{A}{t that time:} Jesus said unto his disciples: There shall be signs in the sun, and in the moon, and in the stars; and upon the earth distress of nations, with perplexity; the sea and the waves roaring; Men's hearts failing them for fear, and for looking after those things which are coming on the earth: for the powers of heaven shall be shaken. And then shall they see the Son of man coming in a cloud with power and great glory. And when these things begin to come to pass, then look up, and lift up your heads; for your redemption draweth nigh. And he spake to them a parable; Behold the fig tree, and all the trees; When they now shoot forth, ye see and know of your own selves that summer is now nigh at hand. So likewise ye, when ye see these things come to pass, know ye that the kingdom of God is nigh at hand. Verily I say unto you, This generation shall not pass away, till all be fulfilled. Heaven and earth shall pass away: but my words shall not pass away.

\subby{Second Sunday of Advent}
\fancyhead[RE,LO]{Advent II}
\begin{inhead}
{Second Class Semidouble}
\end{inhead}

\collect
\lett{S}{tir} up our hearts, O Lord, to prepare the ways of thine only-begotten Son: that through his advent we may be worthy to serve thee with pure minds. Who liveth.
\readingcitation{Gospel}{Matthew 11:2}
\lett{A}{t that time:} When John had heard in the prison the works of Christ, he sent two of his disciples, And said unto him, Art thou he that should come, or do we look for another? Jesus answered and said unto them, Go and shew John again those things which ye do hear and see: The blind receive their sight, and the lame walk, the lepers are cleansed, and the deaf hear, the dead are raised up, and the poor have the gospel preached to them. And blessed is he, whosoever shall not be offended in me. And as they departed, Jesus began to say unto the multitudes concerning John, What went ye out into the wilderness to see? A reed shaken with the wind? But what went ye out for to see? A man clothed in soft raiment? behold, they that wear soft clothing are in kings' houses. But what went ye out for to see? A prophet? yea, I say unto you, and more than a prophet. For this is he, of whom it is written, Behold, I send my messenger before thy face, which shall prepare thy way before thee.

\subby{Third Sunday of Advent}
\fancyhead[RE,LO]{Advent III}
\begin{inhead}
{Second Class Semidouble}
\end{inhead}

\collect
\lett{I}{ncline} thine ear to our prayers, we beseech thee, O Lord: and by the grace of thy visitation enlighten the darkness of our minds. Who livest.
\begin{rubric}
    Additional Collects are as on \nth{1} Sunday of Advent (p. \pageref{AdventI}).
\end{rubric}
\readingcitation{Epistle}{Philippians 4:4}
\lett{B}{rethren:} Rejoice in the Lord alway: and again I say, Rejoice. Let your moderation be known unto all men. The Lord is at hand. Be careful for nothing; but in every thing by prayer and supplication with thanksgiving let your requests be made known unto God. And the peace of God, which passeth all understanding, shall keep your hearts and minds through Christ Jesus our Lord.

\readingcitation{Gospel}{John 1:19}
\lett{A}{t that time:} The Jews sent priests and Levites from Jerusalem to ask him, Who art thou? And he confessed, and denied not; but confessed, I am not the Christ. And they asked him, What then? Art thou Elias? And he saith, I am not. Art thou that prophet? And he answered, No. Then said they unto him, Who art thou? that we may give an answer to them that sent us. What sayest thou of thyself? He said, I am the voice of one crying in the wilderness, Make straight the way of the Lord, as said the prophet Esaias. And they which were sent were of the Pharisees. And they asked him, and said unto him, Why baptizest thou then, if thou be not that Christ, nor Elias, neither that prophet? John answered them, saying, I baptize with water: but there standeth one among you, whom ye know not; He it is, who coming after me is preferred before me, whose shoe's latchet I am not worthy to unloose. These things were done in Bethabara beyond Jordan, where John was baptizing.

\subby{Fourth Sunday of Advent}
\fancyhead[RE,LO]{Advent IV}
\begin{inhead}
{Second Class Semidouble}
\end{inhead}

\collect
\lett{S}{tir} up, we beseech thee, O Lord, thy power, and come: and with great might succour us; that whereas through our sins we are hindered, we may through the help of thy grace speedily obtain thy merciful pardon. Who livest.
\begin{rubric}
    Additional Collects are as on \nth{1} Sunday of Advent (p. \pageref{AdventI}).
\end{rubric}

\readingcitation{Epistle}{1 Corinthians 4:1}
\lett{B}{rethren:} Let a man so account of us, as of the ministers of Christ, and stewards of the mysteries of God. Moreover it is required in stewards, that a man be found faithful. But with me it is a very small thing that I should be judged of you, or of man's judgment: yea, I judge not mine own self. For I know nothing by myself; yet am I not hereby justified: but he that judgeth me is the Lord. Therefore judge nothing before the time, until the Lord come, who both will bring to light the hidden things of darkness, and will make manifest the counsels of the hearts: and then shall every man have praise of God.

\readingcitation{Gospel}{Luke 3:1}
\lett{I}{n} the fifteenth year of the reign of Tiberius Caesar, Pontius Pilate being governor of Judaea, and Herod being tetrarch of Galilee, and his brother Philip tetrarch of Ituraea and of the region of Trachonitis, and Lysanias the tetrarch of Abilene, Annas and Caiaphas being the high priests, the word of God came unto John the son of Zacharias in the wilderness. And he came into all the country about Jordan, preaching the baptism of repentance for the remission of sins; As it is written in the book of the words of Esaias the prophet, saying, The voice of one crying in the wilderness, Prepare ye the way of the Lord, make his paths straight. Every valley shall be filled, and every mountain and hill shall be brought low; and the crooked shall be made straight, and the rough ways shall be made smooth; And all flesh shall see the salvation of God.

\subby{Nativity of Our Lord}
\fancyhead[RE,LO]{Christmas}
\begin{inhead}
{First Class Double, Third Class Octave}
\end{inhead}
\centerline{\small{(The First Mass at Night)}}

\introit
\lett{T}{he} Lord hath said unto me: Thou art my Son, this day have I begotten thee. \textit{Ps.} Why do the heathen so furiously rage together: and why do the people imagine a vain thing?
\collect
\lett{O}{God,} who hast made this most sacred night to shine with the brightness of the true light: grant, we beseech thee; that we, who have known the mysteries of his light on earth, may also attain to the fruition of his joys in heaven. Who liveth.

\subby{St. Stephen}
\fancyhead[RE,LO]{St. Stephen}
\begin{inhead}
{26 December}\par
{Second Class Double, Simple Octave}
\end{inhead}

\collect
%Roman Collect without intercession of the saint
\lett{G}{rant} us, we beseech thee, O Lord, so to imitate that which we revere: that we may learn to love even our enemies; for as much as we celebrate the birthday of him who knew how to pray even for his persecutors to our Lord Jesus Christ thy Son. Who liveth.

\readingcitation{Epistle}{Acts 6:8}
\lett{I}{n those days:} Stephen, full of faith and power, did great wonders and miracles among the people. Then there arose certain of the synagogue, which is called the synagogue of the Libertines, and Cyrenians, and Alexandrians, and of them of Cilicia and of Asia, disputing with Stephen. And they were not able to resist the wisdom and the spirit by which he spake. When they heard these things, they were cut to the heart, and they gnashed on him with their teeth. But Stephen, being full of the Holy Ghost, looked up stedfastly into heaven, and saw the glory of God, and Jesus standing on the right hand of God, And said, Behold, I see the heavens opened, and the Son of man standing on the right hand of God. Then they cried out with a loud voice, and stopped their ears, and ran upon him with one accord, And cast him out of the city, and stoned him: and the witnesses laid down their clothes at a young man's feet, whose name was Saul. And they stoned Stephen, calling upon God, and saying, Lord Jesus, receive my spirit. And he kneeled down, and cried with a loud voice, Lord, lay not this sin to their charge. And when he had said this, he fell asleep.

\subby{St. John}
\fancyhead[RE,LO]{St. John}
\begin{inhead}
{27 December}\par
{Second Class Double, Simple Octave}
\end{inhead}

\collect
%Roman collect without intercession of the Saints
\lett{O}{Lord,} graciously illumine thy Church: that it, being enlightened by the doctrine of blessed John thy Apostle and Evangelist, may attain unto everlasting gifts. Through.
\begin{rubric}
    Commemoration is made of the Nativity from the Third Mass (p. \pageref{NativityMassIII}).
\end{rubric}

\readingcitation{Epistle}{Ecclesiasticus 15:1}
\lett{H}{e} that feareth the Lord will do good, and he that hath the knowledge of the law shall obtain her. And as a mother shall she meet him, and receive him as a wife married of a virgin. With the bread of understanding shall she feed him, and give him the water of wisdom to drink. He shall be stayed upon her, and shall not be moved; and shall rely upon her, and shall not be confounded. She shall exalt him above his neighbours, and in the midst of the congregation shall she open his mouth. He shall find joy and a crown of gladness, and she shall cause him to inherit an everlasting name.

\readingcitation{Gospel}{John 21:19}
\lett{A}{t that time:} Jesus said unto Peter: Follow me. Then Peter, turning about, seeth the disciple whom Jesus loved following; which also leaned on his breast at supper, and said, Lord, which is he that betrayeth thee? Peter seeing him saith to Jesus, Lord, and what shall this man do? Jesus saith unto him, If I will that he tarry till I come, what is that to thee? follow thou me. Then went this saying abroad among the brethren, that that disciple should not die: yet Jesus said not unto him, He shall not die; but, If I will that he tarry till I come, what is that to thee? This is the disciple which testifieth of these things, and wrote these things: and we know that his testimony is true.


\subby{Holy Innocents}
\fancyhead[RE,LO]{Holy Innocents}
\begin{inhead}
{28 December}\par
{Second Class Double, Simple Octave}
\end{inhead}

\collect
\lett{O}{God,} whose praise the Innocents, thy Martyrs did this day, not by speaking, but by dying, confess: mortify all evil vices in us; that, as our tongues do confess thy faith, so our lives may by their actions set forth the same. Through.
\begin{rubric}
    Commemoration is made of the Nativity from the Third Mass (p. \pageref{NativityMassIII}).
\end{rubric}

\subby{Sunday within the Nativity Octave}
\fancyhead[RE,LO]{Christmas I}
\begin{inhead}
{Semidouble}
\end{inhead}

\collect
\lett{A}{lmighty} and everlasting God, direct our actions according to thy good pleasure: that in the name of thy well-beloved Son we may be made worthy to abound in good works. Who liveth and reigneth with thee.

\readingcitation{Gospel}{Luke 2:33}
\lett{A}{t that time:} Joseph and Mary, the Mother of Jesus, marvelled at those things which were spoken of him. And Simeon blessed them, and said unto Mary his mother, Behold, this child is set for the fall and rising again of many in Israel; and for a sign which shall be spoken against; (Yea, a sword shall pierce through thy own soul also,) that the thoughts of many hearts may be revealed. And there was one Anna, a prophetess, the daughter of Phanuel, of the tribe of Aser: she was of a great age, and had lived with an husband seven years from her virginity; And she was a widow of about fourscore and four years, which departed not from the temple, but served God with fastings and prayers night and day. And she coming in that instant gave thanks likewise unto the Lord, and spake of him to all them that looked for redemption in Jerusalem. And when they had performed all things according to the law of the Lord, they returned into Galilee, to their own city Nazareth. And the child grew, and waxed strong in spirit, filled with wisdom: and the grace of God was upon him.

\subby{Circumcision of Our Lord}
\fancyhead[RE,LO]{Circumcision of Our Lord}
\begin{inhead}
{1 January}\par
{Second Class Double}
\end{inhead}

\collect
\lett{O}{God,} who by the virgin child-bearing of blessed Mary hast bestowed upon mankind the rewards of eternal salvation: grant, we beseech thee; that we may perceive her intercession for us, through whom we have been counted worthy to receive the author of life, Jesus Christ thy Son, our Lord. Who liveth.

%CHECK:
\readingcitation{Epistle}{Titus 2:11}
\lett{B}{eloved:} The grace of God that bringeth salvation hath appeared to all men, Teaching us that, denying ungodliness and worldly lusts, we should live soberly, righteously, and godly, in this present world; Looking for that blessed hope, and the glorious appearing of the great God and our Saviour Jesus Christ; Who gave himself for us, that he might redeem us from all iniquity, and purify unto himself a peculiar people, zealous of good works. These things speak, and exhort, and rebuke with all authority. Let no man despise thee.

\readingcitation{Gospel}{Luke 2:15}
\lett{A}{t that time:} When eight days were accomplished for the circumcising of the child, his name was called \divineName{Jesus}, which was so named of the angel before he was conceived in the womb.


\subby{Epiphany of Our Lord}\label{EpiphanyMass}
\fancyhead[RE,LO]{Epiphany}
\begin{inhead}
{First Class Double, Second Class Octave}
\end{inhead}

\collect
\lett{O}{God,} who on this day by the leading of a star didst manifest thy only-begotten Son to the Gentiles: mercifully grant; that we, which know thee now by faith, may be brought even to the contemplation of the beauty of thy majesty. Through the same.
\begin{rubric}
    This Collect to be said daily throughout the Octave.
\end{rubric}

\readingcitation{Epistle}{Isaiah 60:1}
\lett{A}{rise,} shine; for thy light is come, and the glory of the \divineName{Lord} is risen upon thee. For, behold, the darkness shall cover the earth, and gross darkness the people: but the \divineName{Lord} shall arise upon thee, and his glory shall be seen upon thee. And the Gentiles shall come to thy light, and kings to the brightness of thy rising. Lift up thine eyes round about, and see: all they gather themselves together, they come to thee: thy sons shall come from far, and thy daughters shall be nursed at thy side. Then thou shalt see, and flow together, and thine heart shall fear, and be enlarged; because the abundance of the sea shall be converted unto thee, the forces of the Gentiles shall come unto thee. The multitude of camels shall cover thee, the dromedaries of Midian and Ephah; all they from Sheba shall come: they shall bring gold and incense; and they shall shew forth the praises of the \divineName{Lord}.


\subby{Sunday within the Epiphany Octave}
\fancyhead[RE,LO]{Epiphany I}
\begin{inhead}
{Semidouble}
\end{inhead}

%CHECK:
\readingcitation{Epistle}{Romans 12:1}
\lett{B}{rethren:} I beseech you therefore, brethren, by the mercies of God, that ye present your bodies a living sacrifice, holy, acceptable unto God, which is your reasonable service. And be not conformed to this world: but be ye transformed by the renewing of your mind, that ye may prove what is that good, and acceptable, and perfect, will of God. For I say, through the grace given unto me, to every man that is among you, not to think of himself more highly than he ought to think; but to think soberly, according as God hath dealt to every man the measure of faith. For as we have many members in one body, and all members have not the same office: So we, being many, are one body in Christ, and every one members one of another: in Christ Jesus our Lord.

\readingcitation{Gospel}{Luke 2:41}
\lett{W}{hen} Jesus was twelve years old, they went up to Jerusalem after the custom of the feast. And when they had fulfilled the days, as they returned, the child Jesus tarried behind in Jerusalem; and Joseph and his mother knew not of it. But they, supposing him to have been in the company, went a day's journey; and they sought him among their kinsfolk and acquaintance. And when they found him not, they turned back again to Jerusalem, seeking him. And it came to pass, that after three days they found him in the temple, sitting in the midst of the doctors, both hearing them, and asking them questions. And all that heard him were astonished at his understanding and answers. And when they saw him, they were amazed: and his mother said unto him, Son, why hast thou thus dealt with us? behold, thy father and I have sought thee sorrowing. And he said unto them, How is it that ye sought me? wist ye not that I must be about my Father's business? And they understood not the saying which he spake unto them. And he went down with them, and came to Nazareth, and was subject unto them: but his mother kept all these sayings in her heart. And Jesus increased in wisdom and stature, and in favour with God and man.


\subby{Second Sunday after Epiphany}\label{EpiphanyII}
\fancyhead[RE,LO]{Epiphany II}
\begin{inhead}
{Semidouble}
\end{inhead}

\readingcitation{Gospel}{John 2:1}
\lett{A}{t that time:} There was a marriage in Cana of Galilee; and the mother of Jesus was there: And both Jesus was called, and his disciples, to the marriage. And when they wanted wine, the mother of Jesus saith unto him, They have no wine. Jesus saith unto her, Woman, what have I to do with thee? mine hour is not yet come. His mother saith unto the servants, Whatsoever he saith unto you, do it. And there were set there six waterpots of stone, after the manner of the purifying of the Jews, containing two or three firkins apiece. Jesus saith unto them, Fill the waterpots with water. And they filled them up to the brim. And he saith unto them, Draw out now, and bear unto the governor of the feast. And they bare it. When the ruler of the feast had tasted the water that was made wine, and knew not whence it was: (but the servants which drew the water knew;) the governor of the feast called the bridegroom, And saith unto him, Every man at the beginning doth set forth good wine; and when men have well drunk, then that which is worse: but thou hast kept the good wine until now. This beginning of miracles did Jesus in Cana of Galilee, and manifested forth his glory; and his disciples believed on him.


\subby{Fourth Sunday after Epiphany}
\fancyhead[RE,LO]{Epiphany IV}
\begin{inhead}
{Semidouble}
\end{inhead}

%CHECK:
\collect
%Is the English collect sufficiently similar?
\lett{O}{God,} who knowest us to be set in the midst of so many and great dangers, that by reason of the frailty of our nature we cannot always stand upright; Grant to us such strength and protection, as may support us in all dangers, and carry us through all temptations. Through.
\begin{rubric}
    If it be after 2 February, \nth{2} Collect is \emph{Of the Saints} (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.\par
    Otherwise, additional Collects are as in the Second Sunday after Epiphany (p. \pageref{EpiphanyII}).
\end{rubric}

\readingcitation{Epistle}{Romans 13:8}
\lett{B}{rethren:} Owe no man any thing, but to love one another: for he that loveth another hath fulfilled the law. For this, Thou shalt not commit adultery, Thou shalt not kill, Thou shalt not steal, Thou shalt not bear false witness, Thou shalt not covet; and if there be any other commandment, it is briefly comprehended in this saying, namely, Thou shalt love thy neighbour as thyself Love worketh no ill to his neighbour: therefore love is the fulfilling of the law.

\readingcitation{Gospel}{Matthew 8:23}
\lett{A}{t that time:} When Jesus was entered into a ship, his disciples followed him. And, behold, there arose a great tempest in the sea, insomuch that the ship was covered with the waves: but he was asleep. And his disciples came to him, and awoke him, saying, Lord, save us: we perish. And he saith unto them, Why are ye fearful, O ye of little faith? Then he arose, and rebuked the winds and the sea; and there was a great calm. But the men marvelled, saying, What manner of man is this, that even the winds and the sea obey him!


\subby{Fifth Sunday after Epiphany}\label{EpiphanyV}
\fancyhead[RE,LO]{Epiphany V}
\begin{inhead}
{Semidouble}
\end{inhead}
%CHECK:
\collect
\lett{O}{Lord,} we beseech thee to keep thy Church and household continually in thy true religion; that they who do lean only upon the hope of thy heavenly grace may evermore be defended by thy mighty power. Through.
\begin{rubric}
    \nth{2} Collect is \emph{Of the Saints} (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}
%CHECK:
\readingcitation{Epistle}{Colossians 3:12}
\lett{B}{rethren:} Put on therefore, as the elect of God, holy and beloved, bowels of mercies, kindness, humbleness of mind, meekness, longsuffering; Forbearing one another, and forgiving one another, if any man have a quarrel against any: even as Christ forgave you, so also do ye. And above all these things put on charity, which is the bond of perfectness. And let the peace of God rule in your hearts, to the which also ye are called in one body; and be ye thankful. Let the word of Christ dwell in you richly in all wisdom; teaching and admonishing one another in psalms and hymns and spiritual songs, singing with grace in your hearts to the Lord. And whatsoever ye do in word or deed, do all in the name of the Lord Jesus, giving thanks to God and the Father by %
Our Lord Jesus Christ.


\subby{Sixth Sunday after Epiphany}
\fancyhead[RE,LO]{Epiphany VI}
\begin{inhead}
{Semidouble}
\end{inhead}

\collect
\lett{G}{rant,} we beseech thee, almighty God: that we, ever meditating those things that be reasonable and pleasing unto thee, may both in word and deed fulfil the same. Through.
\begin{rubric}
    Additional Collects are as in the Fifth Sunday after Epiphany (p. \pageref{EpiphanyV}).
\end{rubric}

%Double-check introduction.
\readingcitation{Epistle}{1 Thessalonians 1:2}
\lett{B}{rethren:} We give thanks to God always for you all, making mention of you in our prayers; Remembering without ceasing your work of faith, and labour of love, and patience of hope in our Lord Jesus Christ, in the sight of God and our Father; Knowing, brethren beloved, your election of God. For our gospel came not unto you in word only, but also in power, and in the Holy Ghost, and in much assurance; as ye know what manner of men we were among you for your sake. And ye became followers of us, and of the Lord, having received the word in much affliction, with joy of the Holy Ghost: So that ye were ensamples to all that believe in Macedonia and Achaia. For from you sounded out the word of the Lord not only in Macedonia and Achaia, but also in every place your faith to God-ward is spread abroad; so that we need not to speak any thing. For they themselves shew of us what manner of entering in we had unto you, and how ye turned to God from idols to serve the living and true God; And to wait for his Son from heaven, whom he raised from the dead, even Jesus, which delivered us from the wrath to come.

\readingcitation{Gospel}{Matthew 13:31}
\lett{A}{t that time:} Jesus spake this parable unto the crowds: The kingdom of heaven is like to a grain of mustard seed, which a man took, and sowed in his field: Which indeed is the least of all seeds: but when it is grown, it is the greatest among herbs, and becometh a tree, so that the birds of the air come and lodge in the branches thereof. Another parable spake he unto them; The kingdom of heaven is like unto leaven, which a woman took, and hid in three measures of meal, till the whole was leavened. All these things spake Jesus unto the multitude in parables; and without a parable spake he not unto them: That it might be fulfilled which was spoken by the prophet, saying, I will open my mouth in parables; I will utter things which have been kept secret from the foundation of the world.


\subby{Septuagesima Sunday}
\fancyhead[RE,LO]{Septuagesima}
\begin{inhead}
{Second Class Semidouble}
\end{inhead}

%CHECK:
\collect
\lett{O}{Lord,} we beseech thee favourably to hear the prayers of thy people; that we, who are justly punished for our offences, may be mercifully delivered by thy goodness, for the glory of thy Name; through Jesus Christ our Saviour, who liveth and reigneth with thee and the Holy Ghost ever, one God, world without end. \textit{Amen.}
\begin{rubric}
    Additional Collects are as in the Fifth Sunday after Epiphany (p. \pageref{EpiphanyV}).
\end{rubric}

\readingcitation{Epistle}{1 Corinthians 9:24}
\lett{B}{rethren:} Know ye not that they which run in a race run all, but one receiveth the prize? So run, that ye may obtain. And every man that striveth for the mastery is temperate in all things. Now they do it to obtain a corruptible crown; but we an incorruptible. I therefore so run, not as uncertainly; so fight I, not as one that beateth the air: But I keep under my body, and bring it into subjection: lest that by any means, when I have preached to others, I myself should be a castaway. Moreover, brethren, I would not that ye should be ignorant, how that all our fathers were under the cloud, and all passed through the sea; And were all baptized unto Moses in the cloud and in the sea; And did all eat the same spiritual meat; And did all drink the same spiritual drink: for they drank of that spiritual Rock that followed them: and that Rock was Christ. But with many of them God was not well pleased.


\subby{Sexagesima Sunday}
\fancyhead[RE,LO]{Sexagesima}
\begin{inhead}
{Second Class Semidouble}
\end{inhead}

%Collect used in BCP:
%\collect
%\lett{O}{God,} who seest that we put not our trust in any thing that we do: mercifully grant; that by the protection of the Doctor of the Gentiles we may be defended against all adversity. Through.
\begin{rubric}
    Additional Collects are as in the Fifth Sunday after Epiphany (p. \pageref{EpiphanyV}).
\end{rubric}

\readingcitation{Epistle}{2 Corinthians 11:19}
\lett{B}{rethren:} Ye suffer fools gladly, seeing ye yourselves are wise. For ye suffer, if a man bring you into bondage, if a man devour you, if a man take of you, if a man exalt himself, if a man smite you on the face. I speak as concerning reproach, as though we had been weak. Howbeit whereinsoever any is bold, (I speak foolishly,) I am bold also. Are they Hebrews? so am I. Are they Israelites? so am I. Are they the seed of Abraham? so am I. Are they ministers of Christ? (I speak as a fool) I am more; in labours more abundant, in stripes above measure, in prisons more frequent, in deaths oft. Of the Jews five times received I forty stripes save one. Thrice was I beaten with rods, once was I stoned, thrice I suffered shipwreck, a night and a day I have been in the deep; In journeyings often, in perils of waters, in perils of robbers, in perils by mine own countrymen, in perils by the heathen, in perils in the city, in perils in the wilderness, in perils in the sea, in perils among false brethren; In weariness and painfulness, in watchings often, in hunger and thirst, in fastings often, in cold and nakedness. Beside those things that are without, that which cometh upon me daily, the care of all the churches. Who is weak, and I am not weak? who is offended, and I burn not? If I must needs glory, I will glory of the things which concern mine infirmities. The God and Father of our Lord Jesus Christ, which is blessed for evermore, knoweth that I lie not. In Damascus the governor under Aretas the king kept the city of the Damascenes with a garrison, desirous to apprehend me: And through a window in a basket was I let down by the wall, and escaped his hands. It is not expedient for me doubtless to glory. I will come to visions and revelations of the Lord. I knew a man in Christ above fourteen years ago, (whether in the body, I cannot tell; or whether out of the body, I cannot tell: God knoweth;) such an one caught up to the third heaven. And I knew such a man, (whether in the body, or out of the body, I cannot tell: God knoweth;) How that he was caught up into paradise, and heard unspeakable words, which it is not lawful for a man to utter. Of such an one will I glory: yet of myself I will not glory, but in mine infirmities. For though I would desire to glory, I shall not be a fool; for I will say the truth: but now I forbear, lest any man should think of me above that which he seeth me to be, or that he heareth of me. And lest I should be exalted above measure through the abundance of the revelations, there was given to me a thorn in the flesh, the messenger of Satan to buffet me, lest I should be exalted above measure. For this thing I besought the Lord thrice, that it might depart from me. And he said unto me, My grace is sufficient for thee: for my strength is made perfect in weakness. Most gladly therefore will I rather glory in my infirmities, that the power of Christ may rest upon me.


\subby{Quinquagesima Sunday}
\fancyhead[RE,LO]{Quinquagesima}
\begin{inhead}
{Second Class Semidouble}
\end{inhead}

\collect
\lett{W}{e} beseech thee, O Lord, graciously hear our prayers: that we, being loosed from the bonds of our sins, may be defended against all adversity. Through.
\begin{rubric}
    Additional Collects are as in the Fifth Sunday after Epiphany (p. \pageref{EpiphanyV}).
\end{rubric}


\subby{Ash Wednesday}
\fancyhead[RE,LO]{Ash Wednesday}
\begin{inhead}
{First Class Feria}
\end{inhead}

\collect
\lett{G}{rant,} O Lord, to thy faithful people: that they may both enter upon the venerable solemnity of these fasts with fitting devotion, and perform the same in godly quietness. Through.
\begin{rubric}
    \nth{2} Collect is of the Saints (p. \pageref{SPSaints}) \& \nth{3} for the Living and Departed (p. \pageref{SPLivingDeparted}).
\end{rubric}

\readingcitation{Epistle}{Joel 2:12}
\lett{T}{hus} saith the Lord: Turn ye even to me with all your heart, and with fasting, and with weeping, and with mourning: And rend your heart, and not your garments, and turn unto the \divineName{Lord} your God: for he is gracious and merciful, slow to anger, and of great kindness, and repenteth him of the evil. Who knoweth if he will return and repent, and leave a blessing behind him; even a meat offering and a drink offering unto the \divineName{Lord} your God? Blow the trumpet in Zion, sanctify a fast, call a solemn assembly: Gather the people, sanctify the congregation, assemble the elders, gather the children, and those that suck the breasts: let the bridegroom go forth of his chamber, and the bride out of her closet. Let the priests, the ministers of the \divineName{Lord}, weep between the porch and the altar, and let them say, Spare thy people, O \divineName{Lord}, and give not thine heritage to reproach, that the heathen should rule over them: wherefore should they say among the people, Where is their God? Then will the \divineName{Lord} be jealous for his land, and pity his people. Yea, the \divineName{Lord} will answer and say unto his people, Behold, I will send you corn, and wine, and oil, and ye shall be satisfied therewith: and I will no more make you a reproach among the heathen.


\subby{First Sunday of Lent}
\fancyhead[RE,LO]{Lent I}
\begin{inhead}
{First Class Semidouble}
\end{inhead}

\collect
\lett{O}{God,} who year by year dost purify the Church by this Lenten observance: grant unto thy family; that those things which they strive to obtain from thee by abstinence, they may by good works fulfil. Through.
\begin{rubric}
    Additional Collects are as on Ash Wednesday (p. \pageref{AshWednesdayMass}).
\end{rubric}


\subby{Second Sunday of Lent}
\fancyhead[RE,LO]{Lent II}
\begin{inhead}
{First Class Semidouble}
\end{inhead}

%CHECK:
\readingcitation{Epistle}{1 Thessalonians 4:1}
\lett{B}{rethren:} we beseech you, and exhort you by the Lord Jesus, that as ye have received of us how ye ought to walk and to please God, so ye would abound more and more. For ye know what commandments we gave you by the Lord Jesus. For this is the will of God, even your sanctification, that ye should abstain from fornication: That every one of you should know how to possess his vessel in sanctification and honour; Not in the lust of concupiscence, even as the Gentiles which know not God That no man go beyond and defraud his brother in any matter: because that the Lord is the avenger of all such, as we also have forewarned you and testified. For God hath not called us unto uncleanness, but unto holiness. He therefore that despiseth, despiseth not man, but God, who hath also given unto us his holy Spirit.

\readingcitation{Gospel}{Matthew 17:1}
\lett{A}{t that time:} Jesus taketh Peter, James, and John his brother, and bringeth them up into an high mountain apart, And was transfigured before them: and his face did shine as the sun, and his raiment was white as the light. And, behold, there appeared unto them Moses and Elias talking with him. Then answered Peter, and said unto Jesus, Lord, it is good for us to be here: if thou wilt, let us make here three tabernacles; one for thee, and one for Moses, and one for Elias. While he yet spake, behold, a bright cloud overshadowed them: and behold a voice out of the cloud, which said, This is my beloved Son, in whom I am well pleased; hear ye him. And when the disciples heard it, they fell on their face, and were sore afraid. And Jesus came and touched them, and said, Arise, and be not afraid. And when they had lifted up their eyes, they saw no man, save Jesus only. And as they came down from the mountain, Jesus charged them, saying, Tell the vision to no man, until the Son of man be risen again from the dead.

\subby{Third Sunday of Lent}
\fancyhead[RE,LO]{Lent III}
\begin{inhead}
{First Class Semidouble}
\end{inhead}

%CHECK:
\collect
\lett{W}{e} beseech thee, Almighty God, look upon the hearty desires of thy humble servants, and stretch forth the right hand of thy Majesty, to be our defence. Through.
\begin{rubric}
    Additional Collects are as on Ash Wednesday (p. \pageref{AshWednesdayMass}).
\end{rubric}

\readingcitation{Epistle}{Ephesians 5:1}
\lett{B}{rethren:} Be ye therefore followers of God, as dear children; And walk in love, as Christ also hath loved us, and hath given himself for us an offering and a sacrifice to God for a sweetsmelling savour. But fornication, and all uncleanness, or covetousness, let it not be once named among you, as becometh saints; Neither filthiness, nor foolish talking, nor jesting, which are not convenient: but rather giving of thanks. For this ye know, that no whoremonger, nor unclean person, nor covetous man, who is an idolater, hath any inheritance in the kingdom of Christ and of God. Let no man deceive you with vain words: for because of these things cometh the wrath of God upon the children of disobedience. Be not ye therefore partakers with them. For ye were sometimes darkness, but now are ye light in the Lord: walk as children of light: (For the fruit of the Spirit is in all goodness and righteousness and truth;).


\subby{Fourth Sunday of Lent}
\fancyhead[RE,LO]{Lent IV}
\begin{inhead}
{First Class Semidouble}
\end{inhead}

%CHECK:
\readingcitation{Epistle}{Galatians 4:21}
%Anglican Addition:
\lett{B}{rethren:} tell me, ye that desire to be under the law, do ye not hear the law? For it is written, that Abraham had two sons, the one by a bondmaid, the other by a freewoman. But he who was of the bondwoman was born after the flesh; but he of the freewoman was by promise. Which things are an allegory: for these are the two covenants; the one from the mount Sinai, which gendereth to bondage, which is Agar. For this Agar is mount Sinai in Arabia, and answereth to Jerusalem which now is, and is in bondage with her children. But Jerusalem which is above is free, which is the mother of us all. For it is written, Rejoice, thou barren that bearest not; break forth and cry, thou that travailest not: for the desolate hath many more children than she which hath an husband. Now we, brethren, as Isaac was, are the children of promise. But as then he that was born after the flesh persecuted him that was born after the Spirit, even so it is now. Nevertheless what saith the scripture? Cast out the bondwoman and her son: for the son of the bondwoman shall not be heir with the son of the freewoman. So then, brethren, we are not children of the bondwoman, but of the free.

\readingcitation{Gospel}{John 6:1}
\lett{A}{t that time:} Jesus went over the sea of Galilee, which is the sea of Tiberias. And a great multitude followed him, because they saw his miracles which he did on them that were diseased. And Jesus went up into a mountain, and there he sat with his disciples. And the passover, a feast of the Jews, was nigh. When Jesus then lifted up his eyes, and saw a great company come unto him, he saith unto Philip, Whence shall we buy bread, that these may eat? And this he said to prove him: for he himself knew what he would do. Philip answered him, Two hundred pennyworth of bread is not sufficient for them, that every one of them may take a little. One of his disciples, Andrew, Simon Peter's brother, saith unto him, There is a lad here, which hath five barley loaves, and two small fishes: but what are they among so many? And Jesus said, Make the men sit down. Now there was much grass in the place. So the men sat down, in number about five thousand. And Jesus took the loaves; and when he had given thanks, he distributed to the disciples, and the disciples to them that were set down; and likewise of the fishes as much as they would. When they were filled, he said unto his disciples, Gather up the fragments that remain, that nothing be lost. Therefore they gathered them together, and filled twelve baskets with the fragments of the five barley loaves, which remained over and above unto them that had eaten. Then those men, when they had seen the miracle that Jesus did, said, This is of a truth that prophet that should come into the world. When Jesus therefore perceived that they would come and take him by force, to make him a king, he departed again into a mountain himself alone.

\subby{Passion Sunday}\label{PassionSunday}
\fancyhead[RE,LO]{Passion Sunday}
\begin{inhead}
{First Class Semidouble}
\end{inhead}

%CHECK:
\readingcitation{Epistle}{Hebrews 9:11}
\lett{B}{rethren:} Christ being come an high priest of good things to come, by a greater and more perfect tabernacle, not made with hands, that is to say, not of this building; Neither by the blood of goats and calves, but by his own blood he entered in once into the holy place, having obtained eternal redemption for us. For if the blood of bulls and of goats, and the ashes of an heifer sprinkling the unclean, sanctifieth to the purifying of the flesh: How much more shall the blood of Christ, who through the eternal Spirit offered himself without spot to God, purge your conscience from dead works to serve the living God? And for this cause he is the mediator of the new testament, that by means of death, for the redemption of the transgressions that were under the first testament, they which are called might receive the promise of eternal inheritance, in Christ Jesus our Lord.


\subby{Palm Sunday}\label{PalmSunday}
\fancyhead[RE,LO]{Palm Sunday}
\begin{inhead}
{First Class Semidouble}
\end{inhead}

\collect
\lett{A}{lmighty} and everlasting God, who didst cause our Saviour to take upon him our flesh, and to suffer on the cross, that mankind might follow the example of his humility: mercifully grant; that we may both keep in mind the lessons of his patience, and be made partakers of his resurrection. Through the same.
\begin{rubric}
    This Collect only is said, even if a Commemoration has been made in the Office.
\end{rubric}
\begin{rubric}
    This Collect is to be said every day, after the Collect appointed for the day, until Good Friday.
\end{rubric}


\subby{Easter Day}
\fancyhead[RE,LO]{Easter Day}
\begin{inhead}
    {First Class Double, First Class Octave}
\end{inhead}

\collect
\lett{O}{God,} who through thine only-begotten Son hast this day overcome death, and opened unto us the gate of eternity: prosper by thine assistance our desires, which by thy preventing grace thou dost put into our minds. Through the same.
\begin{rubric}
    This Collect is to be said daily throughout the Easter Octave.
\end{rubric}

\readingcitation{Epistle}{1 Corinthians 5:6}
\lett{B}{rethren:} Purge out therefore the old leaven, that ye may be a new lump, as ye are unleavened. For even Christ our passover is sacrificed for us: Therefore let us keep the feast, not with old leaven, neither with the leaven of malice and wickedness; but with the unleavened bread of sincerity and truth.


\readingcitation{Gospel}{Mark 16:1}
\lett{A}{t that time:} Mary Magdalene, and Mary the mother of James, and Salome, bought sweet spices, that they might come and anoint him. And very early in the morning the first day of the week, they came unto the sepulchre at the rising of the sun. And they said among themselves, Who shall roll us away the stone from the door of the sepulchre? And when they looked, they saw that the stone was rolled away: for it was very great. And entering into the sepulchre, they saw a young man sitting on the right side, clothed in a long white garment; and they were affrighted. And he saith unto them, Be not affrighted: Ye seek Jesus of Nazareth, which was crucified: he is risen; he is not here: behold the place where they laid him. But go your way, tell his disciples and Peter that he goeth before you into Galilee: there shall ye see him, as he said unto you. And they went out quickly, and fled from the sepulchre; for they trembled and were amazed: neither said they any thing to any man; for they were afraid.


\subby{Easter Monday}
\fancyhead[RE,LO]{Easter Monday}
\begin{inhead}
    {First Class Double}
\end{inhead}

\collect
\lett{O}{God,} who by the paschal solemnity hast bestowed healing upon the world: prosper, we beseech thee, thy people with thy heavenly gift; that they may be worthy to attain unto perfect freedom, and may likewise be profited unto everlasting life. Through.

\readingcitation{Epistle}{Acts 10:37}
\lett{I}{n those days:} Peter standing in the midst of the people, said: Men and brethren, ye know the word which was published throughout all Jud{\ae}a: and began from Galilee, after the baptism which John preached: how God anointed Jesus of Nazareth with the Holy Ghost, and with power, who went about doing good, and healing all that were oppressed of the devil, for God was with him. And we are witnesses of all things, which he did both in the land of the Jews and in Jerusalem, whom they slew and hanged on a tree. Him God raised up the third day, and shewed him openly, not to all the people, but unto witnesses chosen before of God: even to us, who did eat and drink with him, after he rose from the dead. And he commanded us to preach unto the people, and to testify, that it is he who was ordained of God to be the judge of quick and dead. To him give all the Prophets witness, that through his name whosoever believeth in him shall receive remission of sins.

\subby{Easter Tuesday}
\fancyhead[RE,LO]{Easter Tuesday}
\begin{inhead}
    {First Class Double}
\end{inhead}

\collect
\lett{O}{God,} who dost ever multiply thy Church with new offspring; grant unto thy servants; that they may hold fast in their lives the sacrament which by faith they have received. Through.

\readingcitation{Epistle}{Acts 13:26}
\lett{I}{n those days:} Paul standing up, and beckoning with his hand to keep silence, said: Men and brethren, children of the stock of Abraham, and whosoever among you feareth God, to you is the word of this salvation sent. For they that dwell at Jerusalem, and their rulers, because they knew not Jesus, nor yet the voices of the Prophets which are read every sabbath-day, they have fulfilled them in condemning him: and though they found no cause of death in him, yet desired they Pilate that he should be slain. And when they had fulfilled all that was written of him, they took him down from the tree, and laid him in a sepulchre But God raised him from the dead: and he was seen many days of them which came up with him from Galilee to Jerusalem, who are his witnesses unto the people. And we declare unto you glad tidings, how that the promise which was made unto the fathers, God, hath fulfilled the same unto us their children, in that he hath raised up Jesus Christ, our Lord.

\subby{Low Sunday}
\fancyhead[RE,LO]{Easter I}
\begin{inhead}
    {First Class Double, Octave Day of Easter}
\end{inhead}


\collect
\lett{G}{rant,} we beseech thee, almighty God: that we who have celebrated the paschal feasts may, through thy bounty, ever keep them in our life and conversation. Through.

\readingcitation{Epistle}{1 John 1:4}
\lett{D}{early beloved:} Whatsoever is born of God overcometh the world: and this is the victory that overcometh the world, even our faith. Who is he that overcometh the world, but he that believeth that Jesus is the Son of God? This is he that came by water and blood, even Jesus Christ; not by water only, but by water and blood. And it is the Spirit that beareth witness, because the Spirit is truth. For there are three that bear record in heaven, the Father, the Word, and the Holy Ghost: and these three are one. And there are three that bear witness in earth, the spirit, and the water, and the blood: and these three agree in one. If we receive the witness of men, the witness of God is greater: for this is the witness of God which he hath testified of his Son. He that believeth on the Son of God hath the witness in himself.

\readingcitation{Gospel}{John 20:19}
\lett{A}{t that time:} The same day at evening, being the first day of the week, when the doors were shut where the disciples were assembled for fear of the Jews, came Jesus and stood in the midst, and saith unto them, Peace be unto you. And when he had so said, he shewed unto them his hands and his side. Then were the disciples glad, when they saw the Lord. Then said Jesus to them again, Peace be unto you: as my Father hath sent me, even so send I you. And when he had said this, he breathed on them, and saith unto them, Receive ye the Holy Ghost: Whose soever sins ye remit, they are remitted unto them; and whose soever sins ye retain, they are retained. But Thomas, one of the twelve, called Didymus, was not with them when Jesus came. The other disciples therefore said unto him, We have seen the Lord. But he said unto them, Except I shall see in his hands the print of the nails, and put my finger into the print of the nails, and thrust my hand into his side, I will not believe. And after eight days again his disciples were within, and Thomas with them: then came Jesus, the doors being shut, and stood in the midst, and said, Peace be unto you. Then saith he to Thomas, Reach hither thy finger, and behold my hands; and reach hither thy hand, and thrust it into my side: and be not faithless, but believing. And Thomas answered and said unto him, My Lord and my God. Jesus saith unto him, Thomas, because thou hast seen me, thou hast believed: blessed are they that have not seen, and yet have believed. And many other signs truly did Jesus in the presence of his disciples, which are not written in this book: But these are written, that ye might believe that Jesus is the Christ, the Son of God; and that believing ye might have life through his name.


\subby{Second Sunday after Easter}\label{EasterII}
\fancyhead[RE,LO]{Easter II}
\begin{inhead}
    {Semidouble}
\end{inhead}

\collect
\lett{O}{God,} who in the humility of thy Son hast raised up a fallen world: grant unto thy faithful abiding gladness: that they whom thou hast delivered from the peril of eternal death, may be brought to the fruition of everlasting joys. Through the same.
\begin{rubric}
    \nth{2} Collect is of St. Mary in Eastertide (p. \pageref{SPMaryInEaster}) \& \nth{3} Against the Persecutors of the Church (p. \pageref{SPAgainst}) or for the Chief Bishop (p. \pageref{SPChiefBishop}).
\end{rubric}


\readingcitation{Epistle}{1 Peter 2:19}
\lett{D}{early beloved:} Christ suffered for us, leaving us an example, that ye should follow his steps: Who did no sin, neither was guile found in his mouth: Who, when he was reviled, reviled not again; when he suffered, he threatened not; but committed himself to him that judgeth righteously: Who his own self bare our sins in his own body on the tree, that we, being dead to sins, should live unto righteousness: by whose stripes ye were healed. For ye were as sheep going astray; but are now returned unto the Shepherd and Bishop of your souls.


\subby{Third Sunday after Easter}
\fancyhead[RE,LO]{Easter III}
\begin{inhead}
    {Semidouble}
\end{inhead}

\readingcitation{Epistle}{1 Peter 2:11}
\lett{D}{early beloved:} I beseech you as strangers and pilgrims, abstain from fleshly lusts, which war against the soul; Having your conversation honest among the Gentiles: that, whereas they speak against you as evildoers, they may by your good works, which they shall behold, glorify God in the day of visitation. Submit yourselves to every ordinance of man for the Lord's sake: whether it be to the king, as supreme; Or unto governors, as unto them that are sent by him for the punishment of evildoers, and for the praise of them that do well. For so is the will of God, that with well doing ye may put to silence the ignorance of foolish men: As free, and not using your liberty for a cloke of maliciousness, but as the servants of God. Honour all men. Love the brotherhood. Fear God. Honour the king. Servants, be subject to your masters with all fear; not only to the good and gentle, but also to the froward. For this is thankworthy: in Christ Jesus, our Lord.

\subby{Fourth Sunday after Easter}
\fancyhead[RE,LO]{Easter IV}
\begin{inhead}
    {Semidouble}
\end{inhead}

\collect
\lett{O}{God,} who makest the minds of the faithful to be of one will: grant unto thy people, that they may love the thing which thou commandest, and desire that which thou dost promise; that amid the changes of the world our hearts may there be fixed, where true joys are to be found. Through.
\begin{rubric}
    Additional Collects are as on Easter II (p. \pageref{EasterII}).
\end{rubric}

\subby{Rogation Sunday}
\fancyhead[RE,LO]{Easter V}
\begin{inhead}
    {Semidouble}
\end{inhead}

\collect
\lett{O}{God,} from whom all good things do come, grant to us thy humble servants: that by thy inspiration we may think those things that be right; and by thy guiding may perform the same. Through.

\readingcitation{Gospel}{John 16:23}
\lett{A}{t that time:} Jesus said unto his disciples: Verily, verily, I say unto you, Whatsoever ye shall ask the Father in my name, he will give it you. Hitherto have ye asked nothing in my name: ask, and ye shall receive, that your joy may be full. These things have I spoken unto you in proverbs: but the time cometh, when I shall no more speak unto you in proverbs, but I shall shew you plainly of the Father. At that day ye shall ask in my name: and I say not unto you, that I will pray the Father for you: For the Father himself loveth you, because ye have loved me, and have believed that I came out from God. I came forth from the Father, and am come into the world: again, I leave the world, and go to the Father. His disciples said unto him, Lo, now speakest thou plainly, and speakest no proverb. Now are we sure that thou knowest all things, and needest not that any man should ask thee: by this we believe that thou camest forth from God.


\subby{Rogation Days}
\fancyhead[RE,LO]{Rogation Days}
\subsubsec{Daily Office Propers}\par\noindent
\begin{inhead}
    Monday\par
 \nth{2} Feria
\end{inhead}

\begin{inhead}
    Wednesday
\end{inhead}
%CHECK:

\collect
\lett{G}{rant,} we beseech thee, almighty God: that we, who in our affliction do put our trust in thy mercy: may ever be defended by thy protection against all adversities. Through.
\begin{rubric}
    Additional Collects are as on Easter II (p. \pageref{EasterII}).
\end{rubric}

\readingcitation{Epistle}{James 5:16}
\lett{D}{early beloved:} Confess your faults one to another, and pray one for another, that ye may be healed. The effectual fervent prayer of a righteous man availeth much. Elias was a man subject to like passions as we are, and he prayed earnestly that it might not rain: and it rained not on the earth by the space of three years and six months. And he prayed again, and the heaven gave rain, and the earth brought forth her fruit. Brethren, if any of you do err from the truth, and one convert him; Let him know, that he which converteth the sinner from the error of his way shall save a soul from death, and shall hide a multitude of sins.

\subby{Ascension Thursday}
\fancyhead[RE,LO]{Ascension Thursday}
\begin{inhead}
    {First Class Double, Third Class Octave}
\end{inhead}

\collect
\lett{G}{rant,} we beseech thee almighty God: that as we believe thy only-begotten Son our Redeemer to have this day ascended into the heavens: so we may also in mind dwell in heavenly places. Through the same.

%CHECK:
\readingcitation{Gospel}{Mark 16:14}
\lett{A}{t that time:} Jesus appeared unto the eleven as they sat at meat, and upbraided them with their unbelief and hardness of heart, because they believed not them which had seen him after he was risen. And he said unto them, Go ye into all the world, and preach the gospel to every creature. He that believeth and is baptized shall be saved; but he that believeth not shall be damned. And these signs shall follow them that believe; In my name shall they cast out devils; they shall speak with new tongues; They shall take up serpents; and if they drink any deadly thing, it shall not hurt them; they shall lay hands on the sick, and they shall recover. So then after the Lord had spoken unto them, he was received up into heaven, and sat on the right hand of God. And they went forth, and preached every where, the Lord working with them, and confirming the word with signs following.

\subby{Sunday within the Ascension Octave}
\fancyhead[RE,LO]{Sunday within Ascension}
\begin{inhead}
    {Semidouble}
\end{inhead}

\collect
\lett{A}{lmighty} and everlasting God: grant that our wills may ever be subjected unto thee; and that we may serve thy majesty in sincerity of heart. Through.
\lett{G}{rant,} we beseech thee, almighty God: that like as we do believe thy only-begotten Son our Lord Jesus Christ to have ascended into the heavens; so we may also in heart and mind thither ascend and with him continually dwell. Who liveth and reigneth with thee.

%CHECK:
\readingcitation{Epistle}{1 Peter 4:7}

\lett{D}{early beloved:} The end of all things is at hand:
%
be ye therefore sober, and watch unto prayer. And above all things have fervent charity among yourselves: for charity shall cover the multitude of sins. Use hospitality one to another without grudging. As every man hath received the gift, even so minister the same one to another, as good stewards of the manifold grace of God. If any man speak, let him speak as the oracles of God; if any man minister, let him do it as of the ability which God giveth: that God in all things may be glorified through Jesus Christ,
%
to whom be praise and dominion for ever and ever. Amen.

\subby{Whitsunday}
\fancyhead[RE,LO]{Whitsunday}
\begin{inhead}
    {First Class Double, First Class Octave}
\end{inhead}

%CHECK:
\readingcitation{Gospel}{John 14:15}
\lett{A}{t that time:} Jesus said unto his disciples: If ye love me, keep my commandments. And I will pray the Father, and he shall give you another Comforter, that he may abide with you for ever; Even the Spirit of truth; whom the world cannot receive, because it seeth him not, neither knoweth him: but ye know him; for he dwelleth with you, and shall be in you. I will not leave you comfortless: I will come to you. Yet a little while, and the world seeth me no more; but ye see me: because I live, ye shall live also. At that day ye shall know that I am in my Father, and ye in me, and I in you. He that hath my commandments, and keepeth them, he it is that loveth me: and he that loveth me shall be loved of my Father, and I will love him, and will manifest myself to him. Judas saith unto him, not Iscariot, Lord, how is it that thou wilt manifest thyself unto us, and not unto the world? Jesus answered and said unto him,
%
If a man love me, he will keep my words: and my Father will love him, and we will come unto him, and make our abode with him. He that loveth me not keepeth not my sayings: and the word which ye hear is not mine, but the Father's which sent me. These things have I spoken unto you, being yet present with you. But the Comforter, which is the Holy Ghost, whom the Father will send in my name, he shall teach you all things, and bring all things to your remembrance, whatsoever I have said unto you. Peace I leave with you, my peace I give unto you: not as the world giveth, give I unto you. Let not your heart be troubled, neither let it be afraid. Ye have heard how I said unto you, I go away, and come again unto you. If ye loved me, ye would rejoice, because I said, I go unto the Father: for my Father is greater than I. And now I have told you before it come to pass, that, when it is come to pass, ye might believe. Hereafter I will not talk much with you: for the prince of this world cometh, and hath nothing in me. But that the world may know that I love the Father; and as the Father gave me commandment, even so I do.


\subby{Whit-Monday}
\fancyhead[RE,LO]{Whit-Monday}
\begin{inhead}
    {First Class Double}
\end{inhead}

\collect
\lett{O}{God,} who didst give the Holy Spirit unto thine Apostles: grant to thy people the fulfilment of their godly petitions; and as thou hast given to them faith, so also bestow upon them peace. Through.

%CHECK:
\readingcitation{Epistle}{Acts 10:34}
%The English and Roman epistles are so similar, we make no distinction.
\lett{I}{n those days:} Peter opened his mouth, and said, Of a truth I perceive that God is no respecter of persons: But in every nation he that feareth him, and worketh righteousness, is accepted with him. The word which God sent unto the children of Israel, preaching peace by Jesus Christ: (he is Lord of all:) That word, I say, ye know, which was published throughout all Judaea, and began from Galilee, after the baptism which John preached; How God anointed Jesus of Nazareth with the Holy Ghost and with power: who went about doing good, and healing all that were oppressed of the devil; for God was with him. And we are witnesses of all things which he did both in the land of the Jews, and in Jerusalem; whom they slew and hanged on a tree: Him God raised up the third day, and shewed him openly; Not to all the people, but unto witnesses chosen before of God, even to us, who did eat and drink with him after he rose from the dead.
%
And he commanded us to preach unto the people, and to testify that it is he which was ordained of God to be the Judge of quick and dead. To him give all the prophets witness, that through his name whosoever believeth in him shall receive remission of sins. While Peter yet spake these words, the Holy Ghost fell on all them which heard the word. And they of the circumcision which believed were astonished, as many as came with Peter, because that on the Gentiles also was poured out the gift of the Holy Ghost. For they heard them speak with tongues, and magnify God. Then answered Peter, Can any man forbid water, that these should not be baptized, which have received the Holy Ghost as well as we? And he commanded them to be baptized in the name of the Lord. Then prayed they him to tarry certain days.

\subby{Whit-Tuesday}
\fancyhead[RE,LO]{Whit-Tuesday}
\begin{inhead}
    {First Class Double}
\end{inhead}


\collect
\lett{W}{e} beseech thee, O Lord, let the power of the Holy Ghost be with us: that it may both mercifully cleanse our hearts, and defend us from all adversities. Through.


\subby{Trinity Sunday}
\fancyhead[RE,LO]{Pentecost Sunday}
\begin{inhead}
    {First Class Double}
\end{inhead}
\readingcitation{Epistle}{Romans 11:33}
\lett{O}{the} depth of the riches both of the wisdom and knowledge of God! how unsearchable are his judgments, and his ways past finding out! For who hath known the mind of the Lord? or who hath been his counsellor? Or who hath first given to him, and it shall be recompensed unto him again? For of him, and through him, and to him, are all things: to whom be glory for ever. Amen.


\readingcitation{Gospel}{Matthew 28:18}
\lett{A}{t that time:} Jesus said unto his disciples: All power is given unto me in heaven and in earth. Go ye therefore, and teach all nations, baptizing them in the name of the Father, and of the Son, and of the Holy Ghost: Teaching them to observe all things whatsoever I have commanded you: and, lo, I am with you alway, even unto the end of the world.


\subby{Sunday in the Octave of Corpus Christi}
\fancyhead[RE,LO]{Pentecost I}
\centerline{\small{(First Sunday after Trinity)}}
\begin{inhead}
    {Semidouble}
\end{inhead}

\readingcitation{Epistle}{1 John 4:7}
\lett{D}{early beloved:} God is love. In this was manifested the love of God toward us, because that God sent his only begotten Son into the world, that we might live through him. Herein is love, not that we loved God, but that he loved us, and sent his Son to be the propitiation for our sins. Beloved, if God so loved us, we ought also to love one another. No man hath seen God at any time. If we love one another, God dwelleth in us, and his love is perfected in us. Hereby know we that we dwell in him, and he in us, because he hath given us of his Spirit. And we have seen and do testify that the Father sent the Son to be the Saviour of the world. Whosoever shall confess that Jesus is the Son of God, God dwelleth in him, and he in God. And we have known and believed the love that God hath to us. God is love; and he that dwelleth in love dwelleth in God, and God in him. Herein is our love made perfect, that we may have boldness in the day of judgment: because as he is, so are we in this world. There is no fear in love; but perfect love casteth out fear: because fear hath torment. He that feareth is not made perfect in love. We love him, because he first loved us. If a man say, I love God, and hateth his brother, he is a liar: for he that loveth not his brother whom he hath seen, how can he love God whom he hath not seen? And this commandment have we from him, That he who loveth God love his brother also.


\subby{Third Sunday after Pentecost}\label{PentecostIII}
\fancyhead[RE,LO]{Pentecost III}
\begin{inhead}
    {Semidouble}
\end{inhead}

\collect
\lett{O}{God,} the protector of all that trust in thee, without whom nothing is strong, nothing is holy: increase and multiply upon us thy mercy; that, thou being our ruler and guide, we may so pass through things temporal, that we finally lose not the things eternal. Through.
\begin{rubric}
    \nth{2} Collect is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

%CHECK:
\readingcitation{Epistle}{1 Peter 5:5}
\lett{D}{early beloved:} All of you be subject one to another, and be clothed with humility: for God resisteth the proud, and giveth grace to the humble. %
Humble yourselves therefore under the mighty hand of God, that he may exalt you in due time: casting all your care upon him; for he careth for you. Be sober, be vigilant; because your adversary the devil, as a roaring lion, walketh about, seeking whom he may devour: whom resist stedfast in the faith, knowing that the same afflictions are accomplished in your brethren that are in the world. But the God of all grace, who hath called us unto his eternal glory by Christ Jesus, after that ye have suffered a while, make you perfect, stablish, strengthen, settle you. To him be glory and dominion for ever and ever. Amen.

\secret
\lett{R}{egard,} O Lord, the gifts of thy suppliant Church: and grant that the partaking thereof may avail for the salvation and continual sanctification of them that believe. Through.
\begin{rubric}
    \nth{2} Secret is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\postcommunion
\lett{M}{ay} the holy things, which we have received, quicken us, O Lord: that we being cleansed thereby may be made ready for thine everlasting mercy. Through.
\begin{rubric}
    \nth{2} Postcommunion is of the Saints (p. \pageref{SPSaints}) \& \nth{3} is \emph{ad libitum}.
\end{rubric}

\subby{Fourth Sunday after Pentecost}
\fancyhead[RE,LO]{Pentecost IV}
\begin{inhead}
    {Semidouble}
\end{inhead}

\collect
\lett{G}{rant,} O Lord, we beseech thee, that the course of this world may be so peaceably ordered by thy governance, that thy Church may joyfully serve thee in all godly quietness. Through.
\begin{rubric}
    Additional Collects are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\readingcitation{Gospel}{Luke 5:1}
\lett{A}{t that time:} It came to pass, that as the people pressed upon Jesus to hear the word of God, he stood by the lake of Gennesaret, and saw two ships standing by the lake: but the fishermen were gone out of them, and were washing their nets. And he entered into one of the ships, which was Simons, and prayed him that he would thrust out a little from the land. And he sat down, and taught the people out of the ship. Now when he had left speaking, he said unto Simon, Launch out into the deep, and let down your nets for a draught. And Simon answering said unto him, Master, we have toiled all the night, and have taken nothing: nevertheless at thy word I will let down the net. And when they had this done, they inclosed a great multitude of fishes: and their net brake. And they beckoned unto their partners, which were in the other ship, that they should come and help them. And they came, and filled both the ships, so that they began to sink. When Simon Peter saw it, he fell down at Jesus knees, saying, Depart from me; for I am a sinful man, O Lord. For he was astonished, and all that were with him, at the draught of the fishes which they had taken: and so was also James, and John, the sons of Zebedee, which were partners with Simon. And Jesus said unto Simon, Fear not; from henceforth thou shalt catch men. And when they had brought their ships to land, they forsook all, and followed him.

\secret
\lett{O}{Lord,} we beseech thee mercifully to receive our oblations: and graciously turn our rebel wills to thee. Through.
\begin{rubric}
    Additional Secrets are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\postcommunion
\lett{W}{e} beseech thee, O Lord: that the mysteries which we have received may purify us: and defend us by the gift which they bestow. Through.
\begin{rubric}
    Additional Postcommunions are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\subby{Fifth Sunday after Pentecost}
\fancyhead[RE,LO]{Pentecost V}
\begin{inhead}
    {Semidouble}
\end{inhead}

\collect
\lett{O}{God,} who hast prepared for them that love thee such good things as pass man's understanding: pour into our hearts such love toward thee: that we, loving thee above all things, may obtain thy promises, which exceed all that we can desire. Through.
\begin{rubric}
    Additional Collects are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}


\readingcitation{Gospel}{Matthew 5:20}
\lett{A}{t that time:} Jesus said unto his disciples: Except your righteousness shall exceed the righteousness of the scribes and Pharisees, ye shall in no case enter into the kingdom of heaven. Ye have heard that it was said by them of old time, Thou shalt not kill; and whosoever shall kill shall be in danger of the judgment: But I say unto you, That whosoever is angry with his brother without a cause shall be in danger of the judgment: and whosoever shall say to his brother, Raca, shall be in danger of the council: but whosoever shall say, Thou fool, shall be in danger of hell fire. Therefore if thou bring thy gift to the altar, and there rememberest that thy brother hath ought against thee; Leave there thy gift before the altar, and go thy way; first be reconciled to thy brother, and then come and offer thy gift.

\secret
\lett{B}{e} merciful, O Lord, unto our supplications: and graciously receive these oblations of thy servants and handmaids; that those things which each hath offered to the honour of thy name may be profitable unto all for their salvation. Through.
\begin{rubric}
    Additional Secrets are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\postcommunion
\lett{O}{Lord} who hast satisfied us with thy heavenly gift: grant, we beseech thee; that we may be cleansed from our secret faults, and delivered from the snares of our enemies. Through.
\begin{rubric}
    Additional Postcommunions are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\subby{Sixth Sunday after Pentecost}
\fancyhead[RE,LO]{Pentecost VI}
\begin{inhead}
    {Semidouble}
\end{inhead}

\collect
\lett{G}{od} of all power and might, who art the author of all good things: graft in our hearts the love of thy name and increase in us true religion; that, nourishing all that is good, thou mayest mercifully guard that which thou hast nourished. Through.
\begin{rubric}
    Additional Collects are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\readingcitation{Gospel}{Mark 8:1}
\lett{I}{n those days:} The multitude being very great, and having nothing to eat, Jesus called his disciples unto him, and saith unto them, I have compassion on the multitude, because they have now been with me three days, and have nothing to eat: And if I send them away fasting to their own houses, they will faint by the way: for divers of them came from far. And his disciples answered him, From whence can a man satisfy these men with bread here in the wilderness? And he asked them, How many loaves have ye? And they said, Seven. And he commanded the people to sit down on the ground: and he took the seven loaves, and gave thanks, and brake, and gave to his disciples to set before them; and they did set them before the people. And they had a few small fishes: and he blessed, and commanded to set them also before them. So they did eat, and were filled: and they took up of the broken meat that was left seven baskets. And they that had eaten were about four thousand: and he sent them away.


\secret
\lett{B}{e} merciful, O Lord, unto our supplications, and graciously receive these oblations of thy people: and that none may pray amiss or fail in his petition, grant; that what we ask faithfully, we may obtain effectually. Through.
\begin{rubric}
    Additional Secrets are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\postcommunion
\lett{W}{e} have been fulfilled with thy gifts, O Lord: grant, we beseech thee; that we may both be cleansed by their operation, and defended by their succour. Through.
\begin{rubric}
    Additional Postcommunions are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\subby{Seventh Sunday after Pentecost}
\fancyhead[RE,LO]{Pentecost VII}
\begin{inhead}
    {Semidouble}
\end{inhead}

\collect
\lett{O}{God,} whose never-failing providence ordereth all things both in heaven and earth: we humbly beseech thee; to put away from us all hurtful things, and to give us all things which be profitable for us. Through.
\begin{rubric}
    Additional Collects are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\readingcitation{Gospel}{Matthew 7:15}
\lett{A}{t that time:} Jesus said unto his disciples: Beware of false prophets, which come to you in sheep's clothing, but inwardly they are ravening wolves. Ye shall know them by their fruits. Do men gather grapes of thorns, or figs of thistles? Even so every good tree bringeth forth good fruit; but a corrupt tree bringeth forth evil fruit. A good tree cannot bring forth evil fruit, neither can a corrupt tree bring forth good fruit. Every tree that bringeth not forth good fruit is hewn down, and cast into the fire. Wherefore by their fruits ye shall know them. Not every one that saith unto me, Lord, Lord, shall enter into the kingdom of heaven; but he that doeth the will of my Father which is in heaven.

\secret
\lett{O}{God} who by the one perfect sacrifice didst ratify the divers offerings of the law: accept the sacrifice of thy bounden servants, and sanctify it with that blessing wherewith thou didst bless the gifts of Abel: that what each hath offered to the honour of thy majesty may be profitable unto all for their salvation. Through.
\begin{rubric}
    Additional Secrets are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\postcommunion
\lett{L}{et} the operation of thy healing, O Lord, both mercifully set us free from our perversities; and lead us to the things that be right. Through.
\begin{rubric}
    Additional Postcommunions are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\subby{Eighth Sunday after Pentecost}
\fancyhead[RE,LO]{Pentecost VIII}
\begin{inhead}
    {Semidouble}
\end{inhead}

\collect
\lett{G}{rant} to us, Lord, we beseech thee, the spirit to think and do always such things as be rightful: that we, who cannot exist without thee; may by thee be enabled to live according to thy will. Through.
\begin{rubric}
    Additional Collects are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\readingcitation{Gospel}{Luke 16:1}
\lett{A}{t that time:} Jesus spake this parable unto his disciples: There was a certain rich man, which had a steward; and the same was accused unto him that he had wasted his goods. And he called him, and said unto him, How is it that I hear this of thee? give an account of thy stewardship; for thou mayest be no longer steward. Then the steward said within himself, What shall I do? for my lord taketh away from me the stewardship: I cannot dig; to beg I am ashamed. I am resolved what to do, that, when I am put out of the stewardship, they may receive me into their houses. So he called every one of his lord's debtors unto him, and said unto the first, How much owest thou unto my lord? And he said, An hundred measures of oil. And he said unto him, Take thy bill, and sit down quickly, and write fifty. Then said he to another, And how much owest thou? And he said, An hundred measures of wheat. And he said unto him, Take thy bill, and write fourscore. And the lord commended the unjust steward, because he had done wisely: for the children of this world are in their generation wiser than the children of light. And I say unto you, Make to yourselves friends of the mammon of unrighteousness; that, when ye fail, they may receive you into everlasting habitations.

\secret
\lett{A}{ccept,} we beseech thee, O Lord, the gifts which of thine own bounty we offer unto thee: that, by the mighty working of thy grace, these most sacred mysteries may both sanctify our conversation in this life, and bring us to everlasting joys. Through.
\begin{rubric}
    Additional Secrets are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\postcommunion
\lett{M}{ay} this heavenly mystery be unto us, O Lord, the renewing both of mind and of body: that like as we do offer unto thee our worship, so we may perceive the benefit of the same. Through.
\begin{rubric}
    Additional Postcommunions are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\subby{Ninth Sunday after Pentecost}
\fancyhead[RE,LO]{Pentecost IX}
\begin{inhead}
{Semidouble}
\end{inhead}

\collect
\lett{L}{et} thy merciful ears, O Lord, be open to the prayers of thy humble servants: and that they may obtain their petitions; make them to ask such things as shall please thee. Through.
\begin{rubric}
    Additional Collects are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\readingcitation{Gospel}{Luke 19:41}
\lett{A}{t that time:} When Jesus was come near to Jerusalem, he beheld the city, and wept over it, Saying, If thou hadst known, even thou, at least in this thy day, the things which belong unto thy peace! but now they are hid from thine eyes. For the days shall come upon thee, that thine enemies shall cast a trench about thee, and compass thee round, and keep thee in on every side, And shall lay thee even with the ground, and thy children within thee; and they shall not leave in thee one stone upon another; because thou knewest not the time of thy visitation. And he went into the temple, and began to cast out them that sold therein, and them that bought; Saying unto them, It is written, My house is the house of prayer: but ye have made it a den of thieves. And he taught daily in the temple. But the chief priests and the scribes and the chief of the people sought to destroy him.

\secret
\lett{G}{rant} to us, we beseech thee, O Lord, worthily to frequent these mysteries: for as oft as the commemoration of this victim is celebrated, the work of our redemption is performed. Through.
\begin{rubric}
    Additional Secrets are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\postcommunion
\lett{W}{e} beseech thee, O Lord, that the communion of thy sacrament may both bestow upon us purification, and grant us purity. Through.
\begin{rubric}
    Additional Postcommunions are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\subby{Tenth Sunday after Pentecost}
\fancyhead[RE,LO]{Pentecost X}
\begin{inhead}
{Semidouble}
\end{inhead}

\collect
\lett{O}{God,} who declarest thy almighty power most chiefly in shewing mercy and pity: multiply upon us thy loving-kindness; that we, running toward thy promises, may by thee be made partakers of thy heavenly gifts. Through.
\begin{rubric}
    Additional Collects are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\readingcitation{Gospel}{Luke 18:9}
\lett{A}{t that time:} Jesus spake this parable unto certain which trusted in themselves that they were righteous, and despised others: Two men went up into the temple to pray; the one a Pharisee, and the other a publican. The Pharisee stood and prayed thus with himself, God, I thank thee, that I am not as other men are, extortioners, unjust, adulterers, or even as this publican. I fast twice in the week, I give tithes of all that I possess. And the publican, standing afar off, would not lift up so much as his eyes unto heaven, but smote upon his breast, saying, God be merciful to me a sinner. I tell you, this man went down to his house justified rather than the other: for every one that exalteth himself shall be abased; and he that humbleth himself shall be exalted.

\secret
\lett{L}{et} the sacrifices which we dedicate be offered unto thee, O Lord: which thou hast granted us so to present for the honour of thy name, that they may also be for our healing. Through.
\begin{rubric}
    Additional Secrets are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\postcommunion
\lett{W}{e} beseech thee, O Lord our God: that as thou failest not to renew us with thy divine sacraments, so thou wouldest not leave us destitute of thy gracious help. Through.
\begin{rubric}
    Additional Postcommunions are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\subby{Eleventh Sunday after Pentecost}
\fancyhead[RE,LO]{Pentecost XI}
\begin{inhead}
{Semidouble}
\end{inhead}

\collect
\lett{A}{lmighty} and everlasting God, who in the abundance of thy loving-kindness dost give more than thy suppliants either desire or deserve: pour down upon us thy mercy; forgiving us those things whereof our conscience is afraid, and giving us those things which we do not presume to ask. Through.
\begin{rubric}
    Additional Collects are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\readingcitation{Gospel}{Mark 7:31}
\lett{A}{t that time:} Jesus, departing from the coasts of Tyre and Sidon, he came unto the sea of Galilee, through the midst of the coasts of Decapolis. And they bring unto him one that was deaf, and had an impediment in his speech; and they beseech him to put his hand upon him. And he took him aside from the multitude, and put his fingers into his ears, and he spit, and touched his tongue; And looking up to heaven, he sighed, and saith unto him, Ephphatha, that is, Be opened. And straightway his ears were opened, and the string of his tongue was loosed, and he spake plain. And he charged them that they should tell no man: but the more he charged them, so much the more a great deal they published it; And were beyond measure astonished, saying, He hath done all things well: he maketh both the deaf to hear, and the dumb to speak.

\secret
\lett{L}{ook} favourably, O Lord, we beseech thee, on this our service: that the gift which we offer may be acceptable to thee, and avail for the succour of our frailty. Through.
\begin{rubric}
    Additional Secrets are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\postcommunion
\lett{W}{e} beseech thee, O Lord, that we who have received thy sacrament may in such wise feel the succour of the same: that being preserved both in body and soul, we may glory in the fulness of thy heavenly healing. Through.
\begin{rubric}
    Additional Postcommunions are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\subby{Twelfth Sunday after Pentecost}
\fancyhead[RE,LO]{Pentecost XII}
\begin{inhead}
{Semidouble}
\end{inhead}

\collect
\lett{A}{lmighty} and merciful God, of whose gift it cometh that thy faithful people do unto thee worthy and laudable service: grant, we beseech thee; that we may so faithfully serve thee in this life, that we may without stumbling run the way of thy promises. Through.
\begin{rubric}
    Additional Collects are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\readingcitation{Gospel}{Luke 10:23}
\lett{A}{t that time:} Jesus said unto his disciples: Blessed are the eyes which see the things that ye see: For I tell you, that many prophets and kings have desired to see those things which ye see, and have not seen them; and to hear those things which ye hear, and have not heard them. And, behold, a certain lawyer stood up, and tempted him, saying, Master, what shall I do to inherit eternal life? He said unto him, What is written in the law? how readest thou? And he answering said, Thou shalt love the Lord thy God with all thy heart, and with all thy soul, and with all thy strength, and with all thy mind; and thy neighbour as thyself. And he said unto him, Thou hast answered right: this do, and thou shalt live. But he, willing to justify himself, said unto Jesus, And who is my neighbour? And Jesus answering said, A certain man went down from Jerusalem to Jericho, and fell among thieves, which stripped him of his raiment, and wounded him, and departed, leaving him half dead. And by chance there came down a certain priest that way: and when he saw him, he passed by on the other side. And likewise a Levite, when he was at the place, came and looked on him, and passed by on the other side. But a certain Samaritan, as he journeyed, came where he was: and when he saw him, he had compassion on him, And went to him, and bound up his wounds, pouring in oil and wine, and set him on his own beast, and brought him to an inn, and took care of him. And on the morrow when he departed, he took out two pence, and gave them to the host, and said unto him, Take care of him; and whatsoever thou spendest more, when I come again, I will repay thee. Which now of these three, thinkest thou, was neighbour unto him that fell among the thieves? And he said, He that shewed mercy on him. Then said Jesus unto him, Go, and do thou likewise.

\secret
\lett{M}{ercifully} look, we beseech thee, O Lord, upon the sacrifices which we present on thy sacred altars: that while they do bring us pardon, they may minister unto the honour of thy name. Through.
\begin{rubric}
    Additional Secrets are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\postcommunion
\lett{Q}{uicken} us, O Lord, we pray thee, who have received these holy mysteries: that they may obtain for us both pardon and protection. Through.
\begin{rubric}
    Additional Postcommunions are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\subby{Thirteenth Sunday after Pentecost}
\fancyhead[RE,LO]{Pentecost XIII}
\begin{inhead}
{Semidouble}
\end{inhead}

\collect
\lett{A}{lmighty} and everlasting God, give unto us the increase of faith, hope, and charity: and, that we may be worthy to obtain that which thou dost promise, make us to love that which thou dost command. Through.
\begin{rubric}
    Additional Collects are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\readingcitation{Gospel}{Luke 17:11}
\lett{A}{t that time:} As Jesus went to Jerusalem, he passed through the midst of Samaria and Galilee. And as he entered into a certain village, there met him ten men that were lepers, which stood afar off: And they lifted up their voices, and said, Jesus, Master, have mercy on us. And when he saw them, he said unto them, Go shew yourselves unto the priests. And it came to pass, that, as they went, they were cleansed. And one of them, when he saw that he was healed, turned back, and with a loud voice glorified God, And fell down on his face at his feet, giving him thanks: and he was a Samaritan. And Jesus answering said, Were there not ten cleansed? but where are the nine? There are not found that returned to give glory to God, save this stranger. And he said unto him, Arise, go thy way: thy faith hath made thee whole.

\secret
\lett{B}{e} favourable, O Lord, to thy people, be favourable unto their offerings: that, being well-pleased with this oblation, thou mayest both bestow upon us pardon and grant us our petitions. Through.
\begin{rubric}
    Additional Secrets are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\postcommunion
\lett{W}{e} beseech thee, O Lord, that we who have received thy heavenly sacraments, may thereby grow in grace to the attainment of eternal redemption. Through.
\begin{rubric}
    Additional Secrets are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\subby{Fourteenth Sunday after Pentecost}
\fancyhead[RE,LO]{Pentecost XIV}
\begin{inhead}
{Semidouble}
\end{inhead}

\collect
\lett{K}{eep,} we beseech thee, O Lord, thy Church with thy perpetual mercy: and, because the frailty of man without thee cannot but fall; keep us ever by thy help from things hurtful, and lead us to those things profitable to our salvation. Through.
\begin{rubric}
    Additional Collects are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\readingcitation{Gospel}{Matthew 6:24}
\lett{A}{t that time:} Jesus said unto his disciples: No man can serve two masters: for either he will hate the one, and love the other; or else he will hold to the one, and despise the other. Ye cannot serve God and mammon. Therefore I say unto you, Take no thought for your life, what ye shall eat, or what ye shall drink; nor yet for your body, what ye shall put on. Is not the life more than meat, and the body than raiment? Behold the fowls of the air: for they sow not, neither do they reap, nor gather into barns; yet your heavenly Father feedeth them. Are ye not much better than they? Which of you by taking thought can add one cubit unto his stature? And why take ye thought for raiment? Consider the lilies of the field, how they grow; they toil not, neither do they spin: And yet I say unto you, That even Solomon in all his glory was not arrayed like one of these. Wherefore, if God so clothe the grass of the field, which to day is, and to morrow is cast into the oven, shall he not much more clothe you, O ye of little faith? Therefore take no thought, saying, What shall we eat? or, What shall we drink? or, Wherewithal shall we be clothed? (For after all these things do the Gentiles seek:) for your heavenly Father knoweth that ye have need of all these things. But seek ye first the kingdom of God, and his righteousness; and all these things shall be added unto you.

\secret
\lett{G}{rant} to us, O Lord, we beseech thee, that this saving victim may avail, both for the cleansing of our offences, and for the propitiation of thy power. Through.
\begin{rubric}
    Additional Secrets are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\postcommunion
\lett{M}{ay} thy sacraments, O God, evermore cleanse and defend us: and bring us to the attainment of eternal salvation. Through.
\begin{rubric}
    Additional Postcommunions are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\subby{Fifteenth Sunday after Pentecost}
\fancyhead[RE,LO]{Pentecost XV}
\begin{inhead}
{Semidouble}
\end{inhead}

\collect
\lett{O}{Lord,} we beseech thee, let thy continual pity cleanse and defend thy Church: and, because it cannot continue in safety without thee; govern it evermore by thy goodness. Through.
\begin{rubric}
    Additional Collects are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\readingcitation{Epistle}{Galatians 5:25}
\lett{B}{rethren:} If we live in the Spirit, let us also walk in the Spirit. Let us not be desirous of vain glory, provoking one another, envying one another. Brethren, if a man be overtaken in a fault, ye which are spiritual, restore such an one in the spirit of meekness; considering thyself, lest thou also be tempted. Bear ye one another's burdens, and so fulfil the law of Christ. For if a man think himself to be something, when he is nothing, he deceiveth himself. But let every man prove his own work, and then shall he have rejoicing in himself alone, and not in another. For every man shall bear his own burden. Let him that is taught in the word communicate unto him that teacheth in all good things. Be not deceived; God is not mocked: for whatsoever a man soweth, that shall he also reap. For he that soweth to his flesh shall of the flesh reap corruption; but he that soweth to the Spirit shall of the Spirit reap life everlasting. And let us not be weary in well doing: for in due season we shall reap, if we faint not. As we have therefore opportunity, let us do good unto all men, especially unto them who are of the household of faith.

\readingcitation{Gospel}{Luke 7:11}
\lett{A}{t that time:} Jesus went into a city called Nain; and many of his disciples went with him, and much people. Now when he came nigh to the gate of the city, behold, there was a dead man carried out, the only son of his mother, and she was a widow: and much people of the city was with her. And when the Lord saw her, he had compassion on her, and said unto her, Weep not. And he came and touched the bier: and they that bare him stood still. And he said, Young man, I say unto thee, Arise. And he that was dead sat up, and began to speak. And he delivered him to his mother. And there came a fear on all: and they glorified God, saying, That a great prophet is risen up among us; and, That God hath visited his people.

\secret
\lett{M}{ay} thy sacraments guard us, O Lord: and ever protect us against the assaults of the devil. Through.
\begin{rubric}
    Additional Secrets are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\postcommunion
\lett{W}{e} beseech thee, O Lord, that the operation of thy heavenly gift may so possess our souls and bodies: that not our own desires, but the effectual working of the same may ever prevail within us. Through.
\begin{rubric}
    Additional Postcommunions are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\subby{Sixteenth Sunday after Pentecost}
\fancyhead[RE,LO]{Pentecost XVI}
\begin{inhead}
{Semidouble}
\end{inhead}

\collect
\lett{L}{ord,} we pray thee that thy grace may always prevent and follow us: and make us continually to be given to good works. Through.
\begin{rubric}
    Additional Collects are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\readingcitation{Gospel}{Luke 14:1}
\lett{A}{t that time:} When Jesus went into the house of one of the chief Pharisees to eat bread on the sabbath day, that they watched him. And, behold, there was a certain man before him which had the dropsy. And Jesus answering spake unto the lawyers and Pharisees, saying, Is it lawful to heal on the sabbath day? And they held their peace. And he took him, and healed him, and let him go; And answered them, saying, Which of you shall have an ass or an ox fallen into a pit, and will not straightway pull him out on the sabbath day? And they could not answer him again to these things. And he put forth a parable to those which were bidden, when he marked how they chose out the chief rooms; saying unto them, When thou art bidden of any man to a wedding, sit not down in the highest room; lest a more honourable man than thou be bidden of him; And he that bade thee and him come and say to thee, Give this man place; and thou begin with shame to take the lowest room. But when thou art bidden, go and sit down in the lowest room; that when he that bade thee cometh, he may say unto thee, Friend, go up higher: then shalt thou have worship in the presence of them that sit at meat with thee. For whosoever exalteth himself shall be abased; and he that humbleth himself shall be exalted.

\secret
\lett{C}{leanse} us, O Lord, we pray thee, by the effectual working of this our sacrifice: and so accomplish in us the work of thy mercy; that we may be found worthy to be made partakers of the same. Through.
\begin{rubric}
    Additional Secrets are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\postcommunion
\lett{O}{Lord,} we beseech thee, mercifully to purify our minds and renew them with thy heavenly sacraments: that we may thereby have help also for our bodies both now and for evermore. Through.
\begin{rubric}
    Additional Postcommunions are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\subby{Seventeenth Sunday after Pentecost}
\fancyhead[RE,LO]{Pentecost XVII}
\begin{inhead}
{Semidouble}
\end{inhead}

\collect
\lett{L}{ord,} we beseech thee: grant thy people grace to avoid the infections of the devil: and with pure minds to follow thee, the only God. Through.
\begin{rubric}
    Additional Collects are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\readingcitation{Gospel}{Matthew 22:34}
\lett{A}{t that time:} The Pharisees came to Jesus: and one of them, which was a lawyer, asked him a question, tempting him, and saying, Master, which is the great commandment in the law? Jesus said unto him, Thou shalt love the Lord thy God with all thy heart, and with all thy soul, and with all thy mind. This is the first and great commandment. And the second is like unto it, Thou shalt love thy neighbour as thyself. On these two commandments hang all the law and the prophets. While the Pharisees were gathered together, Jesus asked them, Saying, What think ye of Christ? whose son is he? They say unto him, The Son of David. He saith unto them, How then doth David in spirit call him Lord, saying, The \divineName{Lord} said unto my Lord, Sit thou on my right hand, till I make thine enemies thy footstool? If David then call him Lord, how is he his son? And no man was able to answer him a word, neither durst any man from that day forth ask him any more questions.

\secret
\lett{W}{e} humbly entreat thy majesty, O Lord: that these holy mysteries which we perform may deliver us both from past and future transgressions. Through.
\begin{rubric}
    Additional Secrets are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\postcommunion
\lett{A}{lmighty} God, let thy holy mysteries both heal our vices, and bestow upon us everlasting remedies. Through.
\begin{rubric}
    Additional Postcommunions are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\subby{Eighteenth Sunday after Pentecost}
\fancyhead[RE,LO]{Pentecost XVIII}
\begin{inhead}
{Semidouble}
\end{inhead}

\collect
\lett{L}{et} the operation of thy mercy, O Lord, direct our hearts: forasmuch as without thee we are not able to please thee. Through.
\begin{rubric}
    Additional Collects are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\readingcitation{Gospel}{Matthew 9:1}
\lett{A}{t that time:} Jesus entered into a ship, and passed over, and came into his own city. And, behold, they brought to him a man sick of the palsy, lying on a bed: and Jesus seeing their faith said unto the sick of the palsy; Son, be of good cheer; thy sins be forgiven thee. And, behold, certain of the scribes said within themselves, This man blasphemeth. And Jesus knowing their thoughts said, Wherefore think ye evil in your hearts? For whether is easier, to say, Thy sins be forgiven thee; or to say, Arise, and walk? But that ye may know that the Son of man hath power on earth to forgive sins, (then saith he to the sick of the palsy,) Arise, take up thy bed, and go unto thine house. And he arose, and departed to his house. But when the multitudes saw it, they marvelled, and glorified God, which had given such power unto men.

\secret
\lett{O}{God,} who, by communion in this venerable sacrifice, hast made us partakers of the one supreme Godhead: grant, we beseech thee; that as we have the knowledge of thy truth, so by worthy conversation we may attain unto the same. Through.
\begin{rubric}
    Additional Secrets are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\postcommunion
\lett{W}{e} render thanks unto thee, O Lord, for that thou hast quickened us with thy sacred gift: humbly beseeching thy mercy; that thou wouldest make us worthy partakers of the same. Through.
\begin{rubric}
    Additional Postcommunions are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}


\subby{Nineteenth Sunday after Pentecost}
\fancyhead[RE,LO]{Pentecost XIX}
\begin{inhead}
{Semidouble}
\end{inhead}

\collect
\lett{O}{Almighty} and most merciful God: of thy bountiful goodness keep us from all things that may hurt us; that we, being ready both in body and soul, may cheerfully accomplish those things that thou wouldest have done. Through.
\begin{rubric}
    Additional Collects are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

%CHECK:
\readingcitation{Epistle}{Ephesians 4:17}
\lett{B}{rethren:} This I say therefore, and testify in the Lord, that ye henceforth walk not as other Gentiles walk, in the vanity of their mind, Having the understanding darkened, being alienated from the life of God through the ignorance that is in them, because of the blindness of their heart: Who being past feeling have given themselves over unto lasciviousness, to work all uncleanness with greediness. But ye have not so learned Christ; If so be that ye have heard him, and have been taught by him, as the truth is in Jesus: That ye put off concerning the former conversation the old man, which is corrupt according to the deceitful lusts;
%
Be renewed in the spirit of your mind; And that ye put on the new man, which after God is created in righteousness and true holiness. Wherefore putting away lying, speak every man truth with his neighbour: for we are members one of another. Be ye angry, and sin not: let not the sun go down upon your wrath: Neither give place to the devil. Let him that stole steal no more: but rather let him labour, working with his hands the thing which is good, that he may have to give to him that needeth.
%
Let no corrupt communication proceed out of your mouth, but that which is good to the use of edifying, that it may minister grace unto the hearers. And grieve not the holy Spirit of God, whereby ye are sealed unto the day of redemption. Let all bitterness, and wrath, and anger, and clamour, and evil speaking, be put away from you, with all malice: And be ye kind one to another, tenderhearted, forgiving one another, even as God for Christ's sake hath forgiven you.

\readingcitation{Gospel}{Matthew 22:1}
\lett{A}{t that time:} Jesus spake unto the chief priests and Pharisees in parables, saying: The kingdom of heaven is like unto a certain king, which made a marriage for his son, And sent forth his servants to call them that were bidden to the wedding: and they would not come. Again, he sent forth other servants, saying, Tell them which are bidden, Behold, I have prepared my dinner: my oxen and my fatlings are killed, and all things are ready: come unto the marriage. But they made light of it, and went their ways, one to his farm, another to his merchandise: And the remnant took his servants, and entreated them spitefully, and slew them. But when the king heard thereof, he was wroth: and he sent forth his armies, and destroyed those murderers, and burned up their city. Then saith he to his servants, The wedding is ready, but they which were bidden were not worthy. Go ye therefore into the highways, and as many as ye shall find, bid to the marriage. So those servants went out into the highways, and gathered together all as many as they found, both bad and good: and the wedding was furnished with guests. And when the king came in to see the guests, he saw there a man which had not on a wedding garment: And he saith unto him, Friend, how camest thou in hither not having a wedding garment? And he was speechless. Then said the king to the servants, Bind him hand and foot, and take him away, and cast him into outer darkness; there shall be weeping and gnashing of teeth. For many are called, but few are chosen.

\secret
\lett{G}{rant,} we beseech thee, O Lord, that these gifts which we offer in the sight of thy majesty, may be for our salvation. Through.
\begin{rubric}
    Additional Secrets are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\postcommunion
\lett{M}{ay} the operation of thy healing, O Lord, both mercifully cleanse us from our perversities, and make us ever to cleave unto thy commandments. Through.
\begin{rubric}
    Additional Postcommunions are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\subby{Twentieth Sunday after Pentecost}
\fancyhead[RE,LO]{Pentecost XX}
\begin{inhead}
{Semidouble}
\end{inhead}

\collect
\lett{G}{rant,} we beseech thee, merciful Lord, to thy faithful people pardon and peace: that they may be cleansed from all their sins; and serve thee with a quiet mind. Through.
\begin{rubric}
    Additional Collects are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\readingcitation{Gospel}{John 4:46}
\lett{A}{t that time:} There was a certain nobleman, whose son was sick at Capernaum. When he heard that Jesus was come out of Judaea into Galilee, he went unto him, and besought him that he would come down, and heal his son: for he was at the point of death. Then said Jesus unto him, Except ye see signs and wonders, ye will not believe. The nobleman saith unto him, Sir, come down ere my child die. Jesus saith unto him, Go thy way; thy son liveth. And the man believed the word that Jesus had spoken unto him, and he went his way. And as he was now going down, his servants met him, and told him, saying, Thy son liveth. Then enquired he of them the hour when he began to amend. And they said unto him, Yesterday at the seventh hour the fever left him. So the father knew that it was at the same hour, in the which Jesus said unto him, Thy son liveth: and himself believed, and his whole house.

\secret
\lett{W}{e} beseech thee, O Lord, that these mysteries may bestow upon us thy heavenly healing: and purge away the vices of our hearts. Through.
\begin{rubric}
    Additional Secrets are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\postcommunion
\lett{W}{e} beseech thee, O Lord, make us ever to obey thy commandments: that we may be rendered worthy of thy sacred gifts. Through.
\begin{rubric}
    Additional Postcommunions are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

%CHANGE LABEL?
\subby{Twenty-First Sunday after Pentecost}\label{trinity}
\fancyhead[RE,LO]{Pentecost XXI}
\begin{inhead}
{Semidouble}
\end{inhead}

\collect
\lett{L}{ord,} we beseech thee to keep thy household in continual godliness: that through thy protection it may be free from all adversities; and in good works be devoted to thy name. Through.
\begin{rubric}
    Additional Collects are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

%CHECK:
\readingcitation{Epistle}{Ephesians 6:10}
\lett{B}{rethren:} Be strong in the Lord, and in the power of his might. Put on the whole armour of God, that ye may be able to stand against the wiles of the devil. For we wrestle not against flesh and blood, but against principalities, against powers, against the rulers of the darkness of this world, against spiritual wickedness in high places. Wherefore take unto you the whole armour of God, that ye may be able to withstand in the evil day, and having done all, to stand. Stand therefore, having your loins girt about with truth, and having on the breastplate of righteousness; And your feet shod with the preparation of the gospel of peace; Above all, taking the shield of faith, wherewith ye shall be able to quench all the fiery darts of the wicked. And take the helmet of salvation, and the sword of the Spirit, which is the word of God:
%
Praying always with all prayer and supplication in the Spirit, and watching thereunto with all perseverance and supplication for all saints; And for me, that utterance may be given unto me, that I may open my mouth boldly, to make known the mystery of the gospel, For which I am an ambassador in bonds: that therein I may speak boldly, as I ought to speak.

\readingcitation{Gospel}{Matthew 18:23}
\lett{A}{t that time:} Jesus spake unto his disciples this parable: The kingdom of heaven is likened unto a certain king, which would take account of his servants. And when he had begun to reckon, one was brought unto him, which owed him ten thousand talents. But forasmuch as he had not to pay, his lord commanded him to be sold, and his wife, and children, and all that he had, and payment to be made. The servant therefore fell down, and worshipped him, saying, Lord, have patience with me, and I will pay thee all. Then the lord of that servant was moved with compassion, and loosed him, and forgave him the debt. But the same servant went out, and found one of his fellowservants, which owed him an hundred pence: and he laid hands on him, and took him by the throat, saying, Pay me that thou owest. And his fellowservant fell down at his feet, and besought him, saying, Have patience with me, and I will pay thee all. And he would not: but went and cast him into prison, till he should pay the debt. So when his fellowservants saw what was done, they were very sorry, and came and told unto their lord all that was done. Then his lord, after that he had called him, said unto him, O thou wicked servant, I forgave thee all that debt, because thou desiredst me: Shouldest not thou also have had compassion on thy fellowservant, even as I had pity on thee? And his lord was wroth, and delivered him to the tormentors, till he should pay all that was due unto him. So likewise shall my heavenly Father do also unto you, if ye from your hearts forgive not every one his brother their trespasses.

\secret
\lett{M}{ercifully} receive, O Lord, these offerings: wherewith thou hast willed that atonement should be made unto thee, and that salvation should be restored to us by the power of thy loving kindness. Through.
\begin{rubric}
    Additional Secrets are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\postcommunion
\lett{O}{Lord,} who hast bestowed upon us the food of immortality, we beseech thee: that those things which we have taken with our lips we may seek after in purity of heart. Through.
\begin{rubric}
    Additional Postcommunions are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\subby{Twenty-Second Sunday after Pentecost}
\fancyhead[RE,LO]{Pentecost XXII}
\begin{inhead}
{Semidouble}
\end{inhead}

\collect
\lett{O}{God,} our refuge and strength, who art the author of all godliness: hear the devout prayers of thy Church, and grant; that those things which we ask faithfully we may obtain effectually. Through.
\begin{rubric}
    Additional Collects are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\readingcitation{Gospel}{Matthew 22:15}
\lett{A}{t that time:} The Pharisees went, and took counsel how they might entangle Jesus in his talk. And they sent out unto him their disciples with the Herodians, saying, Master, we know that thou art true, and teachest the way of God in truth, neither carest thou for any man: for thou regardest not the person of men. Tell us therefore, What thinkest thou? Is it lawful to give tribute unto Caesar, or not? But Jesus perceived their wickedness, and said, Why tempt ye me, ye hypocrites? Shew me the tribute money. And they brought unto him a penny. And he saith unto them, Whose is this image and superscription? They say unto him, Caesar's. Then saith he unto them, Render therefore unto Caesar the things which are Caesar's; and unto God the things that are God's.

\secret
\lett{W}{e} offer thee, O Lord, this sacrifice of praise for the increase of our dutiful service: that thou wouldest mercifully accomplish that which thou hast bestowed upon us beyond our deserving. Through.
\begin{rubric}
    Additional Secrets are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\postcommunion
\lett{W}{e} beseech thee, almighty God: that we, whom thou makest to rejoice in the partaking of heavenly things, may by thee be defended against all earthly perils. Through.
\begin{rubric}
    Additional Postcommunions are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\subby{Twenty-Third Sunday after Pentecost}
\fancyhead[RE,LO]{Pentecost XXIII}
\begin{inhead}
{Semidouble}
\end{inhead}
\begin{rubric}
	If this Sunday be hindered by the supervention of the last after Pentecost, it is anticipated on Saturday with all the privileges proper to a Sunday occurring, and in it are said \emph{Gloria in excelsis}, the Nicene Creed, Preface of the Trinity, and \emph{Ite, Missa est}.
\end{rubric}

\collect
\lett{O}{Lord,} we beseech thee absolve thy people from their offences; that through thy bountiful goodness we may be delivered from the bonds of those sins, which by our frailty we have committed. Through.
\begin{rubric}
    Additional Collects are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\readingcitation{Epistle}{Philippians 3:17}
\lett{B}{rethren:} Be followers together of me, and mark them which walk so as ye have us for an ensample. (For many walk, of whom I have told you often, and now tell you even weeping, that they are the enemies of the cross of Christ: Whose end is destruction, whose God is their belly, and whose glory is in their shame, who mind earthly things.) For our conversation is in heaven; from whence also we look for the Saviour, the Lord Jesus Christ: Who shall change our vile body, that it may be fashioned like unto his glorious body, according to the working whereby he is able even to subdue all things unto himself. Therefore, my brethren dearly beloved and longed for, my joy and crown, so stand fast in the Lord, my dearly beloved. I beseech Euodias, and beseech Syntyche, that they be of the same mind in the Lord. And I intreat thee also, true yokefellow, help those women which laboured with me in the gospel, with Clement also, and with other my fellowlabourers, whose names are in the book of life.

\readingcitation{Gospel}{Matthew 9:18}
\lett{A}{t that time:} While Jesus spake unto the multitudes, Behold, there came a certain ruler, and worshipped him, saying, My daughter is even now dead: but come and lay thy hand upon her, and she shall live. And Jesus arose, and followed him, and so did his disciples. And, behold, a woman, which was diseased with an issue of blood twelve years, came behind him, and touched the hem of his garment: For she said within herself, If I may but touch his garment, I shall be whole. But Jesus turned him about, and when he saw her, he said, Daughter, be of good comfort; thy faith hath made thee whole. And the woman was made whole from that hour. And when Jesus came into the ruler's house, and saw the minstrels and the people making a noise, He said unto them, Give place: for the maid is not dead, but sleepeth. And they laughed him to scorn. But when the people were put forth, he went in, and took her by the hand, and the maid arose. And the fame hereof went abroad into all that land.

\subby{Twenty-Fourth---and last---Sunday after Pentecost}
\fancyhead[RE,LO]{Pentecost XXIV}
\begin{inhead}
{Semidouble}
\end{inhead}

\collect
\lett{S}{tir} up, we beseech thee, O Lord, the wills of thy faithful people; that they, plenteously bringing forth the fruit of good works, may by thee be plenteously rewarded. Through.
\begin{rubric}
    Additional Collects are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\readingcitation{Epistle}{Colossians 1:9}
\lett{B}{rethren:} We do not cease to pray for you, and to desire that ye might be filled with the knowledge of his will in all wisdom and spiritual understanding; That ye might walk worthy of the Lord unto all pleasing, being fruitful in every good work, and increasing in the knowledge of God; Strengthened with all might, according to his glorious power, unto all patience and longsuffering with joyfulness; Giving thanks unto the Father, which hath made us meet to be partakers of the inheritance of the saints in light: Who hath delivered us from the power of darkness, and hath translated us into the kingdom of his dear Son: In whom we have redemption through his blood, even the forgiveness of sins.

\readingcitation{Gospel}{Matthew 24:15}
\lett{A}{t that time:} Jesus said unto his disciples: When ye shall see the abomination of desolation, spoken of by Daniel the prophet, stand in the holy place, (whoso readeth, let him understand:) Then let them which be in Judaea flee into the mountains: Let him which is on the housetop not come down to take any thing out of his house: Neither let him which is in the field return back to take his clothes. And woe unto them that are with child, and to them that give suck in those days! But pray ye that your flight be not in the winter, neither on the sabbath day: For then shall be great tribulation, such as was not since the beginning of the world to this time, no, nor ever shall be. And except those days should be shortened, there should no flesh be saved: but for the elect's sake those days shall be shortened. Then if any man shall say unto you, Lo, here is Christ, or there; believe it not. For there shall arise false Christs, and false prophets, and shall shew great signs and wonders; insomuch that, if it were possible, they shall deceive the very elect. Behold, I have told you before. Wherefore if they shall say unto you, Behold, he is in the desert; go not forth: behold, he is in the secret chambers; believe it not. For as the lightning cometh out of the east, and shineth even unto the west; so shall also the coming of the Son of man be. For wheresoever the carcase is, there will the eagles be gathered together. Immediately after the tribulation of those days shall the sun be darkened, and the moon shall not give her light, and the stars shall fall from heaven, and the powers of the heavens shall be shaken: And then shall appear the sign of the Son of man in heaven: and then shall all the tribes of the earth mourn, and they shall see the Son of man coming in the clouds of heaven with power and great glory. And he shall send his angels with a great sound of a trumpet, and they shall gather together his elect from the four winds, from one end of heaven to the other. Now learn a parable of the fig tree; When his branch is yet tender, and putteth forth leaves, ye know that summer is nigh: So likewise ye, when ye shall see all these things, know that it is near, even at the doors. Verily I say unto you, This generation shall not pass, till all these things be fulfilled. Heaven and earth shall pass away, but my words shall not pass away.

\secret
\lett{B}{e} favourable, O Lord, to our supplications: and, accepting the offerings and prayers of us thy people, turn the hearts of all to thee; that we, being delivered from earthly lusts, may turn to heavenly desires. Through.
\begin{rubric}
    Additional Secrets are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\postcommunion
\lett{G}{rant} to us, we beseech thee, O Lord: that through these sacraments, which we have received, whatsoever is corrupt in our minds may be made whole by the healing power of the same. Through.
\begin{rubric}
    Additional Postcommunions are as on Pentecost III (p. \pageref{PentecostIII}).
\end{rubric}

\begin{rubric}
    If there be more than 24 Sundays after Pentecost, then the Sundays after the \nth{23} after Pentecost are supplied from the unused Sundays after Epiphany (p. \pageref{epiphany}). Then, the \nth{24} Sunday after Pentecost.
\end{rubric}
