\phantomsection
\addcontentsline{toc}{section}{Holy Week}
\fancyhead[C]{\LARGE Holy Week}

\subby{Holy Monday}
\fancyhead[RO,LE]{\textit{Holy Monday}}
\fancyhead[RE,LO]{}
\begin{inhead}
    {First Class Feria}
\end{inhead}

	\properantiphon{Ben.}{O Father, glorify thou me {\dag} with thine own self, with the glory which I had with thee before the world began.}

    \properantiphon{Mag.}{Thou couldest have no power {\dag} at all against me, except it were given thee from above.}
    
   \introit
   \lett{P}{lead} thou my cause, O Lord, with them that strive with me, and fight thou against them that fight against me: lay hand upon the shield and buckler, and stand up to help me, O Lord, thou strength of my salvation. \textit{Ps.} Bring forth the spear, and stop the way against them that persecute me: say unto my soul: I am thy salvation.
   
   \collect
	\lett{A}{lmighty} God, whose most dear Son went not up to joy but first he suffered pain, and entered not into glory before he was crucified; Mercifully grant that we, walking in the way of the cross, may find it none other than the way of life and peace. Through the same.
	
\begin{rubric}
    \nth{2} Collect is \emph{Against the Persecutors of the Church} (p. \SPAgainst) or \emph{for the Chief Bishop} (p. \SPChiefBishop).
\end{rubric}
	
\readingcitation{Epistle}{Isaiah 63:1}
\lett{I}{n those days:} Said Isaiah: Who is this that cometh from Edom, with dyed garments from Bozrah? this that is glorious in his apparel, travelling in the greatness of his strength? I that speak in righteousness, mighty to save. Wherefore art thou red in thine apparel, and thy garments like him that treadeth in the winefat? I have trodden the winepress alone; and of the people there was none with me: for I will tread them in mine anger, and trample them in my fury; and their blood shall be sprinkled upon my garments, and I will stain all my raiment. For the day of vengeance is in mine heart, and the year of my redeemed is come. And I looked, and there was none, to help; and I wondered that there was none to uphold: therefore mine own arm brought salvation unto me; and my fury it upheld me. And I will tread down the people in mine anger, and make them drunk in my fury, and I will bring down their strength to the earth. I will mention the lovingkindnesses of the Lord, and the praises of the Lord according to all that the Lord hath bestowed on us, and the great goodness toward the house of Israel, which he hath bestowed on them according to his mercies, and according to the multitude of his lovingkindnesses. For he said, Surely they are my people, children that will not lie: so he was their Saviour. In all their affliction he was afflicted, and the angel of his presence saved them: in his love and in his pity he redeemed them; and he bare them, and carried them all the days of old. But they rebelled, and vexed his holy Spirit; therefore he was turned to be their enemy, and he fought against them. Then he remembered the days of old, Moses, and his people, saying, Where is he that brought them up out of the sea with the shepherd of his flock? where is he that put his holy Spirit within him? that led them by the right hand of Moses with his glorious arm, dividing the water before them, to make himself an everlasting name? that led them through the deep, as an horse in the wilderness., that they should not stumble? As a beast goeth down into the valley, the spirit of the Lord caused him to rest: so didst thou lead thy people, to make thyself a glorious name. Look down from heaven, and behold from the habitation of thy holiness and of thy glory: where is thy zeal and thy strength, the sounding of thy bowels and of thy mercies toward me? are they restrained? Doubtless thou art our father, though Abraham be ignorant of us, and Israel acknowledge us not, thou, O Lord, art our father our redeemer; thy name is from everlasting. O Lord, why hast thou made us to err from thy ways, and hardened our heart from thy fear? Return for thy servants' sake, the tribes of thine inheritance. The people of thy holiness have possessed it but a little while: our adversaries have trodden down thy sanctuary. We are thine: thou never barest rule over them; they were not called by thy name.

\gradtr{Awake, O Lord, and stand up to judge my quarrel, avenge thou my cause, my God, and my Lord. ℣. Bring forth the spear, and stop the way against them that persecute me.}{O Lord, deal not with us after our sins which we have committed: nor reward us according to our wickednesses. ℣. Lord, remember not our old sins, but have mercy upon us, and that soon: for we are come to great misery. \inrub{Here genuflect.} ℣. Help us, O God of our salvation: and for the glory of thy name, O Lord, deliver us: and be merciful unto our sins, for thy name's sake. \inrub{Arise.}}

\readingcitation{Gospel}{Mark 14:1}
\lett{A}{t that time:} After two days was the feast of the passover, and of unleavened bread: and the chief priests and the scribes sought how they might take him by craft, and put him to death. But they said, Not on the feast day, lest there be an uproar of the people. And being in Bethany in the house of Simon the leper, as he sat at meat, there came a woman having an alabaster box of ointment of spikenard very precious; and she brake the box, and poured it on his head. And there were some that had indignation within themselves, and said, Why was this waste of the ointment made? for it might have been sold for more than three hundred pence, and have been given to the poor. And they murmured against her. And Jesus said, Let her alone; why trouble ye her? she hath wrought a good work on me. For ye have the poor with you always, and whensoever ye will ye may do them good: but me ye have not always. She hath done what she could: she is come aforehand to anoint my body to the burying. Verily I say unto you, Wheresoever this gospel shall be preached throughout the whole world, this also that she hath done shall be spoken of for a memorial of her. And Judas Iscariot, one of the twelve, went unto the chief priests, to betray him unto them. And when they heard it, they were glad, and promised to give him money. And he sought how he might conveniently betray him. And the first day of unleavened bread, when they killed the passover, his disciples said unto him, Where wilt thou that we go and prepare that thou mayest eat the passover? And he sendeth forth two of his disciples, and saith unto them, Go ye into the city, and there shall meet you a man bearing a pitcher of water: follow him. And wheresoever he shall go in, say ye to the good-man of the house, The Master saith, Where is the guestchamber, where I shall eat the passover with my disciples? And he will shew you a large upper room furnished and prepared: there make ready for us. And his disciples went forth, and came into the city, and found as he had said unto them: and they made ready the passover. And in the evening he cometh with the twelve. And as they sat and did eat, Jesus said, Verily I say unto you, One of you which eateth with me shall betray me. And they began to be sorrowful, and to say unto him one by one, Is it I? and another said, Is it I? And he answered and said unto them, It is one of the twelve, that dippeth with me in the dish. The Son of man indeed goeth, as it is written of him: but woe to that man by whom the Son of man is betrayed! good were it for that man if he had never been born. And as they did eat, Jesus took bread, and blessed, and brake it, and gave to them, and said, Take, eat; this is my body. And he took the cup, and when he had given thanks, he gave it to them: and they all drank of it. And he said unto them, This is my blood of the new testament, which is shed for many. Verily I say unto you, I will drink no more of the fruit of the vine, until that day that I drink it new in the kingdom of God. And when they had sung an hymn, they went out into the mount of Olives. And Jesus saith unto them, All ye shall be offended because of me this night: for it is written, I will smite the shepherd, and the sheep shall be scattered. But after that I am risen I will go before you into Galilee. But Peter said unto him, Although all shall be offended, yet will not I. And Jesus saith unto him, Verily I say unto thee, that this day even in this night, before the cock crow twice, thou shalt deny me thrice. But he spake the more vehemently, If I should die with thee, I will not deny thee in any wise. Likewise also said they all. And they came to a place which was named Gethsemane: and he saith to his disciples, Sit ye here, while I shall pray. And he taketh with him Peter and James and John, and began to be sore amazed, and to be very heavy; and saith unto them, My soul is exceeding sorrowful unto death: tarry ye here, and watch. And he went forward a little, and fell on the ground, and prayed that, if it were possible, the hour might pass from him. And he said, Abba, Father, all things are possible unto thee; take away this cup from me: nevertheless not what I will, but what thou wilt. And he cometh. and findeth them sleeping, and saith unto Peter, Simon, sleepest thou? couldest not thou watch one hour? Watch ye and pray, lest ye enter into temptation. The spirit truly is ready, but the flesh is weak. And again lie went away, and prayed, and spake the same words. And when ho returned, he found them asleep again, (for their eyes were heavy,) neither wist they what to answer him. And he cometh the third time, and saith unto them, Sleep on now, and take your rest: it is enough, the hour is come; behold, the Son of man is betrayed into the hands of sinners. Rise up, let us go; lo, he that betrayeth me is at hand. And immediately, while he yet spake, cometh Judas, one of the twelve, and with him a great multitude with swords and staves, from the chief priests and the scribes and the elders. And he that betrayed him had given them a token, saying, Whomsoever I shall kiss, that same is he; take him, and lead him away safely. And as soon as he was come, he goeth straightway to him, and saith, Master, master; and kissed him. And they laid their hands on him, and took him. And one of them that stood by drew a sword, and smote a servant of the high priest, and cut off his ear. And Jesus answered and said unto them, Are ye come out, as against a thief, with swords and with staves, to take me? I was daily with you in the temple teaching, and ye took me not: but the scriptures must be fulfilled. And they all forsook him, and fled. And there followed him a certain young man, having a linen cloth cast about his naked body; and the young men laid hold on him: and he left the linen cloth, and fled from them naked. And they led Jesus away to the high priest: and with him were assembled all the chief priests and the elders and the scribes. And Peter followed him afar off, even into the palace of the high priest: and he sat with the servants, and warmed himself at the fire. And the chief priests and all the council sought for witness against Jesus to put him to death; and found none. For many bare false witness against him, but their witness agreed not together. And there arose certain, and bare false witness against him, saying, We heard him say, I will destroy this temple that is made with hands, and within three days I will build another made without hands. But neither so did their witness agree together. And the high priest stood up in the midst, and asked Jesus, saying, Answerest thou nothing? what is it which these witness against thee? But he held his peace, and answered nothing. Again the high priest asked him, and said unto him, Art thou the Christ, the Son of the Blessed? And Jesus said, I am: and ye shall see the Son of man sitting on the right hand of power, and coming in the clouds of heaven. Then the high priest rent his clothes, and saith, What need we any further witnesses? ye have heard the blasphemy: what think ye? And they all condemned him to be guilty of death. And some began to spit on him, and to cover his face, and to buffet him, and to say unto him, Prophesy: and the servants did strike him with the palms of their hands. And as Peter was beneath in the palace, there cometh one of the maids of the high priest: and when she saw Peter warming himself, she looked upon him, and said, And thou also wast with Jesus of Nazareth. But he denied, saying, I know not, neither understand I what thou sayest. And he went out into the porch; and the cock crew. And a maid saw him again, and began to say to them that stood by, This is one of them. And he denied it again. And a little after, they that stood by said again to Peter, Surely thou art one of them: for thou art a Galilean, and thy speech agreeth thereto. But he began to curse and to swear, saying, I know not this man of whom ye speak. And the second time the cock crew. And Peter called to mind the word that Jesus said unto him, Before the cock crow twice, thou shalt deny me thrice. And when he thought thereon, he wept.

\offertory{Deliver me, O Lord, from mine enemies: for I flee unto thee to hide me, teach me to do the thing that pleaseth thee: for thou art my God.}

\secret
\lett{A}{lmighty} God, let these sacrifices cleanse us by their effectual power, that we may thereby attain in purity unto the author of the same. Through.

\begin{rubric}
    \nth{2} Secret is \emph{Against the Persecutors of the Church} (p. \SPAgainst) or \emph{for the Chief Bishop} (p. \SPChiefBishop).
\end{rubric}

\vspace{-1ex}

\communion{Let them be put to confusion and shame together, that rejoice at my trouble: let them be clothed with rebuke and dishonour that boast themselves against me.}

\vspace{-1ex}

\postcommunion
\lett{M}{ay} thy holy mysteries, O Lord, impart unto us heavenly fervour: that we may delight in their performance, and likewise in bringing forth the fruit thereof. Through.

\begin{rubric}
    \nth{2} Postcommunion is \emph{Against the Persecutors of the Church} (p. \SPAgainst) or \emph{for the Chief Bishop} (p. \SPChiefBishop).
\end{rubric}

\textsc{Priest.} Let us pray.\par
\textsc{Deacon.} Humble your heads before God.\par
\lett{H}{elp} us, O God of our salvation: and grant, that we may come with joy to the celebration of those benefits by which thou hast vouchsafed to redeem us. Through.


\subby{Holy Tuesday}
\fancyhead[RO,LE]{\textit{Holy Tuesday}}
\fancyhead[RE,LO]{}
\begin{inhead}
    {First Class Feria}
\end{inhead}

	\properantiphon{Ben.}{Now before the feast {\dag} of the Passover, when Jesus knew that his hour was come, having loved his own, he loved them unto the end.}

    \properantiphon{Mag.}{I have power {\dag} to lay down my life, and I have power to take it again.}

\vspace{-1ex}

\introit
\lett{B}{ut} it behoveth us to glory in the Cross of our Lord Jesus Christ: in whom is our salvation, life, and resurrection: by whom we are saved and set free. \textit{Ps.} God be merciful unto us, and bless us: and shew us the light of his countenance, and be merciful unto us.

\collect
	\lett{O}{Lord} God, whose blessed Son, our Saviour, gave his back to the smiters and hid not his face from shame; Grant us grace to take joyfully the sufferings of the present time, in full assurance of the glory that shall be revealed. Through the same.
	
	\begin{rubric}
    \nth{2} Collect is \emph{Against the Persecutors of the Church} (p. \SPAgainst) or \emph{for the Chief Bishop} (p. \SPChiefBishop).
	\end{rubric}
	
	\readingcitation{Epistle}{Isaiah 50:5}
	\lett{I}{n those days:} Said Isaiah: The Lord \divineName{God} hath opened mine ear, and I was not rebellious, neither turned away back. I gave my back to the smiters, and my cheeks to them that plucked off the hair: I hid not my face from shame and spitting. For the Lord God will help me; therefore shall I not be confounded: therefore have I set my face like a flint and I know that I shall not be ashamed. He is near that justifieth me; who will contend with me? let us stand together: who is mine adversary? let him come near to me. Behold, the Lord God will help me; who is he that shall condemn me? lo, they all shall wax old as a garment; the moth shall eat them up. Who is among you that feareth the Lord that obeyeth the voice of his servant, that walketh in darkness, and hath no light? let him trust in the name of the Lord, and stay upon his God. Behold, all ye that kindle a fire, that compass yourselves about with sparks: walk in the light of your fire, and in the sparks that ye have kindled. This shall ye have of mine hand; ye shall lie down in sorrow.
	
	\gradual{Nevertheless, when they were sick, I put on sackcloth, and humbled my soul with fasting: and my prayer shall turn into mine own bosom. ℣. Plead thou my cause, O Lord, with them that strive with me, and fight thou against them that fight against me: lay hand upon the shield and buckler, and stand up to help me.}
	
	\readingcitation{Gospel}{Mark 15:1}
	\lett{A}{t that time:} Straightway in the morning the chief priests held a consultation with the elders and scribes and the whole council, and bound Jesus, and carried him away, and delivered him to Pilate. And Pilate asked him, Art thou the King of the Jews? And he answering said unto him, Thou sayest it. And the chief priests accused him of many things: but he answered nothing. And Pilate asked him again, saying, Answerest thou nothing? behold how many things they witness against thee. But Jesus yet answered nothing; so that Pilate marvelled. Now at that feast he released unto them one prisoner, whomsoever they desired. And there was one named Barabbas, which lay bound with them that had made insurrection with him, who had committed murder in the insurrection. And the multitude crying aloud began to desire him to do as he had ever done unto them. But Pilate answered them, saying, Will ye that I release unto you the King of the Jews? For he knew that the chief priests had delivered him for envy. But the chief priests moved the people, that he should rather release Barabbas unto them. And Pilate answered and said again unto them, What will ye then that I shall do unto him whom ye call the King of the Jews? And they cried out again, Crucify him. Then Pilate said unto them, Why, what evil hath he done? And they cried out the more exceedingly, Crucify him, And so Pilate, willing to content the people, released Barabbas unto them, and delivered Jesus, when he had scourged him, to be crucified. And the soldiers led him away into the hall, called Pr{\ae}torium; and. they called together the whole band. And they clothed him with purple, and platted a crown of thorns, and put it about his head, and began to salute him, Hail, King of the Jews! And they smote him on the head with a reed, and. did spit upon him, and bowing their knees worshipped him. And when they had mocked. him, they took off the purple from him, and put his own clothes on him, and led him out to. crucify him. And they compel one Simon a Cyrenian, who passed by, coming out of the country, the father of Alexander and Rufus, to bear his cross. And they bring him unto the place Golgotha, which is, being interpreted, The place of a skull. And they gave him to drink wine mingled with myrrh: but he received it not. And when they had crucified him, they parted his garments, casting lots upon them, what every man should take. And it was the third hour, and they crucified him. And the superscription of his accusation was written over, \divineName{the King of the Jews}. And with him they crucify two thieves; the one on his right hand, and the other on his left. And the scripture was fulfilled, which saith, And he was numbered with the transgressors. And they that passed by railed on him, wagging their heads, and saying, Ah, thou that destroyest the temple, and buildest it in three days, save thy self, and come down from the cross. Likewise also the chief priests mocking said among them selves with the scribes, He saved others; himself he cannot save. Let Christ the King of Israel descend now from the cross, that we may see and believe. And they that were crucified with him reviled him. And when the sixth hour was come, there was darkness over the whole land until the ninth hour. And at the ninth hour Jesus cried with a loud voice, saying, Eloi, Eloi, lama sabachthani? which is, being interpreted, My God, my God, why hast thou forsaken me? And some of them that stood by, when they heard it, said, Behold, he calleth Elias. And one ran and filled a spunge full of vinegar, and put it on a reed, and gave him to drink, saying, Let alone; let us see whether Elias will come to take him down. And Jesus cried with a loud voice, and gave up the ghost. \inrub{Here genuflect, and pause awhile.} And the veil of the temple was rent in twain from the top to the bottom. And when the centurion, which stood over against him, saw that he so cried out, and gave up the ghost, he said, Truly this man was the Son of God.
	
	\offertory{Keep me, O Lord, from the hands of the ungodly: preserve me from the wicked men.}
	
	\secret
	\lett{W}{e} beseech thee, O Lord: that these sacrifices which we offer unto thee in this time of fasting, ordained for the healing of our souls, may effectually renew us. Through.

	\begin{rubric}
    \nth{2} Secret is \emph{Against the Persecutors of the Church} (p. \SPAgainst) or \emph{for the Chief Bishop} (p. \SPChiefBishop).
	\end{rubric}
	
	\communion{They that sit in the gate speak against me: and the drunkards make songs upon me: but, Lord, I make my prayer unto thee: in an acceptable time, O God, in the multitude of thy mercy.}
	
	\postcommunion
	\lett{A}{lmighty} God, let thy holy mysteries both heal our vices and bestow upon us everlasting remedies. Through.
	
	\begin{rubric}
    \nth{2} Postcommunion is \emph{Against the Persecutors of the Church} (p. \SPAgainst) or \emph{for the Chief Bishop} (p. \SPChiefBishop).
	\end{rubric}
	
	\textsc{Priest.} Let us pray.\par
	\textsc{Deacon.} Humble your heads before God.\par
	\lett{L}{et} thy merciful kindness, O God, cleanse us throughly from all the corruption of our former nature, and make us to be effectually renewed unto holiness of life. Through.


\subby{Spy Wednesday}
\fancyhead[RO,LE]{\textit{Spy Wednesday}}
\fancyhead[RE,LO]{}
\begin{inhead}
    {First Class Double}
\end{inhead}

	\properantiphon{Ben.}{Simon, sleepest thou? {\dag} couldest not thou watch with me one hour?}

    \properantiphon{Mag.}{A maid said unto Peter, {\dag} Surely thou art one of them, for thy speech betrayeth thee.}

\introit
\lett{A}{t} the name of Jesus every knee should bow, of things in heaven, and things in earth, and things under the earth: for the Lord became obedient unto death, even the death of the cross: wherefore Jesus Christ is Lord, to the glory of God the Father. \textit{Ps.} Hear my prayer, O Lord: and let my crying come unto thee.

\begin{rubric}
    After the Kyrie,
\end{rubric}
\elcol{℣. Let us pray.}{℣. Orémus.}
\elcol{℣. Let us bow the knee.}{℣. Flectámus génua.}
\elcol{℣. Arise.}{℣. Leváte.}

\collect
\lett{G}{rant,} we beseech thee, almighty God: that we, who are continually afflicted by reason of our transgressions, may be delivered by the passion of thine only-begotten Son. Who liveth.

\readingcitation{Epistle}{Isaiah 62:11}
\lett{T}{hus saith the Lord God:} Say ye to the daughter of Zion, Behold, thy salvation cometh; behold, his reward is with him. Who is this that cometh from Edom, with dyed garments from Bozrah? this that is glorious in his apparel, travelling in the greatness of his strength? I that speak in righteousness, mighty to save. Wherefore art thou red in thine apparel, and thy garments like him that treadeth in the winefat? I have trodden the winepress alone; and of the people there was none with me: for I will tread them in mine anger, and trample them in my fury; and their blood shall be sprinkled upon my garments, and I will stain all my raiment. For the day of vengeance is in mine heart, and the year of my redeemed is come. And I looked, and there was none to help; and I wondered that there was none to uphold: therefore mine own arm brought salvation unto me; and my fury, it upheld me. And I will tread down the people in mine anger, and make them drunk in my fury, and I will bring down their strength to the earth. I will mention the lovingkindnesses of the \divineName{Lord}, and the praises of the \divineName{Lord}, according to all that the \divineName{Lord} hath bestowed on us.

\gradual{Hide not thy face from thy servant, for I am in trouble: O haste thee, and hear me. ℣. Save me, O God, for the waters are come in, even unto my soul: I stick fast in the deep mire, where no ground is.}

\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}
\elcol{℣. Let us pray.}{℣. Orémus.}

\collect
\lett{A}{ssist} us mercifully with thy help, O Lord God of our salvation; that we may enter with joy upon the meditation of those mighty acts, whereby thou hast given unto us life and immortality. Through.
	
	\begin{rubric}
    \nth{2} Collect is \emph{Against the Persecutors of the Church} (p. \SPAgainst) or \emph{for the Chief Bishop} (p. \SPChiefBishop).
	\end{rubric}

\readingcitation{Epistle}{Hebrews 9:16}
\lett{B}{rethren:} Where a testament is, there must also of necessity be the death of the testator. For a testament is of force after men are dead: otherwise it is of no strength at all while the testator liveth. Whereupon neither the first testament was dedicated without blood. For when Moses had spoken every precept to all the people according to the law, he took the blood of calves and of goats, with water, and. scarlet wool, and hyssop, and sprinkled both the book and all the people, saying, This is the blood of the testament which God hath enjoined unto you. Moreover, he sprinkled likewise with blood both the tabernacle, and all the vessels of the ministry. And almost all things are by the law purged with blood; and without shedding of blood is no remission. It was therefore necessary that the patterns of things in the heavens should be purified with these; but the heavenly things themselves with better sacrifices than these. For Christ is not entered into the holy places made with hands, which are the figures of the true; but into heaven itself, now to appear in the presence of God for us: nor yet that he should offer himself often, as the high priest entereth into the holy place every year with blood of others; for then must he often have suffered since the foundation of the world: but now once in the end of the world hath he appeared to put away sin by the sacrifice of himself. And as it is appointed unto men once to die, but after this the judgment: so Christ was once offered to bear the sins of many; and unto them that look for him shall he appear the second time without sin unto salvation.

\tract{Hear my prayer, O Lord, and let my crying come unto thee. ℣. Hide not thy face from me in the time of my trouble. ℣. Incline thine ear unto me when I call, O hear me, and that right soon. ℣. For my days are consumed away like smoke: and my bones are burnt up as it were a fire-brand. ℣. My heart is smitten down, and withered like grass: so that I forget to eat my bread. ℣. Thou shalt arise, O Lord, and have mercy upon Sion: for it is time that thou have mercy upon her.}

\readingcitation{Gospel}{Luke 22:1}
\lett{A}{t that time:} The feast of unleavened bread drew nigh, which is called the Passover. And the chief priests and scribes sought how they might kill him; for they feared the people. Then entered Satan into Judas surnamed Iscariot, being of the number of the twelve. And he went his way, and communed with the chief priests and captains, how he might betray him unto them. And they were glad, and covenanted to give him money. And he promised, and sought opportunity to betray him unto them in the absence of the multitude. Then came the day of unleavened bread, when the passover must be killed. And he sent Peter and John, saying, Go and prepare us the passover, that we may eat. And they said unto him, Where wilt thou that we prepare? And he said unto them, Behold, when ye are entered into the city, there shall a man meet you, bearing a pitcher of water; follow him into the house where he entereth in. And ye shall say unto the goodman of the house, The Master saith unto thee, Where is the guest-chamber, where I shall eat the passover with my disciples? And he shall shew you a large upper room furnished: there make ready. And they went, and found as he had said unto them: and they made ready the passover. And when the hour was come, he sat down, and the twelve apostles with him. And he said unto them, With desire I have desired to eat this passover with you before I suffer: for I say unto you, I will not any more eat thereof, until it be fulfilled in the kingdom of God. And he took the cup, and gave thanks, and said, Take this, and divide it among yourselves: for I say unto you I will not drink of the fruit of the vine, until the kingdom of God shall come. And he took bread, and gave thanks, and brake it, and gave unto them, saying, This is my body which is given for you: this do in remembrance of me. Likewise also the cup after supper, saying, This cup is the new testament in my blood, which is shed for you. But, behold, the hand of him that betrayeth me is with me on the table. And truly the Son of man goeth, as it was determined: but woe unto that man by whom he is betrayed! And they began to enquire among themselves, which of them it was that should do this thing. And there was also a strife among them, which of them should be accounted the greatest And he said unto them, The kings of the Gentiles exercise lordship over them; and they that exercise authority upon them are called benefactors. But ye shall not be so: but he that is greatest among you, let him be as the younger; and he that is chief, as he that doth serve. For whether is greater, he that sitteth at meat, or he that serveth? is not he that sitteth at meat? but I am among you as he that serveth. Ye are they which have continued with me in my temptations. And I appoint unto you a kingdom, as my Father hath appointed unto me; that ye may eat and drink at my table in my kingdom. and sit on thrones, judging the twelve tribes of Israel. And the Lord said, Simon, Simon, behold, Satan hath desired to have you, that he may sift you as wheat: but I have prayed for thee, that thy faith fail not; and when thou art converted, strengthen thy brethren. And he said unto him, Lord, I am ready to go with thee, both into prison, and to death. And he said, I tell thee, Peter, the cock shall not crow this day, before that thou shalt thrice deny that thou knowest me. And he said unto them, When I sent you without purse, and scrip, and shoes, lacked ye any thing? And they said, Nothing. Then said he unto them, But now, he that hath a purse, let him take it, and likewise his scrip: and he that hath no sword, let him sell his garment, and buy one. For I say unto you, that this that is written must yet be accomplished in me, And he was reckoned among the transgressors: for the things concerning me have an end. And they said, Lord, behold, here are two swords. And he said unto them, It is enough. And he came out, and went, as he was wont, to the mount of Olives; and his disciples also followed him. And when he was at the place, he said unto them, Pray that ye enter not into temptation. And he was withdrawn from them about a stone's cast, and kneeled down, and prayed, saying, Father, if thou be willing, remove this cup from me: nevertheless not my will, but thine, be done. And there appeared an angel unto him from heaven, strengthening him. And being in an agony he prayed more earnestly: and his sweat was as it were great drops of blood falling down to the ground. And when he rose up from prayer, and was come to his disciples, he found them sleeping for sorrow, and said unto them, Why sleep ye? rise and pray, lest ye enter into temptation. And while he yet spake behold a multitude, and he that was called Judas, one of the twelve, went before them, and drew near unto Jesus to kiss him. But Jesus said unto him. Judas, betrayest thou the Son of man with a kiss? When they which were about him saw what would follow, they said unto him, Lord, shall we smite with the sword? And one of them smote the servant of the high priest, and cut off his right ear. And Jesus answered and said, Suffer ye thus far. And he touched his ear, and healed him. Then Jesus said unto the chief priests, and captains of the temple, and the elders, which were come to him, Be ye come out, as against a thief, with swords and staves? When I was daily with you in the temple, ye stretched forth no hands against me: but this is your hour, and the power of darkness. Then took they him, and led him, and brought him into the high priest's house. And Peter followed afar off. And when they had kindled a fire in the midst of the hall, and were set down together, Peter sat down among them. But a certain maid beheld him as he sat by the fire, and earnestly looked upon him, and said, This man was also with him. And he denied him, saying, Woman, I know him not. And after a little while another saw him, and said, Thou art also of them. And Peter said, Man, I am not. And about the space of one hour after another confidently affirmed, saying, Of a truth this fellow also was with him: for he is a Galilean. And Peter said, Man, I know not what thou sayest. And immediately, while he yet spake, the cock crew. And the Lord turned, and looked upon Peter. And Peter remembered the word of the Lord, how he had said unto him, Before the cock crow, thou shalt deny me thrice. And Peter went out, and wept bitterly. And the men that held Jesus mocked him, and smote him. And when they had blindfolded him, they struck him on the face, and asked him, saying, Prophesy, who is it that smote thee? And many other things blasphemously spake they against him. And as soon as it was day the elders of the people and the chief priests and the scribes came together, and led him into their council, saying, Art thou the Christ? tell us. And he said unto them, If I tell you, ye will not believe: and if I also ask you, ye will not answer me, nor let me go. Hereafter shall the Son of man sit on the right hand of the power of God. Then said they all, Art thou then the Son of God? And he said unto them, Ye say that I am. And they said, What need we any further witness? for we ourselves have heard of his own mouth.

\offertory{Hear my prayer, O Lord, and let my crying come unto thee: hide not thy face from me.}

\secret
\lett{A}{ccept,} we beseech thee, O Lord, the gift which we offer, and vouchsafe so to work in us: that as we set forth in a mystery the passion of thy Son our Lord, so by devout affection we may attain unto the same. Through the same.

	\begin{rubric}
    \nth{2} Secret is \emph{Against the Persecutors of the Church} (p. \SPAgainst) or \emph{for the Chief Bishop} (p. \SPChiefBishop).
	\end{rubric}

\communion{I have mingled my drink with weeping: for thou hast taken me up, and cast me down: and I am withered like grass: but thou, O Lord, shalt endure for ever: thou shalt arise, and have mercy upon Sion, for it is time that thou have mercy upon her.}

\postcommunion
\lett{A}{lmighty} God, enlighten our understanding: that, as these wondrous mysteries do testify to the temporal death of thy Son, so thereby we may have assurance that thou hast bestowed upon us life eternal. Through the same.

	\begin{rubric}
    \nth{2} Postcommunion is \emph{Against the Persecutors of the Church} (p. \SPAgainst) or \emph{for the Chief Bishop} (p. \SPChiefBishop).
	\end{rubric}
	
	\textsc{Priest.} Let us pray.\par
	\textsc{Deacon.} Humble your heads before God.\par
	\lett{A}{lmighty} God, we beseech thee graciously to behold this thy family, for which our Lord Jesus Christ was contented to be betrayed, and given up into the hands of wicked men, and to suffer death upon the Cross. Who liveth.
	

\subby{Maundy Thursday}
\fancyhead[RO,LE]{\textit{Maundy Thursday}}
\fancyhead[RE,LO]{}
\begin{inhead}
    {First Class Double}
\end{inhead}

    \properantiphon{Ben.}{Now he that betrayed him {\dag} gave them a sign, saying: Whomsoever I shall kiss, that same is he; hold him fast.}

    \properantiphon{Mag.}{As they were eating, {\dag} Jesus took bread, and blessed it, and brake it, and gave it to his disciples.}
    
    \introit
    \lett{B}{ut} it behoveth us to glory in the Cross of our Lord Jesus Christ: in whom is our salvation, life, and resurrection: by whom we are saved, and set free. \textit{Ps.} God be merciful unto us, and bless us: and shew us the light of his countenance, and be merciful unto us.
    
    \begin{rubric}
    	\emph{Gloria in excelsis} is said, and after the intonation thereof the bells are rung: and, the Hymn being ended, they are silent until Holy Saturday, as is there noted.
    \end{rubric}
    
    \collect
    \lett{A}{lmighty} Father, whose dear Son, on the night before he suffered, did institute the Sacrament of his Body and Blood; Mercifully grant that we may thankfully receive the same in remembrance of him, who in these holy mysteries giveth us a pledge of life eternal; through the same thy Son Jesus Christ our Lord, who now liveth and reigneth with thee and the Holy Spirit ever, one God, world without end. \textit{Amen.}
    
    \readingcitation{Epistle}{1 Corinthians 11:17}
    \lett{B}{rethren:} In this that I declare unto you I praise you not, that ye come together not for the better, but for the worse. For first of all, when ye come together in the church, I hear that there be divisions among you; and I partly believe it. For there must be also heresies among you, that they which are approved may be made manifest among you. When ye come together therefore into one place, this is not to eat the Lord's supper. For in eating every one taketh before other his own supper: and one is hungry, and another is drunken. What? have ye not houses to eat and to drink in? or despise ye the church of God, and shame them that have not? What shall I say to you? shall I praise you in this? I praise you not. For I have received of the Lord that which also I delivered unto you, That the Lord Jesus the same night in which he was betrayed took bread: and when he had given thanks, he brake it, and said, Take, eat: this is my body, which is broken for you: this do in remembrance of me After the same manner also he took the cup, when he had supped, saying, This cup is the new testament in my blood: this do ye, as oft as ye drink it, in remembrance of me. For as often as ye eat this bread, and drink this cup, ye do shew the Lord's death till he come. Wherefore whosoever shall eat this bread, and drink this cup of the Lord, unworthily, shall be guilty of the body and blood of the Lord. But let a man examine himself, and so let him eat of that bread, and drink of that cup. For he that eateth and drinketh unworthily, eateth and drinketh damnation to himself, not discerning the Lord's body. For this cause many are weak and sickly among you, and many sleep. For if we would judge ourselves, we should not be judged. But when we are judged, we are chastened of the Lord, that we should not be condemned with the world. Wherefore, my brethren, when ye come together to eat, tarry one for another. And if any man hunger, let him eat at home; that ye come not together unto condemnation. And the rest will I set in order when I come.
    
    \gradual{Christ for us became obedient unto death, even the death of the cross. ℣. Wherefore God also hath highly exalted him: and given him a name which is above every name.}
    
    \readingcitation{Gospel}{Luke 23:1}
    \lett{A}{t that time:} The whole multitude of them arose, and led him unto Pilate. And they began to accuse him, saying, We found this fellow perverting the nation, and forbidding to give tribute to Cæsar, saying that he himself is Christ a King. And Pilate asked him, saying, Art thou the King of the Jews? And he answered him and said, Thou sayest it. Then said Pilate to the chief priests and to the people, I find no fault in this man. And they were the more fierce, saying, He stirreth up the people, teaching through-out all Jewry, beginning from Galilee to this place. When Pilate heard of Galilee, he asked whether the man were a Galilean. And as soon as he knew that he belonged unto Herod's jurisdiction, he sent him to Herod, who himself also was at Jerusalem at that time. And when Herod saw Jesus, he was exceeding glad: for he was desirous to see him of a long season, because he had heard many things of him; and he hoped to have seen some miracle done by him. Then he questioned with him in many words; but he answered him nothing. And the chief priests and scribes stood and vehemently accused him. And Herod with his men of war set him at nought, and mocked him, and arrayed him in a gorgeous robe, and sent him again to Pilate. And the same day Pilate and Herod were made friends together: for before they were at enmity between themselves. And Pilate, when he had called together the chief priests and the rulers and the people, said unto them, Ye have brought this man unto me, as one that perverted the people: and, behold, I, having examined him before you, have found no fault in this man touching those things whereof ye accuse him: no, nor yet Herod: for I sent you to him; and, lo, nothing worthy of death is done unto him. I will therefore chastise him, and release him. (For of necessity he must release one unto them at the feast.) And they cried out all at once, saying, Away with this man, and release unto us Barabbas: (who for a certain sedition made in the city, and for murder, was cast into prison.) Pilate therefore, willing to release Jesus, spake again to them. But they cried, saying, Crucify him, crucify him. And he said unto them the third time, Why, what evil hath he done? I have found no cause of death in him: I will therefore chastise him, and let him go. And they were instant with loud voices, requiring that he might be crucified. And the voices of them and of the chief priests prevailed. And Pilate gave sentence that it should be as they required. And he released unto them him that for sedition and murder was cast into prison, whom they had desired; but he delivered Jesus to their will. And as they led him away, they laid hold upon one Simon, a Cyrenian, coming out of the country, and on him they laid the cross, that he might bear it after Jesus. And there followed him a great company of people, and of women, which also bewailed and lamented him. But Jesus turning unto them said, Daughters of Jerusalem, weep not for me, but weep for yourselves, and for your children. For, behold, the days are coming, in the which they shall say, Blessed are the barren, and the wombs that never bare, and the paps which never gave suck. Then shall they begin to say to the mountains, Fall on us; and to the hills, Cover us. For if they do these things in a green tree, what shall be done in the dry? And there were also two other, malefactors, led with him to be put to death. And when they were come to the place which is called Calvary, there they crucified him, and the malefactors, one on the right hand, and the other on the left. Then said Jesus, Father, forgive them; for they know not what they do. And they parted his raiment, and cast lots. And the people stood beholding. And the rulers also with them derided him, saying, He saved others; let him save himself, if he be Christ, the chosen of God. And the soldiers also mocked him, coming to him, and offering him vinegar, and saying, If thou be the king of the Jews, save thyself. And a superscription also was written over him, in letters of Greek, and Latin, and Hebrew, \textsc{This is the King of the Jews}. And one of the malefactors which were hanged railed on him, saying, If thou be Christ, save thyself and us. But the other answering rebuked him, saying, Dost not thou fear God, seeing thou art in the same condemnation? And we indeed justly; for we receive the due reward of our deeds: but this man hath done nothing amiss. And he said unto Jesus, Lord, remember me when thou comest into thy kingdom. And Jesus said unto him, Verily I say unto thee, To day shalt thou be with me in paradise. And it was about the sixth hour, and there was a darkness over all the earth until the ninth hour. And the sun was darkened, and the veil of the temple was rent in the midst. And when Jesus had cried with a loud voice, he said, Father, into thy hands I commend my spirit: and having said thus, he gave up the ghost. \inrub{Here genuflect, and pause awhile.} Now when the centurion saw what was done, he glorified God, saying, Certainly this was a righteous man. And all the people that came together to that sight, beholding the things which were done, smote their breasts, and returned. And all his acquaintance, and the women that followed him from Galilee, stood afar off, beholding these things.
    
    \offertory{The right hand of the Lord bringeth mighty things to pass, the right of the Lord hath exalted me: I shall not die, but live, and declare the works of the Lord.}
    
    %MANUAL ADJUSTMENT:
    \clearpage
    \secret
    \lett{W}{e} beseech thee, O Lord holy, Father almighty, everlasting God: that he may render our sacrifice acceptable unto thee, who on this day commanded and taught his disciples to do this in remembrance of him, even Jesus Christ thy Son our Lord. Who liveth.
    
    \begin{rubric}
    	After the Comfortable Words, the \emph{Pax} is not given.
    \end{rubric}
    
    \begin{rubric}
    	\textsc{Note,} To-day the Priest consecrates two Hosts, of which he consumes one, and reserves the other (touched with the Blood) for the following day, on which the Sacrament is not consecrated; he also reserves some consecrated particles, in case of need, for the sick.\par
    	But be consumes the whole of the Blood, and before washing his fingers he places the reserved Host in another Chalice, which the Deacon covers with the Pall and the Paten inverted, and spreads a Veil over it, and places it in the midst of the Altar.\par
    	Then the Communion is said and the Mass finished. The Priest genuflects, however, whenever he approaches or departs from the middle of the Altar, or crosses before the Sacrament reserved in the Chalice: and when he is to say \emph{The Lord be with you}, he does not turn to the people in the middle of the Altar, lest be turn his back on the Sacrament, but from the Gospel side: and at the end he gives the blessing from the same place and does not complete the circle.
    \end{rubric}
    
    \communion{The Lord Jesus, after he had supped with his disciples, washed their feet, and said to them: Know ye what I, your Lord and Master, have done to you? I have given you an example, that ye also should so do.}
    
    \postcommunion
    \lett{O}{Lord} our God, who hast refreshed us with life-giving sustenance, we beseech thee: that we who observe this institution in the time of our mortal life, may obtain the benefit thereof in thy gift of immortality. Through.
    
    \begin{rubric}
    	\emph{Ite, Missa est} is said, and the Blessing is given, and the Gospel of St. John is read, at the beginning of which the Priest does not sign the Altar, but himself only.
    \end{rubric}
    
    \subby{Altar of Repose \& Stripping of the Altar}
    
    \begin{rubric}
    	To-day a suitable place is to be prepared in some Chapel of the Church, or Altar: and adorned as decently as may be, with veils and lights: where the Chalice with the Host, reserved as above, shall be set. When the Mass is over, torches are lighted, and a Procession is formed in the usual manner, but with another Subdeacon vested carrying the Cross. The Celebrant, vested in a white Cope, standing before the Altar, places incense in two thuribles without a blessing: then kneeling in the middle, with one he censes the Sacrament thrice: and receiving the Chalice with the Sacrament from the hand of the Deacon, who stands, and having it covered with the ends of the Veil with which his shoulders are covered, he proceeds between the same Deacon on his right, and the Subdeacon on his left, under a baldachino with two Acolytes continually censing the Sacrament, to the place prepared, where it is to be kept for the morrow. Meanwhile, during the Procession, the following hymn is sung.
    \end{rubric}
    
    \subbysub{Pange Lingua}
    
    \gregorioscore{resources/gabc/PangeLingua.gabc}
    
    
    \begin{rubric}
    	When they arrive at the place prepared, the Deacon kneels down and receives from the Priest, who stands, the Chalice with the Sacrament and places it first upon the Altar, where it is censed by the Priest kneeling, as above: then be places it in the receptacle.\par
    	Afterwards Evensong is said in the Choir without singing. Which being finished, the Priest, vested in an Alb and violet stole, with the Ministers shall lay bare the Altars, reading the following:
    \end{rubric}

    \subbysub{Psalm 22. \textit{Deus, Deus meus}}
    
        \antiphon{}{They parted my garments among them: and upon my vesture did they cast lots.}

\lett{M}{y} God, my God, look upon me; why hast thou forsaken me : and art so far from my health, and from the words of my complaint?\par
\secondline{\psanum{2}O my God, I cry in the day-time, but thou hearest not : and in the night-season also I take no rest.}
\thirdline{\psanum{3}And thou continuest holy : O thou worship of Israel.}
\psanum{4}Our fathers hoped in thee : they trusted in thee, and thou didst deliver them.\par
\psanum{5}They called upon thee, and were holpen : they put their trust in thee, and were not confounded.\par
\psanum{6}But as for me, I am a worm, and no man : a very scorn of men, and the outcast of the people.\par
\psanum{7}All they that see me laugh me to scorn : they shoot our their lips, and shake their heads, saying,\par
\psanum{8}He trusted in God, that he would deliver him : let him deliver him, if he will have him.\par
\psanum{9}But thou art he that took me out of my mother's womb : thou wast my hope, when I hanged yet upon my mother's breasts.\par
\psanum{10}I have been left unto thee ever since I was born : thou art my God, even from my mother's womb.\par
\psanum{11}O go not from me, for trouble is hard at hand : and there is none to help me.\par
\psanum{12}Many oxen are come about me : fat bulls of Basan close me in on every side.\par
\psanum{13}They gape upon me with their mouths : as it were a ramping and a roaring lion.\par
\psanum{14}I am poured out like water, and all my bones are out of joint : my heart also in the midst of my body is even like melting wax.\par
\psanum{15}My strength is dried up like a potsherd, and my tongue cleaveth to my gums : and thou shalt bring me into the dust of death.\par
\psanum{16}For many dogs are come about me : and the council of the wicked layeth siege against me.\par
\psanum{17}They pierced my hands and my feet; I may tell all my bones : they stand staring and looking upon me.\par
\psanum{18}They part my garments among them : and casts lots upon my vesture.\par
\psanum{19}But be not thou far from me, O Lord : thou art my succour, haste thee to help me.\par
\psanum{20}Deliver my soul from the sword : my darling from the power of the dog.\par
\psanum{21}Save me from the lion's mouth : thou hast heard me also from among the horns of the unicorns.\par
\psanum{22}I will declare thy Name unto my brethren : in the midst of the congregation will I praise thee.\par
\psanum{23}O praise the Lord, ye that fear him : magnify him, all ye of the seed of Jacob, and fear him, all ye seed of Israel.\par
\psanum{24}For he hath not despised, nor abhorred, the low estate of the poor : he hath not hid his face from him, but when he called unto him he heard him.\par
\psanum{25}My praise is of thee in the great congregation : my vows will I perform in the sight of them that fear him.\par
\psanum{26}The poor shall eat and be satisfied : they that seek after the Lord shall praise him; your heart shall live for ever.\par
\psanum{27}All the ends of the world shall remember themselves, and be turned unto the Lord : and all the kindreds of the nations shall worship before him.\par
\psanum{28}For the kingdom is the Lord's : and he is the Governor among the people.\par
\psanum{29}All such as be fat upon earth : have eaten and worshipped.\par
\psanum{30}All they that go down into the dust shall kneel before him : and no man hath quickened his own soul.\par
\psanum{31}My seed shall serve him : they shall be counted unto the Lord for a generation.\par
\psanum{32}They shall come, and the heavens shall declare his righteousness : unto a people that shall be born, whom the Lord hath made.\par

    \antiphon{}{They parted my garments among them: and upon my vesture did they cast lots.}
    
    \begin{rubric}
    	After the stripping of the Altars, at the appointed hour a signal is given with a board, and the Clergy come together to perform the Maundy.\par
    	The Prelate or Superior is vested with Stole and violet Cope over Amice and Alb, and in the appointed place with the assistance of the Deacon (who with the Subdeacon assists him in white vestments, as at Mass) puts incense in the thurible: then the Deacon holding the book of the Gospels before his breast, kneels before the Superior and asks a blessing: on receiving this, while two Acolytes with lighted candles stand by, and the Subdeacon holds the book, he signs it and censes it, and sings as usual the following Gospel.
    	\end{rubric}
    
    \readingcitation{Gospel}{John 13:1}
    \lett{B}{efore} the feast of the passover, when Jesus knew that his hour was come that he should depart out of this world unto the Father, having loved his own which were in the world, he loved them unto the end. And supper being ended, the devil having now put into the heart of Judas Iscariot, Simon's son, to betray him; Jesus knowing that the Father had given all things into his hands, and that he was come from God, and went to God; he riseth from supper, and laid aside his garments; and took a towel, and girded himself. After that he poureth water into a bason, and began to wash the disciples' feet, and to wipe them with the towel wherewith he was girded. Then cometh he to Simon Peter: and Peter saith unto him, Lord, dost thou wash my feet? Jesus answered and said unto him, What I do thou knowest not now; but thou shalt know hereafter. Peter saith unto him, Thou shalt never wash my feet. Jesus answered him, If I wash thee not, thou hast no part with me. Simon Peter saith unto him, Lord, not my feet only, but also my hands and my head. Jesus saith to him, He that is washed needeth not save to wash his feet, but is clean every whit: and ye are clean, but not all. For he knew who should betray him; therefore said he, Ye are not all clean. So after he had washed their feet, and had taken his garments, and was set down again, he said unto them, Know ye what I have done to you? Ye call me Master and Lord: and ye say well; for so I am. If I then, your Lord and Master, have washed your feet; ye also ought to wash one another's feet. For I have given you an example, that ye should do as I have done to you.
    
    \begin{rubric}
    	When this is finished, the Subdeacon carries the open book to be kissed by the Superior, and the Deacon censes him as usual.\par
    	Then the Superior puts off the Cope, and is girded by the Deacon and Subdeacon with a linen cloth, and thus girded, the Deacon and Subdeacon assisting, proceeds to the washing of feet. Those who are to be washed are arranged in order, and while Clerks minister a bason and water, and the Subdeacon holds the right foot of each, the Superior kneels before each, washes his foot, wipes and kisses it, the Deacon giving the towel to wipe them. Meanwhile the following are sung:
    \end{rubric}
    
    \antiphon{}{A new commandment I give unto you: that ye love one another, as I have loved you, saith the Lord. \textit{Ps. 119.} Blessed are those that are undefiled in the way: and walk in the law of the Lord. ℣. A new commandment I give unto you: that ye love one another, as I have loved you, saith the Lord.}
    
    \antiphon{}{After the Lord had risen from supper, he poured water into a bason, and began to wash his disciples' feet: this example left he unto them. \textit{Ps. 48.} Great is the Lord, and highly to be praised: in the city of our God, even upon his holy hill. ℣. After the Lord had risen from supper, he poured water into a bason, and began to wash his disciples' feet: this example left he unto them.}
    
    \antiphon{}{The Lord Jesus, after he had supped with his disciples, washed their feet, and said to them: Know ye what I, your Lord and Master, have done to you? I have given you an example, that ye also should so do. \textit{Ps. 85.} Lord, thou art become gracious unto thy land: thou hast turned away the captivity of Jacob. ℣. The Lord Jesus, after he had supped with his disciples, washed their feet, and said to them: Know ye what I, your Lord and Master, have done to you? I have given you an example, that ye also should so do.}
    
    \antiphon{}{Lord, dost thou wash my feet? Jesus answered, and said unto him: If I wash not thy feet, thou hast no part with me. ℣. Then cometh he to Simon Peter, and Peter saith unto him.}
    
    \antiphon{}{Lord, dost thou wash my feet? Jesus answered, and said unto him: If I wash not thy feet, thou hast no part with me. ℣. What I do, thou knowest not now: but thou shalt know hereafter.}
    
    \antiphon{}{Lord, dost thou wash my feet? Jesus answered, and said unto him: If I wash not thy feet, thou hast no part with me. ℣. If I, your Lord and Master, have washed your feet: how much more ought ye to wash one another's feet? \textit{Ps. 49.} O hear ye this, all ye people: ponder it with your ears, all ye that dwell in the world. ℣. If I, your Lord and Master, have washed your feet: how much more ought ye to wash one another's feet?}
    
    \antiphon{}{By this shall all men know that ye are my disciples, if ye have love one to another. ℣. Jesus said to his disciples. ℣. By this shall all men know that ye are my disciples, if ye have love one to another.}
    
    \antiphon{}{Let there abide in you faith, hope, charity, these three: but the greatest of these is charity. ℣. And now abideth faith, hope, charity, these three, but the greatest of these is charity. ℣. Let there abide in you faith, hope, charity, these three: but the greatest of these is charity.}
    
    \antiphon{}{Blessed be the holy Trinity and the undivided Unity: we will give thanks unto him, because he hath shewed his mercy upon us. ℣. Let us bless the Father and the Son with the Holy Ghost. \textit{Ps. 84.} O how amiable are thy dwellings, thou Lord of hosts! My soul hath a desire and longing to enter into the courts of the Lord. ℣. Blessed be the holy Trinity and the undivided Unity: we will give thanks unto him, because he hath shewed his mercy upon us.}
    
    \antiphon{}{Where charity and love are, there is God. ℣. The love of Christ hath gathered us together in one. ℣. Let us rejoice and be glad in him. ℣. Let us fear and love the living God. ℣. And love one another in sincerity of heart.}
    
    \antiphon{}{Where charity and love are, there is God. ℣. When therefore we are gathered together in one: ℣. Take we heed that we be not divided in mind. ℣. Let malicious quarrels, let contentions cease. ℣. And let Christ our God be in the midst of us.}
    
    \antiphon{}{Where charity and love are, there is God. ℣. So may we also with the Blessed see ℣. In glory thy countenance, O Christ our God: ℣. Joy, that is infinite and undefiled, ℣. For ever and for evermore. Amen.}
    
    \begin{rubric}
    	After the washing, the Superior, or he that hath washed the feet, washes his hands, and wipes then with another towel: then returning to the place where he was at first he puts on the Cope, and standing with head uncovered says:
    \end{rubric}
    
    \begin{rubric}
    	\emph{Our Father}, secretly until,
    \end{rubric}
    
    ℣. And lead us not into temptation.
    
    ℟. But deliver us from evil.
    
    ℣. Thou hast charged, O Lord.
    
    ℟. That we shall diligently keep thy commandments.
    
    ℣. Thou didst wash the feet of thy disciples.
    
    ℟. Despise not thou the works of thine hands.
    
    ℣. O Lord, hear my prayer.
    
    ℟. And let my cry come unto thee.
    
    ℣. The Lord be with you.
    
    ℟. And with thy spirit.
    
    \letuspray
    
    \lett{A}{ssist} us, we beseech thee, O Lord, in this our bounden service: and since thou wast pleased to wash the feet of thy disciples, despise not the works of thine own hands which thou hast bidden us to follow: that as here outward stains by us and from us are washed away; so also the secret sins of us all may be blotted out by thee. Which do thou vouchsafe to grant, who livest and reignest God: throughout all ages, world without end. \textit{Amen.}

\subby{Good Friday}
\fancyhead[RO,LE]{\textit{Good Friday}}
\fancyhead[RE,LO]{}
\begin{inhead}
    {First Class Double}
\end{inhead}

\properantiphon{Ben.}{They set up {\dag} over his head his accusation written: Jesus of Nazareth, the King of the Jews.}

    \properantiphon{Mag.}{When he had received {\dag} the vinegar, he said, It is finished; and he bowed his head, and gave up the ghost.}

\begin{rubric}
	In Choir, None being ended, the Priest and Ministers in black vestments, without lights and incense, proceed to the Altar: and pray awhile prostrate before it. Meanwhile the Acolytes spread one cloth only upon the Altar. The Priest with the Ministers, having finished their prayer, go up to the Altar, and he kisses it in the middle: then the Reader proceeds to read the Prophecy in the place where the Epistle is read, and begins it without any announcement: the Priest meanwhile reads the same in a low voice at the Epistle corner.
\end{rubric}

\readingcitation{Lesson}{Hosea 6:1}
\lett{T}{hus saith the Lord:} In their affliction they will seek me early: Come, and let us return unto the \divineName{Lord}: for he hath torn, and he will heal us; he hath smitten, and he will bind us up. After two days will he revive us: in the third day he will raise us up, and we shall live in his sight. Then shall we know, if we follow on to know the \divineName{Lord}: his going forth is prepared as the morning; and he shall come unto us as the rain, as the latter and former rain unto the earth. O Ephraim, what shall I do unto thee? O Judah, what shall I do unto thee? for your goodness is as a morning cloud, and as the early dew it goeth away. Therefore have I hewed them by the prophets; I have slain them by the words of my mouth: and thy judgments are as the light that goeth forth. For I desired mercy, and not sacrifice; and the knowledge of God more than burnt offerings.

\begin{rubric}
	\emph{Thanks be to God} is not answered here, nor after the following Lesson.
\end{rubric}

\tract{O Lord, I have heard thy speech, and was afraid: I have considered thy works, and was confounded. ℣. O Lord, revive thy work in the midst of the years: in the midst of the years make it known. ℣. In the time of confusion of my soul: in wrath, remember mercy. ℣. God came from Teman, and his Holy One from the thick woods of the mountains. ℣. His glory covered the heavens: and the earth was full of his praise.}

\begin{rubric}
	The Tract being finished, the Priest at the Epistle corner says: \emph{Let us pray}. The Deacon: \emph{Let us bow the knee}, and the Subdeacon: \emph{Arise}.
\end{rubric}

\collect
\lett{O}{God,} from whom Judas received the punishment of his guilt, and the thief the reward of his confession, grant unto us the effects of thy propitiation: that as in his passion Jesus Christ, our Lord, gave unto both the divers rewards of their merits; so he may take away the transgressions of our old nature, and bestow upon us the grace of his resurrection. Who liveth.

\begin{rubric}
	The Subdeacon in the tone of the Epistle, likewise without announcement, sings the following Lesson:
\end{rubric}

\readingcitation{Lesson}{Hebrews 10:1}
\lett{B}{rethren:} The law having a shadow of good things to come, and not the very image of the things, can never with those sacrifices which they offered year by year continually make the corners thereunto perfect. For then would they not have ceased to be offered? because that the worshippers once purged should have had no more conscience of sins. But in those sacrifices there is a remembrance again made of sins every year. For it is not possible that the blood of bulls and of goats should take away sins. Wherefore when he cometh into the world, he saith, Sacrifice and offering thou wouldest not, but a body hast thou prepared me: in burnt offerings and sacrifices for sin thou hast had no pleasure. Then said I, Lo, I come (in the volume of the book it is written of me,) to do thy will, O God. Above when he said, Sacrifice and offering and burnt offerings and offering for sin thou wouldest not, neither hadst pleasure therein; which are offered by the law; then said he, Lo, I come to do thy will, O God. He taketh away the first, that he may establish the second. By the which will we are sanctified through the offering of the body of Jesus Christ once for all. And every priest standeth daily ministering and offering oftentimes the same sacrifices, which can never take away sins: but this man, after he had offered one sacrifice for sins for ever, sat down on the right hand of God; from hence-forth expecting till his enemies be made his footstool. For by one offering he hath perfected for ever them that are sanctified. Whereof the Holy Ghost also is a witness to us: for after that he had said before, This is the covenant that I will make with them after those days, saith the Lord, I will put my laws into their hearts, and in their minds will I write them; and their sins and iniquities will I remember no more. Now where remission of these is, there is no more offering for sin. Having therefore, brethren, boldness to enter into the holiest by the blood of Jesus, by a new and living way, which he hath consecrated for us, through the veil, that is to say, his flesh; and having an high priest over the house of God; let us draw near with a true heart in full assurance of faith, having our hearts sprinkled from an evil conscience, and our bodies washed with pure water. Let us hold fast the profession of our faith without wavering; (for he is faithful that promised;) and let us consider one another to provoke unto love and to good works: not forsaking the assembling of ourselves together, as the manner of some is; but exhorting one another: and so much the more, as ye see the day approaching.

\tract{Deliver me, O Lord, from the evil man: and preserve me from the wicked man. ℣. Who imagine mischief in their hearts: and stir up strife all the day long. ℣. They have sharpened their tongues like a serpent: adders' poison is under their lips. ℣. Keep me, O Lord, from the hands of the ungodly: and preserve me from the wicked men. ℣. Who are purposed to overthrow my goings: the proud have laid a snare for me. ℣. And spread a net abroad with cords: yea, and set traps in my way. ℣. I said unto the Lord: Thou art my God: hear the voice of my prayers, O Lord. ℣. O Lord God, thou strength of my health: thou hast covered my head in the day of battle. ℣. Let not the ungodly have his desire, O Lord: let not his mischievous imagination prosper, lest they be too proud. ℣. Let the mischief of their own lips fall upon the head of them: that compass me about. ℣. The righteous also shall give thanks unto thy name: and the just shall continue in thy sight.}

\begin{rubric}
	The Tract ended, the Gospel is read upon a bare pulpit: the Celebrant meanwhile reads it in a low voice at the Epistle corner.
\end{rubric}

\readingcitation{Gospel}{John 19:1}
\lett{A}{t that time:} Pilate took Jesus, and scourged him. And the soldiers platted a crown of thorns, and put it on his head, and they put on him a purple robe, and said, Hail, King of the Jews! and they smote him with their hands. Pilate therefore went forth again, and saith unto them, Behold, I bring him forth to you, that ye may know that I find no fault in him. Then came Jesus forth, wearing the crown of thorns, and the purple robe. And Pilate saith unto them, Behold the man! When the chief priests therefore and officers saw him, they cried out, saying, Crucify him, crucify him., Pilate saith unto them, Take ye him, and crucify him: for I find no fault in him. The Jews answered him, We have a law, and by our law he ought to die, because he made himself the Son of God. When Pilate therefore heard that saying, he was the more afraid; and went again into the judgment hall, and saith unto Jesus, Whence art thou? But Jesus gave him no answer. Then saith Pilate unto him, Speakest thou not unto me? knowest thou not that I have power to crucify thee, and have power to release thee? Jesus answered, Thou couldest have no power at all against me, except it were given thee from above: therefore he that delivered me unto thee hath the greater sin. And from thenceforth Pilate sought to release him: but the Jews cried out, saying, If thou let this man go, thou art not C{\ae}sar's friend: whosoever maketh himself a king, speaketh against C{\ae}sar. When Pilate therefore heard that saying, he brought Jesus forth, and sat down in the judgment seat in a place that is called the Pavement, but in the Hebrew, Gabbatha. And it was the preparation of the passover, and about the sixth hour: and he saith unto the Jews, Behold your King! But they cried out, Away with him, away with him, crucify him. Pilate saith unto them, Shall I crucify your King? The chief priests answered, We have no king but Cæsar. Then delivered he him therefore unto them to be crucified. And they took Jesus, and led him away. And he bearing his cross went forth into a place called the place of a skull, which is called in the Hebrew Golgotha: where they crucified him, and two other with him, on either side one, and Jesus in the midst. And Pilate wrote a title and put it on the cross. And the writing was, \divineName{Jesus of Nazareth the King of the Jews}. This title then read many of the Jews: for the place where Jesus was crucified was nigh to the city; and it was written in Hebrew, and Greek, and Latin. Then said the chief priests of the Jews to Pilate, Write not, The King of the Jews; but that he said, I am King of the Jews. Pilate answered, What I have written I have written. Then the soldiers, when they had crucified Jesus, took his garments, and made four parts, to every soldier a part; and also his coat: now the coat was without seam, woven from the top throughout. They said therefore among themselves, Let us not rend it, but cast lots for it, whose it shall be: that the scripture might be fulfilled, which saith, They parted my raiment among them, and for my vesture they did cast lots. These things therefore the soldiers did. Now there stood by the cross of Jesus his mother, add his mother's sister, Mary the wife of Cleophas, and Mary Magdalene. When Jesus therefore saw his mother, and the disciple standing by, whom he loved, he saith unto his mother, Woman, behold thy son! Then saith he to the disciple, Behold thy mother! And from that hour that disciple took her unto his own home. After this, Jesus knowing that all things were now accomplished, that the scripture might be fulfilled, saith, I thirst. Now there was set a vessel full of vinegar: and they filled a spunge with vinegar, and put it upon hyssop, and put it to his mouth. When Jesus therefore had received the vinegar, he said, It is finished: and he bowed his head, and gave up the ghost. \inrub{Here genuflect, and pause awhile.} The Jews therefore, because it was the preparation, that the bodies should not remain upon the cross on the sabbath day, (for that sabbath day was an high day,) besought Pilate that their legs might be broken, and that they might be taken away. Then came the soldiers, and brake the legs of the first, and of the other which was crucified with him. But when they came to Jesus, and saw that he was dead already, the brake not his legs: but one of the soldiers with a spear pierced his side, and forthwith came there out blood and water. And he that saw it bare record, and his record is true: and he knoweth that he saith true, that ye might believe. For these things were done, that the scripture should be fulfilled, A bone of him shall not be broken. And again another scripture saith, They shall look on him whom they pierced.

\begin{rubric}
	Then the Priest, standing at the Epistle side of the Altar, begins at once with joined hands:
\end{rubric}
\par\noindent
Let us pray dearly beloved, for the holy Church of God: that our God and Lord would vouchsafe to give her peace and unity, and preserve her throughout all the world: making subject unto her principalities and powers: and grant that, leading a quiet and peaceful life, we may glorify God the Father almighty.

\elcol{℣. Let us pray.}{℣. Orémus.}
\elcol{℣. Let us bow the knee.}{℣. Flectámus génua.}
\elcol{℣. Arise.}{℣. Leváte.}

\begin{rubric}
	The Collect is sung in the ferial tone of the Collect of the Mass, with hands extended. And the same manner is observed in all that follow.
\end{rubric}

\lett{A}{lmighty} and everlasting God, who in Christ hast revealed thy glory to all nations: preserve the works of thy mercy; that thy Church, spread abroad over the whole world, may with steadfast faith persevere in the confession of thy name. Through the same.\\
\par\noindent
Let us pray also for our most blessed Father \textit{N.}: that our God and Lord, who hath chosen him unto the order of bishops, may preserve him in health and safety to his holy Church, for the governance of the holy people of God.

\elcol{℣. Let us pray.}{℣. Orémus.}
\elcol{℣. Let us bow the knee.}{℣. Flectámus génua.}
\elcol{℣. Arise.}{℣. Leváte.}
\needspace{4\baselineskip}
\lett{A}{lmighty} and everlasting God, by whose judgement all things are established: mercifully regard our prayers, and in thy goodness preserve him whom thou hast chosen to be our Bishop; that the Christian people who are governed by thine authority may under so great a Pontiff increase in the merits of their faith. Through.\\
\par\noindent
Let us pray also for all Bishops, Priests, and Deacons, for all Subdeacons, Acolytes. Exorcists, Door-keepers, Confessors, Readers, Virgins, and Widows: and for all the holy people of God.

\elcol{℣. Let us pray.}{℣. Orémus.}
\elcol{℣. Let us bow the knee.}{℣. Flectámus génua.}
\elcol{℣. Arise.}{℣. Leváte.}
\needspace{4\baselineskip}
\lett{A}{lmighty} and everlasting God, by whose Spirit the whole body of the Church is governed and sanctified: receive our supplications, which we offer before thee for all orders of the same; that by the bounty of thy grace they may faithfully serve thee in their several estates. Through . . . in the unity of the same.\\
\par\noindent
Let us pray also for our most Christian Emperor (elect) \textit{N.}: that our God and Lord would make subject unto him all the nations of the heathen, for our perpetual peace.

\elcol{℣. Let us pray.}{℣. Orémus.}
\elcol{℣. Let us bow the knee.}{℣. Flectámus génua.}
\elcol{℣. Arise.}{℣. Leváte.}
\needspace{4\baselineskip}
\lett{A}{lmighty} and everlasting God, in whose hand is the dominion of all things and the government of all kingdoms: graciously look upon the Roman Empire; that the nations which trust in their own fierceness may be overthrown by the right hand of thy power. Through.\\
\par\noindent
Let us pray also for our catechumens: that our God and Lord would open the ears of their hearts, and the gate of mercy: that, receiving in the waters of regeneration the remission of all their sins, they also may be found in Christ Jesus our Lord.

\elcol{℣. Let us pray.}{℣. Orémus.}
\elcol{℣. Let us bow the knee.}{℣. Flectámus génua.}
\elcol{℣. Arise.}{℣. Leváte.}
\needspace{4\baselineskip}
\lett{A}{lmighty} and everlasting God, who dost continually enrich thy Church with a new offspring: increase the faith and understanding of our catechumens; that they, being born again in the water of baptism, may be numbered among the sons of thine adoption. Through.\\
\par\noindent
Let us pray, dearly beloved, unto God the Father almighty, that he would purge the world from all errors: would take away diseases: drive away famine: open the prisons: loosen the chains: grant unto pilgrims a safe return: to the sick, healing: and to them that travel by sea a haven of safety.

\elcol{℣. Let us pray.}{℣. Orémus.}
\elcol{℣. Let us bow the knee.}{℣. Flectámus génua.}
\elcol{℣. Arise.}{℣. Leváte.}
\needspace{4\baselineskip}
\lett{A}{lmighty} and everlasting God, the comfort of them that mourn, the strength of them that travail: let the prayers of them that cry out of any tribulation ascend unto thee; that in their necessities all may rejoice in the succour of thy loving kindness. Through.\\
\par\noindent
Let us pray also for heretics and schismatics: that our God and Lord would deliver them from all their errors; and vouchsafe to call them back to their holy mother, the Catholic and Apostolic Church.

\elcol{℣. Let us pray.}{℣. Orémus.}
\elcol{℣. Let us bow the knee.}{℣. Flectámus génua.}
\elcol{℣. Arise.}{℣. Leváte.}
\needspace{4\baselineskip}
\lett{A}{lmighty} and everlasting God, who savest all men, and wouldest not that any should perish: look upon the souls that are deceived by the craft of the devil; that the hearts of them that are gone astray being delivered from all perversity of heresy, may turn to wisdom and come again to the unity of thy truth. Through.\\
\par\noindent
Let us pray also for the faithless Jews: that our God and Lord would take away the veil from their hearts; that they also may acknowledge Jesus Christ, our Lord.

\begin{rubric}
	\emph{Amen} is not said, nor \emph{Let us pray}, nor \emph{Let us bow the knee}, nor \emph{Arise}, but at once is said:
\end{rubric}
\needspace{4\baselineskip}
\lett{A}{lmighty} and everlasting God, who deniest not thy mercy even to the faithless Jews: graciously hear our prayers, which we offer for the blindness of this people; that they acknowledging the light of thy truth, which is Christ, may be delivered from their darkness. Through the same.\\
\par\noindent
Let us pray also for the heathen: that God almighty would take away the iniquity from their hearts; that, forsaking their idols, they may be turned unto the living and true God, and to his only Son, Jesus Christ, our God and Lord.

\elcol{℣. Let us pray.}{℣. Orémus.}
\elcol{℣. Let us bow the knee.}{℣. Flectámus génua.}
\elcol{℣. Arise.}{℣. Leváte.}
\needspace{4\baselineskip}
\lett{A}{lmighty} and everlasting God, who desirest not the death of sinners but rather that they should live: mercifully receive our prayer, and deliver them from the worship of idols; and gather them unto thy holy Church, to the praise and glory of thy name. Through.

\begin{rubric}
	Having finished the Prayers, the Priest lays aside his Chasuble and goes to the Epistle side of the Altar, where at the back corner he receives from the Deacon the Cross which has previously been made ready on the Altar. With his face to the people, he uncovers the upper portion, beginning alone \emph{Behold the wood of the Cross}, and from thence he is helped in singing the rest by the Ministers until \emph{O come, let us worship}.\par
When the Choir sing \emph{O come, let us worship}, all prostrate themselves except the Celebrant. He then advances to the front corner of the same Epistle side: and uncovering the right arm of the Cross, and raising it a little, he begins in a higher key than at first: \emph{Behold the wood of the Cross}, the others singing and worshipping as before. Then the Priest proceeds to the middle of the Altar: and uncovering the Cross entirely and raising it, for the third time he begins in a still higher key: \emph{Behold the wood of the Cross}, the others singing and worshipping as before.
\end{rubric}
Behold the wood of the Cross, whereon was hung the world's salvation.

\textit{Choir.} ℟. O come, let us worship.

\begin{rubric}
	Then the Priest alone carries the Cross to the place prepared before the Altar, and kneeling there he puts it down: then laying aside his shoes, he proceeds to worship the Cross %ADDITION TO CLARIFY MANNER OF WORSHIP:
	with the worship of 
	\begin{otherlanguage}{greek}
	προσκύνησις, 
	\end{otherlanguage}
	that is, dulia.
	\par
	And kneeling thrice before, he kisses it. This done, he returns, and puts on his shoes and the Chasuble. Afterwards the Ministers of the Altar, then the other Clerics and laics, two and two, kneeling thrice, as has been said, worship the Cross. Meantime, while the worship of the Cross is proceeding, the Reproaches are sung, together with the other things that follow, either all or in part, according as the many or few that worship the Cross require: which also the Priest sitting at the bench reads with the Ministers, as below:
\end{rubric}

\textsc{Celebrant.} O my people, what have I done unto thee? or wherein have I wearied thee? Answer me.\par
℣. Because I brought thee forth from the land of Egypt: thou hast prepared a Cross for thy Saviour.

\textsc{Deacon.} \begin{otherlanguage}{greek} Ἅγιος ὁ Θεός.\end{otherlanguage}\par
\textsc{Subdeacon.} Holy God.\par
\textsc{Deacon.} \begin{otherlanguage}{greek} Ἅγιος ἰσχυρός.\end{otherlanguage}\par
\textsc{Subdeacon.} Holy Mighty.\par
\textsc{Deacon.} \begin{otherlanguage}{greek} Ἅγιος ἀθάνατος, ἐλέησον ἡμᾶς.\end{otherlanguage}\par
\textsc{Subdeacon.} Holy Immortal, have mercy upon us.\par
\textsc{Celebrant.} Because I led thee through the desert forty years, and fed thee with manna, and brought thee into a land exceeding good: thou hast prepared a Cross for thy Saviour.\par
\textsc{Deacon.} \begin{otherlanguage}{greek} Ἅγιος ὁ Θεός.\end{otherlanguage}\par
\textsc{Subdeacon.} Holy God.\par
\textsc{Deacon.} \begin{otherlanguage}{greek} Ἅγιος ἰσχυρός.\end{otherlanguage}\par
\textsc{Subdeacon.} Holy Mighty.\par
\textsc{Deacon.} \begin{otherlanguage}{greek} Ἅγιος ἀθάνατος, ἐλέησον ἡμᾶς.\end{otherlanguage}\par
\textsc{Subdeacon.} Holy Immortal, have mercy upon us.\par
\textsc{Celebrant.} What more could I have done for thee that I have not done? I indeed did plant thee, my vineyard, exceeding fair: and thou art become very bitter unto me: for vinegar thou gavest to quench my thirst: and hast pierced with a spear the side of thy Saviour.\par
\textsc{Deacon.} \begin{otherlanguage}{greek} Ἅγιος ὁ Θεός.\end{otherlanguage}\par
\textsc{Subdeacon.} Holy God.\par
\textsc{Deacon.} \begin{otherlanguage}{greek} Ἅγιος ἰσχυρός.\end{otherlanguage}\par
\textsc{Subdeacon.} Holy Mighty.\par
\textsc{Deacon.} \begin{otherlanguage}{greek} Ἅγιος ἀθάνατος, ἐλέησον ἡμᾶς.\end{otherlanguage}\par
\textsc{Subdeacon.} Holy Immortal, have mercy upon us.\par
\textsc{Celebrant.} I did scourge Egypt with her first-born for thy sake: and thou hast scourged me and delivered me up.\par
\textsc{Deacon and Subdeacon.} O my people, what have I done unto thee, or wherein have I wearied thee? Answer me.\par 
\textsc{Celebrant.} I led thee out of Egypt, drowning Pharaoh in the Red Sea: and thou hast delivered me unto the chief priests.\par
\textsc{Deacon and Subdeacon.} O my people, what have I done unto thee, or wherein have I wearied thee? Answer me.\par
\textsc{Celebrant.} I opened the sea before thee: and thou hast opened my side with a spear.\par
\textsc{Deacon and Subdeacon.} O my people, what have I done unto thee, or wherein have I wearied thee? Answer me.\par
\textsc{Celebrant.} I went before thee in a pillar of cloud: and thou hast led me unto the judgment hall of Pilate.\par
\textsc{Deacon and Subdeacon.} O my people, what have I done unto thee, or wherein have I wearied thee? Answer me.\par
\textsc{Celebrant.} I fed thee with manna in the desert: and thou hast stricken me with blows and scourges.\par
\textsc{Deacon and Subdeacon.} O my people, what have I done unto thee, or wherein have I wearied thee? Answer me.\par
\textsc{Celebrant.} I gave thee to drink of the water of salvation from the rock: and thou hast given me gall and vinegar to drink.\par
\textsc{Deacon and Subdeacon.} O my people, what have I done unto thee, or wherein have I wearied thee? Answer me.\par
\textsc{Celebrant.} For thee I smote the kings of the Canaan-ites: and thou hast smitten my head with a reed.\par
\textsc{Deacon and Subdeacon.} O my people, what have I done unto thee, or wherein have I wearied thee? Answer me.\par
\textsc{Celebrant.} I gave thee a royal sceptre: and thou hast given unto my head a crown of thorns.\par
\textsc{Deacon and Subdeacon.} O my people, what have I done unto thee, or wherein have I wearied thee? Answer me.\par
\textsc{Celebrant.} I exalted thee with great power: and thou hast hanged me upon the gibbet of the Cross.\par
\textsc{Deacon and Subdeacon.} O my people, what have I done unto thee, or wherein have I wearied thee? Answer me.\par
\textsc{Celebrant.} We worship thy Cross, O Lord: and praise and glorify thy holy resurrection: for behold, by virtue of the tree joy hath come to the whole world. \textit{Ps.} God be merciful unto us and bless us :\par
\textsc{Deacon and Subdeacon.} And shew us the light of his countenance, and be merciful unto us.\par
\textsc{Celebrant.} We worship thy Cross, O Lord: and praise and glority thy holy resurrection: for behold, by virtue of the tree joy hath come to the whole world.\par
\textsc{Deacon and Subdeacon.} Faithful Cross, above all other, one and only noble tree: none in foliage, none in blossom, none in fruit thy peer may be.\par
Sweetest wood and sweetest iron, sweetest weight is hung on thee.\par
\textsc{Celebrant.} \textit{Hymn.} Sing, my tongue, the glorious battle, sing the ending of the fray, now above the Cross, loud triumphant lay: tell how Christ, the world's Redeemer, as a victim won the day.\par
\textsc{Deacon and Subdeacon.} Faithful Cross, above all other, one and only noble tree: none in foliage, none in blossom, none in fruit thy peer may be.\par
\textsc{Celebrant.} God in pity saw man fallen, shamed and sunk in misery, when he fell on death by tasting fruit of the forbidden tree: then another tree was chosen which the world from death should free.\par
\textsc{Deacon and Subdeacon.} Sweetest wood and sweetest iron, sweetest weight is hung on thee.\par
\textsc{Celebrant.} Thus the scheme of our salvation, was of old in order laid: that the manifold deceiver's art by art might be outweighed: and the lure the toe put forward into means of healing made.\par
\textsc{Deacon and Subdeacon.} Faithful Cross, above all other, one and only noble tree: none in foliage, none in blossom, none in fruit thy peer may be.\par
\textsc{Celebrant.} Therefore when the appointed fulness of the holy time was come, he was sent who maketh all things forth from God's eternal home; thus he came to earth, incarnate, offspring of a virgin's womb.\par
\textsc{Deacon and Subdeacon.} Sweetest wood and sweetest iron, sweetest weight is hung on thee.\par
\textsc{Celebrant.} Lo! he lies, an Infant weeping, where the narrow manger stands, while the Mother-Maid his members wraps in mean and lowly bands, and the swaddling clothes is winding round his helpless feet and hands.\par
\textsc{Deacon and Subdeacon.} Faithful Cross, above all other, one and only noble tree: none in foliage, none in blossom, none in fruit thy peer may be.\par
\textsc{Celebrant.} Thirty years among us dwelling, his appointed time fulfilled, born for this, he meets his Passion, for that this he freely willed, on the Cross the Lamb is lifted where his life-blood shall be spilled.\par
\textsc{Deacon and Subdeacon.} Sweetest wood and sweetest iron, sweetest weight is hung on thee.\par
\textsc{Celebrant.} He endured the nails, the spitting, vinegar, and spear, and reed; from that holy body broken blood and water forth proceed: earth, and stars, and sky, and ocean by that flood from stain are freed.\par
\textsc{Deacon and Subdeacon.} Faithful Cross, above all other, one and only noble tree: none in foliage, none in blossom, none in fruit thy peer may be.\par
\textsc{Celebrant.} Bend thy boughs, O tree of glory! thy relaxing sinews bend; for awhile the ancient rigour that thy birth bestowed, suspend; and the King of heavenly beauty on thy bosom gently tend!\par
\textsc{Deacon and Subdeacon.} Sweetest wood and sweetest iron, sweetest weight is hung on thee.\par
\textsc{Celebrant.} Thou alone wast counted worthy this world's ransom to uphold ; for a shipwreck'd race preparing harbour, like the ark of old; with the sacred blood anointed from the smitten Lamb that rolled.\par
\textsc{Deacon and Subdeacon.} Faithful Cross, above all other, one and only noble tree: none in foliage, none in blossom, none in fruit thy peer may be.\par
\textsc{Celebrant.} To the Trinity be glory everlasting, as is meet; equal to the Father, equal to the Son, and Paraclete: Trinal Unity, whose praises all created things repeat. Amen.\par
\textsc{Deacon and Subdeacon.} Sweetest wood and sweetest iron, sweetest weight is hung on thee.\par

\begin{rubric}
	Towards the end of the worship of the Cross candles are lighted upon the Altar: and the Deacon, receiving the Burse with the Corporal, spreads the Corporal in the usual way, and puts a Purificator by it; and when the adoration is over, reverently receives the Cross and carries it back to the Altar. Then a Procession is formed to the place where on the previous day the Sacrament bad been laid. A Subdeacon goes first with the Cross between two Acolytes bearing candlesticks with lighted candles, and then the Clergy in order, last of all the Priest with the Ministers. When they arrive at the place of the Sacrament, torches are kindled, which are not extinguished till after the consumption of the Sacrament. The Priest genuflects before the Sacrament and prays awhile: the Deacon meanwhile opens the receptacle in which lies hidden the Body of the Lord. Then the Priest, rising, puts incense into two thuribles without blessing, the Deacon presenting the boat, and then kneeling he censes the Sacrament.\par
	The Deacon takes the Chalice with the Sacrament from the receptacle and gives it into the hands of the Priest, covering it with the ends of the Veil which is around his shoulders. Then they proceed in the order in which they came: a baldachino is carried over the Sacrament: and two Acolytes with thuribles continually cense the Sacrament. Meanwhile is sung the following hymn:
\end{rubric}

\subbysub{Vexilla Regis prodeunt}

\gregorioscore{resources/gabc/ProperSeason/GoodFriday.gabc}

\begin{rubric}
	When the Priest arrives at the Altar, he places the Chalice upon it, genuflects, and again censes it, and going up takes the Host out of the Chalice and places it upon the Paten which the Deacon holds: and receiving the Paten from the hand of the Deacon, he places the sacred Host upon the Corporal, saying nothing. If he touches the Sacrament, he washes his hands in some vessel. Meanwhile the Deacon puts wine in the Chalice, and the Subdeacon water, which the Priest does not bless, nor does he say over it the usual Prayer: but receiving the Chalice from the Deacon, be places it upon the Altar, saying nothing: and the Deacon covers it with the Pall.\par
Then he puts incense in the thurible without blessing it, and censes the Oblations, Cross, and Altar in the accustomed manner, genuflecting before and after, and whenever be crosses before the Sacrament.
\end{rubric}

\begin{rubric}
	When he censes the Altar, he says:
\end{rubric}
\lett{L}{et} my prayer, O Lord, be set forth in thy sight as the incense: and let the lifting up of my hands be an evening sacrifice. Set a watch, O Lord, before my mouth, and keep the door of my lips: O let not mine heart be inclined to any evil thing, let me not be occupied in ungodly works.

\begin{rubric}
	When he gives back the thurible to the Deacon he says:
\end{rubric}

\lett{T}{he} Lord kindle in us the fire of his love and the flame of everlasting charity. Amen.

\begin{rubric}
	And he himself is not censed.
\end{rubric}

\begin{rubric}
	Afterwards he washes his hands a little beyond the Altar on the Epistle side, saying nothing: then bowing in the middle of the Altar, with hands joined, he says:
\end{rubric}

\lett{I}{n} a humble spirit and with a contrite heart may we be accepted of thee, O Lord: and so let our sacrifice be offered in thy sight this day, that it may be pleasing unto thee, O Lord God.

\begin{rubric}
	Then turning to the people at the Gospel corner, he says as usual:
\end{rubric}

\lett{P}{ray,} brethren, that my sacrifice and yours may be acceptable to God the Father almighty.

\begin{rubric}
	And he turns back by the same way, not completing the circle: and, omitting the rest, proceeds:
\end{rubric}

Let us pray: Commanded by saving precepts, and taught by divine institution, we are bold to say:

\lett{O}{ur} Father, who art in heaven: Hallowed be thy name: Thy kingdom come: Thy will be done, on earth as it is in heaven. Give us this day our daily bread: And forgive us our trespasses, as we forgive those that trespass against us. And lead us not into temptation.\par
℟. But deliver us from evil.

\begin{rubric}
	The Priest says silently \emph{Amen}. In the same voice in which he said \emph{Our Father}, and without \emph{Let us pray}, he says in the tone of the Collect of the ferial Mass:
\end{rubric}

\lett{D}{eliver} us, O Lord, we beseech thee, from all evils, past, present, and to come: and at the intercession of the blessed and glorious ever Virgin Mary Mother of God, with thy blessed Apostles Peter and Paul, and with Andrew, and all the Saints, \inrub{he does not sign himself with the Paten} favourably grant peace in our days: that by the help of thine availing mercy we may ever both be free from sin, and safe from all distress. Through the same thy Son Jesus Christ our Lord: Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, throughout all ages, world without end.\par
℟. Amen.

\subbysub{Confession}
\begin{rubric}
    Then the Deacon turneth to the People and saith to those who come to receive the Holy Communion,
\end{rubric}
\lett{Y}{e} who do truly and earnestly repent you of your sins, and are in love and charity with your neighbours, and intend to lead a new life, following the commandments of God, and walking from henceforth in his holy ways; Draw near with faith, and take this holy Sacrament to your comfort; and make your humble confession to Almighty God, devoutly kneeling.
\begin{rubric}
Then shall this General Confession be made, in an audible voice, by the Priest, bowing, and all those who are minded to receive the Holy Communion, humbly kneeling.
\end{rubric}
\needspace{4\baselineskip}
\lett{A}{lmighty} God, Father of our Lord Jesus Christ, Maker of all things, Judge of all men; We acknowledge and bewail our manifold sins and wickedness, Which we, from time to time, most grievously have committed, By thought, word, and deed, Against thy Divine Majesty, Provoking most justly thy wrath and indignation against us. We do earnestly repent, And are heartily sorry for these our misdoings; The remembrance of them is grievous unto us; The burden of them is intolerable. Have mercy upon us, Have mercy upon us, most merciful Father; For thy Son our Lord Jesus Christ's sake, Forgive us all that is past; And grant that we may ever hereafter Serve and please thee in newness of life, To the honour and glory of thy Name; Through Jesus Christ our Lord. Amen.
\begin{rubric}
	Then shall the Priest stand up and, turning to the People, say,
\end{rubric}
\needspace{4\baselineskip}
\lett{A}{lmighty} God, our heavenly Father, who of his great mercy hath promised forgiveness of sins to all those who with hearty repentance and true faith turn unto him; Have mercy upon you; pardon {\ding{64}} and deliver you from all your sins; confirm and strengthen you in all goodness; and bring you to everlasting life; through Jesus Christ our Lord.

℟. Amen.
\subbysub{Comfortable Words}
\begin{rubric}
Then shall the Priest, facing the People, say,
\end{rubric}\noindent
\begin{center}
	\textsc{Hear what comfortable words our Saviour Christ saith unto all who truly turn to him.}
\end{center}
\par\noindent
Come unto me, all ye that travail and are heavy laden, and I will refresh you. \vr{Matt. 11:28}
\par\noindent
    So God loved the world, that he gave his only-begotten Son, to the end that all that believe in him should not perish, but have everlasting life. \vr{John 3:16}
    \par\noindent
    \begin{center}
		\textsc{Hear also what Saint Paul saith.}
	\end{center}
\par\noindent
    This is a true saying, and worthy of all men to be received, That Christ Jesus came into the world to save sinners. \vr{1 Tim. 1:15}

    \begin{center}
		\textsc{Hear also what Saint John saith.}
	\end{center}
    \par\noindent
    If any man sin, we have an Advocate with the Father, Jesus Christ the righteous; and he is the Propitiation for our sins. \vr{1 John 2:1-2}

\subbysub{Prayer of Humble Access}
\begin{rubric}
    The Priest alone then saith, in an audible voice,
\end{rubric}
\lett{W}{e} do not presume to come to this thy Table, O merciful Lord, trusting in our own righteousness, but in thy manifold and great mercies. We are not worthy so much as to gather up the crumbs under thy Table. But thou art the same Lord, whose property is always to have mercy: Grant us therefore, gracious Lord, so to eat the flesh of thy dear Son Jesus Christ, and to drink his blood, that our sinful bodies may be made clean by his Body, and our souls washed through his most precious Blood, and that we may evermore dwell in him, and he in us.\par
℟. Amen.

\subbysub{Communion of the Priest}

\begin{rubric}
	Then the Celebrant, having made a reverence down to the ground, places the Paten beneath the Sacrament, which he takes in his right hand and elevates so that it can be seen by the people: and immediately he divides it over the Chalice into three parts, the last of which he puts into the Chalice in the usual way, saying nothing. \emph{The Peace of the Lord} is not said, nor \emph{O Lamb of God}, nor is the kiss of peace given. Then, omitting the first two Prayers, he says only the following:
\end{rubric}

\lett{L}{et} the partaking of thy Body, O Lord Jesu Christ, which I, unworthy, presume to receive, turn not to my judgement and condemnation: but of thy goodness let it avail unto me for protection of mind and of body, that I may receive thy healing: Who livest and reignest with God the Father in the unity of the Holy Ghost, God, throughout all ages, world without end. Amen.

\begin{rubric}
	Then he genuflects and receives the Paten with the Body of Christ: and with the deepest humility and reverence says:
\end{rubric}

I will receive the bread of heaven, and call upon the name of the Lord.

\begin{rubric}
	He beats his breast, saying thrice:
\end{rubric}

Lord, I am not worthy that thou shouldest come under my roof: but speak the word only, and my soul shall be healed.

\begin{rubric}
	Then he signs himself with the Sacrament, saying:
\end{rubric}

The Body of our Lord Jesus Christ preserve my soul unto everlasting life. Amen.

\begin{rubric}
	And he reverently consumes the Body.
\end{rubric}

\begin{rubric}
	Then, omitting everything which is wont to be said before the consumption of the Blood, he at once reverently consumes the particle of the Host with the wine from the Chalice.
\end{rubric}

\subbysub{Communion of the People}
\begin{rubric}
	The Priest then, turning towards the Congregation, holds a particle over the Chalice and says:
\end{rubric}
\elcol{℣. Behold the Lamb of God. Behold him who taketh away the sins of the world.}{℣. Ecce Agnus Dei, ecce, qui tollit peccáta mundi.}
\elcol{℟. Lord, I am not worthy that thou shouldest come under my roof, but speak the word only, and my soul shall be healed.}{℟. Dómine, non sum dignus, ut intres sub tectum meum, sed tantum dic verbo, et sanábitur ánima mea.}
\elcol{℟. Lord, I am not worthy that thou shouldest come under my roof, but speak the word only, and my soul shall be healed.}{℟. Dómine, non sum dignus, ut intres sub tectum meum, sed tantum dic verbo, et sanábitur ánima mea.}
\elcol{℟. Lord, I am not worthy that thou shouldest come under my roof, but speak the word only, and my soul shall be healed.}{℟. Dómine, non sum dignus, ut intres sub tectum meum, sed tantum dic verbo, et sanábitur ánima mea.}

\begin{rubric}
	The Priest then distributes Communion, in both kinds, to the People in the usual manner.
\end{rubric}

\begin{rubric}
	After distributing the Communion to the People, he makes the ablution of his fingers in the usual manner and taken the purification. He then says, bowing in the midst of the Altar, with hands joined:
\end{rubric}

\lett{G}{rant,} O Lord, that what we have taken with our mouths we may receive with a pure heart: and from a temporal gift may it become to us an everlasting remedy.

%MANUAL ADJUSTMENT:
\clearpage
\begin{rubric}
	\emph{May thy Body} is not said, nor the Postcommunion, nor \emph{Let this my bounden duty}, nor is the blessing given: but having made a reverence to the Altar, the Priest retires with the Ministers.
\end{rubric}

\subbysub{Thanksgiving}
\begin{rubric}
    Standing with his hands joined, the Priest shall alone say the Thanksgiving, in an audible voice.
\end{rubric}
\letuspray
\lett{A}{lmighty} and everliving God, we most heartily thank thee, for that thou dost vouchsafe to feed us who have duly received these holy mysteries with the spiritual food of the most precious Body and Blood of thy Son our Saviour Jesus Christ; and dost assure us thereby of thy favour and goodness towards us; and that we are very members incorporate in the mystical body of thy Son, which is the blessed company of all faithful people; and are also heirs through hope of thy everlasting kingdom, by the merits of his most precious death and passion. And we humbly beseech thee, O heavenly Father, so to assist us with thy grace, that we may continue in that holy fellowship, and do all such good works as thou hast prepared for us to walk in; through Jesus Christ our Lord, to whom, with thee and the Holy Ghost, be all honour and glory, world without end.\par
℟. Amen.

\begin{rubric}
	Evensong is then said without singing: and the Altar is stripped.
\end{rubric}


\subby{Easter Even}
\fancyhead[RO,LE]{\textit{Easter Even}}
\fancyhead[RE,LO]{}
\begin{inhead}
    {First Class Double}
\end{inhead}

\properantiphon{Ben.}{The women, {\dag} sitting over against the sepulchre, made lamentation, weeping for the Lord.}
    
\subbysub{Blessing of the Paschal Fire}
\begin{rubric}
	At a suitable time the Altars are covered with linen cloths, but the candles remain unlit until the beginning of Mass. Meanwhile fire is struck from a flint outside the church, and from it coals are kindled. After Mattins (or None) has been said in Choir, the Priest, vested in Amice, Alb, Girdle, Stole, and violet Cope, or without a Chasuble, standing with the Ministers, with the Cross, holy water, and incense, before the door of the Church, if it can conveniently be done, or in the entry of the church itself, blesses the new fire, saying:
\end{rubric}

℣. The Lord be with you.

℟. And with thy spirit.

\letuspray
\lett{O}{God,} who through thy Son, the true corner-stone, hast bestowed upon thy faithful people the fire of thy brightness: sanct {\ding{64}} ify this new fire, now struck from the flint-stone, to be profitable to our service: and grant unto us that by this paschal feast we may be so inflamed with heavenly desires; that we may with pure hearts attain unto the feast of thy eternal brightness. Through the same Christ, our Lord. \textit{Amen.}

\letuspray
\lett{O}{Lord} God, Father almighty, who art light unfailing, and the author of all lights: bl {\ding{64}} ess this light which hath here been sanctified and blessed by thee, who hast enlightened the whole world: that we may be enkindled by that light, and enlightened with the fire of thy brightness: and like as thou didst enlighten Moses going forth from Egypt, so do thou enlighten our hearts and understandings; that we may be found worthy to attain unto life and light everlasting. Through Christ, our Lord. \textit{Amen.}

\letuspray
\lett{O}{Lord} holy, Father almighty, everlasting God: vouchsafe to assist us who bless this fire in thy name, and in the name of thy only-begotten Son, Jesus Christ, our God and Lord, and of the Holy Ghost; and assist us against the fiery darts of the enemy, and enlighten us with thy heavenly grace: Who livest and reignest with the same thy only-begotten Son and the Holy Ghost, God: throughout all ages world without end. \textit{Amen.}

\begin{rubric}
	He then blesses the five grains of incense, to be placed in the Candle, saying at once this Collect :
\end{rubric}
\lett{W}{e} beseech thee, almighty God, that the abundance of thy bles {\ding{64}} sing may come down upon this incense; and as thou, thyself unseen, art the regenerator of all things, so kindle the brightness of this night; that not only may our sacrifice, which is offered on this night, shine with the secret mingling of thy light; but that in all places wheresoever aught shall be carried from this mystery here sanctified, the wickedness of the crafts of the devil may be driven forth, and the power of thy majesty ever be present. Through Christ, our Lord. \textit{Amen.}

\subbysub{Procession of the Light of Christ}

\begin{rubric}
	Whilst be blesses the grains of incense, an Acolyte takes some of the blessed coals, and sets them in the thurible: and the Priest, when he has finished the last Prayer, puts incense from the boat into the thurible, giving the usual blessing. Then he sprinkles the aforesaid grains of incense and the fire with holy water, saying: \emph{Thou shalt purge me, O Lord}, without note and without the Psalm, and censes them thrice.

	Meanwhile all the lights in the church are extinguished, to be relighted later from the blessed fire.
	
	Then the Deacon, vested in a white Dalmatic, takes the reed upon the top of which are three candles in the shape of a triangle. The Thurifer goes first with an Acolyte bearing the five grains of incense: the Subdeacon follow's with the Cross, the Clergy in their order: then the Deacon with the reed, after him the Celebrant.
	
	When the Deacon enters the church, he lowers the reed, and an Acolyte, carrying a candle which has been lighted from the new fire, lights one of the three candles placed upon it, and the Deacon, elevating the reed, genuflects, as do all others with him except the Subdeacon bearing the Cross, and sings alone: \emph{The Light of Christ}.
\end{rubric}

℟. Thanks be to God.

\begin{rubric}
	And proceeding to the middle of the church, another candle is lighted: and having again genuflected, he sings in a higher tone: \emph{The Light of Christ}.
\end{rubric}

℟. Thanks be to God.

\begin{rubric}
	For the third time he proceeds before the Altar, where the third candle is lighted: and again, having genuflected as before, he sings still higher: \emph{The Light of Christ}.
\end{rubric}

℟. Thanks be to God.

\begin{rubric}
	The Celebrant then ascends to the Epistle corner of the Altar, and the Deacon, banding the reed with the triple candle to an Acolyte, receives the book, and asks a blessing from the Celebrant as at the Gospel, the Priest saying:
\end{rubric}
\par\noindent
The Lord be in thy heart, and on thy lips: that thou mayest worthily and fitly proclaim his paschal praise: In the name of the Father, and of the Son, {\ding{64}} and of the Holy Ghost. Amen.

\begin{rubric}
	The Deacon then goes to the desk and placing the book upon it, censes it. On the right hand of the Deacon stand the Subdeacon with the Cross, and the Thurifer: on his left an Acolyte holding the reed, and another the five grains of blessed incense to be fixed into the Candle. Then, all rising and standing, as at the Gospel, the Deacon sings:
\end{rubric}

\lett{N}{ow} let the Angelic host of heaven rejoice: let the divine mysteries rejoice: and for the victory of the mighty King let the trumpet of salvation sound forth. Let the earth also be glad, illumined by the rays of this great brightness: and enlightened by the splendour of the eternal King, let her know that she hath put away the darkness of the whole world. Let our mother the Church also rejoice, adorned with the brightness of so great a light: and let this temple resound with the triumphant voices of the peoples. Wherefore, dearly beloved brethren, who are here present in the wondrous clearness of this holy light, join with me, I beseech you, in calling upon the mercy of almighty God. That he who hath been pleased, for no merit of mine, to admit me into the number of his Levites: may pour on me the brightness of his light, and make me meet to proclaim the praises of this Candle. Through Jesus Christ his Son our Lord: Who liveth and reigneth with him in the unity of the Holy Ghost, God. Throughout all ages, world without end. \textit{Amen.}


℣. The Lord be with you.

℟. And with thy spirit.

℣. Lift up your hearts.

℟. We lift them up unto the Lord.

℣. Let us give thanks unto our Lord God.

℟. It is meet and right so to do.
\par\noindent
It is very meet, right, and our bounden duty, that we should, with the whole affection of our heart and mind, and with the service of our lips, give praise unto the invisible God, the Father almighty, and unto his only begotten Son, Jesus Christ our Lord. Who paid for us to the eternal Father the debt of Adam's transgression: and with his dear blood wiped away the reproach of our former offences. For this is the paschal feast wherein the very Lamb is slain, by whose blood the door-posts of the faithful are made holy. This is the night, wherein of old thou didst lead forth our fathers, the children of Israel, out of Egypt, and didst make them to pass as on dry land through the Red Sea. Yea, this is the night, that with the brightness of the pillar hath purged away the darkness of iniquity. This is the night, which throughout the whole world today doth separate them that believe in Christ from the wickedness of the age and from the darkness of transgressions, reneweth them unto grace, restoreth them to holiness. This is the night wherein, breaking the chains of death, Christ ascendeth from hell in triumph. For nought indeed had it profited us to be born, if it had not profited us to be redeemed. O how wonderful the condescension of thy loving kindness! O how inestimable the goodness of thy love: who to redeem a slave didst deliver up thy Son! O truly necessary sin of Adam, which by the death of Christ was done away! O happy fault, which was counted worthy to have such and so great a Redeemer! O night truly blessed, which alone was worthy to know the time and the hour wherein Christ rose again from hell! This is the night whereof it is written: And the night is as clear as the day: And, then shall my night be turned into day. The sanctifying power therefore of this night putteth to flight the deeds of wickedness, washeth away sins: restoreth innocence to the fallen, and joy to them that mourn: casteth out enmities, prepareth concord, and boweth down principalities.

\begin{rubric}
	Here the Deacon fixes the five grains of incense in the Candle, placing them in the form of a cross, in this order:
\end{rubric}

\begin{center}
	1
	
	4 2 5
	
	3
\end{center}
\par\noindent
Therefore in this night of grace accept, O holy Father, the evening sacrifice of this incense: which, by the hands of thy ministers, holy Church doth lay before thee, in the solemn offering of this Candle, made from the work of bees. But we already know the excellency of this pillar, which for the honour of God the sparkling fire doth kindle.

\begin{rubric}
	Here the Deacon lights the Candle with one of the three candles set on the reed.
\end{rubric}
\par\noindent
Which, though it be divided into parts, suffereth not loss by the borrowing of its light. For it is fed by melting wax, which the bee the mother hath wrought into the substance of this precious candle.

\begin{rubric}
	Here the lamps are lighted.
\end{rubric}
\par\noindent
O night truly blessed, which did spoil the Egyptians, and made rich the Hebrews! O night wherein heavenly things are joined unto earthly, things human unto things divine. We therefore pray thee, O Lord: that this Candle, consecrated to the honour of thy name, may continue without ceasing to vanquish the darkness of this night. That, being accepted for a savour of sweetness, it may be mingled with the lights of heaven. May the morning star find it burning. That morning star, I say, which knoweth not his going down. That star, which, rising again from hell, steadfastly giveth light to all mankind. We therefore pray thee, O Lord: that thou wouldest vouchsafe to rule, govern and preserve with thy continual protection us thy servants, the whole clergy and thy most faithful people: together with our most blessed Father \textit{N.} (and our Bishop \textit{N.}), granting us peaceful times in this out paschal joy. Through the same Jesus Christ thy Son our Lord: Who liveth and reigneth with thee in the unity of the Holy Ghost God: throughout all ages, world without end. \textit{Amen.}

\subbysub{The Prophecies}

\begin{rubric}
	The blessing of the Candle ended, the Deacon lays aside his white vestments, and puts on violet, and goes to the Celebrant: who, laying aside his Cope, puts on a Maniple and violet Chasuble. Then the Prophecies are read without title, and at the end the response \emph{Thanks be to God} is not made, and the Celebrant reads them in a low voice at the Epistle corner of the Altar.
	
	At the end of the Prophecies are said the Prayers as given below.
	
	Before, or during the reading of the Prophecies, the Priests catechize any Catechumens about to be baptized, and prepare them for Baptism.
\end{rubric}

\readingcitation{Prophecy I}{Genesis 1:1}
\lett{I}{n} the beginning God created the heaven and the earth. And the earth was without form, and void; and darkness was upon the face of the deep. And the Spirit of God moved upon the face of the waters. And God said, Let there be light: and there was light. And God saw the light, that it was good: and God divided the light from the darkness. And God called the light Day, and the darkness he called Night. And the evening and the morning were the first day.

And God said, Let there be a firmament in the midst of the waters, and let it divide the waters from the waters. And God made the firmament, and divided the waters which were under the firmament from the waters which were above the firmament: and it was so. And God called the firmament Heaven. And the evening and the morning were the second day.

And God said, Let the waters under the heaven be gathered together unto one place, and let the dry land appear: and it was so. And God called the dry land Earth; and the gathering together of the waters called he Seas: and God saw that it was good. And God said, Let the earth bring forth grass, the herb yielding seed, and the fruit tree yielding fruit after his kind, whose seed is in itself, upon the earth: and it was so. And the earth brought forth grass, and herb yielding seed after his kind, and the tree yielding fruit, whose seed was in itself, after his kind: and God saw that it was good. And the evening and the morning were the third day.

And God said, Let there be lights in the firmament of the heaven to divide the day from the night; and let them be for signs, and for seasons, and for days, and years: and let them be for lights in the firmament of the heaven to give light upon the earth: and it was so. And God made two great lights; the greater light to rule the day, and the lesser light to rule the night: he made the stars also. And God set them in the firmament of the heaven to give light upon the earth, and to rule over the day and over the night, and to divide the light from the darkness: and God saw that it was good. And the evening and the morning were the fourth day.

And God said, Let the waters bring forth abundantly the moving creature that hath life, and fowl that may fly above the earth in the open firmament of heaven. And God created great whales, and every living creature that moveth, which the waters brought forth abundantly, after their kind, and every winged fowl after his kind: and God saw that it was good. And God blessed them, saying, Be fruitful, and multiply, and fill the waters in the seas, and let fowl multiply in the earth. And the evening and the morning were the fifth day.

And God said, Let the earth bring forth the living creature after his kind, cattle, and creeping thing, and beast of the earth after his kind: and it was so. And God made the beast of the earth after his kind, and cattle after their kind, and every thing that creepeth upon the earth after his kind: and God saw that it was good.

And God said, Let us make man in our image, after our likeness: and let them have dominion over the fish of the sea, and over the fowl of the air, and over the cattle, and over all the earth, and over every creeping thing that creepeth upon the earth. So God created man in his own image, in the image of God created he him; male and female created he them. And God blessed them, and God said unto them, Be fruitful, and multiply, and replenish the earth, and subdue it: and have dominion over the fish of the sea, and over the fowl of the air, and over every living thing that moveth upon the earth.

And God said, Behold, I have given you every herb bearing seed, which is upon the face of all the earth, and every tree, in the which is the fruit of a tree yielding seed; to you it shall be for meat. And to every beast of the earth, and to every fowl of the air, and to every thing that creepeth upon the earth, wherein there is life, I have given every green herb for meat: and it was so. And God saw every thing that he had made, and, behold, it was very good. And the evening and the morning were the sixth day.

Thus the heavens and the earth were finished, and all the host of them. And on the seventh day God ended his work which he had made; and he rested on the seventh day from all his work which he had made.

\begin{rubric}
	\emph{Thanks be to God} is not said either here or after the other Prophecies.
\end{rubric}

\textsc{Priest.} Let us pray.

\textsc{Deacon.} Let us bow the knee.

\textsc{Subdeacon.} Arise.

\lett{O}{God,} who didst wonderfully create man, and hast more wonderfully redeemed him: grant us, we beseech thee, such strength of mind to withstand the enticements of sin; that we may be found worthy to attain to everlasting joys. Through Jesus Christ thy Son our Lord. Who liveth.

\readingcitation{Prophecy II}{Genesis 5:32}
\lett{A}{nd} Noah was five hundred years old: and Noah begat Shem, Ham, and Japheth. And it came to pass, when men began to multiply on the face of the earth, and daughters were born unto them, that the sons of God saw the daughters of men that they were fair; and they took them wives of all which they chose. And the Lord said, My spirit shall not always strive with man, for that he also is flesh: yet his days shall be an hundred and twenty years. There were giants in the earth in those days; and also after that, when the sons of God came in unto the daughters of men, and they bare children to them, the same became mighty men which were of old, men of renown.

And God saw that the wickedness of man was great in the earth, and that every imagination of the thoughts of his heart was only evil continually. And it repented the Lord that he had made man on the earth, and it grieved him at his heart. And the Lord said, I will destroy man whom I have created from the face of the earth; both man, and beast, and the creeping thing, and the fowls of the air; for it repenteth me that I have made them. But Noah found grace in the eyes of the Lord.

These are the generations of Noah: Noah was a just man and perfect in his generations, and Noah walked with God. And Noah begat three sons, Shem, Ham, and Japheth. The earth also was corrupt before God, and the earth was filled with violence. And God looked upon the earth, and, behold, it was corrupt; for all flesh had corrupted his way upon the earth. And God said unto Noah, The end of all flesh is come before me; for the earth is filled with violence through them; and, behold, I will destroy them with the earth. Make thee an ark of gopher wood; rooms shalt thou make in the ark, and shalt pitch it within and without with pitch. And this is the fashion which thou shalt make it of: The length of the ark shall be three hundred cubits, the breadth of it fifty cubits, and the height of it thirty cubits. A window shalt thou make to the ark, and in a cubit shalt thou finish it above; and the door of the ark shalt thou set in the side thereof; with lower, second, and third stories shalt thou make it. And, behold, I, even I, do bring a flood of waters upon the earth, to destroy all flesh, wherein is the breath of life, from under heaven; and every thing that is in the earth shall die. But with thee will I establish my covenant; and thou shalt come into the ark, thou, and thy sons, and thy wife, and thy sons' wives with thee. And of every living thing of all flesh, two of every sort shalt thou bring into the ark, to keep them alive with thee; they shall be male and female. Of fowls after their kind, and of cattle after their kind, of every creeping thing of the earth after his kind, two of every sort shall come unto thee, to keep them alive. And take thou unto thee of all food that is eaten, and thou shalt gather it to thee; and it shall be for food for thee, and for them. Thus did Noah; according to all that God commanded him, so did he. 

And Noah was six hundred years old when the flood of waters was upon the earth. The same day were all the fountains of the great deep broken up, and the windows of heaven were opened. And the rain was upon the earth forty days and forty nights. In the selfsame day entered Noah, and Shem, and Ham, and Japheth, the sons of Noah, and Noah's wife, and the three wives of his sons with them, into the ark; they, and every beast after his kind, and all the cattle after their kind, and every creeping thing that creepeth upon the earth after his kind, and every fowl after his kind, every bird of every sort. And the ark went upon the face of the waters. And the waters prevailed exceedingly upon the earth; and all the high hills, that were under the whole heaven, were covered. Fifteen cubits upward did the waters prevail; and the mountains were covered. And all flesh died that moved upon the earth, both of fowl, and of cattle, and of beast, and of every creeping thing that creepeth upon the earth. And Noah only remained alive, and they that were with him in the ark. And the waters prevailed upon the earth an hundred and fifty days.

And God remembered Noah, and every living thing, and all the cattle that was with him in the ark: and God made a wind to pass over the earth, and the waters asswaged; the fountains also of the deep and the windows of heaven were stopped, and the rain from heaven was restrained; and the waters returned from off the earth continually: and after the end of the hundred and fifty days the waters were abated.

And it came to pass at the end of forty days, that Noah opened the window of the ark which he had made: and he sent forth a raven, which went forth to and fro, until the waters were dried up from off the earth. Also he sent forth a dove from him, to see if the waters were abated from off the face of the ground; but the dove found no rest for the sole of her foot, and she returned unto him into the ark, for the waters were on the face of the whole earth: then he put forth his hand, and took her, and pulled her in unto him into the ark. And he stayed yet other seven days; and again he sent forth the dove out of the ark; and the dove came in to him in the evening; and, lo, in her mouth was an olive leaf pluckt off: so Noah knew that the waters were abated from off the earth. And he stayed yet other seven days; and sent forth the dove; which returned not again unto him any more.

And God spake unto Noah, saying, Go forth of the ark, thou, and thy wife, and thy sons, and thy sons' wives with thee. Bring forth with thee every living thing that is with thee, of all flesh, both of fowl, and of cattle, and of every creeping thing that creepeth upon the earth; that they may breed abundantly in the earth, and be fruitful, and multiply upon the earth. And Noah went forth, and his sons, and his wife, and his sons' wives with him: every beast, every creeping thing, and every fowl, and whatsoever creepeth upon the earth, after their kinds, went forth out of the ark.

And Noah builded an altar unto the Lord; and took of every clean beast, and of every clean fowl, and offered burnt offerings on the altar. And the Lord smelled a sweet savour.

\textsc{Priest.} Let us pray.

\textsc{Deacon.} Let us bow the knee.

\textsc{Subdeacon.} Arise.

\lett{O}{God,} who art strength unchangeable and light eternal: look down in mercy on the wondrous mystery of thy whole Church, and by the operation of thy continual providence accomplish in tranquillity the work of man's salvation; and let the whole world perceive and know that things cast down are being raised up, things grown old are being made new, and all things are returning to their perfection, through him from whom they took their beginning: Jesus Christ thy Son our Lord: Who liveth and reigneth with thee.

\readingcitation{Prophecy III}{Genesis 22:1}
\lett{I}{n those days:} God did tempt Abraham, and said unto him, Abraham: and he said, Behold, here I am. And he said, Take now thy son, thine only son Isaac, whom thou lovest, and get thee into the land of Moriah; and offer him there for a burnt offering upon one of the mountains which I will tell thee of.

And Abraham rose up early in the morning, and saddled his ass, and took two of his young men with him, and Isaac his son, and clave the wood for the burnt offering, and rose up, and went unto the place of which God had told him. Then on the third day Abraham lifted up his eyes, and saw the place afar off. And Abraham said unto his young men, Abide ye here with the ass; and I and the lad will go yonder and worship, and come again to you. And Abraham took the wood of the burnt offering, and laid it upon Isaac his son; and he took the fire in his hand, and a knife; and they went both of them together. And Isaac spake unto Abraham his father, and said, My father: and he said, Here am I, my son. And he said, Behold the fire and the wood: but where is the lamb for a burnt offering? And Abraham said, My son, God will provide himself a lamb for a burnt offering: so they went both of them together. And they came to the place which God had told him of; and Abraham built an altar there, and laid the wood in order, and bound Isaac his son, and laid him on the altar upon the wood. And Abraham stretched forth his hand, and took the knife to slay his son. And the angel of the Lord called unto him out of heaven, and said, Abraham, Abraham: and he said, Here am I. And he said, Lay not thine hand upon the lad, neither do thou any thing unto him: for now I know that thou fearest God, seeing thou hast not withheld thy son, thine only son from me. And Abraham lifted up his eyes, and looked, and behold behind him a ram caught in a thicket by his horns: and Abraham went and took the ram, and offered him up for a burnt offering in the stead of his son. And Abraham called the name of that place Jehovah-jireh: as it is said to this day, In the mount of the Lord it shall be seen.

And the angel of the Lord called unto Abraham out of heaven the second time, and said, By myself have I sworn, saith the Lord, for because thou hast done this thing, and hast not withheld thy son, thine only son: that in blessing I will bless thee, and in multiplying I will multiply thy seed as the stars of the heaven, and as the sand which is upon the sea shore; and thy seed shall possess the gate of his enemies; and in thy seed shall all the nations of the earth be blessed; because thou hast obeyed my voice. So Abraham returned unto his young men, and they rose up and went together to Beer-sheba; and Abraham dwelt at Beer-sheba.

\textsc{Priest.} Let us pray.

\textsc{Deacon.} Let us bow the knee.

\textsc{Subdeacon.} Arise.

\lett{O}{God,} the sovereign Father of the faithful, who dost spread abroad throughout all the world the grace of adoption, to the increasing of the sons of thy promise: and, according as thou didst swear unto thy servant Abraham, dost through this paschal sacrament make him the father of all nations; grant that thy peoples may worthily enter upon the grace of thy calling. Through.

\readingcitation{Prophecy IV}{Exodus 14:24}
\lett{I}{n those days:} It came to pass, that in the morning watch the Lord looked unto the host of the Egyptians through the pillar of fire and of the cloud, and troubled the host of the Egyptians, and took off their chariot wheels, that they drave them heavily: so that the Egyptians said, Let us flee from the face of Israel; for the Lord fighteth for them against the Egyptians.

 And the Lord said unto Moses, Stretch out thine hand over the sea, that the waters may come again upon the Egyptians, upon their chariots, and upon their horsemen. And Moses stretched forth his hand over the sea, and the sea returned to his strength when the morning appeared; and the Egyptians fled against it; and the Lord overthrew the Egyptians in the midst of the sea. And the waters returned, and covered the chariots, and the horsemen, and all the host of Pharaoh that came into the sea after them; there remained not so much as one of them. But the children of Israel walked upon dry land in the midst of the sea; and the waters were a wall unto them on their right hand, and on their left. Thus the Lord saved Israel that day out of the hand of the Egyptians; and Israel saw the Egyptians dead upon the sea shore. And Israel saw that great work which the Lord did upon the Egyptians: and the people feared the Lord, and believed the Lord, and his servant Moses.

Then sang Moses and the children of Israel this song unto the Lord, and spake, saying,

\tract{We will sing unto the Lord: for he hath triumphed gloriously: the horse and his rider hath he thrown into the sea: the Lord is my strength and song, and he is become my salvation. ℣. He is my God, and I will praise him: my father's God, and I will exalt him. ℣. The Lord is a man of war: the Lord is his name.}

\textsc{Priest.} Let us pray.

\textsc{Deacon.} Let us bow the knee.

\textsc{Subdeacon.} Arise.

\lett{O}{God,} whose miracles of old we perceive to shine forth even in our times: who by the water of regeneration dost now work for the salvation of the nations, as thou didst deliver one people from the pursuit of the Egyptians by the power of thy right hand: vouchsafe; that the fulness of the whole world may be numbered among the sons of Abraham and made partakers of the dignity of Israel. Through.

\readingcitation{Prophecy V}{Isaiah 54:17}
\lett{T}{his} is the heritage of the servants of the \divineName{Lord}: and their righteousness is of me, saith the Lord.

\secondline{Ho, every one that thirsteth, come ye to the waters,
and he that hath no money; come ye, buy, and eat;
yea, come, buy wine and milk without money and without price.
Wherefore do ye spend money for that which is not bread?
and your labour for that which satisfieth not?
hearken diligently unto me, and eat ye that which is good,
and let your soul delight itself in fatness.
Incline your ear, and come unto me:
hear, and your soul shall live;
and I will make an everlasting covenant with you,
even the sure mercies of David.
Behold, I have given him for a witness to the people,
a leader and commander to the people.
Behold, thou shalt call a nation that thou knowest not,
and nations that knew not thee shall run unto thee
because of the Lord thy God, and for the Holy One of Israel;
for he hath glorified thee.}

Seek ye the Lord while he may be found,
call ye upon him while he is near:
let the wicked forsake his way,
and the unrighteous man his thoughts:
and let him return unto the Lord,
and he will have mercy upon him;
and to our God, for he will abundantly pardon.

For my thoughts are not your thoughts,
neither are your ways my ways, saith the Lord.
For as the heavens are higher than the earth,
so are my ways higher than your ways,
and my thoughts than your thoughts.
For as the rain cometh down, and the snow from heaven,
and returneth not thither, but watereth the earth,
and maketh it bring forth and bud,
that it may give seed to the sower, and bread to the eater:
so shall my word be that goeth forth out of my mouth:
it shall not return unto me void,
but it shall accomplish that which I please,
and it shall prosper in the thing whereto I sent it, saith the Lord almighty.

\textsc{Priest.} Let us pray.

\textsc{Deacon.} Let us bow the knee.

\textsc{Subdeacon.} Arise.

\lett{A}{lmighty} and everlasting God, multiply to the honour of thy name what thou didst swear to the faith of the fathers: and spread abroad the sons of promise by holy adoption; that as the Saints of old doubted not thine oath, so thy Church may now see in abundant measure the fulfilment of the same. Through.

\readingcitation{Prophecy VI}{Baruch 3:9}

\lett{H}{ear,} O Israel, the commandments of life: give ear to understand wisdom. How happeneth it Israel, that thou art in thine enemies' land, that thou art waxen old in a strange country, that thou art defiled with the dead, That thou art counted with them that go down into the grave? Thou hast forsaken the fountain of wisdom. For if thou hadst walked in the way of God, thou shouldest have dwelled in peace for ever. Learn where is wisdom, where is strength, where is understanding; that thou mayest know also where is length of days, and life, where is the light of the eyes, and peace.

Who hath found out her place? or who hath come into her treasures? Where are the princes of the heathen become, and such as ruled the beasts upon the earth; They that had their pastime with the fowls of the air, and they that hoarded up silver and gold, wherein men trust, and made no end of their getting? For they that wrought in silver, and were so careful, and whose works are unsearchable, They are vanished and gone down to the grave, and others are come up in their steads. Young men have seen light, and dwelt upon the earth: but the way of knowledge have they not known, Nor understood the paths thereof, nor laid hold of it: their children were far off from that way. It hath not been heard of in Chanaan, neither hath it been seen in Theman. The Agarenes that seek wisdom upon earth, the merchants of Meran and of Theman, the authors of fables, and searchers out of understanding; none of these have known the way of wisdom, or remember her paths.

O Israel, how great is the house of God! and how large is the place of his possession! Great, and hath none end; high, and unmeasurable. There were the giants famous from the beginning, that were of so great stature, and so expert in war. Those did not the Lord choose, neither gave he the way of knowledge unto them: But they were destroyed, because they had no wisdom, and perished through their own foolishness. Who hath gone up into heaven, and taken her, and brought her down from the clouds? Who hath gone over the sea, and found her, and will bring her for pure gold? No man knoweth her way, nor thinketh of her path.

But he that knoweth all things knoweth her, and hath found her out with his understanding: he that prepared the earth for evermore hath filled it with fourfooted beasts: He that sendeth forth light, and it goeth, calleth it again, and it obeyeth him with fear. The stars shined in their watches, and rejoiced: when he calleth them, they say, Here we be; and so with cheerfulness they shewed light unto him that made them. This is our God, and there shall none other be accounted of in comparison of him. He hath found out all the way of knowledge, and hath given it unto Jacob his servant, and to Israel his beloved. Afterward did he shew himself upon earth, and conversed with men.

\textsc{Priest.} Let us pray.

\textsc{Deacon.} Let us bow the knee.

\textsc{Subdeacon.} Arise.

\lett{O}{God,} who by the calling of the nations dost ever multiply thy Church: mercifully grant; that they whom thou dost cleanse by the waters of baptism may be defended by thy continual protection. Through.

\readingcitation{Prophecy VII}{Ezekiel 37:1}
\lett{I}{n those days:} The hand of the \divineName{Lord} was upon me, and carried me out in the spirit of the Lord, and set me down in the midst of the valley which was full of bones, and caused me to pass by them round about: and, behold, there were very many in the open valley; and, lo, they were very dry. And he said unto me, Son of man, can these bones live? And I answered, O Lord God, thou knowest.

Again he said unto me, Prophesy upon these bones, and say unto them, O ye dry bones, hear the word of the Lord. Thus saith the Lord God unto these bones; Behold, I will cause breath to enter into you, and ye shall live: and I will lay sinews upon you, and will bring up flesh upon you, and cover you with skin, and put breath in you, and ye shall live; and ye shall know that I am the Lord.

So I prophesied as I was commanded: and as I prophesied, there was a noise, and behold a shaking, and the bones came together, bone to his bone. And when I beheld, lo, the sinews and the flesh came up upon them, and the skin covered them above: but there was no breath in them. Then said he unto me, Prophesy unto the wind, prophesy, son of man, and say to the wind, Thus saith the Lord God; Come from the four winds, O breath, and breathe upon these slain, that they may live. So I prophesied as he commanded me, and the breath came into them, and they lived, and stood up upon their feet, an exceeding great army.

Then he said unto me, Son of man, these bones are the whole house of Israel: behold, they say, Our bones are dried, and our hope is lost: we are cut off for our parts. Therefore prophesy and say unto them, Thus saith the Lord God; Behold, O my people, I will open your graves, and cause you to come up out of your graves, and bring you into the land of Israel. And ye shall know that I am the Lord, when I have opened your graves, O my people, and brought you up out of your graves, and shall put my spirit in you, and ye shall live, and I shall place you in your own land, faith the Lord almighty.

\textsc{Priest.} Let us pray.

\textsc{Deacon.} Let us bow the knee.

\textsc{Subdeacon.} Arise.

\lett{O}{God,} who dost instruct us in the pages of either Testament in the celebrating of the paschal sacrament: grant unto us so to understand thy mercy; that, through the receiving of these present gifts, we may be established in the expectation of those which are to come. Through.

\readingcitation{Prophecy VIII}{Isaiah 4:1}
\lett{I}{n that day} seven women shall take hold of one man, saying: We will eat our own bread, and wear our own apparel: only let us be called by thy name, to take away our reproach. In that day shall the branch of the Lord be beautiful and glorious, and the fruit of the earth shall be excellent and comely for them that are escaped of Israel. And it shall come to pass, that he that is left in Zion, and he that remaineth in Jerusalem, shall be called holy, even every one that is written among the living in Jerusalem: when the Lord shall have washed away the filth of the daughters of Zion, and shall have purged the blood of Jerusalem from the midst thereof by the spirit of judgment, and by the spirit of burning. And the Lord will create upon every dwelling place of mount Zion, and upon her assemblies, a cloud and smoke by day, and the shining of a flaming fire by night: for upon all the glory shall be a defence. And there shall be a tabernacle for a shadow in the daytime from the heat, and for a place of refuge, and for a covert from storm and from rain.

\tract{My well-beloved hath a vineyard in a very fruitful hill. ℣. And he fenced it, and gathered out the stones thereof: and planted it with the choicest vine, and built a tower in the midst of it. ℣. And also made a wine-press therein: for the vineyard of the Lord of hosts is the house of Israel.}

\textsc{Priest.} Let us pray.

\textsc{Deacon.} Let us bow the knee.

\textsc{Subdeacon.} Arise.

\lett{O}{God,} who by the mouth of thy holy prophets hast manifested thyself in all the children of thy Church, to be in all places of thy dominion the sower of good seed and the husbandman of thine elect branches: grant unto thy peoples, who are named thy vineyard and thy harvest field; that being purged from the filthiness of thorns and briars, they may be made to bring forth worthy fruit in abundance. Through.

\readingcitation{Prophecy IX}{Exodus 12:1}
\lett{I}{n those days:} The \divineName{Lord} spake unto Moses and Aaron in the land of Egypt, saying, This month shall be unto you the beginning of months: it shall be the first month of the year to you.

Speak ye unto all the congregation of Israel, saying, In the tenth day of this month they shall take to them every man a lamb, according to the house of their fathers, a lamb for an house: and if the household be too little for the lamb, let him and his neighbour next unto his house take it according to the number of the souls; every man according to his eating shall make your count for the lamb. Your lamb shall be without blemish, a male of the first year: ye shall take it out from the sheep, or from the goats: and ye shall keep it up until the fourteenth day of the same month: and the whole assembly of the congregation of Israel shall kill it in the evening. And they shall take of the blood, and strike it on the two side posts and on the upper door post of the houses, wherein they shall eat it. And they shall eat the flesh in that night, roast with fire, and unleavened bread; and with bitter herbs they shall eat it. Eat not of it raw, nor sodden at all with water, but roast with fire; his head with his legs, and with the purtenance thereof. And ye shall let nothing of it remain until the morning; and that which remaineth of it until the morning ye shall burn with fire.

And thus shall ye eat it; with your loins girded, your shoes on your feet, and your staff in your hand; and ye shall eat it in haste: it is the Lord’s passover.

\textsc{Priest.} Let us pray.

\textsc{Deacon.} Let us bow the knee.

\textsc{Subdeacon.} Arise.

\lett{A}{lmighty} and everlasting God, who art wondrous in the dispensation of all thy works: let thy redeemed understand that the creation of the world in the beginning was not a work more excellent than the sacrifice, in the end of time, of Christ our Passover: Who liveth and reigneth with thee.

\readingcitation{Prophecy X}{Jonah 3:1}
\lett{I}{n those days:} The word of the Lord came unto Jonah the Prophet the second time, saying: Arise, go unto Nineveh, that great city, and preach unto it the preaching that I bid thee. So Jonah arose, and went unto Nineveh, according to the word of the Lord. Now Nineveh was an exceeding great city of three days’ journey. And Jonah began to enter into the city a day’s journey, and he cried, and said, Yet forty days, and Nineveh shall be overthrown.

So the people of Nineveh believed God, and proclaimed a fast, and put on sackcloth, from the greatest of them even to the least of them. For word came unto the king of Nineveh, and he arose from his throne, and he laid his robe from him, and covered him with sackcloth, and sat in ashes. And he caused it to be proclaimed and published through Nineveh by the decree of the king and his nobles, saying, Let neither man nor beast, herd nor flock, taste any thing: let them not feed, nor drink water: but let man and beast be covered with sackcloth, and cry mightily unto God: yea, let them turn every one from his evil way, and from the violence that is in their hands. Who can tell if God will turn and repent, and turn away from his fierce anger, that we perish not?

And God saw their works, that they turned from their evil way; and he had mercy on his people, even he the Lord, our God.

\textsc{Priest.} Let us pray.

\textsc{Deacon.} Let us bow the knee.

\textsc{Subdeacon.} Arise.

\lett{O}{God,} who hast united the diversity of nations in the confession of thy name: grant us both to will, and to have power to do the things which thou commandest; that the people whom thou hast called to everlasting life may agree in unity of faith and in godliness of conversation. Through.

\readingcitation{Prophecy XI}{Deuteronomy 31:22}
\lett{I}{n those days:} Moses wrote this song, and taught it the children of Israel. And he gave Joshua the son of Nun a charge, and said, Be strong and of a good courage: for thou shalt bring the children of Israel into the land which I sware unto them: and I will be with thee.

And it came to pass, when Moses had made an end of writing the words of this law in a book, until they were finished, that Moses commanded the Levites, which bare the ark of the covenant of the Lord, saying, Take this book of the law, and put it in the side of the ark of the covenant of the Lord your God, that it may be there for a witness against thee. For I know thy rebellion, and thy stiff neck: behold, while I am yet alive with you this day, ye have been rebellious against the Lord; and how much more after my death?

Gather unto me all the elders of your tribes, and your officers, that I may speak these words in their ears, and call heaven and earth to record against them. For I know that after my death ye will utterly corrupt yourselves, and turn aside from the way which I have commanded you; and evil will befall you in the latter days; because ye will do evil in the sight of the Lord, to provoke him to anger through the work of your hands.

And Moses spake in the ears of all the congregation of Israel the words of this song, until they were ended.

\tract{Give ear, O ye heavens, and I will speak: and hear, O earth, the words of my mouth. ℣. My doctrine shall drop as the rain: my speech shall distil as the dew. ℣. As the small rain upon the tender herb, and as the showers upon the grass: because I will publish the name of the Lord: ℣. Ascribe ye greatness unto our God: he is the rock, his work is perfect, for all his ways are judgment. ℣. A God of truth and without iniquity: just and right is the Lord.}

\textsc{Priest.} Let us pray.

\textsc{Deacon.} Let us bow the knee.

\textsc{Subdeacon.} Arise.

\lett{O}{God,} the exaltation of the humble and the strength of them that do stand, who by thy holy servant Moses didst vouchsafe so to instruct thy people in the chanting of thy sacred song, that the repeating of the law by them might likewise be for our instruction: stir up thy power among all the fulness of the nations whom thou hast justified, and grant us gladness, assuaging fear; that the sins of all may be blotted out by thy forgiveness, and that which was denounced in vengeance may turn to our salvation. Through.

\readingcitation{Prophecy XII}{Daniel 3:1}
\lett{I}{n those days:} Nebuchadnezzar the king made an image of gold, whose height was threescore cubits, and the breadth thereof six cubits: he set it up in the plain of Dura, in the province of Babylon. Then Nebuchadnezzar the king sent to gather together the princes, the governors, and the captains, the judges, the treasurers, the counsellors, the sheriffs, and all the rulers of the provinces, to come to the dedication of the image which Nebuchadnezzar the king had set up. Then the princes, the governors, and captains, the judges, the treasurers, the counsellors, the sheriffs, and all the rulers of the provinces, were gathered together unto the dedication of the image that Nebuchadnezzar the king had set up; and they stood before the image that Nebuchadnezzar had set up. Then an herald cried aloud, To you it is commanded, O people, nations, and languages, that at what time ye hear the sound of the cornet, flute, harp, sackbut, psaltery, dulcimer, and all kinds of musick, ye fall down and worship the golden image that Nebuchadnezzar the king hath set up: and whoso falleth not down and worshippeth shall the same hour be cast into the midst of a burning fiery furnace. Therefore at that time, when all the people heard the sound of the cornet, flute, harp, sackbut, psaltery, and all kinds of musick, all the people, the nations, and the languages, fell down and worshipped the golden image that Nebuchadnezzar the king had set up.

Wherefore at that time certain Chaldeans came near, and accused the Jews. They spake and said to the king Nebuchadnezzar, O king, live for ever. Thou, O king, hast made a decree, that every man that shall hear the sound of the cornet, flute, harp, sackbut, psaltery, and dulcimer, and all kinds of musick, shall fall down and worship the golden image: and whoso falleth not down and worshippeth, that he should be cast into the midst of a burning fiery furnace. There are certain Jews whom thou hast set over the affairs of the province of Babylon, Shadrach, Meshach, and Abed-nego; these men, O king, have not regarded thee: they serve not thy gods, nor worship the golden image which thou hast set up.

Then Nebuchadnezzar in his rage and fury commanded to bring Shadrach, Meshach, and Abed-nego. Then they brought these men before the king. Nebuchadnezzar spake and said unto them, Is it true, O Shadrach, Meshach, and Abed-nego, do not ye serve my gods, nor worship the golden image which I have set up? Now if ye be ready that at what time ye hear the sound of the cornet, flute, harp, sackbut, psaltery, and dulcimer, and all kinds of musick, ye fall down and worship the image which I have made; well: but if ye worship not, ye shall be cast the same hour into the midst of a burning fiery furnace; and who is that God that shall deliver you out of my hands? Shadrach, Meshach, and Abed-nego, answered and said to the king, O Nebuchadnezzar, we are not careful to answer thee in this matter. If it be so, our God whom we serve is able to deliver us from the burning fiery furnace, and he will deliver us out of thine hand, O king. But if not, be it known unto thee, O king, that we will not serve thy gods, nor worship the golden image which thou hast set up.

Then was Nebuchadnezzar full of fury, and the form of his visage was changed against Shadrach, Meshach, and Abed-nego: therefore he spake, and commanded that they should heat the furnace one seven times more than it was wont to be heated. And he commanded the most mighty men that were in his army to bind Shadrach, Meshach, and Abed-nego, and to cast them into the burning fiery furnace. Then these men were bound in their coats, their hosen, and their hats, and their other garments, and were cast into the midst of the burning fiery furnace. Therefore because the king’s commandment was urgent, and the furnace exceeding hot, the flame of the fire slew those men that took up Shadrach, Meshach, and Abed-nego. And these three men, Shadrach, Meshach, and Abed-nego, fell down bound into the midst of the burning fiery furnace. And they walked in the midst of the fire, praising God, and blessing the Lord.

\letuspray
\lett{A}{lmighty} and everlasting God, the only hope of the world, who, by the preaching of thy Prophets hast declared the mysteries of these present times: mercifully increase the devotion of thy people; forasmuch as without thine inspiration thy faithful can in no wise advance in virtue. Through.

    \par\noindent
	\centerline{\rule{0.25\textwidth}{0.4pt}}
	\par\noindent

\subbysub{Blessing of the Baptismal Font}\label{EasterFontBlessing}

\begin{rubric}
	These things being finished, if the church has a baptismal Font, the Priest who is to bless the Font puts on a violet Cope, and preceded by the Cross with lights and the blessed Candle lighted, descends with the Clergy and Ministers, vested, to the Font: and meanwhile the following is sung:
\end{rubric}

\tract{Like as the hart desireth the water-brooks: so longeth my soul after thee, O God. ℣. My soul is athirst for God, yea, even for the living God: when shall I come to appear before the presence of God? ℣. My tears have been my meat day and night, while they daily say unto me: Where is now thy God?}

\begin{rubric}
	Then the Priest, before he enters for the blessing of the Font, says this Prayer near by the Font:
\end{rubric}

\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}

\letuspray
\lett{A}{lmighty} and everlasting God, mercifully look upon the devotion of this people called to a new birth, who, like the hart, seek the fountain of thy waters: and mercifully grant; that the thirst of their faith may, by the mystery of Baptism, sanctify them in body and soul. Through.

\begin{rubric}
	Then he proceeds to the blessing of the Font, saying:
\end{rubric}

\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}

\letuspray
\lett{A}{lmighty} and everlasting God, be present at the mysteries, be present at the sacraments of thy great goodness: and send forth the spirit of adoption for the regenerating of the new peoples whom the font of baptism doth bring forth unto thee; that what is to be done by our humble ministry, may be effectually fulfilled by thy power. Through Jesus Christ, thy Son, our Lord: Who liveth and reigneth with thee in the unity of the Holy Ghost, God.

\begin{rubric}
	Then raising his voice to the tone of a Preface, he proceeds with joined hands:
\end{rubric}

Throughout all ages, world without end.

℟. Amen.

℣. The Lord be with you.

℟. And with thy spirit.

℣. Lift up your hearts.

℟. We lift them up unto the Lord.

℣. Let us give thanks unto our Lord God.

℟. It is meet and right so to do.
\par\noindent
It is very meet, right, and our bounden duty, that we should at all times and in all places give thanks unto thee: O Lord holy Father almighty, everlasting God: Who by thy invisible power dost wondrously give effect to the work of thy sacraments: And though we be unworthy to perform mysteries so great: Yet thou dost not leave us destitute of the gifts of thy grace, but mercifully inclinest thine ears even unto these our supplications. O God, whose Spirit, in the first beginnings of the world, moved over the waters: that even then the nature of water might conceive the virtue of sanctification. O God, who didst wash away in the waters the iniquities of a sinful world, and didst even in the outpouring of the flood shew forth in a figure our regeneration: that by the mystery of this same element, there should be an end to sin and likewise a beginning of virtue. Look, O Lord, upon the face of thy Church, and multiply in her the power of thy regeneration, for thou dost by the streams of thine abundant grace make glad thy city : and openest the fount of baptism throughout the whole world for the renewing of the nations: that, by the command of thy majesty, she may receive from the Holy Ghost the grace of thine only begotten Son.

\begin{rubric}
	Here the Priest divides the water in the form of a cross with his extended hand, which wipes at once with a linen cloth, saying:
\end{rubric}
\par\noindent
May he, by the secret mingling of his divine power, make fruitful this water prepared for the regeneration of mankind: that having received sanctification, and being born again a new creature, there may come forth an offspring of heaven from the spotless womb of this divine fountain: And that all, whether sex divide them in body, or age in time, may alike be brought forth into one childhood by grace, their mother. Far hence, O Lord, at thy command let every unclean spirit depart: far hence be all the wickedness of the craft of the devil. Let no power of the enemy here have place or portion: let it not ensnare and compass us around: let it not waylay us in secret: let it not taint us with corruption.

\begin{rubric}
	He touches the water with his hand.
\end{rubric}
\par\noindent
May this holy and undefiled creature be free from every assault of the enemy, and purified by the departure of all iniquity. May it be a living fountain, water that doth regenerate, a purifying stream: that all who shall be washed in this laver of salvation may, by the operation of the Holy Ghost within them, obtain the grace of perfect purification.

\begin{rubric}
	He makes three crosses over the Font, saying:
\end{rubric}
\par\noindent
Wherefore I bless thee, O creature of water, by the living {\ding{64}} God, by the true {\ding{64}} God, by the holy {\ding{64}} God: by God who in the beginning through his word divided thee from the dry land: whose spirit moved upon thee.

\begin{rubric}
	Here he divides the water with his hand, and scatters it towards the four quarters of the world, saying:
\end{rubric}
\par\noindent
Who made thee to flow from the fountain of paradise, and commanded thee to water the whole earth with thy tour rivers. Who in the desert bestowed upon thee sweetness when thou wast bitter, that men might drink, and brought thee forth from the rock for the thirsting people. I ble {\ding{64}} ss thee also by Jesus Christ his only Son, our Lord: who in Cana of Galilee by a wondrous miracle did change thee through his power into wine. Who walked upon thee with his feet: and was baptized in thee by John in Jordan. Who brought thee forth together with blood from his side: and commanded his disciples, that believing they should be baptized in thee, saying: Go ye, teach all nations, baptizing them in the name of the Father, and of the Son, and of the Holy Ghost.

\begin{rubric}
	He changes his voice, and proceeds in the tone of the Lesson:
\end{rubric}
\par\noindent
Do thou, almighty God, mercifully assist us, who keep these thy commandments: do thou graciously breathe upon us.

\begin{rubric}
	He breathes thrice upon the water in the form of a cross, saying:
\end{rubric}
\par\noindent
Do thou with thy mouth bless this element of water: that as by nature it hath power to cleanse and wash the body, so also it may be effectual for the purifying of the soul.

\begin{rubric}
	Here the Priest lowers the Candle into the water a little way: and resuming the tone of the Preface, says:
\end{rubric}
\par\noindent
May the power of the Holy Ghost descend upon the fulness of this font.

\begin{rubric}
	Then withdrawing the Candle from the water, he again dips it more deeply, and repeats in a somewhat higher voice, \emph{May the power}. Then he again withdraws the Candle from the water, and the third time dips it to the bottom, repeating in a still higher voice: \emph{May the power}, as above.
	
	Then breathing three times on the water in this form \textsc{Ψ} he proceeds:
\end{rubric}
\par\noindent
And make the whole substance of this water to be fruitful unto regeneration.

\begin{rubric}
	Here the Candle is taken out of the water, and he proceeds:
\end{rubric}
\par\noindent
Here may the stains of every sin be blotted out: here may nature, created in thine image, be restored to the honour of its first estate, and cleansed from all the defilement of its old condition: that every man who cometh to this sacrament of regeneration may be born again unto the new childhood of true innocence.

\begin{rubric}
	He says that which follows in the reading voice:
\end{rubric}
\par\noindent
Through Jesus Christ thy Son, our Lord: Who shall come to judge the quick and the dead, and the world by fire. \textit{Amen.}

\begin{rubric}
	Then some of the water that has been blessed is sprinkled by the assisting Priests over the People.
	
	And meantime one of the ministers of the church takes some of the water in a vessel, for sprinkling in houses and other places. Which done, the Priest who is blessing the Font, pours some of the Oil of Catechumens into the water in the form of a cross, saying in an audible voice:
\end{rubric}
\par\noindent
May this font be sanctified and made fruitful by the Oil of salvation, for such as shall be born again therefrom, unto life everlasting. \textit{Amen.}

\begin{rubric}
	Then, in the same manner, he pours in some of the Chrism, saying:
\end{rubric}
\par\noindent
May this inpouring of the Chrism of our Lord Jesus Christ, and of the Holy Ghost the Paraclete, be wrought in the name of the holy Trinity. \textit{Amen.}

\begin{rubric}
	Lastly he takes both phials of the said holy Oil and the Chrism and pours in some of both together in the form of a cross, saying:
\end{rubric}
\par\noindent
May this commingling of the Chrism of sanctification, and of the Oil of unction and of the Water of baptism be likewise wrought in the name of the Fa {\ding{64}} ther, and of the S {\ding{64}} on, and of the Holy {\ding{64}} Ghost. \textit{Amen.}

\begin{rubric}
	Then he mingles the Oil itself with the water, and with his hand scatters it over the whole Font. If there be any to be baptized, he baptizes them in the usual manner.
\end{rubric}

\begin{rubric}
	Then as the Priest and Ministers return to the Altar, the Litany is sung by two Cantors, and the Choir repeat it also together, as is directed below.
\end{rubric}

    \par\noindent
	\centerline{\rule{0.25\textwidth}{0.4pt}}
	\par\noindent

\subbysub{Paschal Litany}\label{PaschalLitany}

\begin{rubric}
	But where there is no baptismal Font, when the last Prophecy with its Collect is ended, the Celebrant lays aside his Chasuble, and with the Ministers lies prostrate before the Altar: and all kneeling, the Litany is chanted by two Cantors in the midst of the Choir, and both sides of the Choir respond together.
	
	But when they come to the \emph{℣. We sinners}, the Priest and his Ministers rise, and proceed to the sacristy, where they put on white vestments for the solemn Mass, and meanwhile the candles on the Altar are lighted.
	
	At the end of the Litany, \emph{Kyrie eleison} is sung solemnly, and repeated, as is usual at Mass.
\end{rubric}

\elcoldent{Lord, have mercy upon us.}{Kyrie, eléison.}
\elcol{℟. Christ, have mercy upon us.}{℟. Christe, eléison.}
\elcoldent{Lord, have mercy upon us.}{Kyrie, eléison.}
\elcoldent{O Christ hear us.}{Christe, audi nos.}
\elcol{℟. O Christ graciously hear us.}{℟. Christe, exáudi nos.}
\elcoldent{O God the Father of heaven,}{Pater de c{\ae}lis, Deus,}
\elcol{℟. Have mercy upon us.}{℟. Miserére nobis.}
\elcoldent{O God the Son, Redeemer of the world,}{Fili, Redémptor mundi, Deus,}
\elcol{℟. Have mercy upon us.}{℟. Miserére nobis.}
\elcoldent{O God the Holy Ghost,}{Spíritus Sancte, Deus,}
\elcol{℟. Have mercy upon us.}{℟. Miserére nobis.}
\elcoldent{Holy Trinity one God,}{Sancta Trínitas, unus Deus,}
\elcol{℟. Have mercy upon us.}{℟. Miserére nobis.}
\elcoldent{Holy Mary,}{Sancta María,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{Holy Mother of God,}{Sancta Dei Génitrix,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{Holy Virgin of virgins,}{Sancta Virgo vírginum,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{Holy Michael,}{Sancte Micha{\"e}l,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{Holy Gabriel,}{Sancte Gábri{\"e}l,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{Holy Raphael,}{Sancte Rápha{\"e}l,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{All ye holy Angels and Archangels,}{Omnes sancti Angeli et Archángeli,}
\elcol{℟. Pray for us.}{℟. Oráte pro nobis.}
\elcoldent{All ye holy orders of blessed Spirits,}{Omnes sancti beatórum Spírituum órdines,}
\elcol{℟. Pray for us.}{℟. Oráte pro nobis.}
\elcoldent{St. John Baptist,}{Sancte Joánnes Baptísta,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{St. Joseph,}{Sancte Josephe,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{All ye holy Patriarchs and Prophets,}{Omnes sancti Patriárch{\ae} et Prophét{\ae},}
\elcol{℟. Pray for us.}{℟. Oráte pro nobis.}
\elcoldent{St. Peter,}{Sancte Petre,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{St. Paul,}{Sancte Paule,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{St. Andrew,}{Sancte Andréa,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{St. John,}{Sancte Joánnes,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{All ye holy Apostles and Evangelists,}{Omnes sancti Apóstoli et Evangelíst{\ae},}
\elcol{℟. Pray for us.}{℟. Oráte pro nobis.}
\elcoldent{All ye holy Disciples of the Lord,}{Omnes sancti Discípuli Dómini,}
\elcol{℟. Pray for us.}{℟. Oráte pro nobis.}
\elcoldent{St. Stephen,}{Sancte Stéphane,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{St. Lawrence,}{Sancte Laurénti,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{St. Vincent,}{Sancte Vincénti,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{All ye holy Martyrs,}{Omnes sancti Mártyres,}
\elcol{℟. Pray for us.}{℟. Oráte pro nobis.}
\elcoldent{St. Sylvester,}{Sancte Silvéster,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{St. Gregory,}{Sancte Gregóri,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{St. Augustine,}{Sancte Augustíne,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{All ye holy Bishops and Confessors,}{Omnes sancti Pontífices et Confessóres,}
\elcol{℟. Pray for us.}{℟. Oráte pro nobis.}
\elcoldent{All ye holy Doctors,}{Omnes sancti Doctóres,}
\elcol{℟. Pray for us.}{℟. Oráte pro nobis.}
\elcoldent{Holy St. Anthony,}{Sancte Antóni,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{Holy St. Benedict,}{Sancte Benedícte,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{All ye holy Priests and Levites,}{Omnes sancti Sacerdótes et Levít{\ae},}
\elcol{℟. Pray for us.}{℟. Oráte pro nobis.}
\elcoldent{All ye holy Monks and Hermits,}{Omnes sancti Mónachi et Eremít{\ae},}
\elcol{℟. Pray for us.}{℟. Oráte pro nobis.}
\elcoldent{St. Mary Magdalene,}{Sancta María Magdaléna,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{St. Agnes,}{Sancta Agnes,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{St. Cecilia,}{Sancta C{\ae}cília,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{St. Agatha,}{Sancta Agatha,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{St. Anastasia,}{Sancta Anastásia,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{All ye holy Virgins and Widows,}{Omnes sanct{\ae} Vírgines et Vídu{\ae},}
\elcol{℟. Pray for us.}{℟. Oráte pro nobis.}
\elcoldent{All ye Saints of God,}{Omnes Sancti et Sanct{\ae} Dei,}
\elcol{℟. Intercede for us.}{℟. Intercédite pro nobis.}
\elcoldent{Be thou merciful,}{Propítius esto,}
\elcol{℟. Spare us, O Lord.}{℟. Parce nobis, Dómine.}
\elcoldent{Be thou merciful,}{Propítius esto,}
\elcol{℟. Hear us, O Lord.}{℟. Exáudi nos, Dómine.}
\elcoldent{From all evil,}{Ab omni malo,}
\elcol{℟. Deliver us, O Lord.}{℟. Líbera nos, Dómine.}
\elcoldent{From all sin,}{Ab omni peccáto,}
\elcol{℟. Deliver us, O Lord.}{℟. Líbera nos, Dómine.}
\elcoldent{From everlasting death,}{A morte perpétua,}
\elcol{℟. Deliver us, O Lord.}{℟. Líbera nos, Dómine.}
\elcoldent{By the mystery of thy holy Incarnation,}{Per mystérium sanct{\ae} Incarnatiónis tu{\ae},}
\elcol{℟. Deliver us, O Lord.}{℟. Líbera nos, Dómine.}
\elcoldent{By thine advent,}{Per Advéntum tuum,}
\elcol{℟. Deliver us, O Lord.}{℟. Líbera nos, Dómine.}
\elcoldent{By thy nativity,}{Per Nativitátem tuam,}
\elcol{℟. Deliver us, O Lord.}{℟. Líbera nos, Dómine.}
\elcoldent{By thy baptism and holy fasting,}{Per Baptísmum et sanctum Jejúnium tuum,}
\elcol{℟. Deliver us, O Lord.}{℟. Líbera nos, Dómine.}
\elcoldent{By thy cross and passion,}{Per Crucem et Passiónem tuam,}
\elcol{℟. Deliver us, O Lord.}{℟. Líbera nos, Dómine.}
\elcoldent{By thy death and burial,}{Per Mortem et sepultúram tuam,}
\elcol{℟. Deliver us, O Lord.}{℟. Líbera nos, Dómine.}
\elcoldent{By thy holy resurrection,}{Per sanctam Resurrectiónem tuam,}
\elcol{℟. Deliver us, O Lord.}{℟. Líbera nos, Dómine.}
\elcoldent{By thy wonderful ascension,}{Per admirábilem Ascensiónem tuam,}
\elcol{℟. Deliver us, O Lord.}{℟. Líbera nos, Dómine.}
\elcoldent{By the coming of the Holy Ghost the Paraclete,}{Per advéntum Spíritus Sancti Parácliti,}
\elcol{℟. Deliver us, O Lord.}{℟. Líbera nos, Dómine.}
\elcoldent{In the day of judgment,}{In die judícii,}
\elcol{℟. Deliver us, O Lord.}{℟. Líbera nos, Dómine.}
\elcoldent{We sinners,}{Peccatóres,}
\elcol{℟. Beseech thee, hear us.}{℟. Te rogámus, audi nos.}
\elcoldent{That thou wouldest spare us,}{Ut nobis parcas,}
\elcol{℟. We beseech thee, hear us.}{℟. Te rogámus, audi nos.}
\elcoldent{That thou wouldest vouchsafe to govern and preserve thy holy Church,}{Ut Ecclésiam tuam sanctam régere et conserváre dignéris,}
\elcol{℟. We beseech thee, hear us.}{℟. Te rogamus, audi nos.}
\elcoldent{That thou wouldest vouchsafe to preserve our Apostolic lord and all orders of the Church in holy religion,}{Ut Domnum Apostólicum, et omnes ecclesiásticos órdines in sancta religióne conserváre dignéris,}
\elcol{℟. We beseech thee, hear us.}{℟. Te rogamus, audi nos.}
\elcoldent{That thou wouldest vouchsafe to humble the enemies of holy Church,}{Ut inimícos sanct{\ae} Ecclési{\ae} humiliáre dignéris,}
\elcol{℟. We beseech thee, hear us.}{℟. Te rogamus, audi nos.}
\elcoldent{That thou wouldest vouchsafe to give unto Christian kings and rulers peace and true concord,}{Ut régibus et princípibus christiánis pacem et veram concórdiam donáre dignéris,}
\elcol{℟. We beseech thee, hear us.}{℟. Te rogamus, audi nos.}
\elcoldent{That thou wouldest vouchsafe to strengthen, and preserve us in thy holy service,}{Ut nosmetípsos in tuo sancto servítio confortáre et conserváre dignéris,}
\elcol{℟. We beseech thee, hear us.}{℟. Te rogamus, audi nos.}
\elcoldent{That thou wouldest vouchsafe to reward all our benefactors with everlasting blessings,}{Ut ómnibus benefactóribus nostris sempitérna bona retríbuas,}
\elcol{℟. We beseech thee, hear us.}{℟. Te rogamus, audi nos.}
\elcoldent{That thou wouldest vouchsafe to give and preserve the fruits of the earth,}{Ut fructus terr{\ae} dare et conserváre dignéris,}
\elcol{℟. We beseech thee, hear us.}{℟. Te rogamus, audi nos.}
\elcoldent{That thou wouldest vouchsafe to grant unto all the faithful departed rest eternal,}{Ut ómnibus fidélibus defúnctis réquiem {\ae}térnam donáre dignéris,}
\elcol{℟. We beseech thee, hear us.}{℟. Te rogamus, audi nos.}
\elcoldent{That thou wouldest vouchsafe graciously to hear us,}{Ut nos exáudire dignéris,}
\elcol{℟. We beseech thee, hear us.}{℟. Te rogamus, audi nos.}
\elcoldent{O Lamb of God, who takest away the sins of the world,}{Agnus Dei, qui tollis peccáta mundi,}
\elcol{℟. Spare us, O Lord, hear us.}{℟. Parce nobis, Dómine.}
\elcoldent{O Lamb of God, who takest away the sins of the world,}{Agnus Dei, qui tollis peccáta mundi,}
\elcol{℟. Graciously hear us, O Lord.}{℟. Exáudi nos, Dómine.}
\elcoldent{O Lamb of God, who takest away the sins of the world,}{Agnus Dei, qui tollis peccáta mundi,}
\elcol{℟. Have mercy upon us.}{℟. Miserére nobis.}
\elcoldent{O Christ, hear us.}{Christe, audi nos.}
\elcol{℟. O Christ, graciously hear us.}{℟. Christe, exáudi nos.}

\subbysub{Holy Mass}

\begin{rubric}
	Here the Cantors solemnly intone \emph{Kyrie, eléison}, \emph{Christe, eléison}, \emph{Kyrie, eléison}, repeating each invocation thrice.
	
	Meanwhile the Priest with the Ministers in white vestments approaches the Altar: and having said the Psalm \emph{Give sentence with me, O God}, with \emph{Glory be}, he makes the Confession: then going up, he kisses the Altar, and censes it in the usual manner: and after the Choir has finished \emph{Kyrie, eléison}, he solemnly intones \emph{Gloria in excelsis}, and the bells are rung. Then the Priest says:
\end{rubric}

\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}
\elcol{℣. Let us pray.}{℣. Orémus.}

\collect
\lett{O}{God,} who dost enlighten this most sacred night with the glory of the Resurrection of the Lord: preserve in the new offspring of thy family the spirit of adoption, which thou hast given; that they being renewed in body and soul may render thee pure service. Through the same.

\readingcitation{Epistle}{Colossians 3:1}
\lett{B}{rethren:} If ye be risen with Christ, seek those things which are above, where Christ sitteth on the right hand of God. Set your affection on things above, not on things on the earth. For ye are dead, and your life is hid with Christ in God. When Christ, who is our life, shall appear, then shall ye also appear with him in glory.

\begin{rubric}
	The Epistle ended, the Celebrant begins: \emph{Alleluia}.
	
	And be sings it through thrice, raising his voice each time: the Choir repeats it each time in the same tone.
	
	Afterwards the Choir continues:
\end{rubric}
\par\noindent
℣. O give thanks unto the Lord, for he is gracious: because his mercy endureth for ever.

\tract{O praise the Lord, all ye heathen: praise him, all ye nations. ℣. For his merciful kindness is ever more and more towards us: and the truth of the Lord endureth for ever.}

\begin{rubric}
	At the Gospel, lights are not carried, but only incense. A blessing is asked, and other things are done as usual.
\end{rubric}

\readingcitation{Gospel}{Matthew 28:1}
\lett{I}{n} the end of the sabbath, as it began to dawn toward the first day of the week, came Mary Magdalene and the other Mary to see the sepulchre. And, behold, there was a great earthquake: for the angel of the Lord descended from heaven, and came and rolled back the stone from the door, and sat upon it. His countenance was like lightning, and his raiment white as snow: and for fear of him the keepers did shake, and became as dead men. And the angel answered and said unto the women, Fear not ye: for I know that ye seek Jesus, which was crucified. He is not here: for he is risen, as he said. Come, see the place where the Lord lay. And go quickly, and tell his disciples that he is risen from the dead; and, behold, he goeth before you into Galilee; there shall ye see him: lo, I have told you.

\begin{rubric}
	The Creed is not said, but, the Gospel ended, the Priest says: \emph{The Lord be with you}, then: \emph{Let us pray}. The Offertory Verse is not said. At the Lavabo \emph{Gloria Patri} is said.
\end{rubric}

\secret
\lett{R}{eceive,} we beseech thee O Lord, the prayers of thy people together with the offering of these sacrifices: that those things which we have begun in these paschal mysteries may by thine operation be profitable for our healing in eternity. Through.

\begin{rubric}
    	\emph{The Peace of the Lord} is said, but the Kiss of Peace is not given.
    	
    	\emph{Agnus Dei} is not said, nor the Postcommunion, but the accustomed Prayers before the Communion are said.
\end{rubric}

\begin{rubric}
	After the receiving of the Sacrament, for Vespers in the choir is sung the Antiphon: \emph{Alleluia, alleluia, alleluia}, and
\end{rubric}

\subbysub{Psalm 117. \textit{Laudate Dominum}}

\elcol{\lett{O}{praise} the \textsc{Lord}, all ye heathen : praise him, all ye nations.}{\lett{L}{aud\smash{á}te} Dóminum, omnes gentes: * laudáte eum, omnes pópuli:}
\elcol{For his merciful kindness is ever more and more toward us : and the truth of the \textsc{Lord} endureth for ever. Praise the \textsc{Lord}.}{Quóniam confirmáta est super nos misericórdia ejus: * et véritas Dómini manet in {\ae}térnum.}
\elcol{Glory be to the Father, and to the Son, *
    and to the Holy Ghost:}{Glória Patri, et Fílio, * et Spirítui Sancto.}
\elcol{As it was in the beginning, is now, and ever shall be, *
    world without end. Amen.}{Sicut erat in princípio, et nunc, et semper, * et in sǽcula s{\ae}culórum. Amen.}
    
\antiphon{}{Alleluia, alleluia, alleluia.}

\begin{rubric}
	The Chapter, Hymn, and Verse are not said, but the Celebrant begins at once to sing the Antiphon to the Magnificat:
\end{rubric}

\lett{I}{n the end of the Sabbath.}

\vspace{1ex}

\begin{rubric}
	And the Choir continues:
\end{rubric}

As it began to dawn toward the first day of the week: came Mary Magdalene and the other Mary to see the sepulchre, alleluia.

\begin{rubric}
	Then the Magnificat is sung with \emph{Gloria Patri} at the end, and the censing is done, as at other times at Evensong.
\end{rubric}

\begin{rubric}
	The Antiphon having been repeated, the Celebrant says:
\end{rubric}

\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}

\letuspray
\lett{P}{our} forth upon us, O Lord, the Spirit of thy charity: that as thou hast fulfilled us with these paschal sacraments, so of thy goodness thou wouldest make us to be of one heart and mind. Through ... in the unity of the same.

\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}

\begin{rubric}
	And the Deacon, turning himself to the people, sings:
\end{rubric}

    \elcol{℣. Depart in peace, alleluia, alleluia.}{℣. Ite, Missa est, allelúja, allelúja.}
    \elcol{℟. Thanks be to God, alleluia, alleluia.}{℟. Deo grátias, allelúja, allelúja.}

\begin{rubric}
	And the Priest, having said \emph{Let this my bounden duty}, gives the Blessing in the usual manner.
\end{rubric}