\fancyhead[C]{\LARGE Liturgical Hymns}

\section{Daily Office Services}
	
	\subsection{Office Service One -\\Our Worship, Praise, and Thanksgiving\\unto the Glory of God Alone}\fancyhead[RO,LE]{\textit{Office I}}
	
		\begin{inhead}
			Mattins
		\end{inhead}
		
		\subsubsection{O Come, Let Us Sing}\fancyhead[RE,LO]{O Come, Let Us Sing}
			\gregorioscore{resources/gabc/Extra/Venite.gabc}
			
			\psanum{2}Let us come before his presence with • thanksgiving : and shew ourselves • glad in him with psalms.
			
			\psanum{3}For the Lord is a • great God : and a great • King above all gods.
			
			\psanum{4}In his hand are all the corners • of the earth : and the strength of the hills is • his also.
			
			\psanum{5}The sea is his, and • he made it : and his hands prepared • the dry land.
		
			\psanum{6}O come, let us worship and • fall down : and kneel before the • Lord our Maker.
			
			\psanum{7}For he is the • Lord our God : and we are the people of his pasture, and the • sheep of his hand.
			
			\psanum{8}To-day if ye will hear his voice, harden • not your hearts : as in the provocation, and as in the day of temptation • in the wilderness.
			
			\psanum{9}When your fathers • tempted me : proved me, • and saw my works.
			
			\psanum{10}Forty years long was I grieved with this generation, • and said : It is a people that do err in their hearts, for they • have not known my ways;
			
			\psanum{11}Unto whom I sware • in my wrath : that they should not enter • into my rest.
			
			℣. Glory be to the Father, and • to the Son * and • to the Holy Ghost.
			
			℟. As it was in the beginning, is now, and ever • shall be * world without • end. Amen.
			
	
		\clearpage\subsubsection{Christ Our Passover}\fancyhead[RE,LO]{Christ Our Passover}
			\gregorioscore{resources/gabc/Extra/PaschaNostrum.gabc}
	
			Not with old leaven, neither with the leaven of malice and • wickedness * but with the unleavened bread of sin • cerity and truth.
	
			Christ being raised from the dead dieth • no more * death hath no more do • minion over him.
    
    		For in that he died, he died unto • sin once * but in that he liveth, he • liveth unto God.
    
    		Likewise reckon ye also yourselves to be dead indeed • unto sin * but alive unto God through • Jesus Christ our Lord.
    
    		Christ is risen • from the dead * and become the first • fruits of them that slept.

    		For since by • man came death * by man came also the resur • rection of the dead.
    
    		For as in Adam • all die : even so in Christ shall • all be made alive.
    
    		℣. Glory be to the Father, and • to the Son * and • to the Holy Ghost;
    
    		℟. As it was in the beginning, is now, and ever • shall be * world • without end. Amen.
	
		\clearpage\subsubsection{We Praise Thee, O God}\fancyhead[RE,LO]{We Praise Thee, O God}
			\gregorioscore{resources/gabc/Extra/TeDeum1.gabc}
			
			To thee all Angels cry aloud; the Heavens, and all the • Pow'rs therein;  * To thee Cherubim and Seraphim contin • ually do cry,\par
    		Holy, Holy, Holy, Lord God of • Sabaoth; * Heaven and earth are full of the Majesty • of thy glory.\par
    		The glorious company of the Apostles • praise thee. * The goodly fellowship of the • Prophets praise thee.\par
    		The noble army of Martyrs • praise thee. * The holy Church throughout all the world • doth acknowledge thee;\par
    		The Father of an infinite Majesty; Thine honourable, true and only • Son; * Also the Holy • Ghost the Comforter.\par
    		\gregorioscore{resources/gabc/Extra/TeDeum2.gabc}
    		When thou tookest upon thee • to deliver man, * thou didst not ab • hor the Virgin's womb.\par
    		When thou hadst overcome the • sharpness of death, * thou didst open the Kingdom of Heaven to • all believers.\par
    		Thou sittest at the • right hand of God, * in the glory • of the Father.\par
    		We believe that thou shalt • come to be our Judge. * We therefore pray thee, help thy servants, whom thou hast redeemed • with thy precious blood.\par
    		Make them to be • numbered with thy Saints, * in glory • everlasting.\par
    		\clearpage
    		\gregorioscore{resources/gabc/Extra/TeDeum3.gabc}
			Day by day we magnify • thee; * And we worship thy Name ever, • world without end.\par
			Vouchsafe, O Lord, to keep us this day without • sin. * O Lord, have mercy upon us, have mer • cy upon us.\par
			O Lord, let thy mercy lighten upon us, as our trust is in • thee. * O Lord, in thee have I trusted; let me never • be confounded.
			

		\clearpage\subsubsection{Bless Ye the Lord}\fancyhead[RE,LO]{Bless Ye the Lord}
			\gregorioscore{resources/gabc/Extra/Benedicite.gabc}
			
			O ye Angels of the Lord, bless • ye the Lord: * praise him, and magni • fy him for ever.\par
			
			\lett{O}{ye} Heavens, bless • ye the Lord: * praise him, and magni • fy him for ever.\par
			\secondline{O ye Waters that be above the firmament, bless • ye the Lord: * praise him, and magni • fy him for ever.}
			\thirdline{O all ye Powers of the Lord, bless • ye the Lord: * praise him, and magni • fy him for ever.}
			O ye Sun and Moon, bless • ye the Lord: * praise him, and magni • fy him for ever.\par
    		O ye Stars of heaven, bless • ye the Lord: * praise him, and magni • fy him for ever.\par
    		O ye Showers and Dew, bless • ye the Lord: * praise him, and magni • fy him for ever.\par
    		O ye winds of God, bless • ye the Lord: * praise him, and magni • fy him for ever.\par
    		O ye Fire and Heat, bless • ye the Lord: * praise him, and magni • fy him for ever.\par
    		O ye Winter and Summer, bless • ye the Lord: * praise him, and magni • fy him for ever.\par
    		O ye Dews and Frosts, bless • ye the Lord: * praise him, and magni • fy him for ever.\par
    		O ye Frost and Cold, bless • ye the Lord: * praise him, and magni • fy him for ever.\par
    		O ye Ice and Snow, bless ye the Lord * praise him, and magni • fy him for ever.\par
    		O ye Nights and Days, bless • ye the Lord: * praise him, and magni • fy him for ever.\par
    		O ye Light and Darkness, bless • ye the Lord: * praise him, and magni • fy him for ever.\par
    		O ye Lightnings and Clouds, bless ye the Lord * praise him, and magni • fy him for ever.\par
			\lett{O}{let} the Earth • bless the Lord: * yea, let it praise him, and magni • fy him for ever.\par
			\secondline{O ye Mountains and Hills, bless • ye the Lord: * praise him, and magni • fy him for ever.}
			\clearpage
			\gregorioscore{resources/gabc/Extra/Benedicite_secondpage.gabc}
			%\thirdline{O all ye Green Things upon the earth, bless • ye the Lord: * praise him, and magni • fy him for ever.}
			O ye Wells, bless • ye the Lord: * praise him, and magni • fy him for ever.\par
    		O ye Seas and Floods, bless • ye the Lord: * praise him, and magni • fy him for ever.\par
    		O ye Whales, and all that move in the waters, bless • ye the Lord: * praise him, and magni • fy him for ever.\par
    		O all ye Fowls of the air, bless • ye the Lord: * praise him, and magni • fy him for ever.\par
    		O all ye Beasts and Cattle, bless • ye the Lord: * praise him, and magni • fy him for ever.\par
    		O ye Children of Men, bless • ye the Lord: * praise him, and magni • fy him for ever.\par
    		\lett{O}{let} Israel • bless the Lord: * praise him, and magni • fy him for ever.\par
    		\secondline{O ye Priests of the Lord, bless • ye the Lord: * praise him, and magni • fy him for ever.}
    		\thirdline{O ye Servants of the Lord, bless • ye the Lord: * praise him, and magni • fy him for ever.}
    		O ye Spirits and Souls of the Righteous, bless • ye the Lord: * praise him, and magni • fy him for ever.\par
    		O ye holy and humble Men of heart, bless • ye the Lord: * praise him, and magni • fy him for ever.\par
    		\lett{L}{et} us bless the Father, and the Son, and the • Holy Ghost: * praise him, and magnif • y him for ever.
			
			
		\clearpage\subsubsection{Blessed Be the Lord}\fancyhead[RE,LO]{Blessed Be the Lord}
			\gregorioscore{resources/gabc/Extra/Benedictus.gabc}
			
		\clearpage\subsubsection{Athanasian Creed}\fancyhead[RE,LO]{Athanasian Creed}
			\gregorioscore{resources/gabc/Extra/Athanasian.gabc}
	
			\psanum{2}Which Faith except a man keep whole • and undefiled : without doubt he will perish • eternally.
	
			\psanum{3}Now the • Catholick Faith is this : that we worship one God in Trinity, and the Trinity • in Unity;
	
    		\psanum{4}Neither • confusing the Persons : nor divid • ing the substance.\par
    		\psanum{5}For there is one Person of the Father, • another of the Son : another • of the Holy Ghost;\par
    		\psanum{6}But the Godhead of the Father, and of the Son, and of the • Holy Ghost is all one : the glory equal, the majesty • co-eternal.\par
    		\psanum{7}Such as the Father is, • such is the Son : and such • is the Holy Ghost.\par
    		\psanum{8}The Father uncreated, the Son • uncreated : the Holy Ghost • uncreated;\par
    		\psanum{9}The Father infinite, • the Son infinite : the Holy • Ghost infinite;\par
    		\psanum{10}The Father eternal, the • Son eternal : the Holy • Ghost eternal.\par
    		\psanum{11}And yet there are not • three eternals : • but one eternal\par
    		\psanum{12}As also there are not three uncreated, nor • three infinites : but one infinite, and one • uncreated.\par
    		\psanum{13}So likewise the Father is almighty, • the Son almighty : the Holy • Ghost almighty;\par
    		\psanum{14}And yet there are not • three almighties : • but one almighty.\par
    		\psanum{15}So the Father • is God, the Son God : • the Holy Ghost God;\par
    		\psanum{16}And yet there • are not three Gods : • but one God.\par
    		\psanum{17}So the Father is • Lord, the Son Lord : • the Holy Ghost Lord;\par
    		\psanum{18}And yet there • are not three Lords : • but one Lord.\par
    		\psanum{19}For like as we are compelled by the • Christian verity : to confess each Person by himself to • be both God and Lord;\par
    		\psanum{20}So are we forbidden by the • Catholick religion : to speak of • three Gods or three Lords.\par
    		\psanum{21}The Father • is made of none : nor created, • nor begotten.\par
    		\psanum{22}The Son is • of the Father : not made, nor created, • but begotten.\par
    		\psanum{23}The Holy Ghost is • of the Father : not made, nor created, nor begotten, • but proceeding.\par
    		\psanum{24}There is therefore one Father, not three Fathers; • one Son, not three Sons : one Holy Ghost, • not three Holy Ghosts.\par
    		\psanum{25}And in this Trinity there is no • before or after : • no greater or less;\par
    		\gregorioscore{resources/gabc/Extra/Athanasian2.gabc}\par
    		%\psanum{26}But all three Persons are co • eternal together : • and co-equal.\par
    		\psanum{27}So that in all ways, • as is aforesaid : both the Trinity is to be worshipped in Unity, and the Unity • in Trinity.\par
    		\psanum{28}He therefore that • would be saved : let him thus think • of the Trinity.

    		\lett{F}{urthermore} it is necessary to • eternal salvation : that he also believe faithfully the Incarnation of • our Lord Jesus Christ.\par
    		\secondline{\psanum{30}Now the right faith is that we • believe and confess : that our Lord Jesus Christ, the Son of God, • is both God and man.}
    		\thirdline{\psanum{31}He is God, of the substance of the Father, begotten • before the worlds : and he is man, of the substance of his Mother, • born in the world;}
    		\psanum{32}Perfect • God perfect man : of reasoning soul and human • flesh subsisting;\par
    		\psanum{33}Equal to the Father as • touching his Godhead : less than the Father as • touching his manhood.\par
    		\psanum{34}Who although • he be God and man : yet he is not two, • but is one Christ;\par
    		\psanum{35}One however, not by conversion of • Godhead into flesh : but by taking • manhood into God;\par
    		\psanum{36}One • altogether : not by confusion of substance, but by unit • y of person.\par
    		\psanum{37}For as reasoning soul and • flesh is one man : so God and • man is one Christ;\par
    		\psanum{38}Who suffered for • our salvation : descended into hell, rose • again from the dead;\par
    		\psanum{39}Ascended into heaven, sat down at the • right hand of the Father : from whence he shall come to judge • the quick and the dead.\par
    		\psanum{40}At whose coming all men must rise again • with their bodies : and shall give account • for their own deeds.\par
    		\psanum{41}And they that have done good will go into • life eternal : they that have done evil into • eternal fire.\par

    		\lett{T}{his} is the • Catholick Faith : which except a man do faithfully and steadfastly believe, he • cannot be saved.\par
    		\secondline{℣. Glory be to the Father, • and to the Son : and • to the Holy Ghost;}
    		\thirdline{℟. As it was in the beginning, is now, and • ever shall be : world • without end. Amen.}
    		
    	\clearpage
    	\begin{inhead}
    		Evensong
    	\end{inhead}
    	
    	\subsubsection{Glory Be to God on High}\fancyhead[RE,LO]{Glory Be to God}
    		%Old Scottish Chant
			\gregorioscore{resources/gabc/MassParts/Scottish/ScottishGloria.gabc}
    
		\clearpage\subsubsection{My Soul Doth Magnify the Lord}\fancyhead[RE,LO]{My Soul Doth Magnify}
			\gregorioscore{resources/gabc/Extra/Magnificat.gabc}
			
		\subsubsection{Lord, Now Lettest Thou Thy Servant}\fancyhead[RE,LO]{Lord, Now Lettest}
	
			\gregorioscore{resources/gabc/Extra/NuncDimittis.gabc}

			For mine eyes • have seen * • thy salvation,

			Which thou hast pre • pared * before the face • of all people;

			To be a light to lighten the • Gentiles, * and to be the glory of thy • people Israel.
	
			℣. Glory be to the Father, and to • the Son, * and • to the Holy Ghost.
	
			℟. As it was in the beginning, {\dag} is now, and ever • shall be, * world • without end. Amen.
	
	\clearpage
	
	\selectlanguage{latin}
	
	\subsection{Office Service Two - Officium Omnium Sanctorum Monachorum et Eremitarum}\fancyhead[RO,LE]{\textit{Office II}}
	
		\begin{inhead}
			Matutinum
		\end{inhead}
		
		\subsubsection{Venite}\fancyhead[RE,LO]{Venite}
			\gregorioscore{resources/gabc/Extra/VeniteLatin.gabc}
			
		\clearpage\subsubsection{Pascha Nostrum}
			\gregorioscore{resources/gabc/Extra/PaschaNostrumLatin.gabc}
			
			Non in fermento veteri, neque in fermento maliti{\ae} et ne • quiti{\ae} * sed in Azymis sinceritatis et • veritatis.
	
			Christus, resurgens ex mortuis, jam • non moritur * mors illi ultra non do • minabitur.
			
			Quod enim mortuus est, peccato mortuus • est semel * quod autem vivit, • vivit Deo.
			
			Ita et vos existimate, vos mortuos quidem esse • peccato, * viventes autem Deo, in Christo Jesu, Do • mino nostro.
    
			Nunc Christus resurrexit a • mortuis * primiti{\ae} dorm • ientium.
			
			Quoniam quidem per ho • minem mors * et per hominem resurrectio • mortuorum. 
			
			Et, sicut in Adam omnes mor • iuntur * ita et in Christo omnes vivi • ficabuntur. 
    		
    		℣. Glória Patri, et • Fílio, * et Spirí • tui Sancto:
    
    		℟. ℟. Sicut erat in princípio, et nunc, • et semper, * et in sǽcula s{\ae}cu • lórum. Amen.
			
		\clearpage\subsubsection{Te Deum}\fancyhead[RE,LO]{Te Deum}
			\gregorioscore{resources/gabc/Extra/TeDeumLatin.gabc}
			
		\clearpage\subsubsection{Benedicite}\fancyhead[RE,LO]{Benedicite}
			\gregorioscore{resources/gabc/Extra/BenediciteLatin.gabc}
			
			Benedícite, Ángeli • Dómini, Dómino: * laudáte et superexaltáte e • um in sǽcula.\par
			Benedícite, • c{\ae}li, Dómino:  * laudáte et superexaltáte e • um in sǽcula.\par
			Benedícite, aqu{\ae} omnes, qu{\ae} super • c{\ae}los sunt, Dómino: * laudáte et superexaltáte e • um in sǽcula.\par
			Benedícite, omnes virtútes • Dómini, Dómino: * laudáte et superexaltáte e • um in sǽcula.\par
			Benedícite, sol et • luna, Dómino: * laudáte et superexaltáte e • um in sǽcula.\par
			Benedícite, stell{\ae} • c{\ae}li, Dómino: * laudáte et superexaltáte e • um in sǽcula.\par
			Benedícite, omnis imber • et ros, Dómino: * laudáte et superexaltáte e • um in sǽcula.\par
			Benedícite, omnes spíritus • Dei, Dómino: * laudáte et superexaltáte e • um in sǽcula.\par
			Benedícite, ignis et • {\ae}stus, Dómino: * laudáte et superexaltáte e • um in sǽcula.\par
			Benedícite, frigus et • {\ae}stus, Dómino: * laudáte et superexaltáte e • um in sǽcula.\par
			Benedícite, rores et pru • ína, Dómino: * laudáte et superexaltáte e • um in sǽcula.\par
			Benedícite, gelu et • frigus, Dómino: * laudáte et superexaltáte e • um in sǽcula.\par
			Benedícite, glácies et • nives, Dómino: * laudáte et superexaltáte e • um in sǽcula.\par
			Benedícite, noctes et • dies, Dómino: * laudáte et superexaltáte e • um in sǽcula.\par
			Benedícite, lux et • ténebr{\ae}, Dómino: * laudáte et superexaltáte e • um in sǽcula.\par
			Benedícite, fúlgura et • nubes, Dómino: * laudáte et superexaltáte e • um in sǽcula.\par

			Benedícat • terra Dóminum: * laudet et superexáltet e • um in sǽcula.\par
			Benedícite, montes et • colles, Dómino: * laudáte et superexaltáte e • um in sǽcula.\par
			Benedícite, univérsa germinántia in • terra, Dómino: * laudáte et superexaltáte e • um in sǽcula.\par
			Benedícite, • fontes, Dómino: * laudáte et superexaltáte e • um in sǽcula.\par
			Benedícite, mária et • flúmina, Dómino: * laudáte et superexaltáte e • um in sǽcula.\par
			Benedícite, cete, et ómnia, qu{\ae} movéntur in • aquis, Dómino: * laudáte et superexaltáte e • um in sǽcula.\par
			\gregorioscore{resources/gabc/Extra/BenediciteLatin2.gabc}
			%Benedícite, omnes vólucres • c{\ae}li, Dómino: * laudáte et superexaltáte e • um in sǽcula.\par
			Benedícite, omnes bésti{\ae} et • pécora, Dómino: * laudáte et superexaltáte e • um in sǽcula.\par
			Benedícite, fílii • hóminum, Dómino: * laudáte et superexaltáte e • um in sǽcula.\par
			Benedícat • Israël Dóminum: * laudet et superexáltet e • um in sǽcula.\par
			Benedícite, sacerdótes • Dómini, Dómino: * laudáte et superexaltáte e • um in sǽcula.\par
			Benedícite, servi • Dómini, Dómino: * laudáte et superexaltáte e • um in sǽcula.\par
			Benedícite, spíritus, et ánim{\ae} just • órum, Dómino: * laudáte et superexaltáte e • um in sǽcula.\par
			Benedícite, sancti, et húmiles • corde, Dómino: * laudáte et superexaltáte e • um in sǽcula.\par
			Benedicámus Patrem et Fílium cum • Sancto Spíritu: * laudémus et superexaltémus e • um in sǽcula.
			
		\clearpage\subsubsection{Benedictus}\fancyhead[RE,LO]{Benedictus}
			\gregorioscore{resources/gabc/Extra/BenedictusLatin.gabc}
			
		%\clearpage\subsubsection{Symbolum Apostolorum}\fancyhead[RE,LO]{Symbolum Apostolorum}
		%	\gregorioscore{resources/gabc/Extra/CredoApostolorum.gabc}
		
		\clearpage\subsubsection{Symbolum Athanasianum}\fancyhead[RE,LO]{Athanasianum}
			\gregorioscore{resources/gabc/Extra/AthanasianLatin.gabc}
	
			Quam nisi quisque íntegram inviolatámque serváverit, * absque dúbio in {\ae}térnum períbit.
	
			Fides autem cathólica h{\ae}c est: * ut unum Deum in Trinitáte, et Trinitátem in unitáte venerémur.
	
			Neque confundéntes persónas, * neque substántiam separántes.\par
			Alia est enim persóna Patris, ália Fílii, * ália Spíritus Sancti:\par
			Sed Patris, et Fílii, et Spíritus Sancti una est divínitas, * {\ae}quális glória, co{\ae}térna majéstas.\par
			Qualis Pater, talis Fílius, * talis Spíritus Sanctus.\par
			Increátus Pater, increátus Fílius, * increátus Spíritus Sanctus.\par
			Imménsus Pater, imménsus Fílius, * imménsus Spíritus Sanctus.\par
			{\AE}térnus Pater, {\ae}térnus Fílius, * {\ae}térnus Spíritus Sanctus.\par
			Et tamen non tres {\ae}térni, * sed unus {\ae}térnus.\par
			Sicut non tres increáti, nec tres imménsi, * sed unus increátus, et unus imménsus.\par
			Simíliter omnípotens Pater, omnípotens Fílius, * omnípotens Spíritus Sanctus.\par
			Et tamen non tres omnipoténtes, * sed unus omnípotens.\par
			Ita Deus Pater, Deus Fílius, * Deus Spíritus Sanctus.\par
			Ut tamen non tres Dii, * sed unus est Deus.\par
			Ita Dóminus Pater, Dóminus Fílius, * Dóminus Spíritus Sanctus.\par
			Et tamen non tres Dómini, * sed unus est Dóminus.\par
			Quia, sicut singillátim unamquámque persónam Deum ac Dóminum confitéri christiána veritáte compéllimur: * ita tres Deos aut Dóminos dícere cathólica religióne prohibémur.\par
			Pater a nullo est factus: * nec creátus, nec génitus.\par
			Fílius a Patre est: * non factus, nec creátus, sed génitus.\par
			Spíritus Sanctus a Patre: * non factus, nec creátus, nec génitus, sed procédens.\par
			Unus ergo Pater, non tres Patres: unus Fílius, non tres Fílii: * unus Spíritus Sanctus, non tres Spíritus Sancti.\par
			Et in hac Trinitáte nihil prius aut postérius, nihil majus aut minus: * sed tot{\ae} tres persón{\ae} co{\ae}térn{\ae} sibi sunt et co{\ae}quáles.\par
			Ita ut per ómnia, sicut jam supra dictum est, * et únitas in Trinitáte, et Trínitas in unitáte veneránda sit.\par
			Qui vult ergo salvus esse, * ita de Trinitáte séntiat.\par
			\gregorioscore{resources/gabc/Extra/AthanasianLatin2.gabc}\par
			%\lett{S}{ed} necessárium est ad {\ae}térnam salútem, * ut Incarnatiónem quoque Dómini nostri Jesu Christi fidéliter credat.\par
			Est ergo fides recta ut credámus et confiteámur, * quia Dóminus noster Jesus Christus, Dei Fílius, Deus et homo est.
			Deus est ex substántia Patris ante sǽcula génitus: * et homo est ex substántia matris in sǽculo natus.\par
			Perféctus Deus, perféctus homo: * ex ánima rationáli et humána carne subsístens.\par
			{\AE}quális Patri secúndum divinitátem: * minor Patre secúndum humanitátem.\par
			Qui licet Deus sit et homo, * non duo tamen, sed unus est Christus.\par
			Unus autem non conversióne divinitátis in carnem, * sed assumptióne humanitátis in Deum.\par
			Unus omníno, non confusióne substánti{\ae}, * sed unitáte persón{\ae}.\par
			Nam sicut ánima rationális et caro unus est homo: * ita Deus et homo unus est Christus.\par
			Qui passus est pro salúte nostra: descéndit ad ínferos: * tértia die resurréxit a mórtuis.\par
			Ascéndit ad c{\ae}los, sedet ad déxteram Dei Patris omnipoténtis: * inde ventúrus est judicáre vivos et mórtuos.\par
			Ad cujus advéntum omnes hómines resúrgere habent cum corpóribus suis; * et redditúri sunt de factis própriis ratiónem.\par
			Et qui bona egérunt, ibunt in vitam {\ae}térnam: * qui vero mala, in ignem {\ae}térnum.\par
			\lett{H}{{\ae}c} est fides cathólica, * quam nisi quisque fidéliter firmitérque credíderit, salvus esse non póterit.\par
			\secondline{℣. Glória Patri, et Fílio, * et Spirítui Sancto:}
			\thirdline{℟. Sicut erat in princípio, et nunc, et semper, * et in sǽcula s{\ae}culórum. Amen.}
		
		\clearpage
		
		\begin{inhead}
			Vesper{\ae}
		\end{inhead}
		
		\subsubsection{Magnificat}\fancyhead[RE,LO]{Magnificat}
			\gregorioscore{resources/gabc/Extra/MagnificatLatin.gabc}
			
		\subsubsection{Nunc Dimittis}\fancyhead[RE,LO]{Nunc Dimittis}
			\gregorioscore{resources/gabc/Extra/NuncDimittisLatin_simple.gabc}
			Quia vidérunt • óculi mei * salu • táre tuum,\par
			Quod • parásti * ante fáciem ómnium • populórum,\par
			Lumen ad revela • tiónem géntium, * et glóriam plebis • tu{\ae} Isra{\"e}l.\par
			℣. Glória Patri, • et Fílio, * et Spir • ítui Sancto:\par
			℟. Sicut erat in princípio, et • nunc, et semper, * et in sǽcula • s{\ae}culórum. Amen.
			
\clearpage

\selectlanguage{british}

\section{Marian Anthems}
	%Alma Redemptoris Mater
	\subsection{Alma Redemptoris Mater}\fancyhead[RO,LE]{\textit{Gracious Mother}}
	\begin{inhead}
	From I Evensong of Advent I until II Evensong of the Purification, inclusive.
	\end{inhead}
		\gregorioscore{resources/gabc/Extra/AlmaSimple.gabc}

	%Ave, Regina C{\ae}lorum
	\clearpage\subsection{Ave, Regina C{\ae}lorum}\fancyhead[RO,LE]{\textit{Queen of the Heavens}}
	\begin{inhead}
	From Compline of the Purification until Compline of Holy Wednesday, inclusive.
	\end{inhead}
		\gregorioscore{resources/gabc/Extra/AveSimple.gabc}

	%Regina C{\ae}li
	\clearpage\subsection{Regina C{\ae}li}\fancyhead[RO,LE]{\textit{O Queen of Heaven}}
	\begin{inhead}
	From Compline of Holy Saturday until I Evensong of Trinity Sunday, exclusive.
	\end{inhead}
		\gregorioscore{resources/gabc/Extra/ReginaSimple.gabc}

	%Salve Regina
	\clearpage\subsection{Mary, We Hail Thee, Simple Tone}\fancyhead[RO,LE]{\textit{Mary, We Hail Thee}}\label{SalveReginaEnglish}
	\begin{inhead}
	From I Evensong of Trinity Sunday until I Evensong of Advent I, exclusive.
	\end{inhead}
		\gregorioscore{resources/gabc/Extra/SalveReginaSimple.gabc}

	\clearpage\subsection{Mary, We Hail Thee, Solemn Tone}\fancyhead[RO,LE]{\textit{Mary, We Hail Thee}}
	\begin{inhead}
	From I Evensong of Trinity Sunday until I Evensong of Advent I, exclusive.
	\end{inhead}
		\gregorioscore{resources/gabc/Extra/SalveReginaSolemn.gabc}
		
\clearpage

\section{Communion Services}
	
	%Merbecke:
	\subsection{Communion Service One - Mass of the Grace of God}\fancyhead[RO,LE]{\textit{Communion I}}

		\subsubsection{Ten Commandments}\fancyhead[RE,LO]{10 Commandments}
			\begin{rubric}
				In response to the first nine Commandments, the following is sung.
			\end{rubric}

			\gregorioscore{resources/gabc/MassParts/Merbecke/Decalogue1.gabc}

			\begin{rubric}
				In response to the final Commandment, the following is sung.
			\end{rubric}

			\gregorioscore{resources/gabc/MassParts/Merbecke/Decalogue2.gabc}
			
		\subsubsection{Lord, Have Mercy Upon Us}\fancyhead[RE,LO]{Lord, Have Mercy}
			\gregorioscore{resources/gabc/MassParts/Merbecke/MerbeckeKyrie.gabc}
		
		\clearpage\subsubsection{Glory be to God on High}\fancyhead[RE,LO]{Glory Be to God}
			\gregorioscore{resources/gabc/MassParts/Merbecke/MerbeckeGloria.gabc}
		
		\clearpage\subsubsection{Nicene Creed}\fancyhead[RE,LO]{Nicene Creed}
			\gregorioscore{resources/gabc/MassParts/Merbecke/MerbeckeCreed.gabc}
		
		\clearpage\subsubsection{Holy, Holy, Holy}\fancyhead[RE,LO]{Holy, Holy, Holy}
			\gregorioscore{resources/gabc/MassParts/Merbecke/MerbeckeSanctus.gabc}
		
		\subsubsection{O Lamb of God}\fancyhead[RE,LO]{O Lamb of God}
			\gregorioscore{resources/gabc/MassParts/Merbecke/MerbeckeAgnus.gabc}
		
		
	%\subsection{Communion Service Two - Mass of St. Margaret}\fancyhead[RO,LE]{\textit{Communion II}}
	%	\subsubsection{Ten Commandments}\fancyhead[RE,LO]{10 Commandments}
	%		\begin{rubric}
	%			In response to the first nine Commandments, the following is sung.
	%		\end{rubric}
	%		\gregorioscore{resources/gabc/MassParts/Scottish/ScottishDecalogue1.gabc}
	%
	%		\begin{rubric}
	%			In response to the final Commandment, the following is sung.
	%		\end{rubric}
	%		\gregorioscore{resources/gabc/MassParts/Scottish/ScottishDecalogue2.gabc}
	%		
	%	\clearpage\subsubsection{Kyrie Eleison}\fancyhead[RE,LO]{Kyrie}
	%		\gregorioscore{resources/gabc/MassParts/Scottish/ScottishKyrie.gabc}
	%	
	%	\clearpage\subsubsection{Gloria in excelsis}\fancyhead[RE,LO]{Gloria}
	%		%Old Scottish Chant
	%		\gregorioscore{resources/gabc/MassParts/Scottish/ScottishGloria.gabc}
	%	
	%	\clearpage\subsubsection{Nicene Creed}\fancyhead[RE,LO]{Credo}
	%		
	%	
	%	\clearpage\subsubsection{Sanctus}\fancyhead[RE,LO]{Sanctus}
	%		
	%	
	%	\clearpage\subsubsection{Agnus Dei}\fancyhead[RE,LO]{Agnus}
			
	%Missa Marialis
	\clearpage\subsection{Communion Service Two - Mass of Our Lady}\fancyhead[RO,LE]{\textit{Communion II}}
		\subsubsection{Ten Commandments}\fancyhead[RE,LO]{10 Commandments}
			\begin{rubric}
				In response to the first nine Commandments, the following is sung.
			\end{rubric}
			\gregorioscore{resources/gabc/MassParts/Scottish/ScottishDecalogue1.gabc}
	
			\begin{rubric}
				In response to the final Commandment, the following is sung.
			\end{rubric}
			\gregorioscore{resources/gabc/MassParts/Scottish/ScottishDecalogue2.gabc}
			
		\clearpage\subsubsection{Lord, Have Mercy Upon Us}\fancyhead[RE,LO]{Lord, Have Mercy}
			\gregorioscore{resources/gabc/MassParts/Marialis/MarialisKyrie.gabc}
		
		\clearpage\subsubsection{Glory Be to God on High}\fancyhead[RE,LO]{Glory Be to God}
			\gregorioscore{resources/gabc/MassParts/Marialis/MarialisGloria.gabc}
		
		\clearpage\subsubsection{Nicene Creed}\fancyhead[RE,LO]{Nicene Creed}
			\gregorioscore{resources/gabc/MassParts/Marialis/MarialisCreed.gabc}
		
		\clearpage\subsubsection{Holy, Holy, Holy}\fancyhead[RE,LO]{Holy, Holy, Holy}
			\gregorioscore{resources/gabc/MassParts/Marialis/MarialisSanctus.gabc}
		
		\clearpage\subsubsection{O Lamb of God}\fancyhead[RE,LO]{O Lamb of God}
			\gregorioscore{resources/gabc/MassParts/Marialis/MarialisAgnus.gabc}
			
			
		%Missa de Angelis in English
	\clearpage\subsection{Communion Service Three - Mass of the Angels}\fancyhead[RO,LE]{\textit{Communion III}}
		%\subsubsection{Ten Commandments}\fancyhead[RE,LO]{10 Commandments}
		%	\begin{rubric}
		%		In response to the first nine Commandments, the following is sung.
		%	\end{rubric}
		%
			%
		%
		%	\begin{rubric}
		%		In response to the final Commandment, the following is sung.
		%	\end{rubric}
		%
			%
		%
		\subsubsection{Lord, Have Mercy Upon Us}\fancyhead[RE,LO]{Lord, Have Mercy}
			\gregorioscore{resources/gabc/MassParts/Angels/KyrieVIII.gabc}
		
		\clearpage\subsubsection{Glory Be to God on High}\fancyhead[RE,LO]{Glory Be to God}
			\gregorioscore{resources/gabc/MassParts/Angels/GloriaVIII.gabc}
		
		\subsubsection{Holy, Holy, Holy}\fancyhead[RE,LO]{Holy, Holy, Holy}
			\gregorioscore{resources/gabc/MassParts/Angels/SanctusVIII.gabc}
		
		\clearpage\subsubsection{O Lamb of God}\fancyhead[RE,LO]{O Lamb of God}
			\gregorioscore{resources/gabc/MassParts/Angels/AgnusVIII.gabc}
			

		%Missa Orbis Factor in English
	\clearpage\subsection{Communion Service Four - Mass of the Maker of the World}\fancyhead[RO,LE]{\textit{Communion IV}}
		%\subsubsection{Ten Commandments}\fancyhead[RE,LO]{10 Commandments}
		%	\begin{rubric}
		%		In response to the first nine Commandments, the following is sung.
		%	\end{rubric}
		%
			%
		%
		%	\begin{rubric}
		%		In response to the final Commandment, the following is sung.
		%	\end{rubric}
		%
			%
		%
		\subsubsection{Lord, Have Mercy Upon Us}\fancyhead[RE,LO]{Lord, Have Mercy}
			\gregorioscore{resources/gabc/MassParts/World/KyrieXI.gabc}
			
		\clearpage\subsubsection{Glory Be to God on High}\fancyhead[RE,LO]{Glory Be to God}
			\gregorioscore{resources/gabc/MassParts/World/GloriaXI.gabc}
		
		\clearpage\subsubsection{Holy, Holy, Holy}\fancyhead[RE,LO]{Holy, Holy, Holy}
			\gregorioscore{resources/gabc/MassParts/World/SanctusXI.gabc}
		
		\subsubsection{O Lamb of God}\fancyhead[RE,LO]{O Lamb of God}
			\gregorioscore{resources/gabc/MassParts/World/AgnusXI.gabc}


		%Missa Deus Genitor Alme in English
	\clearpage\clearpage\subsection{Communion Service Five - Mass of God Our Sweet Father}\fancyhead[RO,LE]{\textit{Communion V}}
		%\subsubsection{Ten Commandments}\fancyhead[RE,LO]{10 Commandments}
		%	\begin{rubric}
		%		In response to the first nine Commandments, the following is sung.
		%	\end{rubric}
		%
			%
		%
		%	\begin{rubric}
		%		In response to the final Commandment, the following is sung.
		%	\end{rubric}
		%
			%
		%
		\subsubsection{Lord, Have Mercy Upon Us}\fancyhead[RE,LO]{Lord, Have Mercy}
			\gregorioscore{resources/gabc/MassParts/Sweet/KyrieXVIII.gabc}
		
		\subsubsection{Holy, Holy, Holy}\fancyhead[RE,LO]{Holy, Holy, Holy}
			\gregorioscore{resources/gabc/MassParts/Sweet/SanctusXVIII.gabc}
		
		\clearpage\subsubsection{O Lamb of God}\fancyhead[RE,LO]{O Lamb of God}
			\gregorioscore{resources/gabc/MassParts/Sweet/AgnusXVIII.gabc}
			
	
\clearpage\section{Creeds}\fancyhead[RO,LE]{\textit{}}\fancyhead[RE,LO]{}

	%Credo III in English
	\subsection{Creed III}\fancyhead[RO,LE]{\textit{Creed III}}
		\gregorioscore{resources/gabc/MassParts/Creeds/CredoIII.gabc}
		
	%Credo III
	\clearpage\subsection{Credo III}\fancyhead[RO,LE]{\textit{Credo III}}
		\gregorioscore{resources/gabc/MassParts/Credos/CredoIIILatin.gabc}


\clearpage\section{Sequences}
	%Easter
	\subsection{Christians, To the Paschal Victim}\fancyhead[RO,LE]{\textit{To the Paschal Victim}}\fancyhead[RE,LO]{Easter Sunday}
		\gregorioscore{resources/gabc/Sequences/Victimae.gabc}
		
	\clearpage\subsection{Victimae paschali laudes}\fancyhead[RO,LE]{\textit{Victimae}}
		\gregorioscore{resources/gabc/Sequences/VictimaeLatin.gabc}
	
	%Whitsun
	\clearpage\subsection{Come, Thou Holy Paraclete}\fancyhead[RO,LE]{\textit{Come, Paraclete}}\fancyhead[RE,LO]{Whitsunday}
		\gregorioscore{resources/gabc/Sequences/VeniSancte.gabc}
	
	\clearpage\subsection{Veni, Sancte Spiritus}\fancyhead[RO,LE]{\textit{Veni, Sancte}}
		\gregorioscore{resources/gabc/Sequences/VeniSancteLatin.gabc}

	%All Souls
	\clearpage\subsection{Day of Wrath and Doom Impending}\fancyhead[RO,LE]{\textit{Day of Wrath}}\fancyhead[RE,LO]{Requiem}
		\gregorioscore{resources/gabc/Sequences/DayOfWrath.gabc}
		
	\clearpage\subsection{Dies irae}\fancyhead[RO,LE]{\textit{Dies Irae}}
		\gregorioscore{resources/gabc/Sequences/DiesIrae.gabc}

\clearpage
\section{Paraliturgical Hymns}\fancyhead[RE,LO]{}

	\subsection{Therefore, We Before Him Bending}\fancyhead[RO,LE]{\textit{Before Him Bending}}
		\gregorioscore{resources/gabc/Extra/Tantum.gabc}
	
	\clearpage\subsection{Tantum Ergo}\fancyhead[RO,LE]{\textit{Tantum Ergo}}
		\gregorioscore{resources/gabc/Extra/TantumLatin.gabc}
	
	\clearpage\subsection{O Saving Victim Opening Wide}\fancyhead[RO,LE]{\textit{O Saving Victim}}
		\gregorioscore{resources/gabc/Extra/OSavingVictim.gabc}
	
	\subsection{O Salutaris Hostia}\fancyhead[RO,LE]{\textit{O Salutaris}}
		\gregorioscore{resources/gabc/Extra/OSalutarisLatin.gabc}
	
	\clearpage\subsection{Prostrate, I Adore Thee}\fancyhead[RO,LE]{\textit{Prostrate, I Adore}}
		\gregorioscore{resources/gabc/Extra/Adoro.gabc}
	
	\clearpage\subsection{Adoro Te Devote}\fancyhead[RO,LE]{\textit{Adoro}}
		\gregorioscore{resources/gabc/Extra/AdoroLatin.gabc}


	\clearpage\subsection{Advent Prose}\fancyhead[RO,LE]{\textit{Advent Prose}}\label{AdventProse}

	\gregorioscore{resources/gabc/Extra/Rorate.gabc}