\subby{Tsar St. Nicholas of Russia}
\feastday{{St. Nicholas Russia}}
\fancyhead[RE,LO]{17 July}
\begin{inhead}
    {Simple\\
17 July}
\end{inhead}

\textit{Opening Sentence.} To the Lord our God belong mercies and forgivenesses, though we have rebelled against him: neither have we obeyed the voice of the Lord our God, to walk in his laws which he set before us.\vr{Dan 9:9-10}

\textit{Opening Sentence.} Correct us, O Lord, but with judgement: not in thine anger, lest thou bring us to nothing.\vr{Jer 10:24}

\textit{Opening Sentence.} Enter not into judgement with thy servants, O Lord: for in thy sight shall no man living be justified.\vr{Ps 143:2}

\subbysub{I Evensong}

\begin{rubric}
	Proper Psalms: 79 \& 85
\end{rubric}
\begin{rubric}
	Proper Lessons: Jeremiah 12 \& Hebrews 11:32-12:7
\end{rubric}

With thankful hearts thy glory, O King of Saints, we sing:

Shown in the saintly story Of Nich'las Martyr-King;

Who chose to die, obeying The voice of conscience dear,

Not live on earth, betraying All that he counted dear.\\

Shall, then, his mem'ry perish? His name we venerate;

The Faith he loved we cherish, His spirit emulate.

That so by Christ-like living, With charity resign'd:

Each others' faults forgiving, We may enrich mankind.\\

%Changed to match change of alone to with kin.
For long his foes assailed him, Till friends were so fallen;

And this world's weapons fail'd him, And he was left with kin.

No whit his foes relented, Successful in the strife,

But to their King presented The choice of death or life.\\

For all lives liv'd sincerely In Christ may God be blest,

To mortals thus most clearly In mortals manifest:

The Father, who forgiveth Man's failures in the strife;

The Son, in whom he liveth; The Spirit, Source of Life! Amen.\\

    ℣. Thou, O Lord, art my defender.

	℟. And the lifter up of my head.\\
	
\antiphon{Mag.}{The true glory of princes {\dag} consisteth in advancing God's glory, in the maintenance of true religion and the Church's good.}

\subbysub{Mattins}

\begin{rubric}
	Proper Psalms: 9-11
\end{rubric}
\begin{rubric}
	Proper Lessons: 2 Samuel 1 \& Matthew 27
\end{rubric}

\begin{rubric}
	Instead of the \emph{Venite} is said the following Hymn.
\end{rubric}

\lett{R}{ighteous} art thou, O Lord: and just are thy judgements.

\secondline{\textit{Thou art just, O Lord, in all that is brought upon us: for thou hast done right, but we have done wickedly.}}

\thirdline{Nevertheless, our feet were almost gone: our treadings
had well-nigh slipped.}

\textit{For why? we were grieved at the wicked: we did also see the ungodly in such prosperity.}

The people stood up, and the rulers took counsel together: against the Lord, and against his Anointed.

\textit{They cast their heads together with one consent: and were confederate against him.}

He heard the blasphemy of the multitude, and fear was on every side: while they conspired together againſt him, to take away his life.

\textit{They spoke against him with false tongues, and compassed him about with words of hatred: and fought against him without a cause.}

Yea, his own familiar friends, whom he trusted: they that eat of his bread laid great wait for him.

\textit{They rewarded him evil for good: to the great discomfort of his
soul.}

They took their counsel together, saying, God hath forsaken him: persecute him, and take him, for there is none to deliver him.

\textit{The breath of our nostrils, the Anointed of the Lord was taken in their pits: of whom we said, Under his shadow we shall be safe.}

The adversary and the enemy entered into the gates of Jerusalem: saying, When shall he die, and his name perish?

\textit{Let the sentence of guiltiness proceed against him: and now that he lieth, let him rise up no more.}

False witnesses also did rise up against him: they laid to his charge things that he knew not.

\textit{For the sins of the people, and the iniquities of the priests: they shed the blood of the just in the midst of Jerusalem.}

O my soul, come not thou into their secret; unto their assembly, mine honour, be not thou united: for in their anger they slew a man.

\textit{Even the man of thy right hand: the Son of man, whom thou hadst made so strong for thine own self.}

In the sight of the unwise he seemed to die: and his departure was taken for misery.

\textit{They fools counted his life madness, and his end to be without honour: but he is in peace.}

For though he was punished in the sight of men: yet was his hope full of immortality.

\textit{How is he numbered with the children of God: and his lot is
among the saints!}

But, O Lord God, to whom vengeance belongeth, thou God, to whom vengeance belongeth: be favourable and gracious unto Sion.

\textit{Be merciful, O Lord, unto thy people, whom thou hast redeemed: and lay not innocent blood to our charge.}

O shut not up our souls with sinners: nor our lives with the bloodthirsty.

\textit{Deliver us from blood-guiltiness, O God, thou that art the God of our
salvation: and our tongues shall sing of thy righteousness.}

For thou art the God that hast no pleasure in wickedness: neither shall any evil dwell with thee.

\textit{Thou wilt destroy them that speak leasing: the Lord abhors both the blood-thirsty and deceitful man.}

O how suddenly do they consume: perish, and come to a fearful end!

\textit{Yea, even like a dream, when one awaketh: so didst thou make their image to vanish out of the city.}

Great and marvellous are thy works, O Lord God Almighty: just and true are thy ways, O King of saints!

\textit{Righteous art thou, O Lord: and just are thy judgements.}

Glory to the Father, and to the Son: and to the Holy Ghost;

\textit{As it was in the beginning, is now, and ever shall be: world without end. Amen.}

\begin{rubric}
	The following Hymn may be said after the preceding Hymn.
\end{rubric}
\begin{multicols}{2}
Nicholas, who chose to die

Rather than the Faith deny,

Forfeiting his kingly pride

For the sake of Jesu's Bride;

Lovingly his praise we sing,

Russia's martyr, Russia's king.\\

Mirror fair of courtesy,

Flow'r of wedded chastity,

Humble follow'r day by day,

Of the Church's holy way;

Lovingly his praise we sing,

Russia's martyr, Russia's king.\\

All the way of death he trod

For the glory of his God,

And his dying dignity

Made a bright Epiphany;

Lovingly his praise we sing,

Russia's martyr, Russia's king.\\

Bless we God the Three in One,

For all faithful 'neath the sun,

For the faithful gone before,

And for those our country bore;

Chiefly him whose praise we sing:

Russia's martyr, Russia's king. Amen.\\
\end{multicols}

\begin{rubric}
	The Office Hymn is here provided.
\end{rubric}
\begin{multicols}{2}
Nicholas' praise, our martyr King,

With heart and voice come let us sing:

Who earthly shame with Christ doth bear,

The Victor's laurel now doth wear.\\

The prison's gloom, the battle sore,

Christ's martyr here with gladness bore,

And mounting from the seated shame

To his sure hope, in heav'n he came.\\

For holy Church so he withstood,

By the bullet his lifeblood flow'd:

And where that kingly seed was sown

New harvest unto Christ hath grown.\\

For Russia's cause to peril brought;

Crown, life, and freedom deem'd he not:

For Russia's Church, for Russia's realm

We pray, lest storms our ark should whelm.\\

Jesu, to thee we glory pay,

Ruling earth's kingdoms neath thy sway;

All glory, as is ever meet,

To Father and to Paraclete. Amen.\\

    ℣. Hold fast that which thou hast.

	℟. That no man take thy crown.\\
\end{multicols}

\antiphon{Ben.}{Let my sufferings {\dag} satiate the malice of mine and the Church's enemies, but let their cruelty never exceed the measure of my charity.}

\begin{rubric}
	In the end of the Litany (which shall always on this Day be used) immediately after the Collect \emph{We humbly beseech thee, O Father}, the three Collects next following are to be read.
\end{rubric}

\lett{O}{Lord,} we beseech thee mercifully hear our prayers, and spare all those who confess their sins unto thee; that they whose consciences by sin are accused, by thy merciful pardon may be absolved; through Christ our Lord. \textit{Amen.}

\lett{O}{most} mighty God, and merciful Father, who hast compassion upon all men, and hatest nothing that thou hast made: who wouldest not the death of a sinner, but that he would rather turn from his sin, and be saved: Mercifully forgive us our trespasses; receive and comfort us, who are grieved and wearied with the burden of our sins. Thy property is always to have mercy; to thee only it appertaineth to forgive sins. Spare us therefore, good Lord, spare thy people, whom thou hast redeemed; enter not into judgement with thy servants, who are vile earth and miserable sinners; but so turn thine anger from us, who meekly acknowledge our vileness, and truly repent us of our faults; and so make haste to help us in this world, that we may ever live with thee in the world to come; through Jesus Christ our Lord. \textit{Amen.}

\lett{T}{urn} thou us, O good Lord, and so shall we be turned. Be favourable, O Lord, be favourable to thy people, Who turn to thee in weeping, fasting, and praying. For thou art a merciful God, full of compassion, long-suffering, and of great pity. Thou sparest, when we deserve punishment, And in thy wrath thinkest upon mercy. Spare thy people, good Lord, spare them, And let not thine heritage be brought to confusion. Hear us, O Lord, for thy mercy is great, And after the multitude of thy mercies look upon us, Through the merits and mediation of thy blessed Son Jesus Christ our Lord. \textit{Amen.}

\introit
\lett{L}{et} us rejoice in the Lord, celebrating a festival day in honour of blessed Nicholas the Martyr: at whose passion the Angles rejoice and praise the Son of God. \textit{Ps.} Rejoice in the Lord, O ye righteous: for it becometh well the just to be thankful.

\collect
%\lett{O}{most} mighty God, terrible in thy judgements, and wonderful in thy doings toward the children of men; who in thy heavy displeasure didst suffer the life of our gracious Sovereign Emperor Nicholas the Second, to be (as on this day) taken away by the hands of cruel and bloody men: We thy sinful creatures here assembled before thee, do, in the behalf of all this Nation, which brought down this heavy judgement upon us. But, O gracious, when thou makest inquisition for blood, lay not the guilt of this innocent blood, (the shedding whereof nothing but the blood of thy Son can expiate,) lay it not to the charge of the people of this land; not let it ever be required of us, or our posterity. Be merciful, O Lord, be merciful unto thy people, whom thou hast redeemed; and be not angry with us for ever: But pardon us for thy mercy's sake. through the merits of thy Son Jesus Christ our Lord. \textit{Amen.}
%
\lett{B}{lessed} Lord, in whose sight the death of thy saints is precious; We magnify thy Name for thine abundant grace bestowed upon the martyred Sovereign, Emperor Nicholas of Russia; by which he was enabled so cheerfully to follow the steps of his blessed Master and Saviour, in a constant meek suffering of all barbarous indignities, and at the last resisting unto blood; and even then according to the same pattern, witnessing to his murderers. Let his memory, O Lord, be ever blessed among us; that we may follow the example of his courage and constancy, his meekness and patience, and great charity. And grant, by his merits and prayers, that this our land may be ever defended from all disquietude, and thy mercy glorified in the forgiveness of our sins; and all for Jesus Christ his sake, our only Mediator and Advocate. \textit{Amen.}

\readingcitation{Epistle}{1 Peter 2:13}
\lett{B}{rethren:} Submit yourselves to every ordinance of man for the Lord's sake: whether it be to the king, as supreme; Or unto governors, as unto them that are sent by him for the punishment of evildoers, and for the praise of them that do well. For so is the will of God, that with well doing ye may put to silence the ignorance of foolish men: As free, and not using your liberty for a cloke of maliciousness, but as the servants of God. Honour all men. Love the brotherhood. Fear God. Honour the king. Servants, be subject to your masters with all fear; not only to the good and gentle, but also to the froward. For this is thankworthy, if a man for conscience toward God endure grief, suffering wrongfully. For what glory is it, if, when ye be buffeted for your faults, ye shall take it patiently? but if, when ye do well, and suffer for it, ye take it patiently, this is acceptable with God. For even hereunto were ye called: because Christ also suffered for us, leaving us an example, that ye should follow his steps: Who did no sin, neither was guile found in his mouth:

\gradall{I have found David my servant: with my holy oil have I anointed him. My hand shall hold him fast; and mine arm shall strengthen him. ℣. The enemy shall not be able to do him violence: the son of wickedness shall not hurt him.}{Alleluia, alleluia. ℣. Blessed is the man that endureth temptation: for when he is tried, he shall receive a crown of life.}

\begin{rubric}
	In Septuagesimatide \& Lent, the following Tract is said instead of the Alleluia.
\end{rubric}

\tract{Thou hast given him his heart's desire: and hast not denied him the request of his lips. ℣. For thou shalt prevent him with the blessings of goodness. ℣. Thou shalt see a crown of pure gold upon his head.}

\subbysub{Sequence}
\begin{multicols}{2}
Heav'nly King, of Kings the Pastor,

Giv'r of laws, of justice master,

Ruling all by thy behest,

Unto thee to-day we render,

Praise for him, to mem'ry tender,

Nicholas, of kings the best.\\

Traitors shedding blood like water

Fill'd the land with crime and slaughter,

Law was trampl'd in the mud,

Noble churches left forsaken

And the White Rose, overtaken

By the gun, was red with blood.\\

Thus the bardic verse fulfilling

There shall be a time of killing

When the ravens shall be fed,

And a King without pollution

Midst a realm in revolution

Shall be number'd with the dead.\\

Violent men without compassion

Proudly spurned the ancient fashion

Of the sacred right divine;

From his friends by madmen riven

Was our King to slaughter driven

Stain'd with blood his Royal line.\\

Faithful son of Mother holy,

To the Church devoted solely,

He to keep her laws was fain.

He her champion ever glorious,

Was beaten still victorious,

Robb'd of life, but conqueror slain.\\

%`He nothing common did nor mean,

%Upon that memorable scene,'

`With simple and sincere faith blest,

Self-sacrificing until rest'

Against the wicked stood his face
%When on the block he laid his head;

Nor call'd the gods with vulgar spite

To vindicate his helpless right,

But went to death as to his bed.\\

Fair exchange Tsar Nich'las making
%Fair exchange King Charles was making

When, the crown immortal taking

For the earthly crown he wore,

By bullet he follow'd faster

To the realm of Christ his master,

And the Cross behind him bore.\\

Woe to pretended justiciers,

Why, like assassins, did they fear?

To efface their wicked deed?

From which men, then, were they hiding

But from God no favour finding

And their Tsar ascending high.\\

%ORIGINAL:
%Lo, the priest who shares his glory
%(Laud his name and laud his story),
%For his fellow-martyr waits,
%And the white-robed host upraising,
%Heart and voice their Saviour
%Greets him at the heavenly gates.

He by dying brought salvation

To the torn and shatter'd nation,

Life restor'd and liberty;

For the Martyr's blood was sowing

Seed from which the Church is growing

Seed of immortality.\\

Six words said before him they slew,
%Ere his death one word was spoken

%That 'Remember' was the token

He said, `You know not what you do.'

Of his coming victory.

So his blood brought life and healing,

And the Church's triumph sealing,

Never shall forgotten be. Amen.
\end{multicols}
\readingcitation{Gospel}{Matthew 21:33}
\lett{A}{t that time:} Jesus spake a parable: There was a certain householder, which planted a vineyard, and hedged it round about, and digged a winepress in it, and built a tower, and let it out to husbandmen, and went into a far country: And when the time of the fruit drew near, he sent his servants to the husbandmen, that they might receive the fruits of it. And the husbandmen took his servants, and beat one, and killed another, and stoned another. Again, he sent other servants more than the first: and they did unto them likewise. But last of all he sent unto them his son, saying, They will reverence my son. But when the husbandmen saw the son, they said among themselves, This is the heir; come, let us kill him, and let us seize on his inheritance. And they caught him, and cast him out of the vineyard, and slew him. When the lord therefore of the vineyard cometh, what will he do unto those husbandmen? They say unto him, He will miserably destroy those wicked men, and will let out his vineyard unto other husbandmen, which shall render him the fruits in their seasons.

\offertory{In the sight of the unwise he seemed to die and his departure was taken for misery; but he is in peace.}

\secret
\lett{M}{ercifully} regard, we beseech thee, almighty God, this our atoning sacrifice, and give ear to the prayer of the blessed King Nicholas thy Martyr; graciously receive it in propitiation for the sins of this thine household. Through.

\communion{The people stood up, and the rulers took counsel together; against the Lord, and against his Anointed.}

\postcommunion
\lett{M}{ay} the bread of heaven imbue us, O Lord, with the spirit of fortitude; for this was the Bread which gave blessed King Nicholas thy Martyr strength to fight victoriously for thy Church. Through.